### 2021-01-20 multivariate.Rmd
multivariate analysis rmd for Team X

---
title: "Multivariate Analysis"
author: "Scott Coffin"
date: "1/19/2021"
output: html_document
---
# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load packages
library(tidyverse) #General everything
library(RColorBrewer)
library(ggplot2) #General plotting
library(ggrepel) #For adding text labels that repel away from data points
library(calecopal) #Color palette
library(shiny) #Runs shiny
library(shinythemes) #Shiny theme for the page
library(shinyWidgets) #Widgets
library(scales) #SSD - Use the percent format
library(reshape2) #Overview tab - melts bars together
library(ssdtools) #SSD package
library(DT) #Build HTML data tables
#library(plotly) #Make plots interactive
library(viridis) #Colors
library(scales) #To use "percent" function
library(shinyjs) #Exploration tab - reset button
library(tigerstats) #turns things into percents
library(ggbeeswarm) #plot all points
library(fitdistrplus) #alt SSD 
library(RColorBrewer) #colors
library(pheatmap) #pretty heat maps
library(rpart)  #for trees
#library(rattle)    # Fancy tree plot This is a difficult library to install (https://gist.github.com/zhiyzuo/a489ffdcc5da87f28f8589a55aa206dd) 
library(rpart.plot)             # Enhanced tree plots
library(RColorBrewer)       # Color selection for fancy tree plot
library(party)                  # Alternative decision tree algorithm
# library(partykit)               # Convert rpart object to BinaryTree
library(pROC)   #for ROC curves

library(ISLR)  #for the Carseat Data
```
## Data import
```{r data import}
# Load finalized dataset.
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)

#### Introduction Setup ####

# All text inputs below.

#### Overview AO Setup ####

#Final_effect_dataset <- read_csv("Final_effect_dataset.csv")%>%
  #mutate(plot_f = case_when(
    #plot_f == "Polymer" ~ "Polymer",
    #plot_f == "Size" ~ "Size",
    #plot_f == "Shape" ~ "Shape",
    #plot_f == "Organism" ~ "Organism",
    #plot_f == "Lvl1" ~ "Endpoint Category",
    #plot_f == "Life.stage" ~ "Life Stage",
    #plot_f == "Invivo.invivo" ~ "In Vivo or In Vitro",
    #plot_f == "Exposure.route" ~ "Exposure Route"))%>%
  #mutate(plot_f = factor(plot_f))%>%
  #mutate(logEndpoints = log(Endpoints))%>%
  #rename(Percent = Freq)

polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
polyf<-as.data.frame(polydf)%>% #Makes data frame 
  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
  mutate(polymer = case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurathane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid"))%>%
  mutate_if(is.numeric, round,0) #rounds percents 
Endpoints<-xtabs(~polymer +effect ,aoc) #Pulls all study obs. for polymer from dataset
polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
  rename(Endpoints='Freq.1')%>% #renames column
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
sizef<-as.data.frame(sizedf)%>%
  filter(effect %in% c("Y","N"))%>%
  mutate(size.category = case_when(
    size.category == 1 ~ "<1µm",
    size.category == 2 ~ "1µm < 10µm",
    size.category == 3 ~ "10µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "unavailable"))%>%
  rename(Type = "size.category")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Size")
study_s<-xtabs(~size.category +effect ,aoc)
sizefinal<- data.frame(cbind(sizef, study_s))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='size.category')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

shapedf<-rowPerc(xtabs(~shape + effect, aoc))
shapef<-as.data.frame(shapedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type="shape")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Shape")%>%
  mutate(Type = case_when(
    Type == "cube" ~ "Cube",
    Type == "sphere" ~ "Sphere",
    Type == "fragment" ~ "Fragment",
    Type == "fiber" ~ "Fiber"))
study_sh<-xtabs(~shape + effect,aoc)
shapefinal<- data.frame(cbind(shapef, study_sh))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='shape')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

taxdf<-rowPerc(xtabs(~organism.group +effect, aoc))
taxf<-as.data.frame(taxdf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "organism.group")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Organism")
study_t<-xtabs(~organism.group +effect,aoc)
taxfinal<- data.frame(cbind(taxf, study_t))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='organism.group')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

lvl1df<-rowPerc(xtabs(~lvl1 +effect, aoc))
lvl1f<-as.data.frame(lvl1df)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "lvl1")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Lvl1")%>%
  mutate(Type = case_when(
    Type == "alimentary.excretory" ~ "Alimentary, Excretory",
    Type == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    Type == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    Type == "community" ~ "Community",
    Type == "fitness" ~ "Fitness",
    Type == "immune" ~ "Immune",
    Type == "metabolism" ~ "Metabolism",
    Type == "microbiome" ~ "Microbiome",
    Type == "stress" ~ "Stress")) 
study_l<-xtabs(~lvl1 +effect,aoc)
lvl1final<- data.frame(cbind(lvl1f, study_l))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='lvl1')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

lifedf<-rowPerc(xtabs(~life.stage +effect, aoc))
lifef<-as.data.frame(lifedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "life.stage")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Life.stage")
studyli<-xtabs(~life.stage +effect ,aoc)
lifefinal<- data.frame(cbind(lifef, studyli))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='life.stage')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

vivodf<-rowPerc(xtabs(~invitro.invivo +effect, aoc))
vivof<-as.data.frame(vivodf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "invitro.invivo")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Invivo.invivo")%>%
  mutate(Type = case_when(
    Type=="invivo"~"In Vivo",
    Type=="invitro"~"In Vitro"))
study_v<-xtabs(~invitro.invivo +effect,aoc)
vivofinal<- data.frame(cbind(vivof, study_v))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='invitro.invivo')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

routedf<-rowPerc(xtabs(~exposure.route +effect, aoc))
routef<-as.data.frame(routedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "exposure.route")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Exposure.route")%>%
  mutate(Type = case_when(
    Type == "coparental.exposure" ~"Co-Parental Exposure",
    Type == "paternal.exposure" ~ "Paternal Exposure",
    Type == "maternal.exposure" ~ "Maternal Exposure",
    Type == "food" ~ "Food",
    Type == "water" ~ "Water",
    Type == "sediment" ~ "Sediment",
    Type == "media" ~ "Media"))
study_r<-xtabs(~exposure.route +effect,aoc)
routefinal<- data.frame(cbind(routef, study_r))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='exposure.route')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

#### Exploration AO Setup ####

# Master dataset for scatterplots - for Heili's tab.
aoc_v1 <- aoc %>% # start with original dataset
   # full dataset filters.
  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
    effect == "N" ~ "No"),
    levels = c("No", "Yes"))) %>%
  # removing NAs to make data set nicer
  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported"))

aoc_setup <- aoc_v1 %>% # start with original dataset
  mutate(size_f = factor(case_when(
    size.category == 1 ~ "1nm < 100nm",
    size.category == 2 ~ "100nm < 1µm",
    size.category == 3 ~ "1µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
  # shape category data tidying.
  mutate(shape_f = factor(case_when(
    shape == "fiber" ~ "Fiber",
    shape == "fragment" ~ "Fragment",
    shape == "sphere" ~ "Sphere",
    shape == "Not Reported" ~ "Not Reported"),
    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% # order our different shapes.
  # polymer category data tidying.
  mutate(poly_f = factor(case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurathane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid",
    polymer == "Not Reported" ~ "Not Reported"))) %>%
  # taxonomic category data tidying.
  mutate(org_f = factor(organism.group, levels = c("Algae", "Annelida", "Bacterium", "Cnidaria", "Crustacea",
                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% # order our different organisms.
  mutate(lvl1_f = factor(case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
    lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    lvl1 == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    lvl1 == "community" ~ "Community",
    lvl1 == "fitness" ~ "Fitness",
    lvl1 == "immune" ~ "Immune",
    lvl1 == "metabolism" ~ "Metabolism",
    lvl1 == "microbiome" ~ "Microbiome",
    lvl1 == "stress" ~ "Stress"))) %>% # creates new column with nicer names.
  # Level 2 Data tidying
  mutate(lvl2_f = factor(case_when(lvl2 == "abundance"~"Abundance",
    lvl2 == "actinobacteria" ~ "Actinobacteria",
    lvl2 == "aggressivity"~"Agressivity",
    lvl2 == "ammonia.excretion" ~ "Ammonia Excretion",
    lvl2 == "bacteroidetes"~ "Bacteriodetes",
    lvl2 == "blood"~"Blood",
    lvl2 == "body.condition"~"Body Condition",
    lvl2 == "boldness"~"Boldness",
    lvl2 == "brain.histo"~"Brain Histological Abnormalities",
    lvl2 == "burrowing"~"Burrowing",
    lvl2 == "carb.metabolism"~"Carb Metabolism",
    lvl2 == "chemokines.cytokines"~"Chemokines",
    lvl2 == "circulatory"~"Circulatory",
    lvl2 == "detoxification"~"Detoxification",
    lvl2 == "development"~"Development",
    lvl2 == "digestion"~"Digestion",
    lvl2 == "digestive.enzymes"~"Digestive Enzymes",
    lvl2 == "digestive.tract.histo"~"Digestive Tract Histological Abnormalities",
    lvl2 == "diversity"~ "Diversity",
    lvl2 == "feeding"~ "Feeding",
    lvl2 == "firmicutes"~ "Firmicutes",
    lvl2 == "gall.bladder.histo" ~ "Gall Bladder Histological Abnormalities",
    lvl2 == "gen.metabolism"~ "General Metabolism",
    lvl2 == "gill.histo"~ "Gill Histological Abnormalities",
    lvl2 == "gonad.histo"~"Gonad Histological Abnormalities",
    lvl2 == "growth"~ "Growth",
    lvl2 == "immune.cells"~"Immune Cells",
    lvl2 == "immune.other"~"Immune Other ",
    lvl2 == "intestinal.permeability"~"Intestinal Permeability",
    lvl2 == "kidney.histo"~"Kidney Histological abnormalities",
    lvl2 == "lipid.metabolism"~"Lipid Metabolism",
    lvl2 == "liver.histo"~"Liver Histological Abnormalities",
    lvl2 == "liver.kidney.products" ~ "Liver and Kidney Products",
    lvl2 == "locomotion"~"Locomotion",
    lvl2 == "mortality"~"Mortality",
    lvl2 == "nervous.system"~"Nervous System",
    lvl2 == "oxidative.stress"~"Oxidative Stress",
    lvl2 == "photosynthesis"~ "Photosynthesis",
    lvl2 == "proteobacteria"~"Protebacteria",
    lvl2 == "reproduction"~"Reproduction",
    lvl2 == "respiration"~"Respiration",
    lvl2 == "sexhormones"~"Sex Hormones",
    lvl2 == "shoaling"~"Shoaling",
    lvl2 == "stress"~"Stress",
    lvl2 == "vision.system"~"Vision System"))) %>% #Renames for widget
  mutate(bio_f = factor(case_when(bio.org == "cell"~"Cell", #Bio Org Data Tidying
    bio.org == "organism"~"Organism",
    bio.org == "population"~ "Population",
    bio.org == "subcell"~"Subcell",
    bio.org == "tissue" ~ "Tissue")))%>%
  mutate(vivo_f = factor(case_when(invitro.invivo == "invivo"~"In Vivo",
    invitro.invivo == "invitro"~"In Vitro")))%>% ##Renames for widget (Not using a widget right now, but saving for human health database)
  mutate(life_f = factor(case_when(life.stage == "Early"~"Early",
    life.stage == "Juvenile"~"Juvenile",
    life.stage == "Adult"~"Adult",
    life.stage == "Not Reported"~"Not Reported")))%>% #Renames for widget
  mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
    environment == "Marine" ~ "Marine",
    environment == "Terrestrial" ~ "Terrestrial"))) %>%
  mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported))

#### SSD AO Setup ####

# Master dataset for SSDs
aoc_z <- aoc_setup %>% # start with Heili's altered dataset (no filtration for terrestrial data)
  # environment category data tidying.
  mutate(environment.noNA = replace_na(environment, "Not Reported")) %>% # replaces NA to better relabel.
  mutate(env_f = factor(environment.noNA, levels = c("Marine", "Freshwater", "Terrestrial", "Not Reported")))

# final cleanup and factoring  
aoc_z$Species <- as.factor(paste(aoc_z$genus,aoc_z$species)) #must make value 'Species" (uppercase)
aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
```
```{r}
# subset data to selected variables

multiVar <- aoc_z %>% dplyr::select(#doi, size.category, 
                                    size_f,
                                    size.length.um.used.for.conversions, 
                                    shape, 
                                    polymer, 
                                    particle.volume.um3, 
                                    density.mg.um.3, 
                                    organism.group,
                                    environment, 
                                    bio.org, #biological level of organization
                                    #af.time, #assessment factor based on exposure time
                                    treatments, #number of doses (no including control)
                                    effect, #yes no
                                    exposure.duration.d, 
                                    exposure.route, #Factor
                                    organism.group, #factor
                                    media.temp, #numeric
                                    lvl1_f, #endpoints
                                    lvl2_f, #endpoints
                                   # lvl3, 
                                     dose.mg.L.master, 
                                    sex, #factor
                                    media.ph, #numeric
                                    media.sal.ppt, #numeric
                                    dose.particles.mL.master,
                                   effect.metric, #NOEC LOEC
                                    functional.group, #factor
                                    charge, #positive or negatibe
                                    zetapotential.mV) %>%  # numeric   
  filter(!size_f == "Not Reported")  #take out not reported

#recode variables
multiVar <- multiVar %>% mutate(effect_10 = case_when(
    effect == "Y" ~ 1,
    effect == "N" ~ 0
  ))# %>% 
  #mutate_all(is.character, ~as.factor())

```

# Analysis
## Data Exploration
### Completeness vby Size
```{r completeness heatmap}
CompletenessSummary_size <- multiVar %>%
  group_by(size_f) %>% 
     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
  mutate(across(is.numeric,~round(., 2))) %>% 
  mutate(across(is.numeric, ~100 *(1 -.)))
CompletenessSummary_size
```

```{r}
require()
#convert to matrix and transpose
transposed_size <- as.data.frame(t(as.matrix(CompletenessSummary_size[2:25]))) %>% #1 is category, 2-6 are variables
  arrange('1nm < 100nm')
#reassign column names
colnames(transposed_size) <- c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm")
#transposedTag
#format as matrix
MissingMatrix_size <- data.matrix(transposed_size)
#build heatmap
pheatmap(MissingMatrix_size,
                           main = "Data Completeness by Size Bin", #title
                           fontsize = 12,
                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
                           display_numbers = TRUE,
                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)

```
### Completeness by Environment
```{r completeness heatmap}
CompletenessSummary_environment <- multiVar %>%
  filter(!environment == "NA") %>% 
  group_by(environment) %>% 
     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
  mutate(across(is.numeric,~round(., 2))) %>% 
  mutate(across(is.numeric, ~100 *(1 -.)))
CompletenessSummary_environment
```

```{r}
require()
#convert to matrix and transpose
transposed_environment <- as.data.frame(t(as.matrix(CompletenessSummary_environment[2:25])))
#reassign column names
colnames(transposed_environment) <- c("Freshwater", "Marine", "Terrestrial")
#transposedTag
#format as matrix
MissingMatrix_environment <- data.matrix(transposed_environment)
#build heatmap
pheatmap(MissingMatrix_environment,
                           main = "Data Completeness by environment", #title
                           fontsize = 12,
                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
                           display_numbers = TRUE,
                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)

```

## Stepwise Regression

```{r}
## Estimate an OLS Regression
fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master, 
             na.action = na.omit, 
             data = multiVar)
summary(fitols)
```

```{r}
require(reshape2)
multiVar %>% 
  dplyr::select(size.length.um.used.for.conversions, particle.volume.um3, exposure.duration.d ,media.temp, dose.particles.mL.master) %>% 
  melt() %>%  #convert wide to long
   mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  ggplot(aes(x = value)) + 
  stat_density() + 
  facet_wrap(~variable, scales = "free")
```

<!-- # ```{r} -->
<!-- # # cdplot(as.factor(effect) ~ particle.volume.um3, data = multiVar, -->
<!-- #        main = "Conditional Density Plot of ", xlab = "Dose", ylab = "Effect") -->
<!-- # ``` -->

## PCA

## Random Forest
### Recursive Partitioning And Regression Trees

The rpart algorithm works by splitting the dataset recursively, which means that the subsets that arise from a split are further split until a predetermined termination criterion is reached.  At each step, the split is made based on the independent variable that results in the largest possible reduction in heterogeneity of the dependent (predicted) variable.

It is important to note that the algorithm works by making the best possible choice at each particular stage, without any consideration of whether those choices remain optimal in future stages. That is, the algorithm makes a locally optimal decision at each stage. It is thus quite possible that such a choice at one stage turns out to be sub-optimal in the overall scheme of things.  In other words,  the algorithm does not find a globally optimal tree.
```{r}
#trim data so effect is always known
multiVar_sub <- multiVar %>% 
  drop_na(effect_10)

# Split data into training and test sets
set.seed(42)
multiVar_sub[,"train"] <- ifelse(runif(nrow(multiVar_sub)) < 0.8, 1, 0)
# Separate trainig and test sets
trainSet <- multiVar_sub[multiVar_sub$train==1,]
testSet <- multiVar_sub[multiVar_sub$train==0,]
#get column index of train flag
trainColNum <- grep("train", names(trainSet))
# Remove train flag column from train and test sets
trainSet <- trainSet[, -trainColNum]
testSet <- testSet[, -trainColNum]
```
Make a classification tree.

```{r, echo = FALSE}
#get column index of predicted variable in dataset
typeColNum <- grep("effect",names(multiVar_sub))

#build tree
require(rpart)
set.seed(15097)
t1 <- rpart(effect ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,
            method = "class", #classification because response is discrete
            control = rpart.control(minbucket = 20, cp=0.008), #requires that there be at least 19 cases (responded + nonrespondents) in the final grouping of variable values of the terminal node of the tre..
            data = trainSet)
print(t1, digits=4)
```

Plot an interpretable tree.
```{r, echo = FALSE}
require(rpart.plot)
cols <- ifelse(t1$frame$yval == 1, "gray50", "black")
prp(t1, main="Tree for Effect",
    extra=106, # display prob of survival and percent of obs
    nn=TRUE, # display node numbers
    fallen.leaves=TRUE, # put leaves on the bottom of page
    branch=.5, # change angle of branch lines
    faclen=0, # do not abbreviate factor levels
    trace=1, # print automatically calculated cex
    shadow.col="gray", # shadows under the leaves
    branch.lty=1, # draw branches using solid lines
    branch.type=5, # branch lines width = weight(frame$wt), no. of cases here
    split.cex=1.2, # make split text larger than node text
    split.prefix="is ", # put "is " before split text
    split.suffix="?", # put "?" after split text
    col=cols, border.col=cols, # cols[2] if survived
    split.box.col="lightgray", # lightgray split boxes (default is white)
    split.border.col="darkgray", # darkgray border on split boxes
    split.round=0.5) # round the split box corners a tad
```
Next we see how good the model is by seeing how it fares against the test data.
```{r}
t1_predict <- predict(t1, newdata = testSet[,-typeColNum],
                      type="class")
mean(t1_predict==testSet$effect)
# [1] 0.7094088
#confusion matrix
table(pred=t1_predict,true=testSet$effect)
```
```{r}
par(mfrow=c(1,2)) # two plots on one page
#plot approximate R-squared and relative error for different splits (2 plots). labels are only appropriate for the "anova" method.
rsq.rpart(t1)
```

Next, we prune the tree using the cost complexity criterion. Basically, the intent is to see if a shallower subtree can give us comparable results. If so, we’d be better of choosing the shallower tree because it reduces the likelihood of overfitting.

As described earlier, we choose the appropriate pruning parameter (aka cost-complexity parameter) \alpha by picking the value that results in the lowest prediction error. Note that all relevant computations have already been carried out by R when we built the original tree (the call to rpart in the code above). All that remains now is to pick the value of \alpha:

#### Pruning

```{r}
#cost-complexity pruning
printcp(t1)
```
It is clear from the above, that the lowest cross-validation error (xerror in the table) occurs for \alpha =0.008 (this is CP in the table above).   One can find CP programatically like so:
```{r}
# get index of CP with lowest xerror
opt <- which.min(t1$cptable[,"xerror"])
#get its value
cp <- t1$cptable[opt, "CP"]
```
Next, we prune the tree based on this value of CP:

```{r}
# prune the tree
pt1 <- prune(t1,cp)
#pt1<- prune(t1, cp= t1$cptable[which.min(t1$cptable[,"xerror"]),"CP"])

# plot the pruned tree
plot(pt1, uniform=TRUE,
   main="Pruned Classification Tree");text(pt1, use.n=TRUE, all=TRUE, cex=.8)

#post(pfit, file = "c:/ptree.ps",
 #  title = "Pruned Classification Tree for Kyphosis")
```
```{r}
#find proportion of correct predictions using test set
t1_pruned_predict <- predict(pt1, testSet, type="class")
mean(t1_pruned_predict == testSet$effect)
#
```
This is not an improvement over an unprunend tree.. We need to check that this holds up for different training and test sets. This is easily done by creating multiple random partitions of the dataset and checking the efficacy of pruning for each. To do this efficiently, I’ll create a function that takes the training fraction, number of runs (partitions) and the name of the dataset as inputs and outputs the proportion of correct predictions for each run. It also optionally prunes the tree.

```{r}
#function to do multiple runs
multiple_runs_classification <- function(train_fraction,n,dataset,prune_tree=FALSE){
fraction_correct <- rep(NA,n)
set.seed(42)
for (i in 1:n){
  dataset[,"train"] <- ifelse(runif(nrow(dataset))<0.8,1,0)
  trainColNum <- grep("train",names(dataset))
  typeColNum <- grep("effect",names(dataset))
  trainSet <- dataset[dataset$train==1,-trainColNum]
  testSet <- dataset[dataset$train==0,-trainColNum]
  rpart_model <- rpart(effect~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,data = trainSet, method="class")
if(prune_tree==FALSE) {
  rpart_test_predict <- predict(rpart_model,testSet[,-typeColNum],type="class")
  fraction_correct[i] <- mean(rpart_test_predict==testSet$effect)
  }else{
    opt <- which.min(rpart_model$cptable[,"xerror"])
    cp <- rpart_model$cptable[opt, "CP"]
    pruned_model <- prune(rpart_model,cp)
    rpart_pruned_predict <- predict(pruned_model,testSet[,-typeColNum],type="class")
    fraction_correct[i] <- mean(rpart_pruned_predict == testSet$effect)
  }
  }
return(fraction_correct)
}
```

Note that in the above,  I have set the default value of the prune_tree to FALSE, so the function will execute the first branch of the if statement unless the default is overridden.

OK, so let’s do 50 runs with and without pruning, and check the mean and variance of the results for both sets of runs.
```{r}
#50 runs, no pruning
unpruned_set <- multiple_runs_classification(0.8,50,multiVar_sub)
mean(unpruned_set)
#[1] 0.7261882
sd(unpruned_set)
#[1] 0.01554347
#50 runs, with pruning
pruned_set <- multiple_runs_classification(0.8,50,multiVar_sub,prune_tree=TRUE)
mean(pruned_set)
#[1] 0.7261875
sd(pruned_set)
#[1] 0.01552285
```

### CForest
```{r run model}
require(party)
crf <- cforest(as.factor(effect_10) ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d +
                 media.temp + dose.particles.mL.master,
               controls = cforest_control(ntree = 500,
                                          mincriterion = qnorm(0.8), 
                                          trace = TRUE), # adds project bar because it's very slow
               data = multiVar_sub)

crf
```

```{r fit results}
train2 <- multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master)
train3 <-multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master,
                                  effect_10)

fitted <- predict(crf, train2, OOB = TRUE, type ="response")
#rpart.prob <- predict(t1, newdata=imputedSmalls.requested.voluntary,type="prob")

misClasificError <- mean(fitted != train3$effect_10)
print(paste('Training Accuracy', 1 - misClasificError))
```

```{r}
print(crf)
```

```{r}
#Obtaining predicted probabilites for Test data
tree.probs=predict(crf,
                 newdata = train2,
                 type="prob")

#Calculate ROC curve
rocCurve.tree <- roc(train3$effect_10,tree.probs[,'1'])
#plot the ROC curve
plot(rocCurve.tree,col=c(4))
```

Alternative ROC Curve
```{r}
require(pROC)
predicted <- predict(crf, train2, OOB=TRUE, type= "response")
auc(as.numeric(train3$effect_10), as.numeric(predicted))
#Calculate ROC curve
rocCurve.tree <- roc(train3$effect_10,as.numeric(predicted))
#plot the ROC curve
plot(rocCurve.tree,col=c(4))
```
```{r}
#plot reciever operating curve
perf <- performance(preds,"tpr","fpr")
```

```{r}
# compute in-sample results
caret::confusionMatrix(fitted,as.factor(train3$effect_10))
```

```{r}
#plot feature importance
cforestImpPlot <- function(x) {
  cforest_importance <<- v <- varimp(x)
  dotchart(v[order(v)])
}

importancePlot <- cforestImpPlot(crf)
importancePlot
```

```{r}
crf.prob <- matrix(unlist(crfsrvy.prob), ncol=2, byrow=TRUE)
apply(crf.prob,2,mean)
#[1] 0.2524514 0.7475486
tab <- round(cbind(by(rpart.prob[,2], INDICES=t1$where, mean), by(crf.prob[,2], INDICES=t1$where, mean)), 4)
colnames(tab) <- c("rpart", "cforest")
kable(tab)
```

### 2021-01-20 Concentration data/Concentration data.R

@@ -517,7 +517,7 @@ ggsave(ECDF.Location.Type.Season,
        scale = 2,
        dpi = 500)
 
-#modelling
+###### --modelling ####
 sample_dists <- ssd_fit_dists(samplesSFEI, #data frame
                            left = "Conc", #string of the column in data with the concentrations
                            # right = left, #string of the column with the right concentration values. If different from left, then the data are considerd to be censored
@@ -542,7 +542,7 @@ datatable(sample_gof,
 #there are multiple fitting distributions, so check which fits best
 sample_gof <- ssd_gof(sample_dists)
 sample_gof[order(sample_gof$delta), ] #orders by delta. Use the aicc (Akaike's Information Criterion corrected for sample size) for model selection 
-write.csv(sample_gof,"Concentration data/sample_gof.csv")
+#write.csv(sample_gof,"Concentration data/sample_gof.csv")
 set.seed(99)
 sample_pred <- predict(sample_dists,
                                 average = TRUE,
@@ -591,12 +591,14 @@ ECDF_model_occurrence_white <- ECDF_model_occurrence +
   geom_vline(xintercept = 75.6, linetype = 'dashed', color = 'red') +
   geom_vline(xintercept = 11, linetype = 'dashed', color = 'red') +
   geom_vline(xintercept = 521, linetype = 'dashed', color = 	'red') +
+  geom_vline(xintercept = 71.5, linetype = 'dotdash', color = "#50C878", size = 2) +
+  geom_text(label = "Limit of Detection", color = "#50C878", x = 45, y = 0.08)+
   geom_text(label = "5% HC: 95% LCL", color = 'red', x = 15, y = 0)+
   geom_text(label = "5% hazard concentration", color = 'red', x = 110, y = 0.03)+
   geom_text(label = "5% HC: 95% UCL", color = 'red', x = 400, y = 0)+
   geom_text(x = 110, y = 0, label = "75.6 particles/L", color = 'red') +  #label for hazard conc
-  geom_hline(yintercept = 0.925, linetype = 'twodash', color = "#A2A475") +
-  geom_text(label = "92.5% samples below 5% HC Mean", x = 4.5, y = 0.94, color = "#A2A475") +
+  geom_hline(yintercept = 0.925, linetype = 'twodash', color ='red') +
+  geom_text(label = "92.5% samples below 5% HC Mean", x = 4.5, y = 0.94, color = 'red') +
   theme_bw() +
[…]

### 2021-01-21 app.R
added um^3/mL option and added themes and colors to SSD plot

@@ -6,9 +6,6 @@
 
 # Anything that should only happen ONCE should be placed in this setup section, prior to the actual shiny structure.
 
-
- 
-
 # Load packages
 library(tidyverse) #General everything
 library(RColorBrewer)
@@ -29,6 +26,8 @@ library(shinyjs) #Exploration tab - reset button
 library(tigerstats) #turns things into percents
 library(ggbeeswarm) #plot all points
 library(fitdistrplus) #alt SSD 
+library(ggdark) #dark mode ggplot
+library(ggsci) #color palettes
 
 # Load finalized dataset.
 aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
@@ -321,7 +320,8 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
     environment == "Marine" ~ "Marine",
     environment == "Terrestrial" ~ "Terrestrial"))) %>%
   mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
-  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported))
+  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
+  mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) #calculate volume/mL
     
 
 #### SSD AO Setup ####
@@ -623,8 +623,8 @@ column(width = 12,
                        #    multiple = TRUE))
                     
                     radioButtons(inputId = "dose_check", # dosing units
-                                 label = "Particles/mL or mg/L:",
-                                 choices = c("Particles/mL", "mg/L"),
+                                 label = "Particles/mL, mg/L, or um3/mL:",
+                                 choices = c("Particles/mL", "mg/L", "um3/mL"),
                                  selected = "mg/L"),
                     
                     p("Concentrations may be reported in mass/volume or particle #/volume (or sometimes both). Using methods described in", a(href ="https://pubs.acs.org/doi/10.1021/acs.est.0c02982", "Koelmans et. al (2020)"), " units have been converted."),
@@ -758,8 +758,8 @@ column(width = 12,
                            ),#close out column
                     
                     radioButtons(inputId = "particle_mass_check_ssd", # organism checklist
[…]

### 2021-01-22 app.R
filter by effect metric!

@@ -321,6 +321,7 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
     environment == "Terrestrial" ~ "Terrestrial"))) %>%
   mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
   mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
+  mutate(effect.metric = factor(effect.metric)) %>% 
   mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) #calculate volume/mL
     
 
@@ -755,6 +756,11 @@ column(width = 12,
                            #Polymer widget
                            column(width = 4,
                                   htmlOutput("polySelection")),# polymer selection based on other inputs
+                           column(width = 4,
+                                  radioButtons(inputId = "effect.metric_rad_ssd", # organism checklist
+                                              label = "Effect Metric:",
+                                              choices = levels(aoc_z$effect.metric),
+                                              selected = "HONEC")), # allows for multiple inputs
                            ),#close out column
                     
                     radioButtons(inputId = "particle_mass_check_ssd", # organism checklist
@@ -1616,6 +1622,7 @@ server <- function(input, output) {
     lvl1_c_ssd <- input$lvl1_check_ssd #assign broad endpoints
     lvl2_c_ssd <- input$lvl2_check_ssd #assign specific endpoints
     poly_c_ssd <- input$poly_check_ssd #assign polymers
+    
    
     Reported_Converted_rad <- input$Reported_Converted_rad #use nominal or calculated exposure concentrations. Options are TRUE (calculated) or FALSE (reported)
     particle_mass_check_ssd <- input$particle_mass_check_ssd #rename variable
@@ -1633,6 +1640,7 @@ server <- function(input, output) {
     if(Reported_Converted_rad == "all" & particle_mass_check_ssd == "mg/L"){
       aoc_z <- aoc_z %>% 
         mutate(dose_new = dose.mg.L.master)
+      
       #repeat for particles
     }
     if(Reported_Converted_rad == "reported" & particle_mass_check_ssd == "Particles/mL"){
[…]

### 2021-01-22 app.R
text for assessment factor (but no buttons)

@@ -321,8 +321,9 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
     environment == "Terrestrial" ~ "Terrestrial"))) %>%
   mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
   mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
-  mutate(effect.metric = factor(effect.metric)) %>% 
-  mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) #calculate volume/mL
+  mutate(effect.metric = factor(effect.metric)) %>% #factorize
+  mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master)   #calculate volume/mL
+  
     
 
 #### SSD AO Setup ####
@@ -715,6 +716,8 @@ column(width = 12,
                     h3("Species Sensitivity Distribution", align = "center"),
                     p("Species sensitivity distributions (SSDs) are cumulative probability distributions that estimate the percent of species affected by a given concentration of exposure using Maximum Likelihood and model averaging. A useful metric often used for setting risk-based thresholds is the concentration that affects 5% of the species, and is reffered to as the 5% Hazard Concentration (HC). For more information on SSDs, refer to", a(href = "https://bit.ly/2Hy4q10", 'Posthuma, Suter II, and Traas (2001).')),
                     br(), # line break
+                    p("The choice of effect metrics (e.g. NOEC, LOEC, HONEC, ECXX and LCXX) should be carefully considered. Assessment factors are available for converting acute exposures to chronic exposure and estimating NOECs from other effect metrics (e.g. LOEC's), according to the methods described in ", a(href = "https://setac.onlinelibrary.wiley.com/doi/epdf/10.1002/ieam.4214", 'Wigger et al (2019).')),
+                    br(),
                     p("Use the options below to filter the toxicity thresholds dataset. Once complete, hit the 'submit' button"),
                     
                     # widget 1
@@ -768,7 +771,7 @@ column(width = 12,
                                        choices = c("Particles/mL", "mg/L", "um3/mL"),
                                        selected = "mg/L"),
                     
-                                     p("Concentrations may be reported in mass/volume or particle #/volume (or sometimes both). Using methods described in", a(href ="https://pubs.acs.org/doi/10.1021/acs.est.0c02982", "Koelmans et. al (2020)"), " units have been converted."),
[…]

### 2021-01-22 app.R
commented out expand limits on ssd plot

@@ -2023,7 +2023,7 @@ output$downloadSsdPlot <- downloadHandler(
       geom_point(data = aoc_ssd,aes(x = Conc, y =frac, color = Group)) + 
       geom_text_repel(data = aoc_ssd, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 4, segment.alpha = 0.5) + #species labels
       scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
-      expand_limits(x = c(0.000000001, 100000)) + #ensure species labels fit
+      #expand_limits(x = c(0.000000001, 100000)) + #ensure species labels fit
       xlab(particle_mass_check_ssd)+
       coord_trans(x = "log10") +
       scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+

### 2021-01-22 app.R
ggplot mods

@@ -1934,7 +1934,7 @@ output$downloadSsdPlot <- downloadHandler(
   },
   content = function(file) {
     device <- function(..., width, height) {
-      grDevices::png(..., width = 12, height = 16, res = 600, units = "in")
+      grDevices::png(..., width = 14, height = 16, res = 300, units = "in")
     }
     ggsave(file, plot = ssd_ggplot(), device = device)
   })
@@ -2017,6 +2017,7 @@ output$downloadSsdPlot <- downloadHandler(
     
     aochc$est_format <-format(aochc$est, digits = 3, scientific = TRUE)
     
+    #build ggplot
     ggplot(aoc_pred(),aes_string(x = "est")) +
       geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "grey") +
       geom_line(aes_string(y = "percent/100"), color = "gray") +
@@ -2026,15 +2027,16 @@ output$downloadSsdPlot <- downloadHandler(
       #expand_limits(x = c(0.000000001, 100000)) + #ensure species labels fit
       xlab(particle_mass_check_ssd)+
       coord_trans(x = "log10") +
-      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
+      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
+                         labels = trans_format("log10", scales::math_format(10^.x))) + #comma_signif)+
       geom_segment(data = aochc,aes(x = est, y = percent/100, xend = est, yend = est), linetype = 'dashed', color = "red", size = 1) + #hazard conc line vertical
       geom_segment(data = aochc,aes(x = lcl, y = percent/100, xend = est, yend = percent/100), linetype = 'dashed', color = "red", size = 1) + #hazard conc line horizontal
-      geom_text(data = aochc, aes(x = est, y = -0.09, label = paste0(percent, "% Hazard Confidence Level")), color = "red", size = 5) + #label for hazard conc
-      geom_text(data = aochc, aes(x = est, y = -0.05, label = est_format), color = "red", size = 5) + #label for hazard conc
+      geom_text(data = aochc, aes(x = est, y = 0.15, label = paste0(percent, "% Hazard Confidence Level")), color = "red", size = 5) + #label for hazard conc
[…]

### 2021-02-01 app.R
updated logo

@@ -2211,7 +2211,7 @@ output$downloadSsdPlot <- downloadHandler(
   },
   content = function(file) {
     device <- function(..., width, height) {
-      grDevices::png(..., width = 14, height = 16, res = 300, units = "in")
+      grDevices::png(..., width = 12, height = 12, res = 250, units = "in")
     }
     ggsave(file, plot = ssd_ggplot(), device = device)
   })
@@ -2260,8 +2260,8 @@ output$downloadSsdPlot <- downloadHandler(
      
      #Theme type
      theme.type<-switch(input$theme.type,
-                       "light" 	= theme_gray(),
-                       "dark" = dark_theme_bw()) 
+                       "light" 	= theme_gray(base_size = 20),
+                       "dark" = dark_theme_bw(base_size = 20)) 
      #color selection
      fill.type <- switch(input$color.type,
                          "viridis" = scale_fill_viridis(discrete = TRUE),

### 2021-02-02 app.R
added lvl2 endpoint in datatable in ssd and commented out dark mode global theme switch

@@ -28,6 +28,10 @@ library(ggbeeswarm) #plot all points
 library(fitdistrplus) #alt SSD 
 library(ggdark) #dark mode ggplot
 library(ggsci) #color palettes
+#library(bslib) #better themes
+#library(thematic) #complete control over themes (including plots)
+
+
 
 # Load finalized dataset.
 aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
@@ -345,8 +349,17 @@ aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
 aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
 
 #### User Interface ####
-
-ui <- fluidPage(theme = shinytheme("flatly"),  
+#custom themes (EXPERIMENTAL, COMMENT OUT UNTIL SHINY UPDATES BUGS)
+# light <- bs_theme()
+# dark <- bs_theme(bg = "black", fg = "white", primary = "purple")
+# build UI
+ui <- fluidPage(theme = shinytheme("flatly"), #light, 
+                
+                  # div(class = "custom-control custom-switch", 
+                  #   tags$input(id = "dark_mode", type = "checkbox", class = "custom-control-input",
+                  #     onclick = HTML("Shiny.setInputValue('dark_mode', document.getElementById('dark_mode').value);")),
+                  #   tags$label("Dark mode", `for` = "dark_mode", class = "custom-control-label")
+                  #   ),  
   
   # App title
   titlePanel(title=div(img(src = "main_logo.png", width = "10%", height = "10%"), "Toxicity of Microplastics Explorer: Aquatic Organisms")),
@@ -968,7 +981,12 @@ tabPanel("6: Contact",
      
 
 #### Server ####
-server <- function(input, output) {
+server <- function (input, output){  #dark mode: #(input, output, session) {
+
+  # # Theme Switch (comment out until Shiny updates)
+  # observe({session$setCurrentTheme(
+  #   if (isTRUE(input$dark_mode)) dark else light)
+  #   })
 
 #### Introduction S ####
 
@@ -1876,7 +1894,8 @@ server <- function(input, output) {
                 selected = levels(aoc_new$poly_f_new),
                 options = list(`actions-box` = TRUE),
                 multiple = TRUE)}) 
-  
+
+    
[…]

### 2021-02-03 multivariate.Rmd
machine learning

@@
 library(pROC)   #for ROC curves
-
+library(uwot) #umap
 library(ISLR)  #for the Carseat Data
@@
                                     charge, #positive or negatibe
-                                    zetapotential.mV) %>%  # numeric   
+                                    zetapotential.mV, # numeric   
+                                   max.size.ingest.mm) %>%  #max ingestible size
   filter(!size_f == "Not Reported")  #take out not reported
@@
 ## Estimate an OLS Regression
-fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master, 
+fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master +mouth, 
              na.action = na.omit, 
@@
 ## PCA
+
+# UMAP
+
+```{r}
+require(uwot)
+require(Rtsne)
+require(vizier) #devtools::install_github("jlmelville/vizier")
+
+# For some functions we need to strip out non-numeric columns and convert data to matrix
+x2m <- function(X) {
+  if (!methods::is(X, "matrix")) {
+    m <- as.matrix(X[, which(vapply(X, is.numeric, logical(1)))])
+  }
+  else {m <- X} 
+  m}
+
+#choose values with most completeness
+multiVar2 <- multiVar %>% 
+  filter(!environment == "Terrestrial") %>% 
+  dplyr::select(size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na() #drop missing
+
+#convert discrete variables to numeric
+multiVar2[] <- data.matrix(multiVar2)
+
+# build umap for small dataset (<10,000 points)
+multiVar_map  <- umap(multiVar2, pca = 10)
+
+# Remove duplicates for t-SNE
+#multiVar2_noNa_dup <- multiVar2_noNa[-which(duplicated(x2m(multiVar2_noNa))), ]
+
+#build t-SNE
+#multiVar_tsne <- Rtsne::Rtsne(multiVar2_noNa_dup, perplexity = 15, initial_dims = 100, partical_pca = TRUE, exaggeration_factor = 4)
+
+# Non-numeric columns are ignored, so in a lot of cases you can pass a data
+# frame directly to umap
+#iris_umap <- umap(iris, n_neighbors = 50, learning_rate = 0.5, init = "random")
+
+#visualize umap
+embed_img <- function(X, Y, k = 15, ...) {
+  args <- list(...)
+  args$coords <- Y
+  args$x <- X
+
+  do.call(vizier::embed_plot, args)
+}
+
+#plot
+embed_img(multiVar2, multiVar_map, pc_axes = TRUE, equal_axes = TRUE, alpha_scale = 0.5, title = "Tox UMAP", cex = 1)
+```
+
+```{r}
+#PCA
+pca <- stats::prcomp(multiVar2[,-5], retx = TRUE, rank. = 2)
+#build color pallete
+my_colors = colorRampPalette(c("red", "yellow", "green"))(nrow(multiVar2))
+#plot
+embed_plot(pca$x, multiVar2$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), #turbo, #rainbow, #my_colors, 
+           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE)
+```
+
+```{r}
+require(plotly)
+#PCA for discrete variables
+multiVar_discrete <- multiVar %>% 
+  select(size.length.um.used.for.conversions, polymer, dose.mg.L.master, exposure.duration.d) %>% 
+  drop_na %>% 
+  mutate_if(~is.numeric(.) && (.) > 0, log10)
+
+#buildPCA
+pca_discrete <- stats::prcomp(multiVar_discrete[,-2], retx = TRUE, rank. = 2)
+
+embed_plotly(pca_discrete$x, multiVar_discrete$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), 
+           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE,
+           tooltip = paste("Polymer:", multiVar_discrete$polymer))
+```
 
@@
 ```
+# Non-linear Dose Response
+If we hold all paramaters constant except for dose, we should observe a non-linear dose-response behavior that can be best predicted by a 3- or 4-parameter Hill Equation.
+The most available information is likely for polystyrene spheres in crustaceans.There are 21 studies available.

### 2021-02-03 Stepwise Multiple Regressions.Rmd

--- 2021-01-20 multivariate.Rmd
@@
 ---
-title: "Multivariate Analysis"
-author: "Scott Coffin"
-date: "1/19/2021"
+title: "Stepwise Multiple Regressions"
+author: "Leah Thornton Hampton"
+date: "2/1/2021"
 output: html_document
 ---
-# Setup
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-# Load packages
-library(tidyverse) #General everything
-library(RColorBrewer)
-library(ggplot2) #General plotting
-library(ggrepel) #For adding text labels that repel away from data points
-library(calecopal) #Color palette
-library(shiny) #Runs shiny
-library(shinythemes) #Shiny theme for the page
-library(shinyWidgets) #Widgets
-library(scales) #SSD - Use the percent format
-library(reshape2) #Overview tab - melts bars together
-library(ssdtools) #SSD package
-library(DT) #Build HTML data tables
-#library(plotly) #Make plots interactive
-library(viridis) #Colors
-library(scales) #To use "percent" function
-library(shinyjs) #Exploration tab - reset button
-library(tigerstats) #turns things into percents
-library(ggbeeswarm) #plot all points
-library(fitdistrplus) #alt SSD 
-library(RColorBrewer) #colors
-library(pheatmap) #pretty heat maps
-library(rpart)  #for trees
-#library(rattle)    # Fancy tree plot This is a difficult library to install (https://gist.github.com/zhiyzuo/a489ffdcc5da87f28f8589a55aa206dd) 
-library(rpart.plot)             # Enhanced tree plots
-library(RColorBrewer)       # Color selection for fancy tree plot
-library(party)                  # Alternative decision tree algorithm
-# library(partykit)               # Convert rpart object to BinaryTree
-library(pROC)   #for ROC curves
+# Load packages and data import
 
-library(ISLR)  #for the Carseat Data
-```
-## Data import
-```{r data import}
-# Load finalized dataset.
+```{r, echo = FALSE, message = FALSE}
+#Packages
+library(tidyverse)
+library(MASS) #Stepwise regression package
+
+#Data Import
 aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
 
-#### Introduction Setup ####
+```
 
-# All text inputs below.
+# Data tidying
 
-#### Overview AO Setup ####
+```{r}
 
-#Final_effect_dataset <- read_csv("Final_effect_dataset.csv")%>%
-  #mutate(plot_f = case_when(
-    #plot_f == "Polymer" ~ "Polymer",
-    #plot_f == "Size" ~ "Size",
-    #plot_f == "Shape" ~ "Shape",
-    #plot_f == "Organism" ~ "Organism",
-    #plot_f == "Lvl1" ~ "Endpoint Category",
-    #plot_f == "Life.stage" ~ "Life Stage",
-    #plot_f == "Invivo.invivo" ~ "In Vivo or In Vitro",
-    #plot_f == "Exposure.route" ~ "Exposure Route"))%>%
-  #mutate(plot_f = factor(plot_f))%>%
-  #mutate(logEndpoints = log(Endpoints))%>%
-  #rename(Percent = Freq)
-
-polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
-polyf<-as.data.frame(polydf)%>% #Makes data frame 
-  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
-  mutate(polymer = case_when(
-    polymer == "BIO" ~ "Biopolymer",
-    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
-    polymer == "LTX" ~ "Latex",
-    polymer == "PA" ~ "Polyamide",
-    polymer == "PE" ~ "Polyethylene",
-    polymer == "PC" ~ "Polycarbonate",
-    polymer == "PET" ~ "Polyethylene Terephthalate",
-    polymer == "PI" ~ "Polyisoprene",
-    polymer == "PMMA" ~ "Polymethylmethacrylate",
-    polymer == "PP" ~ "Polypropylene",
-    polymer == "PS" ~ "Polystyrene",
-    polymer == "PUR" ~ "Polyurathane",
-    polymer == "PVC" ~ "Polyvinylchloride",
-    polymer == "PLA" ~ "Polylactic Acid"))%>%
-  mutate_if(is.numeric, round,0) #rounds percents 
-Endpoints<-xtabs(~polymer +effect ,aoc) #Pulls all study obs. for polymer from dataset
-polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
-  rename(Endpoints='Freq.1')%>% #renames column
-  mutate(logEndpoints = log(Endpoints))%>%
-  rename(Percent = Freq)#renames column
-
-sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
-sizef<-as.data.frame(sizedf)%>%
-  filter(effect %in% c("Y","N"))%>%
-  mutate(size.category = case_when(
-    size.category == 1 ~ "<1µm",
-    size.category == 2 ~ "1µm < 10µm",
-    size.category == 3 ~ "10µm < 100µm",
-    size.category == 4 ~ "100µm < 1mm",
-    size.category == 5 ~ "1mm < 5mm",
-    size.category == 0 ~ "unavailable"))%>%
-  rename(Type = "size.category")%>%
-  mutate_if(is.numeric, round,0)%>%
-  mutate(plot="Size")
-study_s<-xtabs(~size.category +effect ,aoc)
-sizefinal<- data.frame(cbind(sizef, study_s))%>% 
-  rename(Endpoints='Freq.1')%>%
-  rename(category='size.category')%>%
-  mutate(logEndpoints = log(Endpoints))%>%
-  rename(Percent = Freq)#renames column
-
-shapedf<-rowPerc(xtabs(~shape + effect, aoc))
-shapef<-as.data.frame(shapedf)%>%
-  filter(effect %in% c("Y","N"))%>%
-  rename(Type="shape")%>%
-  mutate_if(is.numeric, round,0)%>%
-  mutate(plot="Shape")%>%
-  mutate(Type = case_when(
-    Type == "cube" ~ "Cube",
-    Type == "sphere" ~ "Sphere",
-    Type == "fragment" ~ "Fragment",
-    Type == "fiber" ~ "Fiber"))
-study_sh<-xtabs(~shape + effect,aoc)
-shapefinal<- data.frame(cbind(shapef, study_sh))%>% 
-  rename(Endpoints='Freq.1')%>%
-  rename(category='shape')%>%
-  mutate(logEndpoints = log(Endpoints))%>%
-  rename(Percent = Freq)#renames column
-
-taxdf<-rowPerc(xtabs(~organism.group +effect, aoc))
-taxf<-as.data.frame(taxdf)%>%
-  filter(effect %in% c("Y","N"))%>%
-  rename(Type= "organism.group")%>%
-  mutate_if(is.numeric, round,0)%>%
-  mutate(plot="Organism")
-study_t<-xtabs(~organism.group +effect,aoc)
-taxfinal<- data.frame(cbind(taxf, study_t))%>% 
-  rename(Endpoints='Freq.1')%>%
-  rename(category='organism.group')%>%
-  mutate(logEndpoints = log(Endpoints))%>%
-  rename(Percent = Freq)#renames column
-
-lvl1df<-rowPerc(xtabs(~lvl1 +effect, aoc))
-lvl1f<-as.data.frame(lvl1df)%>%
-  filter(effect %in% c("Y","N"))%>%
-  rename(Type= "lvl1")%>%
-  mutate_if(is.numeric, round,0)%>%
-  mutate(plot="Lvl1")%>%
-  mutate(Type = case_when(
-    Type == "alimentary.excretory" ~ "Alimentary, Excretory",
-    Type == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
-    Type == "circulatory.respiratory" ~ "Circulatory, Respiratory",
-    Type == "community" ~ "Community",
-    Type == "fitness" ~ "Fitness",
-    Type == "immune" ~ "Immune",
-    Type == "metabolism" ~ "Metabolism",
-    Type == "microbiome" ~ "Microbiome",
-    Type == "stress" ~ "Stress")) 
-study_l<-xtabs(~lvl1 +effect,aoc)
-lvl1final<- data.frame(cbind(lvl1f, study_l))%>% 
-  rename(Endpoints='Freq.1')%>%
-  rename(category='lvl1')%>%
-  mutate(logEndpoints = log(Endpoints))%>%
-  rename(Percent = Freq)#renames column
-
-lifedf<-rowPerc(xtabs(~life.stage +effect, aoc))
-lifef<-as.data.frame(lifedf)%>%
-  filter(effect %in% c("Y","N"))%>%
-  rename(Type= "life.stage")%>%
-  mutate_if(is.numeric, round,0)%>%
-  mutate(plot="Life.stage")
-studyli<-xtabs(~life.stage +effect ,aoc)
-lifefinal<- data.frame(cbind(lifef, studyli))%>% 
-  rename(Endpoints='Freq.1')%>%
-  rename(category='life.stage')%>%
-  mutate(logEndpoints = log(Endpoints))%>%
-  rename(Percent = Freq)#renames column
-
-vivodf<-rowPerc(xtabs(~invitro.invivo +effect, aoc))
-vivof<-as.data.frame(vivodf)%>%
-  filter(effect %in% c("Y","N"))%>%
-  rename(Type= "invitro.invivo")%>%
-  mutate_if(is.numeric, round,0)%>%
-  mutate(plot="Invivo.invivo")%>%
-  mutate(Type = case_when(
-    Type=="invivo"~"In Vivo",
-    Type=="invitro"~"In Vitro"))
-study_v<-xtabs(~invitro.invivo +effect,aoc)
-vivofinal<- data.frame(cbind(vivof, study_v))%>% 
-  rename(Endpoints='Freq.1')%>%
-  rename(category='invitro.invivo')%>%
-  mutate(logEndpoints = log(Endpoints))%>%
-  rename(Percent = Freq)#renames column
-
-routedf<-rowPerc(xtabs(~exposure.route +effect, aoc))
-routef<-as.data.frame(routedf)%>%
-  filter(effect %in% c("Y","N"))%>%
-  rename(Type= "exposure.route")%>%
-  mutate_if(is.numeric, round,0)%>%
-  mutate(plot="Exposure.route")%>%
-  mutate(Type = case_when(
-    Type == "coparental.exposure" ~"Co-Parental Exposure",
-    Type == "paternal.exposure" ~ "Paternal Exposure",
-    Type == "maternal.exposure" ~ "Maternal Exposure",
-    Type == "food" ~ "Food",
-    Type == "water" ~ "Water",
-    Type == "sediment" ~ "Sediment",
-    Type == "media" ~ "Media"))
-study_r<-xtabs(~exposure.route +effect,aoc)
-routefinal<- data.frame(cbind(routef, study_r))%>% 
-  rename(Endpoints='Freq.1')%>%
-  rename(category='exposure.route')%>%
-  mutate(logEndpoints = log(Endpoints))%>%
-  rename(Percent = Freq)#renames column
-
-#### Exploration AO Setup ####
-
-# Master dataset for scatterplots - for Heili's tab.
-aoc_v1 <- aoc %>% # start with original dataset
-   # full dataset filters.
+#This is the same tidying and setup as found in app.R as of 2/1/20
+aoc_setup <- aoc %>% 
   mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
@@
     levels = c("No", "Yes"))) %>%
-  # removing NAs to make data set nicer
-  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported"))
-
-aoc_setup <- aoc_v1 %>% # start with original dataset
+  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported")) %>% 
   mutate(size_f = factor(case_when(
@@
     size.category == 0 ~ "Not Reported"),
-    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
-  # shape category data tidying.
+    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% 
   mutate(shape_f = factor(case_when(
@@
     shape == "Not Reported" ~ "Not Reported"),
-    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% # order our different shapes.
-  # polymer category data tidying.
+    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% 
   mutate(poly_f = factor(case_when(
@@
     polymer == "PS" ~ "Polystyrene",
-    polymer == "PUR" ~ "Polyurathane",
+    polymer == "PUR" ~ "Polyurethane",
     polymer == "PVC" ~ "Polyvinylchloride",
@@
     polymer == "Not Reported" ~ "Not Reported"))) %>%
-  # taxonomic category data tidying.
   mutate(org_f = factor(organism.group, levels = c("Algae", "Annelida", "Bacterium", "Cnidaria", "Crustacea",
-                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% # order our different organisms.
+                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% 
   mutate(lvl1_f = factor(case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
@@
     lvl1 == "microbiome" ~ "Microbiome",
-    lvl1 == "stress" ~ "Stress"))) %>% # creates new column with nicer names.
-  # Level 2 Data tidying
+    lvl1 == "stress" ~ "Stress"))) %>% 
   mutate(lvl2_f = factor(case_when(lvl2 == "abundance"~"Abundance",
@@
     lvl2 == "stress"~"Stress",
-    lvl2 == "vision.system"~"Vision System"))) %>% #Renames for widget
-  mutate(bio_f = factor(case_when(bio.org == "cell"~"Cell", #Bio Org Data Tidying
+    lvl2 == "vision.system"~"Vision System"))) %>% 
+  mutate(bio_f = factor(case_when(bio.org == "cell"~"Cell", 
     bio.org == "organism"~"Organism",
@@
   mutate(vivo_f = factor(case_when(invitro.invivo == "invivo"~"In Vivo",
-    invitro.invivo == "invitro"~"In Vitro")))%>% ##Renames for widget (Not using a widget right now, but saving for human health database)
+    invitro.invivo == "invitro"~"In Vitro")))%>% 
   mutate(life_f = factor(case_when(life.stage == "Early"~"Early",
@@
     life.stage == "Adult"~"Adult",
-    life.stage == "Not Reported"~"Not Reported")))%>% #Renames for widget
+    life.stage == "Not Reported"~"Not Reported")))%>% 
   mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
@@
     environment == "Terrestrial" ~ "Terrestrial"))) %>%
+  mutate(species_f = as.factor(paste(genus,species))) %>% 
   mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
-  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported))
-
-#### SSD AO Setup ####
-
-# Master dataset for SSDs
-aoc_z <- aoc_setup %>% # start with Heili's altered dataset (no filtration for terrestrial data)
-  # environment category data tidying.
-  mutate(environment.noNA = replace_na(environment, "Not Reported")) %>% # replaces NA to better relabel.
-  mutate(env_f = factor(environment.noNA, levels = c("Marine", "Freshwater", "Terrestrial", "Not Reported")))
-
-# final cleanup and factoring  
-aoc_z$Species <- as.factor(paste(aoc_z$genus,aoc_z$species)) #must make value 'Species" (uppercase)
-aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
-aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
-```
-```{r}
-# subset data to selected variables
-
-multiVar <- aoc_z %>% dplyr::select(#doi, size.category, 
-                                    size_f,
-                                    size.length.um.used.for.conversions, 
-                                    shape, 
-                                    polymer, 
-                                    particle.volume.um3, 
-                                    density.mg.um.3, 
-                                    organism.group,
-                                    environment, 
-                                    bio.org, #biological level of organization
-                                    #af.time, #assessment factor based on exposure time
-                                    treatments, #number of doses (no including control)
-                                    effect, #yes no
-                                    exposure.duration.d, 
-                                    exposure.route, #Factor
-                                    organism.group, #factor
-                                    media.temp, #numeric
-                                    lvl1_f, #endpoints
-                                    lvl2_f, #endpoints
-                                   # lvl3, 
-                                     dose.mg.L.master, 
-                                    sex, #factor
-                                    media.ph, #numeric
-                                    media.sal.ppt, #numeric
-                                    dose.particles.mL.master,
-                                   effect.metric, #NOEC LOEC
-                                    functional.group, #factor
-                                    charge, #positive or negatibe
-                                    zetapotential.mV) %>%  # numeric   
-  filter(!size_f == "Not Reported")  #take out not reported
-
-#recode variables
-multiVar <- multiVar %>% mutate(effect_10 = case_when(
-    effect == "Y" ~ 1,
-    effect == "N" ~ 0
-  ))# %>% 
-  #mutate_all(is.character, ~as.factor())
+  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
+  mutate(effect.metric = factor(effect.metric)) %>%
+  mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) %>%  
+  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
+  mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
+                                            af.time_noNA == 1 ~ "Chronic",
+                                            af.time_noNA == "Unavailable" ~ "Unavailable")))
 
@@
 
-# Analysis
-## Data Exploration
-### Completeness vby Size
-```{r completeness heatmap}
-CompletenessSummary_size <- multiVar %>%
-  group_by(size_f) %>% 
-     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
-  mutate(across(is.numeric,~round(., 2))) %>% 
-  mutate(across(is.numeric, ~100 *(1 -.)))
-CompletenessSummary_size
-```
+```{r}
 
-```{r}
-require()
-#convert to matrix and transpose
-transposed_size <- as.data.frame(t(as.matrix(CompletenessSummary_size[2:25]))) %>% #1 is category, 2-6 are variables
-  arrange('1nm < 100nm')
-#reassign column names
-colnames(transposed_size) <- c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm")
-#transposedTag
-#format as matrix
-MissingMatrix_size <- data.matrix(transposed_size)
-#build heatmap
-pheatmap(MissingMatrix_size,
-                           main = "Data Completeness by Size Bin", #title
-                           fontsize = 12,
-                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
-                           display_numbers = TRUE,
-                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
-                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(dose.particles.mL.master, 
+         dose.mg.L.master, 
+         size_f, 
+         size.length.um.used.for.conversions, 
+         shape_f, 
+         poly_f, 
+         dose.um3.mL.master, 
+         max.size.ingest.mm, 
+         exposure.duration.d, 
+         bio_f, 
+         charge, 
+         zetapotential.mV, 
+         particle.source,
+         lvl1_f,
+         lvl2_f,
+         lvl3,
+         effect)
+
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect ~., data = y)
+
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
 
 ```
-### Completeness by Environment
-```{r completeness heatmap}
-CompletenessSummary_environment <- multiVar %>%
-  filter(!environment == "NA") %>% 
-  group_by(environment) %>% 
-     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
-  mutate(across(is.numeric,~round(., 2))) %>% 
-  mutate(across(is.numeric, ~100 *(1 -.)))
-CompletenessSummary_environment
-```
-
-```{r}
-require()
-#convert to matrix and transpose
-transposed_environment <- as.data.frame(t(as.matrix(CompletenessSummary_environment[2:25])))
-#reassign column names
-colnames(transposed_environment) <- c("Freshwater", "Marine", "Terrestrial")
-#transposedTag
-#format as matrix
-MissingMatrix_environment <- data.matrix(transposed_environment)
-#build heatmap
-pheatmap(MissingMatrix_environment,
-                           main = "Data Completeness by environment", #title
-                           fontsize = 12,
-                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
-                           display_numbers = TRUE,
-                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
-                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)
-
-```
-
-## Stepwise Regression
-
-```{r}
-## Estimate an OLS Regression
-fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master, 
-             na.action = na.omit, 
-             data = multiVar)
-summary(fitols)
-```
-
-```{r}
-require(reshape2)
-multiVar %>% 
-  dplyr::select(size.length.um.used.for.conversions, particle.volume.um3, exposure.duration.d ,media.temp, dose.particles.mL.master) %>% 
-  melt() %>%  #convert wide to long
-   mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  ggplot(aes(x = value)) + 
-  stat_density() + 
-  facet_wrap(~variable, scales = "free")
-```
-
-<!-- # ```{r} -->
-<!-- # # cdplot(as.factor(effect) ~ particle.volume.um3, data = multiVar, -->
-<!-- #        main = "Conditional Density Plot of ", xlab = "Dose", ylab = "Effect") -->
-<!-- # ``` -->
-
-## PCA
-
-## Random Forest
-### Recursive Partitioning And Regression Trees
-
-The rpart algorithm works by splitting the dataset recursively, which means that the subsets that arise from a split are further split until a predetermined termination criterion is reached.
-At each step, the split is made based on the independent variable that results in the largest possible reduction in heterogeneity of the dependent (predicted) variable.
-
-It is important to note that the algorithm works by making the best possible choice at each particular stage, without any consideration of whether those choices remain optimal in future stages.
-That is, the algorithm makes a locally optimal decision at each stage.
-It is thus quite possible that such a choice at one stage turns out to be sub-optimal in the overall scheme of things.
-In other words,  the algorithm does not find a globally optimal tree.
-```{r}
-#trim data so effect is always known
-multiVar_sub <- multiVar %>% 
-  drop_na(effect_10)
-
-# Split data into training and test sets
-set.seed(42)
-multiVar_sub[,"train"] <- ifelse(runif(nrow(multiVar_sub)) < 0.8, 1, 0)
-# Separate trainig and test sets
-trainSet <- multiVar_sub[multiVar_sub$train==1,]
-testSet <- multiVar_sub[multiVar_sub$train==0,]
-#get column index of train flag
-trainColNum <- grep("train", names(trainSet))
-# Remove train flag column from train and test sets
-trainSet <- trainSet[, -trainColNum]
-testSet <- testSet[, -trainColNum]
-```
-Make a classification tree.
-
-```{r, echo = FALSE}
-#get column index of predicted variable in dataset
-typeColNum <- grep("effect",names(multiVar_sub))
-
-#build tree
-require(rpart)
-set.seed(15097)
-t1 <- rpart(effect ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,
-            method = "class", #classification because response is discrete
-            control = rpart.control(minbucket = 20, cp=0.008), #requires that there be at least 19 cases (responded + nonrespondents) in the final grouping of variable values of the terminal node of the tre..
-            data = trainSet)
-print(t1, digits=4)
-```
-
-Plot an interpretable tree.
-```{r, echo = FALSE}
-require(rpart.plot)
-cols <- ifelse(t1$frame$yval == 1, "gray50", "black")
-prp(t1, main="Tree for Effect",
-    extra=106, # display prob of survival and percent of obs
-    nn=TRUE, # display node numbers
-    fallen.leaves=TRUE, # put leaves on the bottom of page
-    branch=.5, # change angle of branch lines
-    faclen=0, # do not abbreviate factor levels
-    trace=1, # print automatically calculated cex
-    shadow.col="gray", # shadows under the leaves
-    branch.lty=1, # draw branches using solid lines
-    branch.type=5, # branch lines width = weight(frame$wt), no. of cases here
-    split.cex=1.2, # make split text larger than node text
-    split.prefix="is ", # put "is " before split text
-    split.suffix="?", # put "?" after split text
-    col=cols, border.col=cols, # cols[2] if survived
-    split.box.col="lightgray", # lightgray split boxes (default is white)
-    split.border.col="darkgray", # darkgray border on split boxes
-    split.round=0.5) # round the split box corners a tad
-```
-Next we see how good the model is by seeing how it fares against the test data.
-```{r}
-t1_predict <- predict(t1, newdata = testSet[,-typeColNum],
-                      type="class")
-mean(t1_predict==testSet$effect)
-# [1] 0.7094088
-#confusion matrix
-table(pred=t1_predict,true=testSet$effect)
-```
-```{r}
-par(mfrow=c(1,2)) # two plots on one page
-#plot approximate R-squared and relative error for different splits (2 plots). labels are only appropriate for the "anova" method.
-rsq.rpart(t1)
-```
-
-Next, we prune the tree using the cost complexity criterion.
-Basically, the intent is to see if a shallower subtree can give us comparable results.
-If so, we’d be better of choosing the shallower tree because it reduces the likelihood of overfitting.
-
-As described earlier, we choose the appropriate pruning parameter (aka cost-complexity parameter) \alpha by picking the value that results in the lowest prediction error.
-Note that all relevant computations have already been carried out by R when we built the original tree (the call to rpart in the code above).
-All that remains now is to pick the value of \alpha:
-
-#### Pruning
-
-```{r}
-#cost-complexity pruning
-printcp(t1)
-```
-It is clear from the above, that the lowest cross-validation error (xerror in the table) occurs for \alpha =0.008 (this is CP in the table above).   One can find CP programatically like so:
-```{r}
-# get index of CP with lowest xerror
-opt <- which.min(t1$cptable[,"xerror"])
-#get its value
-cp <- t1$cptable[opt, "CP"]
-```
-Next, we prune the tree based on this value of CP:
-
-```{r}
-# prune the tree
-pt1 <- prune(t1,cp)
-#pt1<- prune(t1, cp= t1$cptable[which.min(t1$cptable[,"xerror"]),"CP"])
-
-# plot the pruned tree
-plot(pt1, uniform=TRUE,
-   main="Pruned Classification Tree");text(pt1, use.n=TRUE, all=TRUE, cex=.8)
-
-#post(pfit, file = "c:/ptree.ps",
- #  title = "Pruned Classification Tree for Kyphosis")
-```
-```{r}
-#find proportion of correct predictions using test set
-t1_pruned_predict <- predict(pt1, testSet, type="class")
-mean(t1_pruned_predict == testSet$effect)
-#
-```
-This is not an improvement over an unprunend tree..
-We need to check that this holds up for different training and test sets.
-This is easily done by creating multiple random partitions of the dataset and checking the efficacy of pruning for each.
-To do this efficiently, I’ll create a function that takes the training fraction, number of runs (partitions) and the name of the dataset as inputs and outputs the proportion of correct predictions for each run.
-It also optionally prunes the tree.
-
-```{r}
-#function to do multiple runs
-multiple_runs_classification <- function(train_fraction,n,dataset,prune_tree=FALSE){
-fraction_correct <- rep(NA,n)
-set.seed(42)
-for (i in 1:n){
-  dataset[,"train"] <- ifelse(runif(nrow(dataset))<0.8,1,0)
-  trainColNum <- grep("train",names(dataset))
-  typeColNum <- grep("effect",names(dataset))
-  trainSet <- dataset[dataset$train==1,-trainColNum]
-  testSet <- dataset[dataset$train==0,-trainColNum]
-  rpart_model <- rpart(effect~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,data = trainSet, method="class")
-if(prune_tree==FALSE) {
-  rpart_test_predict <- predict(rpart_model,testSet[,-typeColNum],type="class")
-  fraction_correct[i] <- mean(rpart_test_predict==testSet$effect)
-  }else{
-    opt <- which.min(rpart_model$cptable[,"xerror"])
-    cp <- rpart_model$cptable[opt, "CP"]
-    pruned_model <- prune(rpart_model,cp)
-    rpart_pruned_predict <- predict(pruned_model,testSet[,-typeColNum],type="class")
-    fraction_correct[i] <- mean(rpart_pruned_predict == testSet$effect)
-  }
-  }
-return(fraction_correct)
-}
-```
-
-Note that in the above,  I have set the default value of the prune_tree to FALSE, so the function will execute the first branch of the if statement unless the default is overridden.
-
-OK, so let’s do 50 runs with and without pruning, and check the mean and variance of the results for both sets of runs.
-```{r}
-#50 runs, no pruning
-unpruned_set <- multiple_runs_classification(0.8,50,multiVar_sub)
-mean(unpruned_set)
-#[1] 0.7261882
-sd(unpruned_set)
-#[1] 0.01554347
-#50 runs, with pruning
-pruned_set <- multiple_runs_classification(0.8,50,multiVar_sub,prune_tree=TRUE)
-mean(pruned_set)
-#[1] 0.7261875
-sd(pruned_set)
-#[1] 0.01552285
-```
-
-### CForest
-```{r run model}
-require(party)
-crf <- cforest(as.factor(effect_10) ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d +
-                 media.temp + dose.particles.mL.master,
-               controls = cforest_control(ntree = 500,
-                                          mincriterion = qnorm(0.8), 
-                                          trace = TRUE), # adds project bar because it's very slow
-               data = multiVar_sub)
-
-crf
-```
-
-```{r fit results}
-train2 <- multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
-                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master)
-train3 <-multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
-                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master,
-                                  effect_10)
-
-fitted <- predict(crf, train2, OOB = TRUE, type ="response")
-#rpart.prob <- predict(t1, newdata=imputedSmalls.requested.voluntary,type="prob")
-
-misClasificError <- mean(fitted != train3$effect_10)
-print(paste('Training Accuracy', 1 - misClasificError))
-```
-
-```{r}
-print(crf)
-```
-
-```{r}
-#Obtaining predicted probabilites for Test data
-tree.probs=predict(crf,
-                 newdata = train2,
-                 type="prob")
-
-#Calculate ROC curve
-rocCurve.tree <- roc(train3$effect_10,tree.probs[,'1'])
-#plot the ROC curve
-plot(rocCurve.tree,col=c(4))
-```
-
-Alternative ROC Curve
-```{r}
-require(pROC)
-predicted <- predict(crf, train2, OOB=TRUE, type= "response")
-auc(as.numeric(train3$effect_10), as.numeric(predicted))
-#Calculate ROC curve
-rocCurve.tree <- roc(train3$effect_10,as.numeric(predicted))
-#plot the ROC curve
-plot(rocCurve.tree,col=c(4))
-```
-```{r}
-#plot reciever operating curve
-perf <- performance(preds,"tpr","fpr")
-```
-
-```{r}
-# compute in-sample results
-caret::confusionMatrix(fitted,as.factor(train3$effect_10))
-```
-
-```{r}
-#plot feature importance
-cforestImpPlot <- function(x) {
-  cforest_importance <<- v <- varimp(x)
-  dotchart(v[order(v)])
-}
-
-importancePlot <- cforestImpPlot(crf)
-importancePlot
-```
-
-```{r}
-crf.prob <- matrix(unlist(crfsrvy.prob), ncol=2, byrow=TRUE)
-apply(crf.prob,2,mean)
-#[1] 0.2524514 0.7475486
-tab <- round(cbind(by(rpart.prob[,2], INDICES=t1$where, mean), by(crf.prob[,2], INDICES=t1$where, mean)), 4)
-colnames(tab) <- c("rpart", "cforest")
-kable(tab)
-```

### 2021-02-03 Stepwise Multiple Regressions.Rmd
Stepwise Regressions + knit

@@
 
-# Data tidying
+# Data Tidying
 
-```{r}
+```{r, echo = FALSE,}
 
@@
 
-```{r}
+#Model
 
-#Select the columns that we want to feed into the model for simplicity
+Doses (count, mass and volume), size (continuous), and exposure duration are all log transformed given histogram results.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+10. Charge (negative/positive)
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity.
 aoc_setup_selected <- aoc_setup %>% 
@@
          dose.um3.mL.master, 
-         max.size.ingest.mm, 
          exposure.duration.d, 
@@
          charge, 
-         zetapotential.mV, 
          particle.source,
-         lvl1_f,
-         lvl2_f,
-         lvl3,
+         org_f,
          effect)
+
+```
+
+```{r, echo = FALSE}
 
@@
 
+summary(Full_Model)
+
 #Stepwise model
@@
 ```
+
+#Model Refining
+
+##Remove Charge
+
+Complete charge data are limited to PS and PC spheres - remove to include for complete data.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+~~10. Charge (negative/positive)~~
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity.
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(dose.particles.mL.master, 
+         dose.mg.L.master, 
+         size_f, 
+         size.length.um.used.for.conversions, 
+         shape_f, 
+         poly_f, 
+         dose.um3.mL.master, 
+         exposure.duration.d, 
+         bio_f, 
+         # charge, 
+         particle.source,
+         org_f,
+         effect)
+
+```
+
+```{r, echo = FALSE}
+
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect ~., data = y)
+
+summary(Full_Model)
+
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+##Remove Charge + Size (binned)
+
+Size (binned) had the highest p value in the last model - remove.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+~~5. Size (binned)~~
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+~~10. Charge (negative/positive)~~
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity.
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(dose.particles.mL.master, 
+         dose.mg.L.master, 
+         # size_f, 
+         size.length.um.used.for.conversions, 
+         shape_f, 
+         poly_f, 
+         dose.um3.mL.master, 
+         exposure.duration.d, 
+         bio_f, 
+         # charge, 
+         particle.source,
+         org_f,
+         effect)
+
+```
+
+```{r, echo = FALSE}
+
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect ~., data = y)
+
+summary(Full_Model)
+
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+##Remove Charge + Size (binned) + Shape
+
+Shape had the highest p value in the last model - remove.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+~~5. Size (binned)~~
+~~6. Shape~~
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+~~10. Charge (negative/positive)~~
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity.
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(dose.particles.mL.master, 
+         dose.mg.L.master, 
+         # size_f, 
+         size.length.um.used.for.conversions, 
+         # shape_f, 
+         poly_f, 
+         dose.um3.mL.master, 
+         exposure.duration.d, 
+         bio_f, 
+         # charge, 
+         particle.source,
+         org_f,
+         effect)
+
+```
+
+```{r, echo = FALSE}
+
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect ~., data = y)
+
+summary(Full_Model)
+
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+#Variable Histograms for Numerical Data
+
+Plot histograms of all numerical continuous values to determine if any need log transformation due to severe skew.
+
+```{r, echo = FALSE}
+
+##################################### Log transform needed #########################################################################
+
+#### Dose - Particles/mL ####
+Dose_Particles_mL <- aoc_setup$dose.particles.mL.master
+
+hist(Dose_Particles_mL)
+
+#Dose - Particles/mL - log transformed
+Dose_Particles_mL_log10 <- log10(aoc_setup$dose.particles.mL.master)
+
+hist(Dose_Particles_mL_log10)
+
+#### Dose - mg/L ####
+Dose_mg_L <- aoc_setup$dose.mg.L.master
+
+hist(Dose_mg_L)
+
+#Dose - mg/L - log transformed
+Dose_mg_L_log10 <- log10(aoc_setup$dose.mg.L.master)
+
+hist(Dose_mg_L_log10)
+
+#### Dose - um3/mL ####
+Dose_um3_mL <- aoc_setup$dose.um3.mL.master
+
+hist(Dose_um3_mL)
+
+#Dose - um3/mL - log transformed
+Dose_um3_mL_log10 <- log10(aoc_setup$dose.um3.mL.master)
+
+hist(Dose_um3_mL_log10)
+
+#### Size - Continuous ####
+Size_Continuous <- aoc_setup$size.length.um.used.for.conversions
+
+hist(Size_Continuous)
+
+#Size - Continuous - log transformed
+Size_Continuous_log10 <- log10(aoc_setup$size.length.um.used.for.conversions)
+
+hist(Size_Continuous_log10)
+
+#### Exposure Duration ####
+Exposure_Duration <- aoc_setup$exposure.duration.d
+
+hist(Exposure_Duration)
+
+#Exposure Duration - log transformed
+Exposure_Duration_log10 <- log10(aoc_setup$exposure.duration.d)
+
+hist(Exposure_Duration_log10)
+
+######################################## Log Transform not needed #################################################################
+
+#### Ingestible Size ####
+Max_Ingest <- aoc_setup$max.size.ingest.mm
+
+hist(Max_Ingest)
+
+#### Zeta Potential ####
+Zeta <- aoc_setup$zetapotential.mV
+
+hist(Zeta)
+
+```
+
+#Model - Log Transformed
+
+Doses (count, mass and volume), size (continuous), and exposure duration are all log transformed given histogram results.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+10. Charge (negative/positive)
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Log transform numeric values 
+aoc_setup <- aoc_setup %>% 
+  mutate(log10(dose.particles.mL.master)) %>% 
+  mutate(log10(dose.mg.L.master)) %>%
+  mutate(log10(dose.um3.mL.master)) %>%
+  mutate(log10(size.length.um.used.for.conversions)) %>%
+  mutate(log10(exposure.duration.d))
+
+#Select the columns that we want to feed into the model for simplicity and log transform data according to histogram results. 
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(`log10(dose.particles.mL.master)`, 
+         `log10(dose.mg.L.master)`, 
+         size_f, 
+         `log10(size.length.um.used.for.conversions)`, 
+         shape_f, 
+         poly_f, 
+         `log10(dose.um3.mL.master)`, 
+        `log10(exposure.duration.d)`, 
+         bio_f, 
+         charge, 
+         particle.source,
+         org_f,
+         effect)
+
+```
+
+```{r, echo = FALSE}
+
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect ~., data = y)
+
+summary(Full_Model)
+
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+#Model Refining
+
+##Remove Charge
+
+Complete charge data are limited to PS and PC spheres - remove to include for complete data.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+~~10. Charge (negative/positive)~~
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Log transform numeric values 
+aoc_setup <- aoc_setup %>% 
+  mutate(log10(dose.particles.mL.master)) %>% 
+  mutate(log10(dose.mg.L.master)) %>%
+  mutate(log10(dose.um3.mL.master)) %>%
+  mutate(log10(size.length.um.used.for.conversions)) %>%
+  mutate(log10(exposure.duration.d))
+
+#Select the columns that we want to feed into the model for simplicity and log transform data according to histogram results. 
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(`log10(dose.particles.mL.master)`, 
+         `log10(dose.mg.L.master)`, 
+         size_f, 
+         `log10(size.length.um.used.for.conversions)`, 
+         shape_f, 
+         poly_f, 
+         `log10(dose.um3.mL.master)`, 
+        `log10(exposure.duration.d)`, 
+         bio_f, 
+         # charge, 
+         particle.source,
+         org_f,
+         effect)
+
+```
+
+```{r, echo = FALSE}
+
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect ~., data = y)
+
+summary(Full_Model)
+
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+##Remove Charge + Size (binned)
+
+Size (binned) had the highest p value in the last model - removed.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+~~5. Size (binned)~~
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+~~10. Charge (negative/positive)~~
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Log transform numeric values 
+aoc_setup <- aoc_setup %>% 
+  mutate(log10(dose.particles.mL.master)) %>% 
+  mutate(log10(dose.mg.L.master)) %>%
+  mutate(log10(dose.um3.mL.master)) %>%
+  mutate(log10(size.length.um.used.for.conversions)) %>%
+  mutate(log10(exposure.duration.d))
+
+#Select the columns that we want to feed into the model for simplicity and log transform data according to histogram results. 
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(`log10(dose.particles.mL.master)`, 
+         `log10(dose.mg.L.master)`, 
+         # size_f, 
+         `log10(size.length.um.used.for.conversions)`, 
+         shape_f, 
+         poly_f, 
+         `log10(dose.um3.mL.master)`, 
+        `log10(exposure.duration.d)`, 
+         bio_f, 
+         # charge, 
+         particle.source,
+         org_f,
+         effect)
+
+```
+
+```{r, echo = FALSE}
+
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect ~., data = y)
+
+summary(Full_Model)
+
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+##Remove Charge + Size (binned) + Particle Source
+
+Particle source had the highest p value in the last model - removed.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+~~5. Size (binned)~~
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+~~10. Charge (negative/positive)~~
+~~11. Particle Source~~
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Log transform numeric values 
+aoc_setup <- aoc_setup %>% 
+  mutate(log10(dose.particles.mL.master)) %>% 
+  mutate(log10(dose.mg.L.master)) %>%
+  mutate(log10(dose.um3.mL.master)) %>%
+  mutate(log10(size.length.um.used.for.conversions)) %>%
+  mutate(log10(exposure.duration.d))
+
+#Select the columns that we want to feed into the model for simplicity and log transform data according to histogram results. 
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(`log10(dose.particles.mL.master)`, 
+         `log10(dose.mg.L.master)`, 
+         # size_f, 
+         `log10(size.length.um.used.for.conversions)`, 
+         shape_f, 
+         poly_f, 
+         `log10(dose.um3.mL.master)`, 
+        `log10(exposure.duration.d)`, 
+         bio_f, 
+         # charge, 
+         # particle.source,
+         org_f,
+         effect)
+
+```
+
+```{r, echo = FALSE}
+
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect ~., data = y)
+
+summary(Full_Model)
+
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```

### 2021-02-03 Stepwise Multiple Regressions.Rmd
Added headings for clairty

@@
 
+### Full
+
 ```{r, echo = FALSE}
-
 #Convert all data to numerical data
@@
 summary(Full_Model)
+```
 
+### Stepwise
+
+```{r, echo = FALSE}
 #Stepwise model
@@
 Full_Model <- lm(effect ~., data = y)
+```
 
+### Stepwise
+
+```{r, echo = FALSE}
 summary(Full_Model)

### 2021-02-04 multivariate.Rmd
new RF model with continuous variables only

@@
 library(ISLR)  #for the Carseat Data
+library(quantregForest)
+library(caret)
+library(tidyverse)
+library(tidymodels)
+library(skimr)
+library(sf)
+library(ggspatial)
+library(nhdplusTools)
+library(patchwork)
+library(Metrics)
+library(gt)
 ```
@@
 aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
+
+#assign row ID
+aoc <- rowid_to_column(aoc, "ID")
+aoc <- aoc %>% 
+  mutate(ID = paste0(ID, "ID"))
 
@@
   mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
-  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported))
+  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
+  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
+  mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
+                                            af.time_noNA == 1 ~ "Chronic",
+                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>%    #factorize assesment factor time into chronic/acute
+  mutate(dose.mg.L.master.AF.noec = dose.mg.L.master * af.noec) %>% #converts ECxx, LC50, HONEC, LOEC, NOEC all to NOEC
+  mutate(dose.particles.mL.master.AF.noec = dose.particles.mL.master * af.noec)  #converts ECxx, LC50, HONEC, LOEC, NOEC all to NOEC
 
@@
 # subset data to selected variables
-
 multiVar <- aoc_z %>% dplyr::select(#doi, size.category, 
+                                    ID,
                                     size_f,
@@
                                     treatments, #number of doses (no including control)
-                                    effect, #yes no
+                                    effect_f, #yes no
                                     exposure.duration.d, 
@@
                                     zetapotential.mV, # numeric   
-                                   max.size.ingest.mm) %>%  #max ingestible size
-  filter(!size_f == "Not Reported")  #take out not reported
-
-#recode variables
-multiVar <- multiVar %>% mutate(effect_10 = case_when(
-    effect == "Y" ~ 1,
-    effect == "N" ~ 0
-  ))# %>% 
-  #mutate_all(is.character, ~as.factor())
-
+                                   max.size.ingest.mm,#max ingestible size
+                                   acute.chronic_f,
+                                   dose.mg.L.master.AF.noec,
+                                   dose.particles.mL.master.AF.noec) %>%  
+  filter(!size_f == "Not Reported") %>%   #take out not reported 
+  mutate_if(is.character, as.factor)
+
+# #recode variables
+# multiVar <- multiVar %>% mutate(effect_10 = case_when(
+#     effect_f == "Y" ~ 1,
+#     effect_f == "N" ~ 0
+#   ))# 
 ```
@@
 #convert to matrix and transpose
-transposed_size <- as.data.frame(t(as.matrix(CompletenessSummary_size[2:25]))) %>% #1 is category, 2-6 are variables
+transposed_size <- as.data.frame(t(as.matrix(CompletenessSummary_size[2:28]))) %>% #1 is category, 2-6 are variables
   arrange('1nm < 100nm')
@@
 #convert to matrix and transpose
-transposed_environment <- as.data.frame(t(as.matrix(CompletenessSummary_environment[2:25])))
+transposed_environment <- as.data.frame(t(as.matrix(CompletenessSummary_environment[2:28])))
 #reassign column names
@@
 #PCA
-pca <- stats::prcomp(multiVar2[,-5], retx = TRUE, rank. = 2)
+pca <- stats::prcomp(multiVar2[,-4], retx = TRUE, rank. = 2)
 #build color pallete
@@
 multiVar_discrete <- multiVar %>% 
-  select(size.length.um.used.for.conversions, polymer, dose.mg.L.master, exposure.duration.d) %>% 
+  select(size.length.um.used.for.conversions, polymer, dose.mg.L.master, exposure.duration.d, effect) %>% 
   drop_na %>% 
-  mutate_if(~is.numeric(.) && (.) > 0, log10)
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  mutate(effect_10 = case_when(
+    effect == "Y" ~ 1,
+    effect == "N" ~ 0
+  )) %>% 
+  select(-effect)
 
 #buildPCA
-pca_discrete <- stats::prcomp(multiVar_discrete[,-2], retx = TRUE, rank. = 2)
+pca_discrete <- stats::prcomp(multiVar_discrete[,-2], retx = TRUE, rank. = 5, scale. = TRUE)
+
+embed_plot(pca_discrete$x, multiVar_discrete$polymer)
 
@@
            tooltip = paste("Polymer:", multiVar_discrete$polymer))
-```
-
-## Random Forest
-### Recursive Partitioning And Regression Trees
+
+#embed_plotly also option
+```
+
+# Random Forest
+## Recursive Partitioning And Regression Trees
 
@@
 In other words,  the algorithm does not find a globally optimal tree.
+
+The size of the data set usually imposes a hard limit on how many features may be considered due to Curse of Dimensionality, i.e.
+your data becomes sparser and sparser as you increase the number of features considered, which usually leads to overfitting.
+While there is no rule of thumb relating to how many features vs.
+the number of observations you should use, I try to keep $e^Nf < No$ (Nf = number of features, No = number of observations) to minimise overfitting [this is not always possible and it does not ensure that we won’t overfit].
+In this case, our training set has 3200, which suggests that using 8 features (e^8 ~ 2980) is not entirely absurd.
+Obviously, this depends on your data, so we will cover some further overfitting checks later on.
 ```{r}
@@
 
-#### Pruning
+### Pruning
 
@@
 
-### CForest
+## CForest
 ```{r run model}
@@
 ```
+# Random Forest Package
+
+## Preparation
+```{r}
+library(quantregForest)
+library(caret)
+library(tidyverse)
+library(tidymodels)
+library(skimr)
+library(sf)
+library(ggspatial)
+library(nhdplusTools)
+library(patchwork)
+library(Metrics)
+library(gt)
+```
+
+```{r}
+multiVar <- multiVar %>% 
+  mutate_if(is.character, as.factor) %>% 
+  mutate(effect_10 = case_when( #convert ordinal to numeric
+     effect_f == "Yes" ~ 1,
+     effect_f == "No" ~ 0))
+
+#choose relevant predictors and log-transform
+multiVar_small_log <- multiVar %>% 
+  dplyr::select(ID, size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na() %>%  #drop missing
+ mutate(effect_10 = case_when( #convert ordinal to numeric
+     effect_f == "Yes" ~ 1,
+     effect_f == "No" ~ 0
+   ))# %>% 
+ #  select(-effect_f)
+
+#no log transform for comparison
+multiVar_small <- multiVar %>% 
+  dplyr::select(ID, size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, #effect_10 , 
+                exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>% 
+  drop_na() #%>%  #drop missing
+# mutate(effect_10 = case_when( #convert ordinal to numeric
+#     effect == "Y" ~ 1,
+#     effect == "N" ~ 0
+#   )) %>% 
+  # select(-effect)
+
+#ensure completeness
+skim(multiVar_small)
+```
+## Training Data
+
+```{r}
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+multiVar_small_split <- multiVar_small %>%
+  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# default is 3/4ths split (but 75% training, 25% testing).
+# Stratification (strata) = grouping training/testing sets by region, state, etc.
+# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+multiVar_small_train <- training(multiVar_small_split)
+multiVar_small_test <- testing(multiVar_small_split)
+# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Create a separate dataset of available IDs that were not used in the training dataset.
+notTrain <- aoc %>% # all COMIDS from StreamCat data, sampled or not
+  filter(!ID %in% aoc$ID) # Removing sites used to train the model. n = 140,097
+```
+## Kitchen Sink Model
+
+```{r}
+# Step Three - Kitchen Sink model -----------------------------------------
+
+# Create finalized training dataset and include all possible variables. 
+rf_dat <- multiVar_small_train %>%
+  select(-ID)
+
+# Random forest -- 
+# a decision tree model, using predictors to answer dichotomous questions to create nested splits.
+# no pruning happens - rather, multiple trees are built (the forest) and then you are looking for consensus across trees
+# training data goes down the tree and ends up in a terminal node.
+# if testing data goes down the same route, then this upholds our conclusions. Or, if it goes awry, this allows us to look for patterns in how it goes awry.
+
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+myrf <- randomForest(y = rf_dat$effect_f, # dependent variable
+  x = rf_dat %>%
+    select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+myrf # examine the results.
+```
+~20% error rate. Let's compare this with the same model with log-transformed values.
+```{r}
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+multiVar_small_log_split <- multiVar_small_log %>%
+  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+multiVar_small_log_train <- training(multiVar_small_log_split)
+multiVar_small_log_test <- testing(multiVar_small_log_split)
+
+# Create finalized training dataset and include all possible variables. 
+rf_dat_log <- multiVar_small_log_train %>%
+  select(-ID, -effect_10)
+
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+myrf_log <- randomForest(y = rf_dat_log$effect_f, # dependent variable
+  x = rf_dat_log %>%
+    select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+myrf_log # examine the results.
+```
+No performance enhancement with log-transformed values.
+
+## Continuous Predictors
+Repeat with continous variables whenever possible, and max of 8 predictors.
+```{r}
+skim(multiVar)
+```
+```{r}
+acute <- multiVar %>% filter(acute.chronic_f == "Chronic") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!exposure.route == "food") %>% 
+  select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d,bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master)) %>%
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na()
+skim(acute)
+```
+With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
+Let's determine if we have enough data for ML.
+```{r}
+exp(1)^6
+```
+$e^6$ is less than n (453), so we may proceed.
+
+```{r}
+plot(myrf)
+# model performance appears to improve most at ~75 trees
+```
+
+```{r}
+varImpPlot(myrf)
+# displays which variables are most important
+# helps to winnow down list of predictors
+# recommended to weigh left pane more
+# right pane also shows how evenly things split based on the list of predictors
+# values close to 0 can be dropped, but don't have to be
+```
+Dose, organism group, exposure duration are important.
+
+```{r}
+importance <- myrf$importance
+View(importance)
+# displays the data plotted in the plot above
+```
+
+```{r}
+# predict()
+# returns out of bag predictions for training data
+# in the bag: every time a tree is built, it uses ~80% of the original 75% we set aside from the original dataset used to create a tree to assure random data selection
+# out of bag: looking at the remaining 20% of the training data to predict, when you want to know what your model does at the training location sites
+
+# # Predict CRAM scores state-wide for all COMIDs.
+# notTrain_prediction <- notTrain %>% # taking all available IDs, that haven't been used in training
+#   na.omit() %>% # remove NAs
+#   mutate(effect_predicted = predict(myrf, newdata = notTrain %>% na.omit())) # using developed model (myrf), inputting predictor variables (notTrain - which contains IDs and associated  data) to predict output/dependent variable (effect_predicted a.k.a.
+effect).
+
+# rePredict CRAM scores for training data.
+multiVar_small_train$effect_predicted <- predict(myrf) # Add column of predicted CRAM values to training dataset.
+
+# Creates new dataset of bound rows for both ...
+# effect_predictions <- bind_rows(nottrain_prediction %>%
+#                             mutate(Set = "Non-training"), # statewide COMIDs (not used for training)
+#                             multiVar_smal_train %>%
+#                             mutate(Set = "Training")) # COMIDS from training dataset
+# This creates the dataset that will be plotted to create a state-wide plot of predicted CRAM scores.
+
+# Plot the data.
+multiVar_small_train %>% 
+  mutate(effect_10 = case_when( #convert ordinal to numeric
+     effect_predicted == "Yes" ~ 1,
+     effect_predicted == "No" ~ 0
+   )) %>% 
+  ggplot(aes(x = log10(dose.mg.L.master), y = effect_10)) +
+  geom_point(alpha = 0.7) +#color = "black") +
+  labs(x = "Density",
+    y = "Predicted CRAM Score") +
+  theme_dark() #+
+  #facet_wrap(.~polymer)
+```
+### Predictor Selector
+```{r}
+# Using caret to select the best predictors
+# What are the parameters you want to use to run recursive feature elimination (rfe)?
+my_ctrl <- rfeControl(functions = rfFuncs,
+                      method = "cv",
+                      verbose = FALSE,
+                      returnResamp = "all")
+
+# rfe = recursive feature elimination
+# THIS STEP TAKES FOR-EV-ER!!!
+set.seed(22)
+my_rfe <- rfe(y = rf_dat$effect_f, # set dependent variable
+              x = rf_dat %>% select(-effect_f), # set predictor variables
+              size = c(1:2, 4, 6, 8, 10, 13), # sets how many variables are in the overall model
+              # I have 13 total possible variables, so I've chosen increments of 3 to look at.
+              rfeControl = my_ctrl) # pull in control from above
+
+# can you make your model even simpler?
+# the following will pick a model with the smallest number of predictor variables based on the tolerance ("tol") that you specify (how much less than the best are you willing to tolerate?)
+my_size <- pickSizeTolerance(my_rfe$results, metric = "Accuracy", tol = 1, maximize = F)
+# higher tol (~10) gives you less variables
+# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
+pickVars(my_rfe$variables, size = my_size)
+```
+
+```{r}
+# Predict scores using the above 2 variables:
+
+# Create re-finalized training dataset and include all possible variables. 
+rf_dat2 <- multiVar_small_train %>%  select(dose.mg.L.master, organism.group, effect_f)
+
+set.seed(4) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+myrf2 <- randomForest(y = rf_dat2$effect_f, # dependent variable
+  x = rf_dat2 %>%
+    select(-effect_f),
+  importance = T, 
+  proximity = T, 
+  ntrees = 500)
+
+myrf2 # examine the results. 
+```
+
+```{r}
+importance <- as.data.frame(as.table(myrf$importance))
+
+# Nicer ggplot variable importance plot.
+vip_plot_a <- importance %>%
+ filter(Var2 == "MeanDecreaseAccuracy") %>%
+  mutate(Var1 = factor(Var1)) %>%
+  mutate(Var1_f = fct_reorder(Var1, Freq)) %>%
+  ggplot(aes(x = Freq, y = Var1_f)) +
+  geom_point(size = 3, alpha = 0.75, color = "black") +
+  labs(x = "Importance",
+    y = "Variables") +
+  theme_bw()
+
+# vip_plot_b <- importance %>%
+#   filter(Var2 == "MeanDecreaseGini") %>%
+#   mutate(Var1 = factor(Var1)) %>%
+#   mutate(Var1_f = fct_reorder(Var1, Freq)) %>%
+#   ggplot(aes(x = Freq, y = Var1_f)) +
+#   geom_point(size = 3, alpha = 0.75, color = "black") +
+#   labs(x = "Node Purity",
+#     y = "Variables") +
+#   theme_bw()
+
+vip_plot <- vip_plot_a #+ vip_plot_b
+
+vip_plot
+# ggsave("cram_vip_plot.png",
+#      path = "/Users/heilil/Desktop/R_figures",
+#      width = 25,
+#      height = 10,
+#      units = "cm"
+#    )
+```
+GINI importance measures the average gain of purity by splits of a given variable.
+If the variable is useful, it tends to split mixed labeled nodes into pure single class nodes.
+Splitting by a permuted variables tend neither to increase nor decrease node purities.
+Permuting a useful variable, tend to give relatively large decrease in mean gini-gain.
+GINI importance is closely related to the local decision function, that random forest uses to select the best available split.
+Therefore, it does not take much extra time to compute.
+On the other hand, mean gini-gain in local splits, is not necessarily what is most useful to measure, in contrary to change of overall model performance.
+Gini importance is overall inferior to (permutation based) variable importance as it is relatively more biased, more unstable and tend to answer a more indirect question.
+
+### Quantile Regression model
+```{r}
+# Quantile random forest regression mode, instead of looking at the mode of trees, can compare to 10th, 50th, 90th percentiles etc.
+
+# Need to make a new dataset taking the above results of pickVars into account.
+# Create finalized training dataset and include all possible variables. 
+qrf_dat <- multiVar_small_train  %>%
+  select(-effect_predicted, -ID)
+#   select(asci, RdCrsWs, PctAgWs, PctUrbWsRp100, PctOpWsRp100, PctOpWs, DamDensWs, RdDensWs, NABD_DensWs, PctUrbWs, PctUrbCatRp100, RdDensWsRp100, PctOpCat, PctUrbCat, RdDensCat, CBNFWs, PctOpCatRp100, PctAgWsRp100, TRIDensWs, AgKffactWs, FertWs)
+
+ set.seed(20)
+ myqrf <- quantregForest(y = qrf_dat$effect_f, # dependent variable
+              x = qrf_dat %>%
+                  select(-effect_f),
+              importance = T,
+              proximity = T,
+              keep.inbag=T,
+              ntrees = 500)
+
+#predict(myqrf) # automatically presents 10th %tile, median, and 90th %tile
+```
+
+```{r}
+#predict(myqrf, what=c(0.2, 0.3, 0.999)) # to print specific quantiles
+```
+
+```{r}
+plot(myqrf) # plots the results.
+```
+# Again appears to improve after ~100 trees.
+
+### Model Validation
+#### Categorical Response
+
+#### Continuous Response
+```{r}
+# Compare predicted vs. actual results, including by size.
+# Adding lines of slope=1 and linear models to each plot.
+val1 <- multiVar_small_train %>% 
+  ggplot(aes(x = effect_predicted, y = effect_f)) +
+  geom_point(color = "#2A3927", alpha = 0.5) +
+  geom_smooth(method = "lm", se = FALSE, color = "#2A3927") +
+  labs(x = "Predicted Effect",
+    y = "Effect measured",
+    title = "Training Data\nn=613") +
+  geom_abline(intercept = 0, slope = 1) +
+  theme_bw()
+
+val1
+
+lm1 <- lm(cram~cram_predicted, data = mydf2_train2)
+summary(lm1)
+
+val2 <- ggplot(mydf2_test2, aes(x = cram_predicted, y = cram)) +
+  geom_point(color = "#3793EC", alpha = 0.5) +
+  geom_smooth(method = "lm", se = FALSE, color = "#3793EC") +
+  scale_x_continuous(breaks = c(0.5, 0.7, 0.9)) +
+  labs(x = "CRAM predicted",
+    y = "CRAM measured",
+    title = "Testing Data\nn=202") +
+  geom_abline(intercept = 0, slope = 1) +
+  theme_bw()
+
+val2
+
+lm2 <- lm(cram~cram_predicted, data = mydf2_test2)
+summary(lm2)
+
+# Create the full testing + training dataset to plot together.
+mydf2_test2$set <- "Testing"
+mydf2_train2$set <- "Training"
+full_train_test <- bind_rows(mydf2_test2, mydf2_train2) %>%
+  mutate(set_f = factor(set, levels = c("Training", "Testing")))
+
+val3 <- ggplot(full_train_test, aes(x = cram_predicted, y = cram, color = set_f)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "lm", se = FALSE) +
+  scale_color_manual(name = "Set", values = c("#2A3927", "#3793EC"), drop = FALSE) +
+  labs(x = "CRAM predicted",
+    y = "CRAM measured",
+    title = "All Data\nn=815") +
+  geom_abline(intercept = 0, slope = 1, color = "black") +
+  facet_wrap(~PSA6) +
+  theme_bw()
+
+val3
+
+val_fig <- (val1 + val2) /
+  (val3)
+
+val_fig + plot_annotation(
+  title = 'CRAM Random Forest Results',
+  subtitle = 'All modeling performed using StreamCAT datasets.',
+  caption = 'Linear models are colored according to dataset. Lines of slope = 1 are denoted in black.'
+)
+
+# Save figure.
+# ggsave("cram_rfmodel_validation.png",
+#      path = "/Users/heilil/Desktop/R_figures",
+#      width = 35,
+#      height = 25,
+#      units = "cm"
+#    )
+
+lm3 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Central_Valley") %>%
+    filter(set_f == "Training"))
+
+lm4 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Central_Valley") %>%
+    filter(set_f == "Testing"))
+
+lm5 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Chaparral") %>%
+    filter(set_f == "Training"))
+
+lm6 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Chaparral") %>%
+    filter(set_f == "Testing"))
+
+lm7 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Deserts_Modoc") %>%
+    filter(set_f == "Training"))
+
+lm8 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Deserts_Modoc") %>%
+    filter(set_f == "Testing"))
+
+lm9 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "North_Coast") %>%
+    filter(set_f == "Training"))
+
+lm10 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "North_Coast") %>%
+    filter(set_f == "Testing"))
+
+lm11 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Sierra") %>%
+    filter(set_f == "Training"))
+
+lm12 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Sierra") %>%
+    filter(set_f == "Testing"))
+
+lm13 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "South_Coast") %>%
+    filter(set_f == "Training"))
+
+lm14 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "South_Coast") %>%
+    filter(set_f == "Testing"))
+
+# Chose not to compute confusion matrix / accuracy score since this is more applicable to categorical ouputs from random forest models -
+# Instead, calculated Root Mean Squared Error (RMSE) of both training and test datasets.
+# If test RMSE values are much greater than training, then possible the model has been over fit.
+
+predtest <- predict(myrf2, mydf2_test2)
+rmse(mydf2_test2$cram,predtest)
+# 9.03
+
+predtrain <- predict(myrf2, mydf2_train2)
+rmse(mydf2_train2$cram,predtrain)
+# 4.20
+
+# Double checking using the original random forest dataset (rf_dat) with all 35 possible variables included to see where the error in number of predictors starts to increase dramatically (to help double check our decision to include 25 parameters).
+dc <- rfcv(rf_dat %>%
+    select(-cram), 
+  rf_dat$cram,
+  step = 0.7, # default is 0.5
+  scale="log")
+
+dc$error.cv
+#34        24        17        12         8         6         4         3         2 
+# 96.28062  97.29366  99.08743 102.93752 115.52580 120.63049 122.09563 123.74425 139.30181 
+# 1 
+# 206.58921
+
+# Appears between 34 and 24 variables, there is an insignificant increase in error.
+# However, this model is much larger than the CSCI (20) and ASCI (10) models, so we may decide to trim this down in the future.
+
+```
+
+# Works Cited
+
 # Non-linear Dose Response

### 2021-02-04 Stepwise Multiple Regressions.Rmd
a typo
simple change

--- 2021-02-03 Stepwise Multiple Regressions.Rmd
@@
 #Fit the full model
-Full_Model <- lm(effect ~., data = y)
+Full_Model <- lm(effect ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
 
+```{r}
 #Stepwise model

### 2021-02-04 Stepwise Multiple Regressions.Rmd
Factoring charge, particle source, and effect

--- 2021-02-03 Stepwise Multiple Regressions.Rmd
@@
     levels = c("No", "Yes"))) %>%
-  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported")) %>% 
+  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported", charge = "Not Reported")) %>% 
   mutate(size_f = factor(case_when(
@@
     environment == "Terrestrial" ~ "Terrestrial"))) %>%
+  mutate(charge_f = factor(case_when(charge == "negative" ~ "Negative",
+                                     charge == "positive" ~ "Positive",
+                                     charge == "Not Reported" ~ "Not Reported"))) %>%
+  mutate(particle_source_f = factor(case_when(particle.source == "commercial" ~ "Commercial",
+                                              particle.source == "lab" ~ "Lab",
+                                              particle.source == "modified" ~ "Modified",
+                                              particle.source == "N" ~ "Not Reported"))) %>% 
   mutate(species_f = as.factor(paste(genus,species))) %>% 
@@
          bio_f, 
-         charge, 
-         particle.source,
+         charge_f, 
+         particle_source_f,
          org_f,
@@
 #Fit the full model
-Full_Model <- lm(effect ~., data = y)
+Full_Model <- lm(effect_f ~., data = y)
 
@@
          bio_f, 
-         # charge, 
-         particle.source,
+         # charge_f, 
+         particle_source_f,
          org_f,
-         effect)
+         effect_f)
 
@@
 #Fit the full model
-Full_Model <- lm(effect ~., data = y)
+Full_Model <- lm(effect_f ~., data = y)
+
+summary(Full_Model)
 ```
@@
 ```{r, echo = FALSE}
-summary(Full_Model)
 
@@
          bio_f, 
-         charge, 
-         particle.source,
+         charge_f, 
+         particle_source_f,
          org_f,
-         effect)
+         effect_f)
 
@@
          bio_f, 
-         # charge, 
-         # particle.source,
+         # charge_f, 
+         # particle_source_f,
          org_f,
-         effect)
+         effect_f)
 
@@
 ```
+
+# GLM Exploration
+
+Since a linear model does not seem to work well for this data set. I'm going to explore other distribution that may fit the data better.
+
+```{r, echo = TRUE}
+
+#Select the columns that we want to feed into the model for simplicity.
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(dose.particles.mL.master, 
+         dose.mg.L.master, 
+         size_f, 
+         size.length.um.used.for.conversions, 
+         shape_f, 
+         poly_f, 
+         dose.um3.mL.master, 
+         exposure.duration.d, 
+         bio_f, 
+         charge_f, 
+         particle_source_f,
+         org_f,
+         effect_f)
+
+str(aoc_setup_selected)
+
+```

### 2021-02-04 Stepwise Multiple Regressions.Rmd
Update including interaction with effect score

@@
 
-Doses (count, mass and volume), size (continuous), and exposure duration are all log transformed given histogram results.
-
 The model starts with the following independent variables: 
@@
          org_f,
-         effect)
+         effect_f)
 
@@
 
-#Model Refining
+##Model Refining
 
@@
 ```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+#Model - Interaction
+
+The response variable is the interaction between effect and effect score
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+10. Charge (negative/positive)
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity.
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(dose.particles.mL.master, 
+         dose.mg.L.master, 
+         size_f, 
+         size.length.um.used.for.conversions, 
+         shape_f, 
+         poly_f, 
+         dose.um3.mL.master, 
+         exposure.duration.d, 
+         bio_f, 
+         charge_f, 
+         particle_source_f,
+         org_f,
+         effect.score,
+         effect_f)
+
+```
+
+### Full
+
+```{r, echo = FALSE}
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y)
+
+summary(Full_Model)
+```
+
+### Stepwise
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+#Model Refining
+
+##Remove Charge
+
+Complete charge data are limited to PS and PC spheres - remove to include for complete data.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+~~10. Charge (negative/positive)~~
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity.
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(dose.particles.mL.master, 
+         dose.mg.L.master, 
+         size_f, 
+         size.length.um.used.for.conversions, 
+         shape_f, 
+         poly_f, 
+         dose.um3.mL.master, 
+         exposure.duration.d, 
+         bio_f, 
+         # charge_f, 
+         particle_source_f,
+         org_f,
+         effect_f,
+         effect.score)
+
+```
+
+### Full
+
+```{r, echo = FALSE}
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y)
+
+summary(Full_Model)
+```
+
+### Stepwise
+
+```{r, echo = FALSE}
+
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## Remove Charge + Dose (volume)
+
+Remove dose(volume) - highest p value
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+~~3. Dose (volume)~~
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+~~10. Charge (negative/positive)~~
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity.
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(dose.particles.mL.master, 
+         dose.mg.L.master, 
+         size_f, 
+         size.length.um.used.for.conversions, 
+         shape_f, 
+         poly_f, 
+         # dose.um3.mL.master, 
+         exposure.duration.d, 
+         bio_f, 
+         # charge_f, 
+         particle_source_f,
+         org_f,
+         effect_f,
+         effect.score)
+
+```
+
+### Full
+
+```{r, echo = FALSE}
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y)
+
+summary(Full_Model)
+```
+
+### Stepwise
+
+```{r, echo = FALSE}
+
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## Remove Charge + Dose (volume) + Dose (mass)
+
+Remove dose(volume) - highest p value
+
+The model starts with the following independent variables: 
+1. Dose (count)
+~~2. Dose (mass)~~
+~~3. Dose (volume)~~
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+~~10. Charge (negative/positive)~~
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity.
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(dose.particles.mL.master, 
+         # dose.mg.L.master, 
+         size_f, 
+         size.length.um.used.for.conversions, 
+         shape_f, 
+         poly_f, 
+         # dose.um3.mL.master, 
+         exposure.duration.d, 
+         bio_f, 
+         # charge_f, 
+         particle_source_f,
+         org_f,
+         effect_f,
+         effect.score)
+
+```
+
+### Full
+
+```{r, echo = FALSE}
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y)
+
+summary(Full_Model)
+```
+
+### Stepwise
+
+```{r, echo = FALSE}
+
 #Stepwise model
@@
 
+#Model - Log Transformed + Interaction
+
+Doses (count, mass and volume), size (continuous), and exposure duration are all log transformed given histogram results. The response variable is the interaction between effect and effect score.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+10. Charge (negative/positive)
+11. Particle Source
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Log transform numeric values 
+aoc_setup <- aoc_setup %>% 
+  mutate(log10(dose.particles.mL.master)) %>% 
+  mutate(log10(dose.mg.L.master)) %>%
+  mutate(log10(dose.um3.mL.master)) %>%
+  mutate(log10(size.length.um.used.for.conversions)) %>%
+  mutate(log10(exposure.duration.d))
+
+#Select the columns that we want to feed into the model for simplicity and log transform data according to histogram results. 
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(`log10(dose.particles.mL.master)`, 
+         `log10(dose.mg.L.master)`, 
+         size_f, 
+         `log10(size.length.um.used.for.conversions)`, 
+         shape_f, 
+         poly_f, 
+         `log10(dose.um3.mL.master)`, 
+        `log10(exposure.duration.d)`, 
+         bio_f, 
+         charge_f, 
+         particle_source_f,
+         org_f,
+         effect_f,
+         effect.score)
+
+```
+
+### Full
+
+```{r, echo = FALSE}
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y)
+
+summary(Full_Model)
+```
+
+### Stepwise
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+#Model Refining
+
+##Remove Particle Source
+
+Doses (count, mass and volume), size (continuous), and exposure duration are all log transformed given histogram results. The response variable is the interaction between effect and effect score.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+10. Charge (negative/positive)
+~~11. Particle Source~~
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity and log transform data according to histogram results. 
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(`log10(dose.particles.mL.master)`, 
+         `log10(dose.mg.L.master)`, 
+         size_f, 
+         `log10(size.length.um.used.for.conversions)`, 
+         shape_f, 
+         poly_f, 
+         `log10(dose.um3.mL.master)`, 
+        `log10(exposure.duration.d)`, 
+         bio_f, 
+         charge_f, 
+         # particle_source_f,
+         org_f,
+         effect_f,
+         effect.score)
+
+```
+
+### Full
+
+```{r, echo = FALSE}
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y)
+
+summary(Full_Model)
+```
+
+### Stepwise
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+##Remove Particle Source + Organism Group
+
+Doses (count, mass and volume), size (continuous), and exposure duration are all log transformed given histogram results. The response variable is the interaction between effect and effect score.
+
+The model starts with the following independent variables: 
+1. Dose (count)
+2. Dose (mass)
+3. Dose (volume)
+4. Size (continuous)
+5. Size (binned)
+6. Shape
+7. Polymer
+8. Exposure Duration
+9. Level of Biological Organization
+10. Charge (negative/positive)
+~~11. Particle Source~~
+12. Organism Group - substitute for body right now
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity and log transform data according to histogram results. 
+aoc_setup_selected <- aoc_setup %>% 
+  dplyr::select(`log10(dose.particles.mL.master)`, 
+         `log10(dose.mg.L.master)`, 
+         size_f, 
+         `log10(size.length.um.used.for.conversions)`, 
+         shape_f, 
+         poly_f, 
+         `log10(dose.um3.mL.master)`, 
+        `log10(exposure.duration.d)`, 
+         bio_f, 
+         charge_f, 
+         # particle_source_f,
+         # org_f,
+         effect_f,
+         effect.score)
+
+```
+
+### Full
+
+```{r, echo = FALSE}
+#Convert all data to numerical data
+aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_selected_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y)
+
+summary(Full_Model)
+```
+
+### Stepwise
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
 # GLM Exploration

### 2021-02-04 multivariate.Rmd
more modeling

@@
                                    dose.mg.L.master.AF.noec,
-                                   dose.particles.mL.master.AF.noec) %>%  
+                                   dose.particles.mL.master.AF.noec,
+                                   effect.score) %>%  #1 = minor, 2 = photosynthesis, feeding, 3 = growth, chlorophyll content, 4 = reproduction, 5  = population growth, 6 = survival
   filter(!size_f == "Not Reported") %>%   #take out not reported 
-  mutate_if(is.character, as.factor)
-
-# #recode variables
-# multiVar <- multiVar %>% mutate(effect_10 = case_when(
-#     effect_f == "Y" ~ 1,
-#     effect_f == "N" ~ 0
-#   ))# 
+  mutate_if(is.character, as.factor) %>% mutate(effect_10 = case_when(
+     effect_f == "Yes" ~ 1,
+     effect_f == "No" ~ 0))
 ```
@@
 ```{r}
-acute <- multiVar %>% filter(acute.chronic_f == "Chronic") %>% 
+acute <- multiVar %>% 
+  filter(acute.chronic_f == "Acute") %>% 
   filter(!environment == "Terrestrial") %>% 
@@
   filter(!exposure.route == "food") %>% 
-  select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d,bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master)) %>%
+  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
   mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
@@
 ```{r}
-plot(myrf)
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+acute_split <- acute %>%
+  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# default is 3/4ths split (but 75% training, 25% testing).
+# Stratification (strata) = grouping training/testing sets by region, state, etc.
+# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+acute_train <- training(acute_split)
+acute_test <- testing(acute_split)
+# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Random forest -- 
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+acuterf <- randomForest(y = acute_train$effect_f, # dependent variable
+  x = acute_train %>%
+    dplyr::select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+acuterf # examine the results.
+```
+
+```{r}
+plot(acuterf)
 # model performance appears to improve most at ~75 trees
@@
 ```{r}
-varImpPlot(myrf)
+varImpPlot(acuterf)
 # displays which variables are most important
@@
 ```{r}
-importance <- myrf$importance
+importance <- acuterf$importance
 View(importance)
 # displays the data plotted in the plot above
+```
+
+###Validation
+
+```{r fit results}
+fitted <- predict(acuterf, acute_test[,-6], OOB = TRUE, type ="response")
+misClasificError <- mean(fitted != acute_test$effect_f)
+print(paste('Training Accuracy', 1 - misClasificError))
+```
+
+Alternative ROC Curve
+```{r}
+require(pROC)
+predicted <- predict(acuterf, acute_test %>%  dplyr::select(-effect_f),
+                       OOB=TRUE, type= "response")
+#Calculate ROC curve
+rocCurve.tree <- roc(as.numeric(acute_test$effect_f),as.numeric(predicted))
+
+##gplot
+# rocks <- roc()
+
+#plot the ROC curve
+plot(rocCurve.tree,col=c(4))
+```
+##### Chronic
+```{r}
+chronic <- multiVar %>% 
+  filter(acute.chronic_f == "Chronic") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!exposure.route == "food") %>% 
+  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na()
+skim(chronic)
+```
+With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
+Let's determine if we have enough data for ML.
+```{r}
+exp(1)^6
+```
+$e^6$ is less than n (453), so we may proceed.
+
+```{r}
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+chronic_split <- chronic %>%
+  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# default is 3/4ths split (but 75% training, 25% testing).
+# Stratification (strata) = grouping training/testing sets by region, state, etc.
+# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+chronic_train <- training(chronic_split)
+chronic_test <- testing(chronic_split)
+# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Random forest -- 
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+chronicrf <- randomForest(y = chronic_train$effect_f, # dependent variable
+  x = chronic_train %>%
+    dplyr::select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+chronicrf # examine the results.
+```
+
+```{r}
+plot(chronicrf)
+# model performance appears to improve most at ~75 trees
+```
+
+```{r}
+varImpPlot(chronicrf)
+# displays which variables are most important
+# helps to winnow down list of predictors
+# recommended to weigh left pane more
+# right pane also shows how evenly things split based on the list of predictors
+# values close to 0 can be dropped, but don't have to be
+```
+Dose, organism group, exposure duration are important.
+
+```{r}
+importance <- chronicrf$importance
+View(importance)
+# displays the data plotted in the plot above
+```
+
+```{r fit results}
+fitted <- predict(chronicrf, chronic_test[,-6], OOB = TRUE, type ="response")
+misClasificError <- mean(fitted != chronic_test$effect_f)
+print(paste('Training Accuracy', 1 - misClasificError))
+```
+
+#### All Data
+```{r}
+all <- multiVar %>% 
+  #filter(acute.chronic_f == "Chronic") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!exposure.route == "food") %>% 
+  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na()
+skim(all)
+```
+With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
+Let's determine if we have enough data for ML.
+```{r}
+exp(1)^6
+```
+$e^6$ is less than n (453), so we may proceed.
+
+```{r}
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+all_split <- all %>%
+  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# default is 3/4ths split (but 75% training, 25% testing).
+# Stratification (strata) = grouping training/testing sets by region, state, etc.
+# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+all_train <- training(all_split)
+all_test <- testing(all_split)
+# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Random forest -- 
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+allrf <- randomForest(y = all_train$effect_f, # dependent variable
+  x = all_train %>%
+    dplyr::select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+allrf # examine the results.
+```
+
+```{r}
+plot(allrf)
+# model performance appears to improve most at ~75 trees
+```
+
+```{r}
+varImpPlot(allrf)
+# displays which variables are most important
+# helps to winnow down list of predictors
+# recommended to weigh left pane more
+# right pane also shows how evenly things split based on the list of predictors
+# values close to 0 can be dropped, but don't have to be
+```
+Dose, organism group, exposure duration are important.
+
+```{r}
+importance <- allrf$importance
+View(importance)
+# displays the data plotted in the plot above
+```
+
+```{r fit results}
+fitted <- predict(allrf, all_test[,-6], OOB = TRUE, type ="response")
+misClasificError <- mean(fitted != all_test$effect_f)
+print(paste('Training Accuracy', 1 - misClasificError))
+```
+
+#### NO FILTERS
+```{r}
+nofilter <- multiVar %>% 
+  #filter(acute.chronic_f == "Chronic") %>% 
+  ##filter(!environment == "Terrestrial") %>% 
+  #filter(bio.org == "organism") %>% 
+  #filter(lvl1_f == "Fitness") %>% 
+  #filter(!exposure.route == "food") %>% 
+  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na()
+skim(nofilter)
+```
+With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
+Let's determine if we have enough data for ML.
+```{r}
+exp(1)^6
+```
+$e^6$ is less than n (453), so we may proceed.
+
+```{r}
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+nofilter_split <- nofilter %>%
+  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# default is 3/4ths split (but 75% training, 25% testing).
+# Stratification (strata) = grouping training/testing sets by region, state, etc.
+# Using the "strata" cnofilter ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+nofilter_train <- training(nofilter_split)
+nofilter_test <- testing(nofilter_split)
+# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Random forest -- 
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+nofilterrf <- randomForest(y = nofilter_train$effect_f, # dependent variable
+  x = nofilter_train %>%
+    dplyr::select(-effect_f), # selecting nofilter predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+nofilterrf # examine the results.
+```
+
+```{r}
+plot(nofilterrf)
+# model performance appears to improve most at ~75 trees
+```
+
+```{r}
+varImpPlot(nofilterrf)
+# displays which variables are most important
+# helps to winnow down list of predictors
+# recommended to weigh left pane more
+# right pane also shows how evenly things split based on the list of predictors
+# values close to 0 can be dropped, but don't have to be
+```
+Dose, organism group, exposure duration are important.
+
+```{r}
+importance <- nofilterrf$importance
+View(importance)
+# displays the data plotted in the plot above
+```
+
+```{r fit results}
+fitted <- predict(nofilterrf, nofilter_test[,-6], OOB = TRUE, type ="response")
+misClasificError <- mean(fitted != nofilter_test$effect_f)
+print(paste('Training Accuracy', 1 - misClasificError))
+```
+
+#### ROC Curve
+yet anothyer try
+```{r}
+###CHRONIC
+chronicpredictions <- as.data.frame(predict(chronicrf, chronic_test[,-6], type = "prob"))
+# predict class and then attach test class
+chronicpredictions$predict <- names(chronicpredictions)[1:2][apply(chronicpredictions[,1:2], 1, which.max)]
+chronicpredictions$observed <- chronic_test$effect_f
+
+####ACUTE
+predictions <- as.data.frame(predict(acuterf, acute_test[,-6], type = "prob"))
+# predict class and then attach test class
+predictions$predict <- names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
+predictions$observed <- acute_test$effect_f
+head(predictions)
+
+#### all
+allpredictions <- as.data.frame(predict(allrf, all_test[,-6], type = "prob"))
+# predict class and then attach test class
+allpredictions$predict <- names(allpredictions)[1:2][apply(allpredictions[,1:2], 1, which.max)]
+allpredictions$observed <- all_test$effect_f
+head(allpredictions)
+
+###nofilter
+nofilterpredictions <- as.data.frame(predict(nofilterrf, nofilter_test[,-6], type = "prob"))
+# predict class and then attach test class
+nofilterpredictions$predict <- names(nofilterpredictions)[1:2][apply(nofilterpredictions[,1:2], 1, which.max)]
+nofilterpredictions$observed <- nofilter_test$effect_f
+```
+```{r}
+# 1 ROC curve, yes vs no for acute
+roc.acute <- roc(ifelse(predictions$observed=="Yes", "Yes", "No"), as.numeric(predictions$Yes))
+
+#chronic
+roc.chronic <- roc(ifelse(chronicpredictions$observed=="Yes", "Yes", "No"), as.numeric(chronicpredictions$Yes))
+#all
+roc.all <- roc(ifelse(allpredictions$observed=="Yes", "Yes", "No"), as.numeric(allpredictions$Yes))
+
+#nofilter
+roc.nofilter <- roc(ifelse(nofilterpredictions$observed=="Yes", "Yes", "No"), as.numeric(nofilterpredictions$Yes))
+
+all <- ggroc(roc.all) + labs(title = "Chronic and Acute",
+                             subtitle = "Accuracy = 81.7%, n = 1510") +
+  theme_minimal()
+acute <- ggroc(roc.acute, col = "red") + labs(title = "Acute",
+                                              subtitle = "Accuracy = 84.8%, n = 883") + theme_minimal()
+chronic <- ggroc(roc.chronic, col = "blue") + labs(title = "Chronic",
+                                                   subtitle = "Accuracy = 67.3%, n = 452") + theme_minimal()
+
+nofilter<- ggroc(roc.nofilter, col = "green") + labs(title = "Entire Dataset",
+                                                   subtitle = "Accuracy = 79.9%, n = 4615") + theme_minimal()
+require(gridExtra)
+grid.arrange(all, acute, chronic, nofilter,
+             ncol = 4)
+```
+
+```{r}
+#plot reciever operating curve
+perf <- performance(preds,"tpr","fpr")
+```
+
+```{r}
+# compute in-sample results
+caret::confusionMatrix(fitted,as.factor(acute_test$effect_f))
+```
+
+```{r}
+# Validation set assessment #1: looking at confusion matrix
+prediction_for_table <- predict(acuterf, acute_test[,-6])
+table(observed = acute_test[,6], predicted = prediction_for_table)
+```
+
+```{r}
+# Validation set assessment #2: ROC curves and AUC
+# Needs to import ROCR package for ROC curve plotting:
+library(ROCR)
+# Calculate the probability of new observations belonging to each class
+# prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
+prediction_for_roc_curve <- predict(acuterf ,acute_test[,-6],type="response")
+# Use pretty colours:
+pretty_colours <- c("#F8766D","#00BA38","#619CFF")
+# Specify the different classes 
+classes <- levels(acute_test$organism.group)
+# For each class
+for (i in 1:13)
+{
+ # Define which observations belong to class[i]
+ true_values <- ifelse(acute_test[,6]==classes[i],"Yes","No")
+ # Assess the performance of classifier for class[i]
+ pred <- prediction(prediction_for_roc_curve[,i],true_values)
+ perf <- performance(pred, "tpr", "fpr")
+ if (i==1)
+ {
+     plot(perf,main="ROC Curve",col=pretty_colours[i]) 
+ }
+ else
+ {
+     plot(perf,main="ROC Curve",col=pretty_colours[i],add=TRUE) 
+ }
+ # Calculate the AUC and print it to screen
+ auc.perf <- performance(pred, measure = "auc")
+ print(auc.perf@y.values)
+}
 ```

### 2021-02-06 Presentation_Figures.R
Mocked up some static figures for another presentation, fixed some errors in effect score column and got rid of old data set

@@ -1,0 +1,1012 @@
+# Microplastics Toxicity Literature Review
+# Southern California Coastal Water Research Project
+# Presentation Figures Mock-Up Script
+# Created on: November 18, 2020 by Leah Thornton Hampton
+
+#### Packages ####
+
+# Attach packages necessary for the actions below.
+library(tidyverse) # For data tidying/viz purposes.
+library(patchwork) # For knitting separate figures together.
+library(devtools) # For installing packages from github.
+library(calecopal) # For palettes used in figures below.
+# Install by running devtools::install_github("an-bui/calecopal") in console.
+# See available palettes by running names(cal_palettes) in console.
+# Display a palette by running cal_palette("sbchannel") in console.
+
+#### Data Import ####
+
+# Import finalized dataset.
+
+# Guess_max ensures columns with lots of NAs are not imported as logical vectors, but as numeric/double.
+aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
+
+#### Set-Up and Cleaning ####
+
+#This is the same tidying and setup as found in app.R as of 2/1/20
+aoc_work <- aoc %>% 
+  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
+                                     effect == "N" ~ "No"),
+                           levels = c("No", "Yes"))) %>%
+  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported")) %>% 
+  mutate(size_f = factor(case_when(
+    size.category == 1 ~ "1nm < 100nm",
+    size.category == 2 ~ "100nm < 1µm",
+    size.category == 3 ~ "1µm < 100µm",
+    size.category == 4 ~ "100µm < 1mm",
+    size.category == 5 ~ "1mm < 5mm",
+    size.category == 0 ~ "Not Reported"),
+    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% 
+  mutate(shape_f = factor(case_when(
+    shape == "fiber" ~ "Fiber",
+    shape == "fragment" ~ "Fragment",
+    shape == "sphere" ~ "Sphere",
+    shape == "Not Reported" ~ "Not Reported"),
+    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% 
[…]

### 2021-02-06 Stepwise Multiple Regressions.Rmd
Modeling with acute/chronic and life stage added

--- 2021-02-04 Stepwise Multiple Regressions.Rmd
@@
 
-```{r}
+```{r, echo = false, message = FALSE}
 
@@
                                             af.time_noNA == 1 ~ "Chronic",
-                                            af.time_noNA == "Unavailable" ~ "Unavailable")))
+                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>% 
+  mutate(charge_f = factor(case_when(charge == "negative" ~ "Negative",
+                                     charge == "positive" ~ "Positive",
+                                     charge == "Not Reported" ~ "Not Reported"))) %>%
+  mutate(particle_source_f = factor(case_when(particle.source == "commercial" ~ "Commercial",
+                                              particle.source == "lab" ~ "Lab",
+                                              particle.source == "modified" ~ "Modified",
+                                              particle.source == "N" ~ "Not Reported")))
 
@@
 
-```{r}
+#Variables to put into model
+
+* Dose (counts)
+* Dose (mass)
+* Dose (volume)
+* Size (continuous)
+* Size (binned)
+* Polymer type
+* Shape
+* Organism Group (this is being used as a temporary substitute for body size)
+* Life Stage
+* Level of Biological Organization
+* Exposure Duration
+* Acute/Chronic (this values are only complete for fish, molluscs, crustaceans and algae)
+* Charge (negative/positive - categorical)
+* Zeta potential 
+* Particle Source (commercial, generated in-house, or mmodified commercial particles e.g., milling)
+
+Response Variables
+* Effect (Y/N) - Binary Data
+* Effect (Y/N) - Binary Data x Effect Score (binned organism level effects)
+
+#Continous Variable Distributions
+
+There are several continuous variables that we want to feed into our model. Before doing so, we're going to check the distribution to see if any of them are skewed and need to be transformed.
+
+```{r, echo = FALSE}
+
+#Dose - Count
+
+Dose_Particles_mL <- aoc_setup$dose.particles.mL.master
+
+hist(Dose_Particles_mL)
+
+hist(log10(Dose_Particles_mL))
+
+#Dose - Mass
+
+Dose_Mass_L <- aoc_setup$dose.mg.L.master
+
+hist(Dose_Mass_L)
+
+hist(log10(Dose_Mass_L))
+
+#Dose - Volume
+
+`Dose_um^3_mL` <- aoc_setup$dose.um3.mL.master
+
+hist(`Dose_um^3_mL`)
+
+hist(log10(`Dose_um^3_mL`))
+
+#Size - Continuous
+
+Size <- aoc_setup$size.length.um.used.for.conversions
+
+hist(Size)
+
+hist(log10(Size))
+
+#Exposure Duration
+
+Exposure_Duration <- aoc_setup$exposure.duration.d
+
+hist(Exposure_Duration)
+
+hist(log10(Exposure_Duration))
+
+#Zeta Potential
+
+Zeta_Potential_mV <- aoc_setup$zetapotential.mV
+
+hist(Zeta_Potential_mV)
+
+#No need to log transform
+
+aoc_setup_select <- aoc_setup %>% 
+  mutate(log10(dose.particles.mL.master)) %>% 
+  mutate(log10(dose.mg.L.master)) %>%
+  mutate(log10(dose.um3.mL.master)) %>%
+  mutate(log10(size.length.um.used.for.conversions)) %>%
+  mutate(log10(exposure.duration.d))
+
+```
+
+Due to skewed data, the following categories are log10 transformed before modeling:
+
+* Dose (counts)
+* Dose (mass)
+* Dose (volume)
+* Size (continuous)
+* Exposure Duration
+
+# Model: All Independent Variables, Response Variable: Effect (Y/N)
+
+Full Model
+
+```{r, echo = FALSE}
 
 #Select the columns that we want to feed into the model for simplicity
-aoc_setup_selected <- aoc_setup %>% 
-  dplyr::select(dose.particles.mL.master, 
-         dose.mg.L.master, 
+aoc_setup_select_1 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
     size_f,     
-         size.length.um.used.for.conversions, 
+    poly_f, 
     shape_f,
-         poly_f, 
-         dose.um3.mL.master, 
-         max.size.ingest.mm, 
-         exposure.duration.d, 
+    org_f,
+    life_f,
     bio_f,
-         charge, 
+    `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    charge_f,
     zetapotential.mV,
-         particle.source,
-         lvl1_f,
-         lvl2_f,
-         lvl3,
-         effect)
+    particle_source_f,
+    effect_f)
 
 #Convert all data to numerical data
-aoc_setup_selected_num <- data.matrix(aoc_setup_selected)
+aoc_setup_select_num <- data.matrix(aoc_setup_select_1)
 
 #Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_selected_num)
+x <- as.data.frame(aoc_setup_select_num)
 
@@
 #Fit the full model
-Full_Model <- lm(effect ~., data = y,
+Full_Model <- lm(effect_f ~., data = y,
                  na.action = na.omit)
@@
 
-```{r}
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
 #Stepwise model
@@
 ```
+
+## Remove Charge and Zeta Potential
+
+Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select2 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    life_f,
+    bio_f,
+    `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    # charge_f,
+    # zetapotential.mV,
+    particle_source_f,
+    effect_f)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select2)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## Remove Exposure Duration
+
+Exposure duration not significant - removed from model.
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select3 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    life_f,
+    bio_f,
+    # `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    # charge_f,
+    # zetapotential.mV,
+    particle_source_f,
+    effect_f)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select3)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## Remove Size (binned)
+
+Size (binned) not significant - removed from model.
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select4 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    # size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    life_f,
+    bio_f,
+    # `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    # charge_f,
+    # zetapotential.mV,
+    particle_source_f,
+    effect_f)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select4)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## Remove Particle Source and Life Stage
+
+Particle Source and Life Stage - removed from model.
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select5 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    # size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    # life_f,
+    bio_f,
+    # `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    # charge_f,
+    # zetapotential.mV,
+    # particle_source_f,
+    effect_f)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select5)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## All Variables: Acute
+
+Add all variables back in but only use acute data (fish, crustaceans, mollucs, and algae)
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select5 <- aoc_setup_select %>% 
+  filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    size_f,
+    poly_f, 
+    shape_f,
+    org_f,
+    life_f,
+    bio_f,
+    `log10(exposure.duration.d)`,
+    charge_f,
+    zetapotential.mV,
+    particle_source_f,
+    effect_f)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select5)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## All Variables: Chronic
+
+Add all variables back in but only use chronic data (fish, crustaceans, mollucs, and algae)
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select6 <- aoc_setup_select %>% 
+  filter(acute.chronic_f == "Chronic") %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    size_f,
+    poly_f, 
+    shape_f,
+    org_f,
+    life_f,
+    bio_f,
+    `log10(exposure.duration.d)`,
+    charge_f,
+    zetapotential.mV,
+    particle_source_f,
+    effect_f)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select6)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+# Model: All Independent Variables, Response Variable: Effect (Y/N) x Effect Score
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select_1 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    life_f,
+    bio_f,
+    `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    charge_f,
+    zetapotential.mV,
+    particle_source_f,
+    effect_f,
+    effect.score)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select_1)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## Remove Charge and Zeta Potential
+
+Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select_2 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    life_f,
+    bio_f,
+    `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    # charge_f,
+    # zetapotential.mV,
+    particle_source_f,
+    effect_f,
+    effect.score)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select_2)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## Remove Size (binned)
+
+Size (binned) not significant - removed from model.
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select_3 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    # size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    life_f,
+    bio_f,
+    `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    # charge_f,
+    # zetapotential.mV,
+    particle_source_f,
+    effect_f,
+    effect.score)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select_3)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## Remove Particle Source
+
+Particle source not significant - removed from model.
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select_4 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    # size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    life_f,
+    bio_f,
+    `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    # charge_f,
+    # zetapotential.mV,
+    # particle_source_f,
+    effect_f,
+    effect.score)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select_4)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```

### 2021-02-10 Stepwise Multiple Regressions.Rmd
Attempt to subset the data by fitness and acute endpoints at the organismal level added

@@
 date: "2/1/2021"
-output: html_document
+output:
+  html_document: 
+    toc: true
+    toc_depth: 4
+    number_sections: true
+    toc_float: true
 ---
-# Load packages and data import
 
-```{r, echo = FALSE, message = FALSE}
+```{r, echo = FALSE, message = FALSE, warning = FALSE}
 #Packages
@@
 
-# Data tidying
+# Set Up
 
-```{r, echo = false, message = FALSE}
+```{r, echo = FALSE, message = FALSE}
 
@@
     org_f,
+    max.size.ingest.mm,
     life_f,
@@
 ```
+
+# Model: All Independent Variables + Maximum Ingestion Size, Response Variable: Effect (Y/N) x Effect Score
+
+There is some data for the maximum ingestible size within the database. It is limited to **only 9 species**, but I wanted to try it out to see if it shows any promise.
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select_1 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    max.size.ingest.mm,
+    life_f,
+    bio_f,
+    `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    charge_f,
+    zetapotential.mV,
+    particle_source_f,
+    effect_f,
+    effect.score)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select_1)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## Remove Charge and Zeta Potential
+
+Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select_2 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    max.size.ingest.mm,
+    life_f,
+    bio_f,
+    `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    # charge_f,
+    # zetapotential.mV,
+    particle_source_f,
+    effect_f,
+    effect.score)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select_2)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```
+
+## Acute, fitness and organismal endpoints + Charge and Zeta Potential Removed
+
+Same as the previous model, but endpoint type is limited to fitness, biological organization is limited to organismal and exposure duration limited to acute
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+aoc_setup_select_3 <- aoc_setup_select %>% 
+  filter(acute.chronic_f == "Acute") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(bio_f == "Organism") %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    size_f,     
+    poly_f, 
+    shape_f,
+    org_f,
+    max.size.ingest.mm,
+    life_f,
+    # bio_f,
+    `log10(exposure.duration.d)`,
+    # acute.chronic_f,
+    # charge_f,
+    # zetapotential.mV,
+    particle_source_f,
+    effect_f,
+    effect.score)
+
+#Convert all data to numerical data
+aoc_setup_select_num <- data.matrix(aoc_setup_select_3)
+
+#Convert data matrix into a data frame
+x <- as.data.frame(aoc_setup_select_num)
+
+#Drop NAs to get model to be happy - this might be sketchy!?
+y<- na.omit(x)
+
+#Fit the full model
+Full_Model <- lm(effect_f*effect.score ~., data = y,
+                 na.action = na.omit)
+summary(Full_Model)
+```
+
+Stepwise Model - Both Directions
+
+```{r, echo = FALSE}
+#Stepwise model
+step.model <- stepAIC(Full_Model, direction = "both", 
+                      trace = FALSE)
+
+summary(step.model)
+
+```

### 2021-02-10 Stepwise Multiple Regressions.Rmd
A few different attempts to plot probabilty but neither are currently working

@@
 
+```{r}
+
+#Custom function that finds factors with only 1 level
+#Source: https://stackoverflow.com/questions/44200195/how-to-debug-contrasts-can-be-applied-only-to-factors-with-2-or-more-levels-er
+
+debug_contr_error <- function (dat, subset_vec = NULL) {
+  if (!is.null(subset_vec)) {
+    ## step 0
+    if (mode(subset_vec) == "logical") {
+      if (length(subset_vec) != nrow(dat)) {
+        stop("'logical' `subset_vec` provided but length does not match `nrow(dat)`")
+        }
+      subset_log_vec <- subset_vec
+      } else if (mode(subset_vec) == "numeric") {
+      ## check range
+      ran <- range(subset_vec)
+      if (ran[1] < 1 || ran[2] > nrow(dat)) {
+        stop("'numeric' `subset_vec` provided but values are out of bound")
+        } else {
+        subset_log_vec <- logical(nrow(dat))
+        subset_log_vec[as.integer(subset_vec)] <- TRUE
+        } 
+      } else {
+      stop("`subset_vec` must be either 'logical' or 'numeric'")
+      }
+    dat <- base::subset(dat, subset = subset_log_vec)
+    } else {
+    ## step 1
+    dat <- stats::na.omit(dat)
+    }
+  if (nrow(dat) == 0L) warning("no complete cases")
+  ## step 2
+  var_mode <- sapply(dat, mode)
+  if (any(var_mode %in% c("complex", "raw"))) stop("complex or raw not allowed!")
+  var_class <- sapply(dat, class)
+  if (any(var_mode[var_class == "AsIs"] %in% c("logical", "character"))) {
+    stop("matrix variables with 'AsIs' class must be 'numeric'")
+    }
+  ind1 <- which(var_mode %in% c("logical", "character"))
+  dat[ind1] <- lapply(dat[ind1], as.factor)
+  ## step 3
+  fctr <- which(sapply(dat, is.factor))
+  if (length(fctr) == 0L) warning("no factor variables to summary")
+  ind2 <- if (length(ind1) > 0L) fctr[-ind1] else fctr
+  dat[ind2] <- lapply(dat[ind2], base::droplevels.factor)
+  ## step 4
+  lev <- lapply(dat[fctr], base::levels.default)
+  nl <- lengths(lev)
+  ## return
+  list(nlevels = nl, levels = lev)
+  }
+
+```
+
 #Variables to put into model
@@
 
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select_1)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
+#Check for single level factors
+debug_contr_error(aoc_setup_select_1)
+
+#Remove polymer and shape because they only have one level
+aoc_setup_select_1 <- aoc_setup_select %>% 
+  dplyr::select(
+    `log10(dose.particles.mL.master)`, 
+    `log10(dose.mg.L.master)`,
+    `log10(dose.um3.mL.master)`,
+    `log10(size.length.um.used.for.conversions)`, 
+    size_f,     
+    # poly_f, 
+    # shape_f,
+    org_f,
+    life_f,
+    bio_f,
+    `log10(exposure.duration.d)`,
+    acute.chronic_f,
+    charge_f,
+    zetapotential.mV,
+    particle_source_f,
+    effect_f)
+
+#Omit missing data before modeling
+aoc_setup_select_1 <- na.omit(aoc_setup_select_1)
+
+#Fit the full model
+Full_Model <- lm(as.numeric(effect_f) ~., data = aoc_setup_select_1)
+
+summary(Full_Model)
+```
+
+```{r, echo = FALSE}
+
+#Code adapted from: https://stackoverflow.com/questions/36942443/plotting-a-multiple-logistic-regression-for-binary-and-continuous-values-in-r
+
+#Predict probabilities based on model
+aoc_setup_select_1$prob = predict(Full_Model, aoc_setup_select_1, type="response")
+
+#Plot probabilities
+test_plot <- aoc_setup_select_1 %>% 
+  ggplot(aes(x = `log10(size.length.um.used.for.conversions)`, y = prob))+
+  geom_point() +
+  stat_smooth(method = "glm") #This doesn't use the model so I should probably get rid of it
+
+plot(test_plot)
+```
+
+```{r}
+
+#https://blogs.uoregon.edu/rclub/2016/04/05/plotting-your-logistic-regression-models/
+
+Full_Model$coefficients
+
+b0 <- Full_Model$coefficients[1]
+dose_counts <- Full_Model$coefficients[2]
+dose_mass <- Full_Model$coefficients[3]
+dose_vol <- Full_Model$coefficients[4]
+size_con <- Full_Model$coefficients[5]
+size_100nm_1um <- Full_Model$coefficients[6]
+size_1um_100um <- Full_Model$coefficients[7]
+org_bacterium <- Full_Model$coefficients[8]
+org_crutacea <- Full_Model$coefficients[9]
+org_echinoderm <- Full_Model$coefficients[10]
+org_fish <- Full_Model$coefficients[11]
+org_mollusc <- Full_Model$coefficients[12]
+org_nematode <- Full_Model$coefficients[13]
+life_early <- Full_Model$coefficients[14]
+life_juv <- Full_Model$coefficients[15]
+bio_organism <- Full_Model$coefficients[17]
+bio_subcell <- Full_Model$coefficients[18]
+bio_tissue <- Full_Model$coefficients[19]
+exp_dur <- Full_Model$coefficients[20]
+acute_chronic <- Full_Model$coefficients[21]
+charge_positive <- Full_Model$coefficients[23]
+zeta <- Full_Model$coefficients[24]
+source_lab <- Full_Model$coefficients[25]
+
+#Set range for x axis variable
+
+X1_range <- seq(from=min(aoc_setup_select_1$`log10(size.length.um.used.for.conversions)`), to=max(aoc_setup_select_1$`log10(size.length.um.used.for.conversions)`), by=.01)
+
+#Not sure what to do here because I'm not running interactions?
+logits <- b0 +
+  size_con*X1_range +
+  dose_counts +
+  dose_mass +
+  dose_vol +
+  size_100nm_1um +
+  size_1um_100um +
+  org_bacterium +
+  org_crutacea +
+  org_echinoderm +
+  org_fish +
+  org_mollusc +
+  org_nematode +
+  life_early +
+  life_juv +
+  bio_organism +
+  bio_subcell +
+  bio_tissue +
+  exp_dur +
+  acute_chronic +
+  charge_positive +
+  zeta +
+  source_lab
+
+# Compute the probibilities (this is what will actually get plotted):
+a_probs <- exp(logits)/(1 + exp(logits))
+
+#plot
+plot(X1_range, a_probs, 
+     ylim=c(0,1),
+     type="l", 
+     lwd=3, 
+     lty=2, 
+     col="gold", 
+     xlab="X1", ylab="P(effect)", main="Probability of Effect")
+
 ```

### 2021-02-10 Stepwise Multiple Regressions.Rmd

--- 2021-02-10 Stepwise Multiple Regressions.Rmd
@@
 ```
-
-## Acute, fitness and organismal endpoints + Charge and Zeta Potential Removed
-
-Same as the previous model, but endpoint type is limited to fitness, biological organization is limited to organismal and exposure duration limited to acute
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select_3 <- aoc_setup_select %>% 
-  filter(acute.chronic_f == "Acute") %>% 
-  filter(lvl1_f == "Fitness") %>% 
-  filter(bio_f == "Organism") %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    max.size.ingest.mm,
-    life_f,
-    # bio_f,
-    `log10(exposure.duration.d)`,
-    # acute.chronic_f,
-    # charge_f,
-    # zetapotential.mV,
-    particle_source_f,
-    effect_f,
-    effect.score)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select_3)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f*effect.score ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```

### 2021-02-10 Stepwise Multiple Regressions.Rmd
this is a hot mess :(

@@
 library(tidyverse)
-library(MASS) #Stepwise regression package
+library(MASS, exclude = 'select') #Stepwise regression package 
+library(ggiraphExtra) #For plotting
 
@@
     levels = c("No", "Yes"))) %>%
+  mutate(effect_10 = case_when(
+     effect_f == "Yes" ~ 1,
+     effect_f == "No" ~ 0)) %>% 
   replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported")) %>% 
@@
 aoc_setup_select <- aoc_setup %>% 
-  mutate(log10(dose.particles.mL.master)) %>% 
-  mutate(log10(dose.mg.L.master)) %>%
-  mutate(log10(dose.um3.mL.master)) %>%
-  mutate(log10(size.length.um.used.for.conversions)) %>%
-  mutate(log10(exposure.duration.d))
+  mutate(log.dose.particles.mL.master = log10(dose.particles.mL.master)) %>% 
+  mutate(log.dose.mg.L.master = log10(dose.mg.L.master)) %>%
+  mutate(log.dose.um3.mL.master = log10(dose.um3.mL.master)) %>%
+  mutate(log.size.length.um.used.for.conversions = log10(size.length.um.used.for.conversions)) %>%
+  mutate(log.exposure.duration.d = log10(exposure.duration.d))
 
@@
 
-# Model: All Independent Variables, Response Variable: Effect (Y/N)
+# Model: Kitchen Sink
+
+All Independent Variables, Response Variable: Effect (Y/N)
 
@@
 aoc_setup_select_1 <- aoc_setup_select %>%
+  filter(life_f != "Not Reported") %>% 
+  filter(acute.chronic_f != "Unavailable") %>% 
+  filter(size_f != "Not Reported") %>% 
+  filter(poly_f != "Not Reported") %>%
+  filter(shape_f != "Not Reported") %>%
+  filter(environment != "Terrestrial") %>%
   dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    size_f,     
-    poly_f, 
+    log.dose.particles.mL.master, 
+    log.dose.mg.L.master,
+    log.dose.um3.mL.master,
+    log.size.length.um.used.for.conversions,
+    # size_f,
+    # poly_f,
     shape_f,
@@
     bio_f,
-    `log10(exposure.duration.d)`,
+    log.exposure.duration.d,
     acute.chronic_f,
-    charge_f,
-    zetapotential.mV,
-    particle_source_f,
-    effect_f)
+    # charge_f,
+    # zetapotential.mV,
+    # particle_source_f,
+    effect_10)
 
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select_1)
+#Check for single level factors
+debug_contr_error(aoc_setup_select_1)
 
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
+#Omit missing data before modeling
+aoc_setup_select_1 <- na.omit(aoc_setup_select_1)
 
 #Fit the full model
-Full_Model <- lm(effect_f ~., data = y,
-                 na.action = na.omit)
+Full_Model <- lm(effect_10 ~., data = aoc_setup_select_1)
+
 summary(Full_Model)
@@
 ```{r, echo = FALSE}
+
 #Stepwise model
@@
 
-## Remove Charge and Zeta Potential
-
-Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.
+# Model: Crustacean Fitness, Organism Level Endpoints Only
 
@@
 #Select the columns that we want to feed into the model for simplicity
-aoc_setup_select2 <- aoc_setup_select %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    life_f,
-    bio_f,
-    `log10(exposure.duration.d)`,
-    acute.chronic_f,
-    # charge_f,
-    # zetapotential.mV,
-    particle_source_f,
-    effect_f)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select2)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
+m1_crust_simple <- aoc_setup %>%
+  filter(!effect_f == "NA") %>%
+  filter(!acute.chronic_f == "Unavailable") %>%
+  filter(bio.org == "organism") %>%
+  filter(!environment == "Terrestrial") %>%
+  filter(lvl1_f == "Fitness") %>%
+  filter(!size_f == "Not Reported") %>%
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master))
 
 #Fit the full model
-Full_Model <- lm(effect_f ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
+m1_crust_model_dose <- glm(effect_10 ~ logdose.mg.L.master + size.length.um.used.for.conversions, data = m1_crust_simple, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_dose)
+```
+
+```{r}
+
+#Plots
+ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
+
+range(m1_crust_simple$logdose.mg.L.master, na.rm = TRUE)
+#[1] -11.638272   8.171552
+# generate sequence of values between range in increments of 100 to give smooth appearance to model
+xdose <- seq(-11.638272 ,8.171552, 0.1)
+ydose <- predict(m1_crust_model_dose, 
+                 list(logdose.mg.L.master = xdose), 
+                 type = "response")
+#ggplot with logistic data
+ plot(m1_crust_simple$logdose.mg.L.master, m1_crust_simple$effect_10, pch = 16, xlab = "log dose (mg/L)", ylab = "effect")
+# lines(xdose, ydose)
+
+ggplot(m1_crust_simple, aes(x = logdose.mg.L.master, y = effect_10))+
+  geom_point() +
+  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)
+
 ```
@@
 ```
-
-## Remove Exposure Duration
-
-Exposure duration not significant - removed from model.
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select3 <- aoc_setup_select %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    life_f,
-    bio_f,
-    # `log10(exposure.duration.d)`,
-    acute.chronic_f,
-    # charge_f,
-    # zetapotential.mV,
-    particle_source_f,
-    effect_f)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select3)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```
-
-## Remove Size (binned)
-
-Size (binned) not significant - removed from model.
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select4 <- aoc_setup_select %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    # size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    life_f,
-    bio_f,
-    # `log10(exposure.duration.d)`,
-    acute.chronic_f,
-    # charge_f,
-    # zetapotential.mV,
-    particle_source_f,
-    effect_f)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select4)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```
-
-## Remove Particle Source and Life Stage
-
-Particle Source and Life Stage - removed from model.
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select5 <- aoc_setup_select %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    # size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    # life_f,
-    bio_f,
-    # `log10(exposure.duration.d)`,
-    acute.chronic_f,
-    # charge_f,
-    # zetapotential.mV,
-    # particle_source_f,
-    effect_f)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select5)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```
-
-## All Variables: Acute
-
-Add all variables back in but only use acute data (fish, crustaceans, mollucs, and algae)
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select5 <- aoc_setup_select %>% 
-  filter(acute.chronic_f == "Acute") %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    size_f,
-    poly_f, 
-    shape_f,
-    org_f,
-    life_f,
-    bio_f,
-    `log10(exposure.duration.d)`,
-    charge_f,
-    zetapotential.mV,
-    particle_source_f,
-    effect_f)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select5)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```
-
-## All Variables: Chronic
-
-Add all variables back in but only use chronic data (fish, crustaceans, mollucs, and algae)
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select6 <- aoc_setup_select %>% 
-  filter(acute.chronic_f == "Chronic") %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    size_f,
-    poly_f, 
-    shape_f,
-    org_f,
-    life_f,
-    bio_f,
-    `log10(exposure.duration.d)`,
-    charge_f,
-    zetapotential.mV,
-    particle_source_f,
-    effect_f)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select6)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```
-
-# Model: All Independent Variables, Response Variable: Effect (Y/N) x Effect Score
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select_1 <- aoc_setup_select %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    max.size.ingest.mm,
-    life_f,
-    bio_f,
-    `log10(exposure.duration.d)`,
-    acute.chronic_f,
-    charge_f,
-    zetapotential.mV,
-    particle_source_f,
-    effect_f,
-    effect.score)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select_1)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f*effect.score ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```
-
-## Remove Charge and Zeta Potential
-
-Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select_2 <- aoc_setup_select %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    life_f,
-    bio_f,
-    `log10(exposure.duration.d)`,
-    acute.chronic_f,
-    # charge_f,
-    # zetapotential.mV,
-    particle_source_f,
-    effect_f,
-    effect.score)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select_2)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f*effect.score ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```
-
-## Remove Size (binned)
-
-Size (binned) not significant - removed from model.
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select_3 <- aoc_setup_select %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    # size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    life_f,
-    bio_f,
-    `log10(exposure.duration.d)`,
-    acute.chronic_f,
-    # charge_f,
-    # zetapotential.mV,
-    particle_source_f,
-    effect_f,
-    effect.score)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select_3)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f*effect.score ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```
-
-## Remove Particle Source
-
-Particle source not significant - removed from model.
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select_4 <- aoc_setup_select %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    # size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    life_f,
-    bio_f,
-    `log10(exposure.duration.d)`,
-    acute.chronic_f,
-    # charge_f,
-    # zetapotential.mV,
-    # particle_source_f,
-    effect_f,
-    effect.score)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select_4)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f*effect.score ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```
-
-# Model: All Independent Variables + Maximum Ingestion Size, Response Variable: Effect (Y/N) x Effect Score
-
-There is some data for the maximum ingestible size within the database. It is limited to **only 9 species**, but I wanted to try it out to see if it shows any promise.
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select_1 <- aoc_setup_select %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    max.size.ingest.mm,
-    life_f,
-    bio_f,
-    `log10(exposure.duration.d)`,
-    acute.chronic_f,
-    charge_f,
-    zetapotential.mV,
-    particle_source_f,
-    effect_f,
-    effect.score)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select_1)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f*effect.score ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```
-
-## Remove Charge and Zeta Potential
-
-Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-aoc_setup_select_2 <- aoc_setup_select %>% 
-  dplyr::select(
-    `log10(dose.particles.mL.master)`, 
-    `log10(dose.mg.L.master)`,
-    `log10(dose.um3.mL.master)`,
-    `log10(size.length.um.used.for.conversions)`, 
-    size_f,     
-    poly_f, 
-    shape_f,
-    org_f,
-    max.size.ingest.mm,
-    life_f,
-    bio_f,
-    `log10(exposure.duration.d)`,
-    acute.chronic_f,
-    # charge_f,
-    # zetapotential.mV,
-    particle_source_f,
-    effect_f,
-    effect.score)
-
-#Convert all data to numerical data
-aoc_setup_select_num <- data.matrix(aoc_setup_select_2)
-
-#Convert data matrix into a data frame
-x <- as.data.frame(aoc_setup_select_num)
-
-#Drop NAs to get model to be happy - this might be sketchy!?
-y<- na.omit(x)
-
-#Fit the full model
-Full_Model <- lm(effect_f*effect.score ~., data = y,
-                 na.action = na.omit)
-summary(Full_Model)
-```
-
-Stepwise Model - Both Directions
-
-```{r, echo = FALSE}
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
-
-summary(step.model)
-
-```

### 2021-02-10 Stepwise Multiple Regressions.Rmd
interaction term!

@@
 #Select the columns that we want to feed into the model for simplicity
-m1_crust_simple <- aoc_setup %>%
+m1_crust <- aoc_setup %>% 
   filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
   filter(!acute.chronic_f == "Unavailable") %>% 
@@
   filter(!size_f == "Not Reported") %>% 
-  mutate(logdose.mg.L.master = log10(dose.mg.L.master))
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.mg.L.master, size.length.um.used.for.conversions)) %>% 
+  drop_na()
 
 #Fit the full model
-m1_crust_model_dose <- glm(effect_10 ~ logdose.mg.L.master + size.length.um.used.for.conversions, data = m1_crust_simple, na.action = "na.exclude", family = "binomial")
+m1_crust_model_dose <- glm(effect_10 ~ logdose.mg.L.master * size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")
 
@@
 ```{r}
+#Plots
+ggPredict(m1_crust_model_dose,colorn=100,jitter=TRUE)
+```
 
-#Plots
-ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
+```{r}
+#alternative method using ggplot
 
-range(m1_crust_simple$logdose.mg.L.master, na.rm = TRUE)
-#[1] -11.638272   8.171552
-# generate sequence of values between range in increments of 100 to give smooth appearance to model
-xdose <- seq(-11.638272 ,8.171552, 0.1)
-ydose <- predict(m1_crust_model_dose, 
-                 list(logdose.mg.L.master = xdose), 
-                 type = "response")
-#ggplot with logistic data
- plot(m1_crust_simple$logdose.mg.L.master, m1_crust_simple$effect_10, pch = 16, xlab = "log dose (mg/L)", ylab = "effect")
-# lines(xdose, ydose)
+#get probabilities
+m1 <- m1_crust %>% 
+  mutate(prob = predict(m1_crust_model_dose,
+                        type = "response"))
 
-ggplot(m1_crust_simple, aes(x = logdose.mg.L.master, y = effect_10))+
-  geom_point() +
-  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)
-
+# plot
+ggplot(m1, aes(x = logdose.mg.L.master, y = effect_10, color = size.length.um.used.for.conversions))+
+  geom_jitter() +
+  geom_smooth(method = 'glm',
+              aes(x = logdose.mg.L.master, y = prob), #use predicted NOEC probability
+              method.args = list(family = binomial(link = 'logit')),
+              se = FALSE, color = 'red') #+
+  # geom_smooth(method = 'glm',
+  #             method.args = list(family = binomial(link = 'probit')),
+  #             se = FALSE, color = 'green') +
+  # stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)
 ```

### 2021-02-10 Stepwise Multiple Regressions.Rmd
survival plots

@@
 #Plots
-ggPredict(m1_crust_model_dose,colorn=100,jitter=TRUE)
+ggPredict(m1_crust_model_dose,colorn=100,jitter=TRUE, interactive = FALSE, digits = 3, se = FALSE)
 ```
@@
 ```
+```{r}
+#alternative method using ggpredict
+require(ggeffects)
+require(ggdark)
+require(ggsci)
+ggpredict(m1_crust_model_dose)
+df <- ggpredict(m1_crust_model_dose, se = TRUE)
+df <- ggpredict(m1_crust_model_dose,terms=c("logdose.mg.L.master", "size.length.um.used.for.conversions"))
+
+#note that faceting works best with discrete variable as interactive term
+ggplot(df, aes(x = x, y = predicted, colour = group)) +
+  geom_line() +
+  geom_ribbon( aes(ymin = conf.low, ymax = conf.high, fill = group, color = NULL), alpha = .15) +
+  facet_wrap(~.) +
+  dark_theme_bw() +
+  scale_color_tron()
+```
+
+```{r}
+#alt method
+plot(df, facet = TRUE)
+```
+
+## Discrete predictor and dose
+```{r, echo = FALSE}
+#Select the columns that we want to feed into the model for simplicity
+m1_crust <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.mg.L.master, size_f)) %>% 
+  drop_na()
+
+#Fit the full model
+m1_crust_model_discrete <- glm(effect_10 ~ logdose.mg.L.master * size_f, data = m1_crust, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_discrete)
+```
+Plot.
+```{r}
+discrete <- ggpredict(m1_crust_model_discrete,terms = c("logdose.mg.L.master", "size_f"))
+plot(discrete, facet = TRUE, add.data = TRUE)
+```
+### Survival package
+```{r}
+#include time
+m1_crust <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.mg.L.master, size_f, exposure.duration.d)) %>% 
+  drop_na()
+
+require(survival)
+survival <- survival::coxph(survival::Surv(exposure.duration.d, effect_10) ~ logdose.mg.L.master + size_f, data = m1_crust)
+#cumulative hazard
+cumHaz <- ggpredict(survival,terms = c("logdose.mg.L.master", "size_f"), type = "cumhaz")
+plot(cumHaz, facet = TRUE, add.data = TRUE, colors = "flat")
+```
+```{r}
+#survival probability over time
+pr <- ggpredict(survival, c("logdose.mg.L.master", "size_f"), type = "surv")
+plot(pr, colors = "social")
+```
 

### 2021-02-10 Stepwise Multiple Regressions.Rmd
Example Plots for the interactions between dose and size

--- 2021-02-10 Stepwise Multiple Regressions.Rmd
@@
 # Model: Crustacean Fitness, Organism Level Endpoints Only
+## Dose (mass)
 
@@
   filter(!size_f == "Not Reported") %>% 
-  mutate(logdose.mg.L.master = log10(dose.mg.L.master))
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
+  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
+  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
+  drop_na()
 
 #Fit the full model
-m1_crust_model_dose <- glm(effect_10 ~ logdose.mg.L.master + size.length.um.used.for.conversions, data = m1_crust_simple, na.action = "na.exclude", family = "binomial")
+m1_crust_model_dose <- glm(effect_10 ~ logdose.mg.L.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")
 
@@
 ```{r}
-
 #Plots
 ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
-
-range(m1_crust_simple$logdose.mg.L.master, na.rm = TRUE)
-#[1] -11.638272   8.171552
-# generate sequence of values between range in increments of 100 to give smooth appearance to model
-xdose <- seq(-11.638272 ,8.171552, 0.1)
-ydose <- predict(m1_crust_model_dose, 
-                 list(logdose.mg.L.master = xdose), 
-                 type = "response")
-#ggplot with logistic data
- plot(m1_crust_simple$logdose.mg.L.master, m1_crust_simple$effect_10, pch = 16, xlab = "log dose (mg/L)", ylab = "effect")
-# lines(xdose, ydose)
-
-ggplot(m1_crust_simple, aes(x = logdose.mg.L.master, y = effect_10))+
-  geom_point() +
-  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)
-
 ```
 
-Stepwise Model - Both Directions
+## Dose (count)
+
+Full Model
 
@@
 
-#Stepwise model
-step.model <- stepAIC(Full_Model, direction = "both", 
-                      trace = FALSE)
+#Select the columns that we want to feed into the model for simplicity
+m1_crust <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
+  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
+  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
+  drop_na()
 
-summary(step.model)
+#Fit the full model
+m1_crust_model_dose <- glm(effect_10 ~ logdose.particles.mL.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")
 
+summary(m1_crust_model_dose)
 ```
+
+```{r}
+#Plots
+ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
+```
+
+## Dose (volume)
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+m1_crust <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
+  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
+  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
+  drop_na()
+
+#Fit the full model
+m1_crust_model_dose <- glm(effect_10 ~ logdose.um3.mL.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_dose)
+```
+
+```{r}
+#Plots
+ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
+```

### 2021-02-10 Stepwise Multiple Regressions.Rmd

@@
 #Plots
-ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
-```
 
-## Dose (count)
-
-Full Model
-
-```{r, echo = FALSE}
-
-#Select the columns that we want to feed into the model for simplicity
-m1_crust <- aoc_setup %>% 
-  filter(!effect_f == "NA") %>% 
-  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
-  #filter(organism.group == "Crustacea") %>% 
-  filter(!acute.chronic_f == "Unavailable") %>% 
-  filter(bio.org == "organism") %>% 
-  filter(!environment == "Terrestrial") %>% 
-  filter(lvl1_f == "Fitness") %>% 
-  filter(!size_f == "Not Reported") %>% 
-  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
-  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
-  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
- # filter(acute.chronic_f == "Acute") %>% 
-  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
-  drop_na()
-
-#Fit the full model
-m1_crust_model_dose <- glm(effect_10 ~ logdose.particles.mL.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")
-
-summary(m1_crust_model_dose)
-```
-
-```{r}
-#Plots
 ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
@@
 ```
+
+## Dose (count)
+
+Full Model
+
+```{r, echo = FALSE}
+
+#Select the columns that we want to feed into the model for simplicity
+m1_crust <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
+  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
+  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
+  drop_na()
+
+#Fit the full model
+m1_crust_model_dose <- glm(effect_10 ~ logdose.particles.mL.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_dose)
+
+ggPredict(m1_crust_model_dose,colorn=100,jitter=TRUE, interactive = FALSE, digits = 3, se = FALSE)
+
+```
+
+```{r}
+#Plots
+ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
+```
+```{r}
+#alternative method using ggpredict
+require(ggeffects)
+require(ggdark)
+require(ggsci)
+ggpredict(m1_crust_model_dose)
+df <- ggpredict(m1_crust_model_dose, se = TRUE)
+df <- ggpredict(m1_crust_model_dose,terms=c("logdose.mg.L.master", "size.length.um.used.for.conversions"))
+
+#note that faceting works best with discrete variable as interactive term
+ggplot(df, aes(x = x, y = predicted, colour = group)) +
+  geom_line() +
+  geom_ribbon( aes(ymin = conf.low, ymax = conf.high, fill = group, color = NULL), alpha = .15) +
+  facet_wrap(~.) +
+  dark_theme_bw() +
+  scale_color_tron()
+```
+
+```{r}
+#alt method
+plot(df, facet = TRUE)
+```
+
+## Discrete predictor and dose
+```{r, echo = FALSE}
+#Select the columns that we want to feed into the model for simplicity
+m1_crust <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.mg.L.master, size_f)) %>% 
+  drop_na()
+
+#Fit the full model
+m1_crust_model_discrete <- glm(effect_10 ~ logdose.mg.L.master * size_f, data = m1_crust, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_discrete)
+```
+Plot.
+```{r}
+discrete <- ggpredict(m1_crust_model_discrete,terms = c("logdose.mg.L.master", "size_f"))
+plot(discrete, facet = TRUE, add.data = TRUE)
+```
+### Survival package
+```{r}
+#include time
+m1_crust <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.mg.L.master, size_f, exposure.duration.d)) %>% 
+  drop_na()
+
+require(survival)
+survival <- survival::coxph(survival::Surv(exposure.duration.d, effect_10) ~ logdose.mg.L.master + size_f, data = m1_crust)
+#cumulative hazard
+cumHaz <- ggpredict(survival,terms = c("logdose.mg.L.master", "size_f"), type = "cumhaz")
+plot(cumHaz, facet = TRUE, add.data = TRUE, colors = "flat")
+```
+```{r}
+#survival probability over time
+pr <- ggpredict(survival, c("logdose.mg.L.master", "size_f"), type = "surv")
+plot(pr, colors = "social")
+```

### 2021-02-10 Stepwise Multiple Regressions.Rmd
particles and volume

--- 2021-02-10 Stepwise Multiple Regressions.Rmd
@@
 #Plots
-ggPredict(m1_crust_model_dose,colorn=100,jitter=TRUE, interactive = FALSE, digits = 3, se = FALSE)
+ggPredict(m1_crust_model_dose,
+          colorn=200,
+ #         jitter=TRUE, 
+point = FALSE,
+          interactive = FALSE, 
+show.summary = TRUE,
+#digits = 3, 
+se = FALSE)
+```
+
+```{r}
+#particle couunt
+#Select the columns that we want to feed into the model for simplicity
+m1_crust_part <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>% 
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.particles.mL.master, size.length.um.used.for.conversions)) %>% 
+  drop_na()
+
+#Fit the full model
+m1_crust_model_particles <- glm(effect_10 ~ logdose.particles.mL.master * size.length.um.used.for.conversions, data = m1_crust_part, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_particles)
+```
+plot
+```{r}
+ggPredict(m1_crust_model_particles, colorn=1000)
+```
+### volume
+```{r}
+#particle couunt
+#Select the columns that we want to feed into the model for simplicity
+m1_crust_volume <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>% 
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
+  drop_na()
+
+#Fit the full model
+m1_crust_model_volume <- glm(effect_10 ~ logdose.um3.mL.master * size.length.um.used.for.conversions, data = m1_crust_volume, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_volume)
+```
+plot
+```{r}
+ggPredict(m1_crust_model_volume, colorn=500)
 ```

### 2021-02-10 Stepwise Multiple Regressions.Rmd
minor change to make sure plots match

--- 2021-02-10 Stepwise Multiple Regressions.Rmd
@@
 #Plots
+ggPredict(m1_crust_model_dose,
+          colorn=200,
+ #         jitter=TRUE, 
+point = FALSE,
+          interactive = FALSE, 
+show.summary = TRUE,
+#digits = 3, 
+se = FALSE)
+```
+
+```{r}
+#particle couunt
+#Select the columns that we want to feed into the model for simplicity
+m1_crust_part <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>% 
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.particles.mL.master, size.length.um.used.for.conversions)) %>% 
+  drop_na()
+
+#Fit the full model
+m1_crust_model_particles <- glm(effect_10 ~ logdose.particles.mL.master * size.length.um.used.for.conversions, data = m1_crust_part, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_particles)
+```
+plot
+```{r}
+ggPredict(m1_crust_model_particles, colorn=1000)
+```
+### volume
+```{r}
+#particle couunt
+#Select the columns that we want to feed into the model for simplicity
+m1_crust_volume <- aoc_setup %>% 
+  filter(!effect_f == "NA") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  #filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>% 
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
+  drop_na()
+
+#Fit the full model
+m1_crust_model_volume <- glm(effect_10 ~ logdose.um3.mL.master * size.length.um.used.for.conversions, data = m1_crust_volume, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_volume)
+```
+plot
+```{r}
+ggPredict(m1_crust_model_volume, colorn=500)
+```
+
+```{r}
+#alternative method using ggplot
+
+#get probabilities
+m1 <- m1_crust %>% 
+  mutate(prob = predict(m1_crust_model_dose,
+                        type = "response"))
+
+# plot
+ggplot(m1, aes(x = logdose.mg.L.master, y = effect_10, color = size.length.um.used.for.conversions))+
+  geom_jitter() +
+  geom_smooth(method = 'glm',
+              aes(x = logdose.mg.L.master, y = prob), #use predicted NOEC probability
+              method.args = list(family = binomial(link = 'logit')),
+              se = FALSE, color = 'red') #+
+  # geom_smooth(method = 'glm',
+  #             method.args = list(family = binomial(link = 'probit')),
+  #             se = FALSE, color = 'green') +
+  # stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)
+```
+
+```{r}
 ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
@@
 ```
+
 ```{r}

### 2021-02-11 app.R
added shape widget to ssd

@@ -763,7 +763,10 @@ column(width = 12,
                                               selected = levels(aoc_z$size_f),
                                               options = list(`actions-box` = TRUE), # option to de/select all
                                               multiple = TRUE)), # allows for multiple inputs
-                          
+                           #Morphology widget
+                           column(width = 3,
+                                  htmlOutput("shapeSelection")), # dependent on sizes
+                           
                             #Endpoint widget
                            column(width = 3,
                                   htmlOutput("lvl1Selection")), # allows for multiple inputs
@@ -922,13 +925,30 @@ column(width = 12,
                               p("Please be patient as maximum likelihood estimations are calculated. If a high number of boostrap simulations are chosen (>100), this may take up to several minutes."),
                               br(),
                               h4("Species Sensitivity Distribution", align = "center"),
-                              plotOutput(outputId = "aoc_ssd_ggplot", width = "160%", height = "500px", hover = hoverOpts(id = "plot_hover")),
+                              plotOutput(outputId = "aoc_ssd_ggplot", width = "140%", height = "500px", hover = hoverOpts(id = "plot_hover")),
                               verbatimTextOutput("info"),
                               br(),
-                              column(width = 12,
+                              # column(width = 12,
+                              #        column(width = 4,
+                              #                pickerInput(inputId = "user_width",
+                              #                            label = "Plot Width (in.)",
+                              #                            value = 8,
+                              #                            min = 2,
+                              #                            max = 18,
+                              #                            step = 1)),
+                              #        column(width = 4,
+                              #               pickerInput(inputId = "user_heigth",
+                              #                           label = "Plot Height (in.)",
+                              #                           value =10,
+                              #                           min = 2,
[…]

### 2021-02-17 app.R
code chunk for possible dynamic plots

@@ -1757,7 +1757,16 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
   }
   print(p)
  
-  })
+  },
+  
+  # #This is based on a slider widget input
+  # width = function() {
+  #   input$cols * 300
+  # }, height = function() {
+  #   input$rows * 400
+  # }
+  
+  )
   
   # Create downloadable csv of filtered dataset.
   # Removed columns created above so the dataset matches Leah's original dataset.

### 2021-02-18 multivariate.Rmd
cleaned up and added an imputed RF

--- 2021-02-03 multivariate.Rmd
@@
 date: "1/19/2021"
-output: html_document
+output: 
+  html_document:
+    code_folding: hide
+    theme: sandstone
+    toc: yes
+    toc_float: yes
+    toc_depth: 4
+    number_sections: true
+  word_document:
+    toc: yes
 ---
+```{r global-options, include=FALSE}
+knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
+                      warning=FALSE, message=FALSE, echo = FALSE)
+```
+---
+
 # Setup
@@
 library(scales) #To use "percent" function
-library(shinyjs) #Exploration tab - reset button
 library(tigerstats) #turns things into percents
 library(ggbeeswarm) #plot all points
-library(fitdistrplus) #alt SSD 
+library(fitdistrplus, exclude = 'select') #alt SSD 
 library(RColorBrewer) #colors
@@
 library(ISLR)  #for the Carseat Data
+library(lme4) #general linearized mixed model GLMM
+library(quantregForest)
+library(caret)
+library(tidyverse)
+library(tidymodels)
+library(skimr)
+library(sf)
+library(ggspatial)
+library(nhdplusTools)
+library(patchwork)
+library(Metrics)
+library(gt)
+library(randomForest)
 ```
@@
 # Load finalized dataset.
-aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
+aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) %>% 
+  mutate(ID = paste0("ID",row_number()))
 
@@
   mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
-  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported))
+  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
+   mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) %>%   #calculate volume/mL
+  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
+  mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
+                                            af.time_noNA == 1 ~ "Chronic",
+                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>%    #factorize assesment factor time into chronic/acute
+  mutate(dose.mg.L.master.AF.noec = dose.mg.L.master * af.noec) %>% 
+  mutate(dose.particles.mL.master.AF.noec = dose.particles.mL.master * af.noec) %>% 
+  mutate(effect_f = factor(effect)) %>% 
+  mutate_if(is.character, as.factor) %>% 
+  mutate(effect_10 = case_when(
+     effect_f == "Y" ~ 1,
+     effect_f == "N" ~ 0))
 
@@
                                     effect, #yes no
+                                    effect_f,
+                                    effect_10,
+                                    size_f,
                                     exposure.duration.d, 
@@
                                     zetapotential.mV, # numeric   
-                                   max.size.ingest.mm) %>%  #max ingestible size
+                                   max.size.ingest.mm,#max ingestible size
+                                   acute.chronic_f,
+                                   dose.mg.L.master.AF.noec,
+                                   dose.particles.mL.master.AF.noec,
+                                   max.size.ingest.mm, #maximum ingestible size range
+                                   effect.score) %>%  #1 = minor, 2 = photosynthesis, feeding, 3 = growth, chlorophyll content, 4 = reproduction, 5  = population growth, 6 = survival
+  filter(!size_f == "Not Reported") %>%   #take out not reported 
+                                 #  max.size.ingest.mm) %>%  #max ingestible size
   filter(!size_f == "Not Reported")  #take out not reported
@@
 #recode variables
-multiVar <- multiVar %>% mutate(effect_10 = case_when(
-    effect == "Y" ~ 1,
-    effect == "N" ~ 0
-  ))# %>% 
-  #mutate_all(is.character, ~as.factor())
+# multiVar <- multiVar %>% mutate(effect_10 = case_when(
+#     effect == "Y" ~ 1,
+#     effect == "N" ~ 0
+#   ))# %>% 
+#   #mutate_all(is.character, ~as.factor())
 
@@
 ### Completeness vby Size
-```{r completeness heatmap}
+```{r completeness heatmap 1}
 CompletenessSummary_size <- multiVar %>%
@@
 ```{r}
-require()
+require(pheatmap)
 #convert to matrix and transpose
@@
 ### Completeness by Environment
-```{r completeness heatmap}
+```{r completeness heatmap 2}
 CompletenessSummary_environment <- multiVar %>%
@@
 ## Estimate an OLS Regression
-fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master +mouth, 
+fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master, 
              na.action = na.omit, 
@@
 
-## PCA
+# PCA
 
@@
 ```{r}
-# prune the tree
-pt1 <- prune(t1,cp)
-#pt1<- prune(t1, cp= t1$cptable[which.min(t1$cptable[,"xerror"]),"CP"])
-
-# plot the pruned tree
-plot(pt1, uniform=TRUE,
-   main="Pruned Classification Tree");text(pt1, use.n=TRUE, all=TRUE, cex=.8)
-
-#post(pfit, file = "c:/ptree.ps",
- #  title = "Pruned Classification Tree for Kyphosis")
-```
-```{r}
-#find proportion of correct predictions using test set
-t1_pruned_predict <- predict(pt1, testSet, type="class")
-mean(t1_pruned_predict == testSet$effect)
+# # prune the tree
+# pt1 <- prune(t1,cp)
+# pt1<- prune(t1, cp= t1$cptable[which.min(t1$cptable[,"xerror"]),"CP"])
 # 
+# # plot the pruned tree
+# plot(pt1, uniform=TRUE,
+#    main="Pruned Classification Tree");text(pt1, use.n=TRUE, all=TRUE, cex=.8)
+# 
+# #post(pfit, file = "c:/ptree.ps",
+#  #  title = "Pruned Classification Tree for Kyphosis")
+```
+```{r}
+# #find proportion of correct predictions using test set
+# t1_pruned_predict <- predict(pt1, testSet, type="class")
+# mean(t1_pruned_predict == testSet$effect)
+# #
 ```
@@
 ```{r}
-#function to do multiple runs
-multiple_runs_classification <- function(train_fraction,n,dataset,prune_tree=FALSE){
-fraction_correct <- rep(NA,n)
-set.seed(42)
-for (i in 1:n){
-  dataset[,"train"] <- ifelse(runif(nrow(dataset))<0.8,1,0)
-  trainColNum <- grep("train",names(dataset))
-  typeColNum <- grep("effect",names(dataset))
-  trainSet <- dataset[dataset$train==1,-trainColNum]
-  testSet <- dataset[dataset$train==0,-trainColNum]
-  rpart_model <- rpart(effect~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,data = trainSet, method="class")
-if(prune_tree==FALSE) {
-  rpart_test_predict <- predict(rpart_model,testSet[,-typeColNum],type="class")
-  fraction_correct[i] <- mean(rpart_test_predict==testSet$effect)
-  }else{
-    opt <- which.min(rpart_model$cptable[,"xerror"])
-    cp <- rpart_model$cptable[opt, "CP"]
-    pruned_model <- prune(rpart_model,cp)
-    rpart_pruned_predict <- predict(pruned_model,testSet[,-typeColNum],type="class")
-    fraction_correct[i] <- mean(rpart_pruned_predict == testSet$effect)
-  }
-  }
-return(fraction_correct)
-}
-```
-
-Note that in the above,  I have set the default value of the prune_tree to FALSE, so the function will execute the first branch of the if statement unless the default is overridden.
-
-OK, so let’s do 50 runs with and without pruning, and check the mean and variance of the results for both sets of runs.
-```{r}
-#50 runs, no pruning
-unpruned_set <- multiple_runs_classification(0.8,50,multiVar_sub)
-mean(unpruned_set)
-#[1] 0.7261882
-sd(unpruned_set)
-#[1] 0.01554347
-#50 runs, with pruning
-pruned_set <- multiple_runs_classification(0.8,50,multiVar_sub,prune_tree=TRUE)
-mean(pruned_set)
-#[1] 0.7261875
-sd(pruned_set)
-#[1] 0.01552285
+# <!-- # #function to do multiple runs -->
+# <!-- # multiple_runs_classification <- function(train_fraction,n,dataset,prune_tree=FALSE){ -->
+# <!-- # fraction_correct <- rep(NA,n) -->
+# <!-- # set.seed(42) -->
+# <!-- # for (i in 1:n){ -->
+# <!-- #   dataset[,"train"] <- ifelse(runif(nrow(dataset))<0.8,1,0) -->
+# <!-- #   trainColNum <- grep("train",names(dataset)) -->
+# <!-- #   typeColNum <- grep("effect",names(dataset)) -->
+# <!-- #   trainSet <- dataset[dataset$train==1,-trainColNum] -->
+# <!-- #   testSet <- dataset[dataset$train==0,-trainColNum] -->
+# <!-- #   rpart_model <- rpart(effect~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,data = trainSet, method="class") -->
+# <!-- # if(prune_tree==FALSE) { -->
+# <!-- #   rpart_test_predict <- predict(rpart_model,testSet[,-typeColNum],type="class") -->
+# <!-- #   fraction_correct[i] <- mean(rpart_test_predict==testSet$effect) -->
+# <!-- #   }else{ -->
+# <!-- #     opt <- which.min(rpart_model$cptable[,"xerror"]) -->
+# <!-- #     cp <- rpart_model$cptable[opt, "CP"] -->
+# <!-- #     pruned_model <- prune(rpart_model,cp) -->
+# <!-- #     rpart_pruned_predict <- predict(pruned_model,testSet[,-typeColNum],type="class") -->
+# <!-- #     fraction_correct[i] <- mean(rpart_pruned_predict == testSet$effect) -->
+# <!-- #   } -->
+# <!-- #   } -->
+# <!-- # return(fraction_correct) -->
+# <!-- # } -->
+# <!-- # ``` -->
+#  
+# <!-- # Note that in the above,  I have set the default value of the prune_tree to FALSE, so the function will execute the first branch of the if statement unless the default is overridden. -->
+# <!-- #  -->
+# <!-- # OK, so let’s do 50 runs with and without pruning, and check the mean and variance of the results for both sets of runs. -->
+# <!-- # ```{r} -->
+# <!-- # #50 runs, no pruning -->
+# <!-- # unpruned_set <- multiple_runs_classification(0.8,50,multiVar_sub) -->
+# <!-- # mean(unpruned_set) -->
+# <!-- # #[1] 0.7261882 -->
+# <!-- # sd(unpruned_set) -->
+# <!-- # #[1] 0.01554347 -->
+# <!-- # #50 runs, with pruning -->
+# <!-- # pruned_set <- multiple_runs_classification(0.8,50,multiVar_sub,prune_tree=TRUE) -->
+# <!-- # mean(pruned_set) -->
+# <!-- # #[1] 0.7261875 -->
+# <!-- # sd(pruned_set) -->
+# <!-- # #[1] 0.01552285 -->
+# <!-- # ``` -->
 ```
@@
 
-```{r fit results}
+```{r fit results 1}
 train2 <- multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
@@
 
-```{r}
-#Obtaining predicted probabilites for Test data
-tree.probs=predict(crf,
-                 newdata = train2,
-                 type="prob")
-
-#Calculate ROC curve
-rocCurve.tree <- roc(train3$effect_10,tree.probs[,'1'])
-#plot the ROC curve
-plot(rocCurve.tree,col=c(4))
-```
+<!-- # ```{r} -->
+<!-- # #Obtaining predicted probabilites for Test data -->
+<!-- # tree.probs=predict(crf, -->
+<!-- #                  newdata = train2, -->
+<!-- #                  type="prob") -->
+<!-- #  -->
+<!-- # #Calculate ROC curve -->
+<!-- # rocCurve.tree <- roc(train3$effect_10,tree.probs[,'1']) -->
+<!-- # #plot the ROC curve -->
+<!-- # plot(rocCurve.tree,col=c(4)) -->
+<!-- # ``` -->
 
@@
 ```
-```{r}
-#plot reciever operating curve
-perf <- performance(preds,"tpr","fpr")
-```
 
@@
 
-```{r}
-crf.prob <- matrix(unlist(crfsrvy.prob), ncol=2, byrow=TRUE)
-apply(crf.prob,2,mean)
-#[1] 0.2524514 0.7475486
-tab <- round(cbind(by(rpart.prob[,2], INDICES=t1$where, mean), by(crf.prob[,2], INDICES=t1$where, mean)), 4)
-colnames(tab) <- c("rpart", "cforest")
-kable(tab)
-```
-# Non-linear Dose Response
-If we hold all paramaters constant except for dose, we should observe a non-linear dose-response behavior that can be best predicted by a 3- or 4-parameter Hill Equation.
-The most available information is likely for polystyrene spheres in crustaceans.There are 21 studies available.
+<!-- # ```{r} -->
+<!-- # crf.prob <- matrix(unlist(crfsrvy.prob), ncol=2, byrow=TRUE) -->
+<!-- # apply(crf.prob,2,mean) -->
+<!-- # #[1] 0.2524514 0.7475486 -->
+<!-- # tab <- round(cbind(by(rpart.prob[,2], INDICES=t1$where, mean), by(crf.prob[,2], INDICES=t1$where, mean)), 4) -->
+<!-- # colnames(tab) <- c("rpart", "cforest") -->
+<!-- # kable(tab) -->
+<!-- # ``` -->
+
+#GLMM
+```{r}
+#tutorial https://aosmith.rbind.io/2020/08/20/simulate-binomial-glmm/
+require(lme4)
+
+mod = glm(effect_f ~ size.length.um.used.for.conversions + polymer + particle.volume.um3 + density.mg.um.3 + organism.group + bio.org + treatments + effect_f + exposure.duration.d + exposure.route + lvl1_f + dose.mg.L.master,
+            data = aoc_z,
+            family = binomial(link = "logit") )
+mod
+```
+
+```{r}
+bin_glmm_fun = function(n_sites = 10,
+                        b0 = 0,
+                        b1 = 1.735,
+                        num_samp = 50,
+                        site_var = 0.5) {
+     site = rep(LETTERS[1:n_sites], each = 2)
+     plot = paste(site, rep(1:2, times = n_sites), sep = "." )
+     treatment = rep( c("treatment", "control"), times = n_sites)
+     dat = data.frame(site, plot, treatment)
+
+     site_eff = rep( rnorm(n = n_sites, mean = 0, sd = sqrt(site_var) ), each = 2)
+
+     log_odds = with(dat, b0 + b1*(treatment == "treatment") + site_eff)
+     prop = plogis(log_odds)
+     dat$num_samp = num_samp
+     dat$y = rbinom(n = n_sites*2, size = num_samp, prob = prop)
+
+     glmer(cbind(y, num_samp - y) ~ treatment + (1|site),
+           data = dat,
+           family = binomial(link = "logit") )
+}
+
+set.seed(16)
+bin_glmm_fun()
+
+```
+
+# Random Forest Modelling
+
+## Preparation
+```{r}
+library(quantregForest)
+library(caret)
+library(tidyverse)
+library(tidymodels)
+library(skimr)
+library(sf)
+library(ggspatial)
+library(nhdplusTools)
+library(patchwork)
+library(Metrics)
+library(gt)
+```
+
+```{r}
+aoc_z_2 <- aoc_z %>% 
+  mutate_if(is.character, as.factor)# %>%
+  # mutate(effect_10 = case_when( #convert ordinal to numeric
+  #    effect_f == "Yes" ~ 1,
+  #    effect_f == "No" ~ 0))
+
+#choose relevant predictors and log-transform
+multiVar_small_log <- aoc_z_2 %>% 
+  dplyr::select(size_f,
+                ID, #row number ID for split/joining by study
+                doi, #need to split studies
+                size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na() %>%  #drop missing
+ mutate(effect_10 = case_when( #convert ordinal to numeric
+     effect_f == "Y" ~ 1,
+     effect_f == "N" ~ 0
+   ))# %>% 
+ #  select(-effect_f)
+skim(multiVar_small_log)
+```
+
+```{r}
+#no log transform for comparison
+multiVar_small <- aoc_z %>% 
+  dplyr::select(size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, life.stage,
+                exposure.duration.d, lvl1_f, dose.mg.L.master, dose.particles.mL.master, dose.um3.mL.master)# %>% 
+  #drop_na() #%>%  #drop missing
+# mutate(effect_10 = case_when( #convert ordinal to numeric
+#     effect == "Y" ~ 1,
+#     effect == "N" ~ 0
+#   )) %>% 
+  # select(-effect)
+
+#ensure completeness
+skim(multiVar_small)
+```
+
+##NON-LOG TRANSFORMED
+### Missing Value Imputation by Rough Fix
+```{r}
+#rough fix
+multiVar_small_roughfix <- na.roughfix(multiVar_small)
+
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+multiVar_small_split <- multiVar_small_roughfix %>%
+  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# default is 3/4ths split (but 75% training, 25% testing).
+# Stratification (strata) = grouping training/testing sets by region, state, etc.
+# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+multiVar_small_train <- training(multiVar_small_split)
+multiVar_small_test <- testing(multiVar_small_split)
+# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Create a separate dataset of available IDs that were not used in the training dataset.
+notTrain <- aoc %>% # all COMIDS from StreamCat data, sampled or not
+  filter(!ID %in% aoc$ID) # Removing sites used to train the model. n = 140,097
+
+count_optimized <- paste0('n = ',nrow(multiVar_small_roughfix))
+
+skim(multiVar_small_roughfix)
+```
+#### Kitchen Sink Model
+#### LINEAR DATA
+```{r}
+# Step Three - Kitchen Sink model -----------------------------------------
+# Random forest -- 
+# a decision tree model, using predictors to answer dichotomous questions to create nested splits.
+# no pruning happens - rather, multiple trees are built (the forest) and then you are looking for consensus across trees
+# training data goes down the tree and ends up in a terminal node.
+# if testing data goes down the same route, then this upholds our conclusions. Or, if it goes awry, this allows us to look for patterns in how it goes awry.
+
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+myrf <- randomForest(y = multiVar_small_roughfix$effect_f, # dependent variable
+  x = multiVar_small_roughfix %>%
+    dplyr::select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  na.action = na.roughfix,
+  ntrees = 100) # 500 trees default.
+
+myrf # examine the results.
+```
+~20% error rate. Let's compare this with the same model with log-transformed values.
+
+##### Predictor Selector/Recursive Feature Elimination
+```{r}
+# Using caret to select the best predictors
+# What are the parameters you want to use to run recursive feature elimination (rfe)?
+my_ctrl <- rfeControl(functions = rfFuncs,
+                      method = "cv",
+                      verbose = FALSE,
+                      returnResamp = "all")
+
+# rfe = recursive feature elimination
+# THIS STEP TAKES FOR-EV-ER!!!
+set.seed(22)
+my_rfe <- rfe(y = multiVar_small_roughfix$effect_f, # set dependent variable
+              x = multiVar_small_roughfix %>% 
+                dplyr::select(-effect_f),
+              rfeControl = my_ctrl,
+               size = c(1:2, 4, 6, 8, 10, 12, 14, 15))#,
+             # na.action = na.roughfix())
+
+# sets how many variables are in the overall model
+#               # I have 13 total possible variables, so I've chosen increments of 3 to look at.
+#               rfeControl = my_ctrl,
+#               na.action = na.roughfix,
+#               testX = multiVar_small_test %>% dplyr::select(-effect_f),
+#               testY = multiVar_small_test$effect_f)
+
+# can you make your model even simpler?
+# the following will pick a model with the smallest number of predictor variables based on the tolerance ("tol") that you specify (how much less than the best are you willing to tolerate?)
+
+#inspect
+my_rfe
+```
+
+```{r}
+trellis.par.set(caretTheme())
+#visualize
+plot(my_rfe,type = c("g", "o"))
+```
+
+```{r}
+my_size <- pickSizeTolerance(my_rfe$results, metric = "Accuracy", tol = 5, maximize = TRUE)
+# higher tol (~10) gives you less variables
+# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
+pickVars(my_rfe$variables, size = my_size)
+```
+The pickSizeTolerance determines the absolute best value then the percent difference of the other points to this value.
+This approach can produce good results for many of the tree based models, such as random forest, where there is a plateau of good performance for larger subset sizes.
+For trees, this is usually because unimportant variables are infrequently used in splits and do not significantly affect performance.
+
+Just use best predictors.
+```{r}
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+myrf_optimized <- randomForest(y = multiVar_small_train$effect_f, # dependent variable
+  x = multiVar_small_train %>%
+    dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master)), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  na.action = na.roughfix,
+  ntrees = 100) # 500 trees default.
+
+myrf_optimized # examine the results.
+```
+```{r}
+varImpPlot(myrf_optimized)
+```
+```{r}
+
+optimized_fitted <- predict(myrf_optimized, multiVar_small_test %>% dplyr::select(-effect_f), OOB = TRUE, type ="response")
+misClasificError_optimized <- mean(optimized_fitted != multiVar_small_test$effect_f)
+accuracy_optimized <- paste0('Accuracy: ', round(100*(1 - misClasificError_optimized), 2), '%')
+accuracy_optimized
+```
+
+```{r}
+df1 <- aoc_z %>% 
+  dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, effect_10)) %>% 
+  drop_na()
+
+m1 <- glm(effect_10 ~ organism.group + exposure.duration.d + lvl1_f + dose.um3.mL.master, 
+   data = df1, na.action = na.omit, family = "binomial")
+summary(m1)
+```
+
+```{r}
+require(ggeffects)
+m1_g <- ggeffects::ggpredict(m1, terms = "dose.um3.mL.master")
+plot(m1_g)
+```
+### Missing Value Imputation by RF
+
+The algorithm starts by imputing NAs using na.roughfix.
+Then randomForest is called with the completed data.
+The proximity matrix from the randomForest is used to update the imputation of the NAs.
+For continuous predictors, the imputed value is the weighted average of the non-missing obervations, where the weights are the proximities.
+For categorical predictors, the imputed value is the category with the largest average proximity.
+This process is iterated iter times.
+
+Note: Imputation has not (yet) been implemented for the unsupervised case.
+Also, Breiman (2003) notes that the OOB estimate of error from randomForest tend to be optimistic when run on the data matrix with imputed values.
+```{r}
+# impute values
+#drop NA's in response
+multiVar_small_noNa <- multiVar_small %>% drop_na(effect_f)
+
+#impute
+set.seed(111)
+multiVar_small_rfImpute <- rfImpute(data = multiVar_small_noNa, 
+                                    effect_f ~., #response value, cannot contain NA's
+                                    iter = 4,
+                                    ntree =75)
+
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+multiVar_small_split_imputed <- multiVar_small_rfImpute %>%
+  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# default is 3/4ths split (but 75% training, 25% testing).
+# Stratification (strata) = grouping training/testing sets by region, state, etc.
+# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+multiVar_small_train_imputed <- training(multiVar_small_split_imputed)
+multiVar_small_test_imputed <- testing(multiVar_small_split_imputed)
+# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Create a separate dataset of available IDs that were not used in the training dataset.
+notTrain <- aoc %>% # all COMIDS from StreamCat data, sampled or not
+  filter(!ID %in% aoc$ID) # Removing sites used to train the model. n = 140,097
+
+count_optimized_imputed <- paste0('n = ',nrow(multiVar_small_rfImpute))
+
+skim(multiVar_small_rfImpute)
+```
+#### Kitchen Sink Model
+#### LINEAR DATA
+```{r}
+# Step Three - Kitchen Sink model -----------------------------------------
+# Random forest -- 
+# a decision tree model, using predictors to answer dichotomous questions to create nested splits.
+# no pruning happens - rather, multiple trees are built (the forest) and then you are looking for consensus across trees
+# training data goes down the tree and ends up in a terminal node.
+# if testing data goes down the same route, then this upholds our conclusions. Or, if it goes awry, this allows us to look for patterns in how it goes awry.
+
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+myrf <- randomForest(y = multiVar_small_rfImpute$effect_f, # dependent variable
+  x = multiVar_small_rfImpute %>%
+    dplyr::select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+myrf # examine the results.
+```
+~20% error rate. Let's compare this with the same model with log-transformed values.
+```{r}
+plot(myrf, log = "y", main = "Random Forest for Imputed Data")
+```
+
+##### Predictor Selector/Recursive Feature Elimination
+```{r}
+# Using caret to select the best predictors
+# What are the parameters you want to use to run recursive feature elimination (rfe)?
+my_ctrl <- rfeControl(functions = rfFuncs,
+                      method = "cv",
+                      verbose = FALSE,
+                      returnResamp = "all")
+
+# rfe = recursive feature elimination
+# THIS STEP TAKES FOR-EV-ER!!!
+set.seed(22)
+my_rfe_imputed <- rfe(y = multiVar_small_rfImpute$effect_f, # set dependent variable
+              x = multiVar_small_rfImpute %>% 
+                dplyr::select(-effect_f),
+              rfeControl = my_ctrl,
+               size = c(1:2, 4, 6, 8, 10, 12, 14, 15))#,
+             # na.action = na.rfImpute())
+
+# sets how many variables are in the overall model
+#               # I have 13 total possible variables, so I've chosen increments of 3 to look at.
+#               rfeControl = my_ctrl,
+#               na.action = na.rfImpute,
+#               testX = multiVar_small_test %>% dplyr::select(-effect_f),
+#               testY = multiVar_small_test$effect_f)
+
+# can you make your model even simpler?
+# the following will pick a model with the smallest number of predictor variables based on the tolerance ("tol") that you specify (how much less than the best are you willing to tolerate?)
+
+#inspect
+my_rfe_imputed
+```
+
+```{r}
+trellis.par.set(caretTheme())
+#visualize
+plot(my_rfe_imputed,type = c("g", "o"))
+```
+
+```{r}
+my_size <- pickSizeTolerance(my_rfe_imputed$results, metric = "Accuracy", tol = 5, maximize = TRUE)
+# higher tol (~10) gives you less variables
+# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
+pickVars(my_rfe_imputed$variables, size = my_size)
+```
+The pickSizeTolerance determines the absolute best value then the percent difference of the other points to this value.
+This approach can produce good results for many of the tree based models, such as random forest, where there is a plateau of good performance for larger subset sizes.
+For trees, this is usually because unimportant variables are infrequently used in splits and do not significantly affect performance.
+
+Just use best predictors.
+```{r}
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+myrf_optimized_imputed <- randomForest(y = multiVar_small_train$effect_f, # dependent variable
+  x = multiVar_small_train %>%
+    dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master)), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+myrf_optimized_imputed # examine the results.
+```
+```{r}
+varImpPlot(myrf_optimized_imputed)
+```
+```{r}
+
+optimized_imputed_fitted <- predict(myrf_optimized_imputed, multiVar_small_test_imputed %>% dplyr::select(-effect_f), OOB = TRUE, type ="response")
+misClasificError_optimized_imputed <- mean(optimized_imputed_fitted != multiVar_small_test_imputed$effect_f)
+
+accuracy_optimized_imputed <- paste0('Accuracy: ', round(100*(1 - misClasificError_optimized_imputed), 2), '%')
+
+accuracy_optimized_imputed
+```
+
+####LOG DATA
+
+```{r}
+#log subset
+## First split data by DOI, then re-join other data
+set.seed(4)
+doi_split <- multiVar_small_log %>% 
+  dplyr::select(doi) %>% 
+  unique() %>% 
+  initial_split(prop = 0.65)
+#split just by doi 
+doi_train <- training(doi_split)
+doi_test <- testing(doi_split)
+
+train_full <- left_join(doi_train, multiVar_small_log, by = "doi") %>% 
+  dplyr::select(-c(doi, ID, effect_10)) %>% 
+  droplevels()
+
+test_full <- left_join(doi_test, multiVar_small_log, by = "doi") %>% 
+    dplyr::select(-c(doi, ID, effect_10)) %>% 
+  droplevels()
+
+#inspect proportion in test and train
+nrow(test_full)
+nrow(train_full)
+```
+Now that we've split the data by study and ensured a good proportion in test and train, let's run the model.
+
+```{r}
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+myrf_log <- randomForest(y = train_full$effect_f, # dependent variable
+  x = train_full %>%
+    dplyr::select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+myrf_log # examine the results.
+```
+<!-- # ```{r fit result 2} -->
+<!-- # #fix levels -->
+<!-- # #levels(test_full$polymer) <- levels(train_full$polymer) -->
+<!-- # levels(test_full$shape) <- levels(train_full$shape) -->
+<!-- # levels(test_full$lvl1_f) <- levels(train_full$lvl1_f) -->
+<!-- # #levels(test_full$organism.group) <- levels(train_full$organism.group) -->
+<!-- # -->
+<!-- # fitted <- predict(myrf_log, test_full %>% dplyr::select(-effect_f), OOB = TRUE, type ="response") -->
+<!-- # misClasificError <- mean(fitted != test_full$effect_f) -->
+<!-- # print(paste('Training Accuracy', 1 - misClasificError)) -->
+<!-- # ``` -->
+
+No performance enhancement with log-transformed values.
+
+## Continuous Predictors
+Repeat with continous variables whenever possible, and max of 8 predictors.
+```{r}
+skim(multiVar)
+```
+```{r}
+acute <- aoc_z %>% 
+  filter(acute.chronic_f == "Acute") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!exposure.route == "food") %>% 
+  filter(!polymer == "Not Reported") %>% 
+  filter(!shape == "Not Reported") %>% 
+  dplyr::select(c(doi, dose.mg.L.master, organism.group, size.length.um.used.for.conversions, polymer, shape, effect.score, dose.particles.mL.master, effect_f)) %>%   
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na() %>% 
+  droplevels()
+
+count_acute <- paste0('n = ',nrow(acute))
+
+skim(acute)
+```
+With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
+Let's determine if we have enough data for ML.
+```{r}
+exp(1)^6
+```
+$e^6$ is less than n (453), so we may proceed.
+
+```{r}
+## First split data by DOI, then re-join other data
+set.seed(4)
+acute_doi_split <- acute %>% 
+  dplyr::select(doi) %>% 
+  unique() %>% 
+  initial_split(prop = 0.65)
+#split just by doi 
+acute_doi_train <- training(acute_doi_split)
+acute_doi_test <- testing(acute_doi_split)
+
+acute_train <- left_join(acute_doi_train, acute, by = "doi") %>% 
+  dplyr::select(-doi) %>% 
+  droplevels()
+
+acute_test <- left_join(acute_doi_test, acute, by = "doi") %>% 
+  dplyr::select(-doi) %>% 
+  droplevels()
+
+#inspect proportion in test and train
+nrow(acute_test)
+nrow(acute_train)
+```
+
+```{r}
+# Create calibration and validation splits with tidymodels initial_split() function.
+# set.seed(4)
+# acute_split <- acute %>%
+#   initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# # default is 3/4ths split (but 75% training, 25% testing).
+# # Stratification (strata) = grouping training/testing sets by region, state, etc.
+# # Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+# 
+# # Create a training data set with the training() function
+# # Pulls from training and testing sets created by initial_split()
+# acute_train <- training(acute_split)
+# acute_test <- testing(acute_split)
+# # Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Random forest -- 
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+acuterf <- randomForest(y = acute_train$effect_f, # dependent variable
+  x = acute_train %>%
+    dplyr::select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+acuterf # examine the results.
+```
+
+```{r}
+plot(acuterf)
+# model performance appears to improve most at ~75 trees
+```
+
+```{r}
+varImpPlot(acuterf)
+# displays which variables are most important
+# helps to winnow down list of predictors
+# recommended to weigh left pane more
+# right pane also shows how evenly things split based on the list of predictors
+# values close to 0 can be dropped, but don't have to be
+```
+Dose, organism group, exposure duration are important.
+
+```{r}
+importance <- acuterf$importance
+#View(importance)
+# displays the data plotted in the plot above
+```
+
+### Performance and Validation
+
+```{r fit results 3}
+# fix levels
+levels(acute_test$polymer) <- levels(acute_train$polymer)
+levels(acute_test$shape) <- levels(acute_train$shape)
+levels(acute_test$organism.group) <- levels(acute_train$organism.group)
+
+fitted <- predict(acuterf, 
+                  newdata = acute_test %>% dplyr::select(-effect_f), 
+                  OOB = TRUE, type ="response")
+
+misClasificError_acute <- mean(fitted != acute_test$effect_f)
+
+accuracy_acute <- paste0('Accuracy: ', round(100*(1 - misClasificError_acute), 2), '%')
+accuracy_acute
+```
+
+Alternative ROC Curve
+```{r}
+require(pROC)
+predicted <- predict(acuterf, acute_test %>%  dplyr::select(-effect_f),
+                       OOB=TRUE, type= "response")
+#Calculate ROC curve
+rocCurve.tree <- roc(as.numeric(acute_test$effect_f),as.numeric(predicted))
+
+##gplot
+# rocks <- roc()
+
+#plot the ROC curve
+plot(rocCurve.tree,col=c(4))
+```
+#### Chronic
+```{r}
+chronic <- aoc_z %>% 
+  filter(acute.chronic_f == "Chronic") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!exposure.route == "food") %>% 
+  dplyr::select(c(doi, dose.mg.L.master, organism.group, size.length.um.used.for.conversions, polymer, shape, effect.score, dose.particles.mL.master, effect_f)) %>%
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na() %>% 
+  droplevels()
+
+count_chronic <- paste0('n = ',nrow(chronic))
+
+skim(chronic)
+```
+With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
+Let's determine if we have enough data for ML.
+```{r}
+exp(1)^6
+```
+$e^6$ is less than n (453), so we may proceed.
+```{r}
+## First split data by DOI, then re-join other data
+set.seed(4)
+chronic_doi_split <- chronic %>% 
+  dplyr::select(doi) %>% 
+  unique() %>% 
+  initial_split(prop = 0.85)
+#split just by doi 
+chronic_doi_train <- training(chronic_doi_split)
+chronic_doi_test <- testing(chronic_doi_split)
+
+chronic_train <- left_join(chronic_doi_train, chronic, by = "doi") %>% 
+  dplyr::select(-doi) %>% 
+  droplevels()
+
+chronic_test <- left_join(chronic_doi_test, chronic, by = "doi") %>% 
+  dplyr::select(-doi) %>% 
+  droplevels()
+
+#inspect proportion in test and train
+nrow(chronic_test)
+nrow(chronic_train)
+```
+
+```{r}
+# # Create calibration and validation splits with tidymodels initial_split() function.
+# set.seed(4)
+# chronic_split <- chronic %>%
+#   initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# # default is 3/4ths split (but 75% training, 25% testing).
+# # Stratification (strata) = grouping training/testing sets by region, state, etc.
+# # Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+# 
+# # Create a training data set with the training() function
+# # Pulls from training and testing sets created by initial_split()
+# chronic_train <- training(chronic_split)
+# chronic_test <- testing(chronic_split)
+# # Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Random forest -- 
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+chronicrf <- randomForest(y = chronic_train$effect_f, # dependent variable
+  x = chronic_train %>%
+    dplyr::select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+chronicrf # examine the results.
+```
+
+```{r}
+plot(chronicrf)
+# model performance appears to improve most at ~75 trees
+```
+
+```{r}
+varImpPlot(chronicrf)
+# displays which variables are most important
+# helps to winnow down list of predictors
+# recommended to weigh left pane more
+# right pane also shows how evenly things split based on the list of predictors
+# values close to 0 can be dropped, but don't have to be
+```
+Dose, organism group, exposure duration are important.
+
+```{r}
+importance <- chronicrf$importance
+#View(importance)
+# displays the data plotted in the plot above
+```
+
+```{r fit results 4}
+# fix levels
+levels(chronic_test$polymer) <- levels(chronic_train$polymer)
+levels(chronic_test$shape) <- levels(chronic_train$shape)
+levels(chronic_test$organism.group) <- levels(chronic_train$organism.group)
+
+fitted <- predict(chronicrf, chronic_test %>% 
+                    dplyr::select(-effect_f), 
+                  OOB = TRUE, type ="response")
+
+misClasificError_chronic <- mean(fitted != chronic_test$effect_f)
+
+accuracy_chronic <- paste0('Accuracy: ', round(100*(1 - misClasificError_chronic), 2), '%')
+accuracy_chronic
+```
+
+#### All Data
+```{r}
+all <- aoc_z %>% 
+  #filter(acute.chronic_f == "Chronic") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!exposure.route == "food") %>% 
+  dplyr::select(c(size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, organism.group, dose.mg.L.master, effect.score, lvl2_f, effect_f, doi)) %>%
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na()
+
+count_all <- paste0('n = ',nrow(all))
+
+skim(all)
+```
+With a complete dataset for just organismal fitness studies in aquatic organisms with aqueous route of exposure, we are left with 1510 complete cases with 8 predictor variables, and 1 response variable (effect y/n).
+Let's determine if we have enough data for ML.
+```{r}
+exp(1)^7
+```
+$e^6$ is less than n (453), so we may proceed.
+
+```{r}
+## First split data by DOI, then re-join other data
+set.seed(4)
+all_doi_split <- all %>% 
+  dplyr::select(doi) %>% 
+  unique() %>% 
+  initial_split(prop = 0.7)
+#split just by doi 
+all_doi_train <- training(all_doi_split)
+all_doi_test <- testing(all_doi_split)
+
+all_train <- left_join(all_doi_train, all, by = "doi") %>% 
+  dplyr::select(-doi) %>% 
+  droplevels()
+
+all_test <- left_join(all_doi_test, all, by = "doi") %>% 
+  dplyr::select(-doi) %>% 
+  droplevels()
+
+#inspect proportion in test and train
+nrow(all_test)
+nrow(all_train)
+```
+
+```{r}
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+# all_split <- all %>%
+#   initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# # default is 3/4ths split (but 75% training, 25% testing).
+# # Stratification (strata) = grouping training/testing sets by region, state, etc.
+# # Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+# 
+# # Create a training data set with the training() function
+# # Pulls from training and testing sets created by initial_split()
+# all_train <- training(all_split)
+# all_test <- testing(all_split)
+# # Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Random forest -- 
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+allrf <- randomForest(y = all_train$effect_f, # dependent variable
+  x = all_train %>%
+    dplyr::select(-effect_f), # selecting all predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+allrf # examine the results.
+```
+
+```{r}
+plot(allrf)
+# model performance appears to improve most at ~75 trees
+```
+
+```{r}
+varImpPlot(allrf)
+# displays which variables are most important
+# helps to winnow down list of predictors
+# recommended to weigh left pane more
+# right pane also shows how evenly things split based on the list of predictors
+# values close to 0 can be dropped, but don't have to be
+```
+Dose, organism group, exposure duration are important.
+
+```{r}
+importance <- allrf$importance
+#View(importance)
+# displays the data plotted in the plot above
+```
+
+```{r fit results 5}
+levels(all_test$polymer) <- levels(all_train$polymer)
+levels(all_test$shape) <- levels(all_train$shape)
+levels(all_test$organism.group) <- levels(all_train$organism.group)
+
+fitted <- predict(allrf, all_test %>% 
+                    dplyr::select(-effect_f), 
+                  OOB = TRUE, type ="response")
+
+misClasificError_all <- mean(fitted != all_test$effect_f)
+accuracy_all <- paste0('Accuracy: ', round(100*(1 - misClasificError_all), 2), '%')
+accuracy_all
+```
+
+#### NO FILTERS
+```{r}
+nofilter <- aoc_z %>% 
+  #filter(acute.chronic_f == "Chronic") %>% 
+  ##filter(!environment == "Terrestrial") %>% 
+  #filter(bio.org == "organism") %>% 
+  #filter(lvl1_f == "Fitness") %>% 
+  #filter(!exposure.route == "food") %>% 
+  filter(!shape == "Not Reported") %>% 
+  filter(!polymer == "Not Reported") %>% 
+  dplyr::select(c(doi, dose.mg.L.master, organism.group, size.length.um.used.for.conversions, polymer, shape,
+                  #effect.score, 
+                  dose.particles.mL.master, 
+                  effect_f, 
+                  particle.volume.um3, 
+                  exposure.duration.d,
+                  lvl1_f)) %>%
+  droplevels() %>% 
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na()
+
+count_nofilter <- paste0('n = ',nrow(nofilter))
+
+skim(nofilter)
+```
+With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
+Let's determine if we have enough data for ML.
+```{r}
+exp(1)^8
+```
+$e^6$ is less than n (453), so we may proceed.
+
+```{r}
+## First split data by DOI, then re-join other data
+set.seed(4)
+nofilter_doi_split <- nofilter %>% 
+  dplyr::select(doi) %>% 
+  unique() %>% 
+  initial_split(prop = 0.84)
+#split just by doi 
+nofilter_doi_train <- training(nofilter_doi_split)
+nofilter_doi_test <- testing(nofilter_doi_split)
+
+nofilter_train <- left_join(nofilter_doi_train, nofilter, by = "doi") %>% 
+  dplyr::select(-doi) %>% 
+  droplevels()
+
+nofilter_test <- left_join(nofilter_doi_test, nofilter, by = "doi") %>% 
+  dplyr::select(-doi) %>% 
+  droplevels()
+
+#inspect proportion in test and train
+nrow(nofilter_test)
+nrow(nofilter_train)
+```
+
+```{r}
+# Create calibration and validation splits with tidymodels initial_split() function.
+# set.seed(4)
+# nofilter_split <- nofilter %>%
+#   initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+# # default is 3/4ths split (but 75% training, 25% testing).
+# # Stratification (strata) = grouping training/testing sets by region, state, etc.
+# # Using the "strata" cnofilter ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
+# 
+# # Create a training data set with the training() function
+# # Pulls from training and testing sets created by initial_split()
+# nofilter_train <- training(nofilter_split)
+# nofilter_test <- testing(nofilter_split)
+# # Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+# Random forest -- 
+set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
+nofilterrf <- randomForest(y = nofilter_train$effect_f, # dependent variable
+  x = nofilter_train %>%
+    dplyr::select(-effect_f), # selecting nofilter predictor variables
+  importance = T, # how useful is a predictor in predicting values (nothing causal)
+  proximity = T, 
+  ntrees = 100) # 500 trees default.
+
+nofilterrf # examine the results.
+```
+
+```{r}
+plot(nofilterrf)
+# model performance appears to improve most at ~75 trees
+```
+
+```{r}
+varImpPlot(nofilterrf)
+# displays which variables are most important
+# helps to winnow down list of predictors
+# recommended to weigh left pane more
+# right pane also shows how evenly things split based on the list of predictors
+# values close to 0 can be dropped, but don't have to be
+```
+Dose, organism group, exposure duration are important.
+
+```{r}
+importance <- nofilterrf$importance
+#View(importance)
+# displays the data plotted in the plot above
+```
+
+```{r fit results 6}
+#fix levels
+levels(nofilter_test$polymer) <- levels(nofilter_train$polymer)
+levels(nofilter_test$shape) <- levels(nofilter_train$shape)
+levels(nofilter_test$lvl1_f) <- levels(nofilter_train$lvl1_f)
+levels(nofilter_test$organism.group) <- levels(nofilter_train$organism.group)
+
+x1 <- nofilter_test %>% dplyr::select(-effect_f)
+fitted <- predict(nofilterrf, 
+                  x1,
+                  OOB = TRUE, type ="response")
+
+misClasificError_nofilter <- mean(fitted != nofilter_test$effect_f)
+accuracy_nofilter <- paste0('Accuracy: ', round(100*(1 - misClasificError_nofilter), 2), '%')
+accuracy_nofilter
+```
+
+#### ROC Curves
+Prep for plotting.
+```{r}
+###CHRONIC
+chronicpredictions <- as.data.frame(predict(chronicrf, chronic_test %>% dplyr::select(-effect_f), type = "prob"))
+# predict class and then attach test class
+chronicpredictions$predict <- names(chronicpredictions)[1:2][apply(chronicpredictions[,1:2], 1, which.max)]
+chronicpredictions$observed <- chronic_test$effect_f
+
+####ACUTE
+predictions <- as.data.frame(predict(acuterf, acute_test%>% dplyr::select(-effect_f), type = "prob"))
+# predict class and then attach test class
+predictions$predict <- names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
+predictions$observed <- acute_test$effect_f
+head(predictions)
+
+#### all
+allpredictions <- as.data.frame(predict(allrf, all_test%>% dplyr::select(-effect_f), type = "prob"))
+# predict class and then attach test class
+allpredictions$predict <- names(allpredictions)[1:2][apply(allpredictions[,1:2], 1, which.max)]
+allpredictions$observed <- all_test$effect_f
+head(allpredictions)
+
+###nofilter
+nofilterpredictions <- as.data.frame(predict(nofilterrf, nofilter_test%>% dplyr::select(-effect_f), type = "prob"))
+# predict class and then attach test class
+nofilterpredictions$predict <- names(nofilterpredictions)[1:2][apply(nofilterpredictions[,1:2], 1, which.max)]
+nofilterpredictions$observed <- nofilter_test$effect_f
+
+###nofilterOptimized
+nofilter.optimizedpredictions <- as.data.frame(predict(myrf_optimized, multiVar_small_test %>%  dplyr::select(-effect_f), type = "prob"))
+# predict class and then attach test class
+nofilter.optimizedpredictions$predict <- names(nofilter.optimizedpredictions)[1:2][apply(nofilter.optimizedpredictions[,1:2], 1, which.max)]
+nofilter.optimizedpredictions$observed <- multiVar_small_test$effect_f
+
+###nofilterOptimizedImputed
+nofilter.optimized.imputedpredictions <- as.data.frame(predict(myrf_optimized_imputed, multiVar_small_test_imputed %>%  dplyr::select(-effect_f), type = "prob"))
+# predict class and then attach test class
+nofilter.optimized.imputedpredictions$predict <- names(nofilter.optimized.imputedpredictions)[1:2][apply(nofilter.optimized.imputedpredictions[,1:2], 1, which.max)]
+nofilter.optimized.imputedpredictions$observed <- multiVar_small_test_imputed$effect_f
+```
+Plot.
+```{r}
+require(ggdark)
+# 1 ROC curve, yes vs no for acute
+roc.acute <- roc(ifelse(predictions$observed=="Y", "Y", "N"), as.numeric(predictions$Y))
+
+#chronic
+roc.chronic <- roc(ifelse(chronicpredictions$observed=="Y", "Y", "N"), as.numeric(chronicpredictions$Y))
+#all
+roc.all <- roc(ifelse(allpredictions$observed=="Y", "Y", "N"), as.numeric(allpredictions$Y))
+
+#nofilter
+roc.nofilter <- roc(ifelse(nofilterpredictions$observed=="Y", "Y", "N"), as.numeric(nofilterpredictions$Y))
+
+#no filter (optimized)
+roc.nofilter.optimized <- roc(ifelse(nofilter.optimizedpredictions$observed=="Y", "Y", "N"), as.numeric(nofilter.optimizedpredictions$Y))
+
+#no filter (optimized; imputed)
+roc.nofilter.optimized.imputed <- roc(ifelse(nofilter.optimized.imputedpredictions$observed=="Y", "Y", "N"), as.numeric(nofilter.optimized.imputedpredictions$Y))
+
+##make ROC curves
+
+#all
+allROC <- ggroc(roc.all, col = "yellow") + 
+  labs(title = "Chronic and Acute",
+       subtitle = paste0(accuracy_all,', ',count_all)) + 
+  dark_theme_bw()
+
+#acute
+acuteROC <- ggroc(roc.acute, col = "green") + 
+  labs(title = "Acute",
+       subtitle = paste0(accuracy_acute,', ',count_acute)) + 
+  dark_theme_bw()
+
+#chronic
+chronicROC <- ggroc(roc.chronic, col = "blue") + 
+  labs(title = "Chronic",
+       subtitle = paste0(accuracy_chronic,', ',count_chronic)) + #auto label
+  dark_theme_bw()
+
+#no filter
+nofilterROC <- ggroc(roc.nofilter, col = "red3") + 
+  labs(title = "Entire Dataset",
+       subtitle = paste0(accuracy_nofilter,', ',count_nofilter)) + 
+  dark_theme_bw()
+
+#optimized (rough fix)
+nofilteroptimizedROC <- ggroc(roc.nofilter.optimized, col = "orangered") + 
+  labs(title = "Entire Dataset (optimized)",
+       subtitle = paste0(accuracy_optimized,', ',count_optimized)) + 
+  dark_theme_bw()
+
+#optimized (multiple imputation)
+nofilteroptimizedimputedROC <- ggroc(roc.nofilter.optimized.imputed, col = "orange") + 
+  labs(title = "Entire Dataset (optimized; imputed)",
+       subtitle = paste0(accuracy_optimized_imputed,', ',count_optimized_imputed)) + 
+  dark_theme_bw()
+
+#arrange together and print
+require(gridExtra)
+grid.arrange(nofilterROC, nofilteroptimizedROC, nofilteroptimizedimputedROC,allROC, acuteROC, chronicROC,
+             ncol = 3)
+```
+ROC curves may also be visualized together
+```{r}
+require(pROC)
+require(tidyverse)
+require(ggdark)
+require(ggsci)
+ggroc(list(all = roc.nofilter, optimized = roc.nofilter.optimized, optimized.imputed = roc.nofilter.optimized.imputed, organisms = roc.all, acute = roc.acute, chronic = roc.chronic), aes = c("linetype", "color")) +
+  labs(title = "ROC Curves for Aquatic Toxicity Random Forest",
+       subtitle = "n = 4615",
+       color = "Dataset",
+       linetype = "Dataset") +
+   scale_color_tron() +
+  # theme_bw(base_size = 20)
+ dark_theme_bw(base_size = 20)# +
+ theme(plot.title.position = element_text(hjust = 0.5),
+     plot.subtitle.position = element_text(hjust = 0.5))
+```
+# Interpretable Tree
+Make a classification tree.
+
+```{r, echo = FALSE}
+#get column index of predicted variable in dataset
+typeColNum <- grep("effect",names(all))
+
+#build tree
+require(rpart)
+set.seed(15097)
+t1 <- rpart(effect_f ~ size.length.um.used.for.conversions + shape + polymer + organism.group + particle.volume.um3 + dose.mg.L.master + effect.score,
+            method = "class", #classification because response is discrete
+            control = rpart.control(minbucket = 20, cp=0.008), #requires that there be at least 19 cases (responded + nonrespondents) in the final grouping of variable values of the terminal node of the tre..
+            data = all_train)
+print(t1, digits=4)
+```
+
+Plot an interpretable tree.
+```{r, echo = FALSE}
+require(rpart.plot)
+cols <- ifelse(t1$frame$yval == 1, "gray50", "black")
+prp(t1, main="Tree for Effect",
+    extra=106, # display prob of survival and percent of obs
+    nn=TRUE, # display node numbers
+    fallen.leaves=TRUE, # put leaves on the bottom of page
+    branch=.5, # change angle of branch lines
+    faclen=0, # do not abbreviate factor levels
+    trace=1, # print automatically calculated cex
+    shadow.col="gray", # shadows under the leaves
+    branch.lty=1, # draw branches using solid lines
+    branch.type=5, # branch lines width = weight(frame$wt), no. of cases here
+    split.cex=1.2, # make split text larger than node text
+    split.prefix="is ", # put "is " before split text
+    split.suffix="?", # put "?" after split text
+    col=cols, border.col=cols, # cols[2] if survived
+    split.box.col="lightgray", # lightgray split boxes (default is white)
+    split.border.col="darkgray", # darkgray border on split boxes
+    split.round=0.5) # round the split box corners a tad
+```
+
+<!-- # ```{r} -->
+<!-- # require(ROCR) -->
+<!-- # #plot reciever operating curve -->
+<!-- # perf <- performance(preds,"tpr","fpr") -->
+<!-- # ``` -->
+
+<!-- # ```{r} -->
+<!-- # # compute in-sample results -->
+<!-- # caret::confusionMatrix(fitted,as.factor(acute_test$effect_f)) -->
+<!-- # ``` -->
+<!-- #  -->
+<!-- # ```{r} -->
+<!-- # # Validation set assessment #1: looking at confusion matrix -->
+<!-- # prediction_for_table <- predict(acuterf, acute_test[,-6]) -->
+<!-- # table(observed = acute_test[,6], predicted = prediction_for_table) -->
+<!-- # ``` -->
+
+<!-- # ```{r} -->
+<!-- # # Validation set assessment #2: ROC curves and AUC -->
+<!-- # # Needs to import ROCR package for ROC curve plotting: -->
+<!-- # library(ROCR) -->
+<!-- # # Calculate the probability of new observations belonging to each class -->
+<!-- # # prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes -->
+<!-- # prediction_for_roc_curve <- predict(acuterf, acute_test %>% dplyr::select(-effect_f), -->
+<!-- #                                     type="response") -->
+<!-- # # Use pretty colours: -->
+<!-- # pretty_colours <- c("#F8766D","#00BA38","#619CFF") -->
+<!-- # # Specify the different classes  -->
+<!-- # classes <- levels(acute_test$organism.group) -->
+<!-- # # For each class -->
+<!-- # for (i in 1:7) -->
+<!-- # { -->
+<!-- #  # Define which observations belong to class[i] -->
+<!-- #  true_values <- ifelse(acute_test %>% dplyr::select(-effect_f) ==classes[i],"Y","N") -->
+<!-- #  # Assess the performance of classifier for class[i] -->
+<!-- #  pred <- prediction(prediction_for_roc_curve[,i],true_values) -->
+<!-- #  perf <- performance(pred, "tpr", "fpr") -->
+<!-- #  if (i==1) -->
+<!-- #  { -->
+<!-- #      plot(perf,main="ROC Curve",col=pretty_colours[i])  -->
+<!-- #  } -->
+<!-- #  else -->
+<!-- #  { -->
+<!-- #      plot(perf,main="ROC Curve",col=pretty_colours[i],add=TRUE)  -->
+<!-- #  } -->
+<!-- #  # Calculate the AUC and print it to screen -->
+<!-- #  auc.perf <- performance(pred, measure = "auc") -->
+<!-- #  print(auc.perf@y.values) -->
+<!-- # } -->
+<!-- # ``` -->
+
+<!-- # ```{r} -->
+<!-- # # predict() -->
+<!-- # # returns out of bag predictions for training data -->
+<!-- # # in the bag: every time a tree is built, it uses ~80% of the original 75% we set aside from the original dataset used to create a tree to assure random data selection -->
+<!-- # # out of bag: looking at the remaining 20% of the training data to predict, when you want to know what your model does at the training location sites -->
+<!-- #  -->
+<!-- # # # Predict CRAM scores state-wide for all COMIDs. -->
+<!-- # # notTrain_prediction <- notTrain %>% # taking all available IDs, that haven't been used in training -->
+<!-- # #   na.omit() %>% # remove NAs -->
+<!-- # #   mutate(effect_predicted = predict(myrf, newdata = notTrain %>% na.omit())) # using developed model (myrf), inputting predictor variables (notTrain - which contains IDs and associated  data) to predict output/dependent variable (effect_predicted a.k.a.
+effect).
+-->
+<!-- #  -->
+<!-- # # rePredict CRAM scores for training data. -->
+<!-- # multiVar_small_train$effect_predicted <- predict(myrf) # Add column of predicted CRAM values to training dataset. -->
+<!-- #  -->
+<!-- # # Creates new dataset of bound rows for both ... -->
+<!-- # # effect_predictions <- bind_rows(nottrain_prediction %>% -->
+<!-- # #                             mutate(Set = "Non-training"), # statewide COMIDs (not used for training) -->
+<!-- # #                             multiVar_smal_train %>% -->
+<!-- # #                             mutate(Set = "Training")) # COMIDS from training dataset -->
+<!-- # # This creates the dataset that will be plotted to create a state-wide plot of predicted CRAM scores. -->
+<!-- #  -->
+<!-- # # Plot the data. -->
+<!-- # multiVar_small_train %>%  -->
+<!-- #   mutate(effect_10 = case_when( #convert ordinal to numeric -->
+<!-- #      effect_predicted == "Yes" ~ 1, -->
+<!-- #      effect_predicted == "No" ~ 0 -->
+<!-- #    )) %>%  -->
+<!-- #   ggplot(aes(x = log10(dose.mg.L.master), y = effect_10)) + -->
+<!-- #   geom_point(alpha = 0.7) +#color = "black") + -->
+<!-- #   labs(x = "Density", -->
+<!-- #     y = "Predicted CRAM Score") + -->
+<!-- #   theme_dark() #+ -->
+<!-- #   #facet_wrap(.~polymer) -->
+<!-- # ``` -->
+<!-- ### Predictor Selector -->
+<!-- ```{r} -->
+<!-- # Using caret to select the best predictors -->
+<!-- # What are the parameters you want to use to run recursive feature elimination (rfe)? -->
+<!-- my_ctrl <- rfeControl(functions = rfFuncs, -->
+<!--                       method = "cv", -->
+<!--                       verbose = FALSE, -->
+<!--                       returnResamp = "all") -->
+
+<!-- # rfe = recursive feature elimination -->
+<!-- # THIS STEP TAKES FOR-EV-ER!!! -->
+<!-- set.seed(22) -->
+<!-- my_rfe <- rfe(y = rf_dat$effect_f, # set dependent variable -->
+<!--               x = rf_dat %>% select(-effect_f), # set predictor variables -->
+<!--               size = c(1:2, 4, 6, 8, 10, 13), # sets how many variables are in the overall model -->
+<!--               # I have 13 total possible variables, so I've chosen increments of 3 to look at. -->
+<!--               rfeControl = my_ctrl) # pull in control from above -->
+
+<!-- # can you make your model even simpler? -->
+<!-- # the following will pick a model with the smallest number of predictor variables based on the tolerance ("tol") that you specify (how much less than the best are you willing to tolerate?) -->
+<!-- my_size <- pickSizeTolerance(my_rfe$results, metric = "Accuracy", tol = 1, maximize = F) -->
+<!-- # higher tol (~10) gives you less variables -->
+<!-- # lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model." -->
+<!-- pickVars(my_rfe$variables, size = my_size) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- # Predict scores using the above 2 variables: -->
+
+<!-- # Create re-finalized training dataset and include all possible variables.  -->
+<!-- rf_dat2 <- multiVar_small_train %>%  select(dose.mg.L.master, organism.group, effect_f)  -->
+
+<!-- set.seed(4) # assures the data pulled is random, but sets it for the run below (makes outcome stable) -->
+<!-- myrf2 <- randomForest(y = rf_dat2$effect_f, # dependent variable -->
+<!--   x = rf_dat2 %>% -->
+<!--     select(-effect_f), -->
+<!--   importance = T,  -->
+<!--   proximity = T,  -->
+<!--   ntrees = 500)   -->
+
+<!-- myrf2 # examine the results.  -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- importance <- as.data.frame(as.table(myrf$importance)) -->
+
+<!-- # Nicer ggplot variable importance plot. -->
+<!-- vip_plot_a <- importance %>% -->
+<!--  filter(Var2 == "MeanDecreaseAccuracy") %>% -->
+<!--   mutate(Var1 = factor(Var1)) %>% -->
+<!--   mutate(Var1_f = fct_reorder(Var1, Freq)) %>% -->
+<!--   ggplot(aes(x = Freq, y = Var1_f)) + -->
+<!--   geom_point(size = 3, alpha = 0.75, color = "black") + -->
+<!--   labs(x = "Importance", -->
+<!--     y = "Variables") + -->
+<!--   theme_bw() -->
+
+<!-- # vip_plot_b <- importance %>% -->
+<!-- #   filter(Var2 == "MeanDecreaseGini") %>% -->
+<!-- #   mutate(Var1 = factor(Var1)) %>% -->
+<!-- #   mutate(Var1_f = fct_reorder(Var1, Freq)) %>% -->
+<!-- #   ggplot(aes(x = Freq, y = Var1_f)) + -->
+<!-- #   geom_point(size = 3, alpha = 0.75, color = "black") + -->
+<!-- #   labs(x = "Node Purity", -->
+<!-- #     y = "Variables") + -->
+<!-- #   theme_bw() -->
+
+<!-- vip_plot <- vip_plot_a #+ vip_plot_b -->
+
+<!-- vip_plot -->
+<!-- # ggsave("cram_vip_plot.png", -->
+<!-- #      path = "/Users/heilil/Desktop/R_figures", -->
+<!-- #      width = 25, -->
+<!-- #      height = 10, -->
+<!-- #      units = "cm" -->
+<!-- #    ) -->
+<!-- ``` -->
+GINI importance measures the average gain of purity by splits of a given variable.
+If the variable is useful, it tends to split mixed labeled nodes into pure single class nodes.
+Splitting by a permuted variables tend neither to increase nor decrease node purities.
+Permuting a useful variable, tend to give relatively large decrease in mean gini-gain.
+GINI importance is closely related to the local decision function, that random forest uses to select the best available split.
+Therefore, it does not take much extra time to compute.
+On the other hand, mean gini-gain in local splits, is not necessarily what is most useful to measure, in contrary to change of overall model performance.
+Gini importance is overall inferior to (permutation based) variable importance as it is relatively more biased, more unstable and tend to answer a more indirect question.
+
+### Quantile Regression model
+<!-- # ```{r} -->
+<!-- # # Quantile random forest regression mode, instead of looking at the mode of trees, can compare to 10th, 50th, 90th percentiles etc. -->
+<!-- #  -->
+<!-- # # Need to make a new dataset taking the above results of pickVars into account. -->
+<!-- # # Create finalized training dataset and include all possible variables.  -->
+<!-- # qrf_dat <- multiVar_small_train  %>% -->
+<!-- #   select(-effect_predicted, -ID) -->
+<!-- # #   select(asci, RdCrsWs, PctAgWs, PctUrbWsRp100, PctOpWsRp100, PctOpWs, DamDensWs, RdDensWs, NABD_DensWs, PctUrbWs, PctUrbCatRp100, RdDensWsRp100, PctOpCat, PctUrbCat, RdDensCat, CBNFWs, PctOpCatRp100, PctAgWsRp100, TRIDensWs, AgKffactWs, FertWs)  -->
+<!-- #  -->
+<!-- #  set.seed(20) -->
+<!-- #  myqrf <- quantregForest(y = qrf_dat$effect_f, # dependent variable -->
+<!-- #               x = qrf_dat %>% -->
+<!-- #                   select(-effect_f), -->
+<!-- #               importance = T, -->
+<!-- #               proximity = T, -->
+<!-- #               keep.inbag=T, -->
+<!-- #               ntrees = 500) -->
+<!-- #  -->
+<!-- # #predict(myqrf) # automatically presents 10th %tile, median, and 90th %tile -->
+<!-- # ``` -->
+
+```{r}
+#predict(myqrf, what=c(0.2, 0.3, 0.999)) # to print specific quantiles
+```
+
+<!-- # ```{r} -->
+<!-- # plot(myqrf) # plots the results. -->
+<!-- # ``` -->
+Again appears to improve after ~100 trees.
+
+### Model Validation
+#### Categorical Response
+
+#### Continuous Response
+```{r eval=FALSE, include=FALSE}
+# Compare predicted vs. actual results, including by size.
+# Adding lines of slope=1 and linear models to each plot.
+val1 <- multiVar_small_train %>% 
+  ggplot(aes(x = effect_predicted, y = effect_f)) +
+  geom_point(color = "#2A3927", alpha = 0.5) +
+  geom_smooth(method = "lm", se = FALSE, color = "#2A3927") +
+  labs(x = "Predicted Effect",
+    y = "Effect measured",
+    title = "Training Data\nn=613") +
+  geom_abline(intercept = 0, slope = 1) +
+  theme_bw()
+
+val1
+
+lm1 <- lm(cram~cram_predicted, data = mydf2_train2)
+summary(lm1)
+
+val2 <- ggplot(mydf2_test2, aes(x = cram_predicted, y = cram)) +
+  geom_point(color = "#3793EC", alpha = 0.5) +
+  geom_smooth(method = "lm", se = FALSE, color = "#3793EC") +
+  scale_x_continuous(breaks = c(0.5, 0.7, 0.9)) +
+  labs(x = "CRAM predicted",
+    y = "CRAM measured",
+    title = "Testing Data\nn=202") +
+  geom_abline(intercept = 0, slope = 1) +
+  theme_bw()
+
+val2
+
+lm2 <- lm(cram~cram_predicted, data = mydf2_test2)
+summary(lm2)
+
+# Create the full testing + training dataset to plot together.
+mydf2_test2$set <- "Testing"
+mydf2_train2$set <- "Training"
+full_train_test <- bind_rows(mydf2_test2, mydf2_train2) %>%
+  mutate(set_f = factor(set, levels = c("Training", "Testing")))
+
+val3 <- ggplot(full_train_test, aes(x = cram_predicted, y = cram, color = set_f)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "lm", se = FALSE) +
+  scale_color_manual(name = "Set", values = c("#2A3927", "#3793EC"), drop = FALSE) +
+  labs(x = "CRAM predicted",
+    y = "CRAM measured",
+    title = "All Data\nn=815") +
+  geom_abline(intercept = 0, slope = 1, color = "black") +
+  facet_wrap(~PSA6) +
+  theme_bw()
+
+val3
+
+val_fig <- (val1 + val2) /
+  (val3)
+
+val_fig + plot_annotation(
+  title = 'CRAM Random Forest Results',
+  subtitle = 'All modeling performed using StreamCAT datasets.',
+  caption = 'Linear models are colored according to dataset. Lines of slope = 1 are denoted in black.'
+)
+
+# Save figure.
+# ggsave("cram_rfmodel_validation.png",
+#      path = "/Users/heilil/Desktop/R_figures",
+#      width = 35,
+#      height = 25,
+#      units = "cm"
+#    )
+
+lm3 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Central_Valley") %>%
+    filter(set_f == "Training"))
+
+lm4 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Central_Valley") %>%
+    filter(set_f == "Testing"))
+
+lm5 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Chaparral") %>%
+    filter(set_f == "Training"))
+
+lm6 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Chaparral") %>%
+    filter(set_f == "Testing"))
+
+lm7 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Deserts_Modoc") %>%
+    filter(set_f == "Training"))
+
+lm8 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Deserts_Modoc") %>%
+    filter(set_f == "Testing"))
+
+lm9 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "North_Coast") %>%
+    filter(set_f == "Training"))
+
+lm10 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "North_Coast") %>%
+    filter(set_f == "Testing"))
+
+lm11 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Sierra") %>%
+    filter(set_f == "Training"))
+
+lm12 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "Sierra") %>%
+    filter(set_f == "Testing"))
+
+lm13 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "South_Coast") %>%
+    filter(set_f == "Training"))
+
+lm14 <- lm(cram~cram_predicted, 
+  data = full_train_test %>%
+    filter(PSA6 == "South_Coast") %>%
+    filter(set_f == "Testing"))
+
+# Chose not to compute confusion matrix / accuracy score since this is more applicable to categorical ouputs from random forest models -
+# Instead, calculated Root Mean Squared Error (RMSE) of both training and test datasets.
+# If test RMSE values are much greater than training, then possible the model has been over fit.
+
+predtest <- predict(myrf2, mydf2_test2)
+rmse(mydf2_test2$cram,predtest)
+# 9.03
+
+predtrain <- predict(myrf2, mydf2_train2)
+rmse(mydf2_train2$cram,predtrain)
+# 4.20
+
+# Double checking using the original random forest dataset (rf_dat) with all 35 possible variables included to see where the error in number of predictors starts to increase dramatically (to help double check our decision to include 25 parameters).
+dc <- rfcv(rf_dat %>%
+    select(-cram), 
+  rf_dat$cram,
+  step = 0.7, # default is 0.5
+  scale="log")
+
+dc$error.cv
+#34        24        17        12         8         6         4         3         2 
+# 96.28062  97.29366  99.08743 102.93752 115.52580 120.63049 122.09563 123.74425 139.30181 
+# 1 
+# 206.58921
+
+# Appears between 34 and 24 variables, there is an insignificant increase in error.
+# However, this model is much larger than the CSCI (20) and ASCI (10) models, so we may decide to trim this down in the future.
+
+```
+
+#3-D plots
+```{r}
+#remotes::install_github("tylermorganwall/rayshader")
+#require(rayshader)
+require(tidyverse)
+require(ggsci)
+require(ggdark)
+
+#scatterplot with size, dose and polymer
+acute3D <- acute %>% 
+  #filter(effect_f == "Yes") %>% 
+  ggplot(aes(x = size.length.um.used.for.conversions, y = dose.mg.L.master, color = effect.score)) + #, shape = polymer)) +
+  geom_point(alpha = 0.6) +
+  geom_smooth() +
+  scale_y_continuous("Dose (mg/L)", limits=c(-4,5)) +
+  scale_color_viridis_c(option = "A") +
+  ggtitle("Toxicity Probability by Length, Dose (mg/L) and Shape") +
+  labs(caption = "Acute Aquatic Organism") +
+  dark_theme_classic() +
+  facet_wrap(shape~.)
+
+acute3D
+
+#3D plot
+#plot_gg(acute3D, multicore = TRUE, width = 5, height = 5, scale =250) #3D plot
+```
+
+```{r eval=FALSE, include=FALSE}
+#3d scatterplot
+require(plotly)
+# size = acute$size.length.um.used.for.conversions
+# dose = acute$dose.mg.L.master
+# effect = acute$effect.score
+# volume = acute$particle.volume.um3
+
+plot_ly(acute, 
+        x = ~size.length.um.used.for.conversions, 
+        y = ~dose.mg.L.master, 
+        z = ~particle.volume.um3, 
+        type = "scatter3d", 
+        mode = "markers", 
+        color = ~effect_f)
+```
+```{r}
+#scatterplot with size, dose and polymer
+mass <-multiVar %>% 
+  filter(!effect_f == "NA") %>% 
+  filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  ggplot(aes(x = size.length.um.used.for.conversions, y = dose.mg.L.master, color = effect_f)) + 
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  scale_y_continuous(limits = c(0, 300))+
+ # scale_y_log10("Dose (mg/L)",limits = c(1e-4, 1e7))+
+  scale_x_log10("Length (um)", limits  = c(1, 1e3)) +
+  #coord_trans(x = "log10") +
+  #scale_x_continuous("Length (um)", breaks = scales::trans_breaks("log10", function(x) 10^x, n = 5),
+                   #      labels = trans_format("log10", scales::math_format(10^.x))) +
+  scale_colour_locuszoom() +
+  ggtitle("Toxicity Probability by Length and Dose (mg/L)") +
+  labs(caption = "Aquatic Organisms, HONEC/LOEC/NOEC") +
+  dark_theme_classic()# + 
+  #facet_wrap(acute.chronic_f ~.)
+mass
+```
+```{r}
+particles <- multiVar %>% 
+  filter(!effect_f == "NA") %>% 
+  filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  ggplot(aes(x = size.length.um.used.for.conversions, y = dose.particles.mL.master, color = effect_f)) + 
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  scale_y_log10("Dose (particles/mL)",limits = c(1e-4, 1e7))+
+  scale_x_log10("Length (um)", limits  = c(1, 1e4)) +
+  #coord_trans(x = "log10") +
+  #scale_x_continuous("Length (um)", breaks = scales::trans_breaks("log10", function(x) 10^x, n = 5),
+                   #      labels = trans_format("log10", scales::math_format(10^.x))) +
+  scale_colour_locuszoom() +
+  ggtitle("Toxicity Probability by Length and Dose (particles/mL)") +
+  labs(caption = "Aquatic Organisms, HONEC/LOEC/NOEC") +
+  dark_theme_classic() #+ 
+  #facet_wrap(acute.chronic_f ~.)
+particles
+```
+```{r}
+volume <- aoc_z %>% 
+  filter(!effect_f == "NA") %>% 
+  filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  ggplot(aes(x = size.length.um.used.for.conversions, y = dose.um3.mL.master, color = effect_f)) + 
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  scale_y_log10("Dose (um3/mL)",limits = c(1e+1, 1e7))+
+  scale_x_log10("Length (um3)", limits  = c(1e-1, 100)) +
+  #coord_trans(x = "log10") +
+  #scale_x_continuous("Length (um)", breaks = scales::trans_breaks("log10", function(x) 10^x, n = 5),
+                   #      labels = trans_format("log10", scales::math_format(10^.x))) +
+  scale_colour_locuszoom() +
+  ggtitle("Toxicity Probability by Length and Dose (um3/mL)") +
+  labs(caption = "Aquatic Organisms, HONEC/LOEC/NOEC") +
+  dark_theme_classic() #+ 
+  #facet_wrap(acute.chronic_f ~.)
+volume
+```
+
+```{r}
+require(gridExtra)
+grid.arrange(particles, mass, volume, nrow = 3)
+```
+
+```{r}
+#scatterplot with size, dose and polymer
+taxa_mass <-aoc_z %>% 
+  filter(!effect_f == "NA") %>% 
+ # filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  ggplot(aes(x = size.length.um.used.for.conversions, y = dose.mg.L.master.AF.noec, color = effect_f)) + 
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  scale_y_continuous(limits = c(0, 300))+
+ # scale_y_log10("Dose (mg/L)",limits = c(1e-4, 1e7))+
+  scale_x_log10("Length (um)", limits  = c(1, 1e3)) +
+  #coord_trans(x = "log10") +
+  #scale_x_continuous("Length (um)", breaks = scales::trans_breaks("log10", function(x) 10^x, n = 5),
+                   #      labels = trans_format("log10", scales::math_format(10^.x))) +
+  scale_colour_locuszoom() +
+  ggtitle("Toxicity Probability by Length and Dose (mg/L)") +
+  labs(caption = "Aquatic Organisms, HONEC/LOEC/NOEC") +
+  dark_theme_classic() + 
+  facet_wrap(organism.group ~.)
+taxa_mass
+```
+
+```{r}
+#Logistic Regression for acute fitness 
+m1_crust <-aoc_z %>% 
+  filter(!effect_f == "NA") %>% 
+ # filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+# filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master))#%>% 
+ # filter(acute.chronic_f == "Acute")
+
+m1_crust_model <- glm(effect_10 ~ (size.length.um.used.for.conversions + log10(dose.mg.L.master) +
+                                     log10(dose.particles.mL.master) + organism.group) ^ 2, #exponent gives all 2-way interactions
+    data = m1_crust, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model)
+```
+```{r}
+m1_crust_model$coef
+```
+
+## Plot binomial
+```{r}
+m1_crust_simple <- m1_crust %>% 
+  dplyr::select(c(effect_10, logdose.mg.L.master, size.length.um.used.for.conversions)) %>% 
+  drop_na
+#simple single parameter
+m1_crust_model_dose <- glm(effect_10 ~ logdose.mg.L.master * size.length.um.used.for.conversions, #exponent gives all 2-way interactions
+    data = m1_crust_simple, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_dose)
+```
+
+```{r eval=FALSE, include=FALSE}
+range(m1_crust_simple$logdose.mg.L.master, na.rm = TRUE)
+range(m1_crust_simple$size.length.um.used.for.conversions, na.rm = TRUE)
+#[1] -11.638272   8.171552
+# generate sequence of values between range in increments of 100 to give smooth appearance to model
+xdose <- seq(-11.638272 ,8.171552, 0.1)
+colorsize <- seq(3.45e-02, 5.00e+03, 1)
+ydose <- predict(m1_crust_model_dose, 
+                 list(logdose.mg.L.master = xdose,
+                      size.length.um.used.for.conversions = colorsize), 
+                 type = "response")
+#plot
+# plot(m1_crust_simple$logdose.mg.L.master, m1_crust_simple$effect_10, pch = 16, xlab = "log dose (mg/L)", ylab = "effect")
+# lines(xdose, ydose)
+
+ggplot(m1_crust_simple, aes(x = logdose.mg.L.master, y = effect_10))+
+  geom_point() +
+  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
+  labs(title = "Crustacea Fitness Binomial",
+       caption = "1,348 observations, dose p=6e-9") +
+  theme_minimal()
+```
+
+```{r}
+#alternative plot with ggpredict
+#devtools::install_github("cardiomoon/ggiraphExtra")
+require(ggiraphExtra)
+
+ggPredict(m1_crust_model_dose,interactive=TRUE,colorn=100,jitter=FALSE)
+```
+What happens if we log-transfrom size?
+
+What about volume?
+```{r}
+m1_crust_simple_volume <- m1_crust %>% 
+  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>% 
+  dplyr::select(c(effect_10, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
+  drop_na
+#simple single parameter
+m1_crust_model_volume<- glm(effect_10 ~ logdose.um3.mL.master * size.length.um.used.for.conversions, #exponent gives all 2-way interactions
+    data = m1_crust_simple_volume, na.action = "na.exclude", family = "binomial")
+
+summary(m1_crust_model_volume)
+```
+```{r}
+#plot logistic volume
+require(ggiraphExtra)
+ggPredict(m1_crust_model_volume,interactive=TRUE,colorn=100,jitter=FALSE)
+```
+
+```{r}
+#scatterplot with size, dose and polymer
+mass_probability_crustacea <-aoc_z %>% 
+  filter(!effect_f == "NA") %>% 
+ # filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(bio.org == "organism") %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  droplevels() %>% 
+  ggplot(aes(x = dose.mg.L.master.AF.noec, y = effect_10, color = effect_f)) + 
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  scale_y_continuous(limits = c(0, 1))+
+ scale_x_log10("Dose (mg/L)",limits = c(1e-4, 1e7))+
+  #scale_x_log10("Length (um)", limits  = c(1, 1e3)) +
+  #coord_trans(x = "log10") +
+  #scale_x_continuous("Length (um)", breaks = scales::trans_breaks("log10", function(x) 10^x, n = 5),
+                   #      labels = trans_format("log10", scales::math_format(10^.x))) +
+  scale_colour_locuszoom() +
+  ggtitle("Toxicity Probability by Length and Dose (mg/L)") +
+  labs(caption = "Aquatic Organisms, HONEC/LOEC/NOEC") +
+  dark_theme_classic() + 
+  facet_wrap(size_f ~.)
+mass_probability_crustacea
+```
+
+```{r eval=FALSE, include=FALSE}
+#3D density plot
+acute_density <- acute %>% 
+  ggplot(aes(x = size.length.um.used.for.conversions, y = dose.mg.L.master)) +
+  stat_density_2d(aes(fill = stat(effect.score)),
+                       geom = "polygon",
+                       n = 100, bins = 10, contour = TRUE) +
+  facet_wrap(shape~.) +
+  scale_fill_viridis_c(option = "A")
+
+plot_gg(acute_density, multicore = TRUE, width = 5, height = 5, scale =250) #3D plot
+```
+
+```{r eval=FALSE, include=FALSE}
+#heatmap
+#meltAcute = reshape2::melt(acute, id.vars = "size.length.um.used.for.conversions")
+
+#meltAcute$age = as.numeric(meltAcute$variable)
+
+acutegg = acute %>% 
+  #scale() or normalize() %>% 
+  ggplot() +
+  geom_raster(aes(x = size.length.um.used.for.conversions,y = dose.mg.L.master, fill = effect_f)) +
+  scale_x_continuous("Size (um)",expand=c(0,0)) + #,breaks = seq(1900,2010,10)) +
+  scale_y_continuous("Dose (mg/L)", expand=c(0,0)) + #,breaks=seq(0,100,10),limits=c(0,100)) +
+  scale_fill_viridis_d("Death\nProbability\nPer Year") +#,trans = "log10",breaks=c(1,0.1,0.01,0.001,0.0001), labels = c("1","1/10","1/100","1/1000","1/10000")) +
+  ggtitle("Acute Toxicity Probability by Size and Dose") +
+  labs(caption = "Acute Aquatic Organism")
+
+acutegg
+#plot_gg(deathgg, multicore=TRUE,height=5,width=6,scale=500)
+```
+
+# Works Cited

### 2021-02-19 DALEX.Rmd
many more models!

--- 2021-02-04 multivariate.Rmd
@@
 ---
-title: "Multivariate Analysis"
+title: "DALEX"
 author: "Scott Coffin"
-date: "1/19/2021"
+date: "2/19/2021"
 output: html_document
 ---
-# Setup
+
 ```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
-# Load packages
-library(tidyverse) #General everything
-library(RColorBrewer)
-library(ggplot2) #General plotting
-library(ggrepel) #For adding text labels that repel away from data points
-library(calecopal) #Color palette
-library(shiny) #Runs shiny
-library(shinythemes) #Shiny theme for the page
-library(shinyWidgets) #Widgets
-library(scales) #SSD - Use the percent format
-library(reshape2) #Overview tab - melts bars together
-library(ssdtools) #SSD package
-library(DT) #Build HTML data tables
-#library(plotly) #Make plots interactive
-library(viridis) #Colors
-library(scales) #To use "percent" function
-library(shinyjs) #Exploration tab - reset button
-library(tigerstats) #turns things into percents
-library(ggbeeswarm) #plot all points
-library(fitdistrplus) #alt SSD 
-library(RColorBrewer) #colors
-library(pheatmap) #pretty heat maps
-library(rpart)  #for trees
-#library(rattle)    # Fancy tree plot This is a difficult library to install (https://gist.github.com/zhiyzuo/a489ffdcc5da87f28f8589a55aa206dd) 
-library(rpart.plot)             # Enhanced tree plots
-library(RColorBrewer)       # Color selection for fancy tree plot
-library(party)                  # Alternative decision tree algorithm
-# library(partykit)               # Convert rpart object to BinaryTree
-library(pROC)   #for ROC curves
-library(uwot) #umap
-library(ISLR)  #for the Carseat Data
-library(quantregForest)
+```
+https://uc-r.github.io/dalex#:~:text=DALEX%20is%20an%20R%20package,performance%20across%20multiple%20models%20convenient.
+```{r}
+# load required packages
+library(rsample)
+library(dplyr)
+library(DALEX)
+library(tidyverse)
+library(tigerstats)
 library(caret)
-library(tidyverse)
-library(tidymodels)
-library(skimr)
-library(sf)
-library(ggspatial)
-library(nhdplusTools)
-library(patchwork)
-library(Metrics)
-library(gt)
 ```
-## Data import
-```{r data import}
+# Data Import
+```{r}
 # Load finalized dataset.
-aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
-
-#assign row ID
-aoc <- rowid_to_column(aoc, "ID")
-aoc <- aoc %>% 
-  mutate(ID = paste0(ID, "ID"))
+aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) #%>% 
+  #mutate(ID = paste0("ID",row_number()))
 
@@
   mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
+   mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) %>%   #calculate volume/mL
   mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
@@
                                             af.time_noNA == "Unavailable" ~ "Unavailable"))) %>%    #factorize assesment factor time into chronic/acute
-  mutate(dose.mg.L.master.AF.noec = dose.mg.L.master * af.noec) %>% #converts ECxx, LC50, HONEC, LOEC, NOEC all to NOEC
-  mutate(dose.particles.mL.master.AF.noec = dose.particles.mL.master * af.noec)  #converts ECxx, LC50, HONEC, LOEC, NOEC all to NOEC
+  mutate(dose.mg.L.master.AF.noec = dose.mg.L.master * af.noec) %>% 
+  mutate(dose.particles.mL.master.AF.noec = dose.particles.mL.master * af.noec) %>% 
+  mutate(effect_f = factor(effect)) %>% 
+  mutate_if(is.character, as.factor) %>% 
+  mutate(effect_10 = case_when(
+     effect_f == "Y" ~ 1,
+     effect_f == "N" ~ 0))
 
@@
 ```
+## Data Selection
 ```{r}
-# subset data to selected variables
-multiVar <- aoc_z %>% dplyr::select(#doi, size.category, 
-                                    ID,
-                                    size_f,
-                                    size.length.um.used.for.conversions, 
-                                    shape, 
-                                    polymer, 
-                                    particle.volume.um3, 
-                                    density.mg.um.3, 
-                                    organism.group,
-                                    environment, 
-                                    bio.org, #biological level of organization
-                                    #af.time, #assessment factor based on exposure time
-                                    treatments, #number of doses (no including control)
-                                    effect_f, #yes no
-                                    exposure.duration.d, 
-                                    exposure.route, #Factor
-                                    organism.group, #factor
-                                    media.temp, #numeric
-                                    lvl1_f, #endpoints
-                                    lvl2_f, #endpoints
-                                   # lvl3, 
-                                     dose.mg.L.master, 
-                                    sex, #factor
-                                    media.ph, #numeric
-                                    media.sal.ppt, #numeric
-                                    dose.particles.mL.master,
-                                   effect.metric, #NOEC LOEC
-                                    functional.group, #factor
-                                    charge, #positive or negatibe
-                                    zetapotential.mV, # numeric   
-                                   max.size.ingest.mm,#max ingestible size
-                                   acute.chronic_f,
-                                   dose.mg.L.master.AF.noec,
-                                   dose.particles.mL.master.AF.noec) %>%  
-  filter(!size_f == "Not Reported") %>%   #take out not reported 
-  mutate_if(is.character, as.factor)
+df <- aoc_z %>% 
+  dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, dose.mg.L.master, dose.particles.mL.master, life.stage, bio.org, polymer, shape, size.length.um.used.for.conversions, treatments, effect)) %>%
+  drop_na() %>%  #drop missing
+mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  mutate(effect_10 = case_when( #convert ordinal to numeric
+      effect == "Y" ~ 1,
+      effect == "N" ~ 0
+    )) %>%
+  mutate(effect_10 = factor(effect_10)) %>% 
+  dplyr::select(-(effect))
 
-# #recode variables
-# multiVar <- multiVar %>% mutate(effect_10 = case_when(
-#     effect_f == "Y" ~ 1,
-#     effect_f == "N" ~ 0
-#   ))# 
-```
-
-# Analysis
-## Data Exploration
-### Completeness vby Size
-```{r completeness heatmap}
-CompletenessSummary_size <- multiVar %>%
-  group_by(size_f) %>% 
-     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
-  mutate(across(is.numeric,~round(., 2))) %>% 
-  mutate(across(is.numeric, ~100 *(1 -.)))
-CompletenessSummary_size
+#ensure completeness
+skim(df)
 ```
@@
 ```{r}
-require()
-#convert to matrix and transpose
-transposed_size <- as.data.frame(t(as.matrix(CompletenessSummary_size[2:28]))) %>% #1 is category, 2-6 are variables
-  arrange('1nm < 100nm')
-#reassign column names
-colnames(transposed_size) <- c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm")
-#transposedTag
-#format as matrix
-MissingMatrix_size <- data.matrix(transposed_size)
-#build heatmap
-pheatmap(MissingMatrix_size,
-                           main = "Data Completeness by Size Bin", #title
-                           fontsize = 12,
-                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
-                           display_numbers = TRUE,
-                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
-                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)
+# create train, validation, and test splits
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+df_split <- df %>%
+  initial_split(prop = 0.75)
+# default is 3/4ths split (but 75% training, 25% testing).
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+train <- training(df_split)
+test <- testing(df_split)
+
+# variable names for resonse & features
+y <- "effect_10"
+x <- setdiff(names(df), y) 
+```
+
+```{r message=FALSE, warning=FALSE, paged.print=FALSE}
+# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+classif_rf <- train(effect_10~., data = train, method = "rf", ntree = 100, tuneLength = 1)
+
+classif_glm <- train(effect_10~., data = train, method = "glm", family = "binomial")
+
+classif_svm <- train(effect_10~., data = train, method = "svmRadial", prob.model = TRUE, tuneLength = 1,
+                     preProcess = c("pca","scale","center"))
+
+classif_decTree <- train(effect_10~., data = train, method = "C5.0", preProcess = c("scale","center"))
+
+classif_nnet <- train(effect_10~., data = train, method = "nnet", preProcess=c("scale","center"))
+
+classif_xgbTree <- train(effect_10~., data = train, method = "xgbTree")
+
+#classif_logicBag <- train(effect_10~., data = train, method = "logicBag", ntree = 100)
+
+# explain
+yTest <- as.numeric(as.character(test$effect_10))
+
+explainer_classif_rf <- DALEX::explain(classif_rf, label = "rf",
+                                       data = test, 
+                                       y = yTest)
+
+explainer_classif_glm <- DALEX::explain(classif_glm, label = "glm", 
+                                        data = test, y = yTest)
+
+explainer_classif_svm <- DALEX::explain(classif_svm,  label = "svm", 
+                                        data = test, y = yTest)
+
+explainer_classif_decTree <- DALEX::explain(classif_rf, label = "Decision Tree",
+                                       data = test, 
+                                       y = yTest)
+
+explainer_classif_nnet <- DALEX::explain(classif_glm, label = "Neural Net", 
+                                        data = test, y = yTest)
+
+explainer_classif_xgbTree <- DALEX::explain(classif_svm,  label = "eXtreme Gradient Boosting Trees", 
+                                        data = test, y = yTest)
+
+#explainer_classif_logicBag <- DALEX::explain(classif_svm,  label = "Bagged Logic Regression", 
+                                       # data = test, y = yTest)
 
 ```
-### Completeness by Environment
-```{r completeness heatmap}
-CompletenessSummary_environment <- multiVar %>%
-  filter(!environment == "NA") %>% 
-  group_by(environment) %>% 
-     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
-  mutate(across(is.numeric,~round(., 2))) %>% 
-  mutate(across(is.numeric, ~100 *(1 -.)))
-CompletenessSummary_environment
+### Model Performance
+```{r}
+mp_classif_rf <- model_performance(explainer_classif_rf)
+mp_classif_glm <- model_performance(explainer_classif_glm)
+mp_classif_svm <- model_performance(explainer_classif_svm)
+mp_classif_decTree <- model_performance(explainer_classif_decTree)
+mp_classif_nnet <- model_performance(explainer_classif_nnet)
+mp_classif_xgbTree <- model_performance(explainer_classif_xgbTree)
+#mp_classif_logicBag <- model_performance(explainer_classif_logicBag)
+
+# compare residuals plots
+resid_dist <- plot(mp_classif_rf, mp_classif_glm, mp_classif_svm, mp_classif_decTree, mp_classif_nnet,
+                   mp_classif_xgbTree) + #, mp_classif_logicBag) +
+  theme_minimal() +
+        theme(legend.position = 'bottom',
+              plot.title = element_text(hjust = 0.5)) + 
+        labs(y = '')
+resid_dist
+```
+#### Residuals
+```{r}
+resid_box <- plot(mp_classif_rf, mp_classif_glm, mp_classif_svm, mp_classif_decTree, mp_classif_nnet, mp_classif_xgbTree,
+     #mp_classif_logicBag, 
+     geom = "boxplot") +
+  theme_minimal() +
+        theme(legend.position = 'bottom',
+              plot.title = element_text(hjust = 0.5)) 
+resid_box
+```
+```{r}
+require(gridExtra)
+grid.arrange(resid_box,resid_dist, ncol=2)
+```
+#### Roc Curves
+```{r}
+plot(mp_classif_rf, 
+     #mp_classif_glm, 
+     mp_classif_svm, 
+     #mp_classif_decTree, 
+     mp_classif_nnet,
+     #mp_classif_xgbTree,
+     geom = "roc") +
+  ggtitle("ROC Curves - All Models",  
+          paste("AUC_rf = ",round(mp_classif_rf$measures$auc,2), 
+                paste("AUC_svm = ",round(mp_classif_svm$measures$auc,2)),
+          paste("AUC_nnet = ",round(mp_classif_nnet$measures$auc,2))
+          )) +
+#,  "AUC_glm = 0.799  AUC_svm = 0.798 AUC_decTree = AUC_nnet = AUC_xgbTree = ") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+```
+
+### Variable Importance
+```{r fig.height=7, fig.width=5}
+vi_classif_rf <- model_parts(explainer_classif_rf, loss_function = loss_root_mean_square)
+vi_classif_glm <- model_parts(explainer_classif_glm, loss_function = loss_root_mean_square)
+vi_classif_svm <- model_parts(explainer_classif_svm, loss_function = loss_root_mean_square)
+vi_classif_decTree <- model_parts(explainer_classif_decTree, loss_function = loss_root_mean_square)
+vi_classif_nnet <- model_parts(explainer_classif_nnet, loss_function = loss_root_mean_square)
+vi_classif_xgbTree <- model_parts(explainer_classif_xgbTree, loss_function = loss_root_mean_square)
+#vi_classif_logicBag <- model_parts(explainer_classif_logicBag, loss_function = loss_root_mean_square)
+
+plot(vi_classif_rf, vi_classif_glm, vi_classif_svm, vi_classif_decTree, vi_classif_nnet, vi_classif_xgbTree)#, #vi_classif_logicBag)
+```
+#### Partial Dependence Plot
+```{r}
+pdp_classif_rf  <- model_profile(explainer_classif_rf, variable = "dose.um3.mL.master", type = "partial")
+pdp_classif_glm  <- model_profile(explainer_classif_glm, variable = "dose.um3.mL.master", type = "partial")
+pdp_classif_svm  <- model_profile(explainer_classif_svm, variable = "dose.um3.mL.master", type = "partial")
+pdp_classif_decTree  <- model_profile(explainer_classif_decTree, variable = "dose.um3.mL.master", type = "partial")
+pdp_classif_nnet  <- model_profile(explainer_classif_nnet, variable = "dose.um3.mL.master", type = "partial")
+pdp_classif_xgbTree  <- model_profile(explainer_classif_xgbTree, variable = "dose.um3.mL.master", type = "partial")
+#pdp_classif_logicBag  <- model_profile(explainer_classif_logicBag, variable = "dose.um3.mL.master", type = "partial")
+
+plot(pdp_classif_rf, pdp_classif_glm, pdp_classif_svm, pdp_classif_decTree, pdp_classif_nnet, pdp_classif_xgbTree)#, pdp_classif_logicBag)
+```
+```{r}
+#Partial Dependence organism type
+pdp_classif_rf  <- model_profile(explainer_classif_rf, variable = "organism.group", type = "partial")
+pdp_classif_glm  <- model_profile(explainer_classif_glm, variable = "organism.group", type = "partial")
+pdp_classif_svm  <- model_profile(explainer_classif_svm, variable = "organism.group", type = "partial")
+pdp_classif_decTree  <- model_profile(explainer_classif_decTree, variable = "organism.group", type = "partial")
+pdp_classif_nnet  <- model_profile(explainer_classif_nnet, variable = "organism.group", type = "partial")
+pdp_classif_xgbTree  <- model_profile(explainer_classif_xgbTree, variable = "organism.group", type = "partial")
+#pdp_classif_logicBag  <- model_profile(explainer_classif_logicBag, variable = "dose.um3.mL.master", type = "partial")
+
+plot(pdp_classif_rf$agr_profiles, pdp_classif_glm$agr_profiles, pdp_classif_svm$agr_profiles, pdp_classif_decTree$agr_profiles, pdp_classif_nnet$agr_profiles, pdp_classif_xgbTree$agr_profiles) +
+  ggtitle("Contrastive Partial Dependence Profiles", "") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+```
+
+#### Breakdown
+
+```{r}
+# create a single observation
+new_cust <- test$effect_10 %>% as.data.frame()
+
+# compute breakdown distances
+new_cust_glm <- predict_parts(explainer_classif_glm, new_observation = test, type = "break_down")
+new_cust_rf <- predict_parts(explainer_classif_rf, new_observation = test, type = "break_down")
+new_cust_svm <- predict_parts(explainer_classif_svm, new_observation = test, type = "break_down")
+new_cust_decTree <- predict_parts(explainer_classif_decTree, new_observation = test, type = "break_down")
+new_cust_nnet <- predict_parts(explainer_classif_nnet, new_observation = test, type = "break_down")
+new_cust_xgbTree <- predict_parts(explainer_classif_xgbTree, new_observation = test, type = "break_down")
+#new_cust_logicBag <- predict_parts(explainer_classif_logicBag, new_observation = test, type = "break_down")
+
+# class of prediction_breakdown output
+class(new_cust_rf)
+
+# check out the top 10 influential variables for this observation
+new_cust_rf[1:10, 1:5]
 ```
@@
 ```{r}
-require()
-#convert to matrix and transpose
-transposed_environment <- as.data.frame(t(as.matrix(CompletenessSummary_environment[2:28])))
-#reassign column names
-colnames(transposed_environment) <- c("Freshwater", "Marine", "Terrestrial")
-#transposedTag
-#format as matrix
-MissingMatrix_environment <- data.matrix(transposed_environment)
-#build heatmap
-pheatmap(MissingMatrix_environment,
-                           main = "Data Completeness by environment", #title
-                           fontsize = 12,
-                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
-                           display_numbers = TRUE,
-                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
-                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)
+plot(new_cust_glm, new_cust_rf, new_cust_svm, new_cust_decTree, new_cust_nnet, new_cust_xgbTree)#,
+     #new_cust_logicBag)
+```
+```{r}
+library(ggplot2)
 
+# filter for top 10 influential variables for each model and plot
+list(new_cust_glm, new_cust_rf, new_cust_svm, new_cust_decTree, new_cust_nnet, 
+     #new_cust_logicBag,
+     new_cust_xgbTree) %>%
+  purrr::map(~ top_n(., 11, wt = abs(contribution))) %>%
+  do.call(rbind, .) %>%
+  mutate(variable = paste0(variable, " (", label, ")")) %>%
+  ggplot(aes(contribution, reorder(variable, contribution))) +
+  geom_point() +
+  geom_vline(xintercept = 0, size = 3, color = "white") +
+  facet_wrap(~ label, scales = "free_y", ncol = 1) +
+  ylab(NULL)
 ```
-
-## Stepwise Regression
+# Final Model
+All in all random forest is my final model of choice: it appears the more balanced and is the most accurate overall.
+I will further refine this model using recursive feature elimination and compare accuracy with the other models.
 
 ```{r}
-## Estimate an OLS Regression
-fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master +mouth, 
-             na.action = na.omit, 
-             data = multiVar)
-summary(fitols)
+
+my_ctrl <- rfeControl(functions = rfFuncs,
+                      method = "repeatedcv",
+                      verbose = FALSE,
+                      repeats = 5,
+                      returnResamp = "all")
+
+rfProfile <- rfe(y = df$effect_10, # set dependent variable
+              x = df %>% 
+                dplyr::select(-effect_10),
+              rfeControl = my_ctrl,
+               size = c(1:2, 4, 6, 8, 10, 12, 13))
+rfProfile
+```
+```{r}
+y <- "effect_10"
+x <- setdiff(names(df), y)
+
+#alt method
+rfRFE <-  list(summary = defaultSummary,
+               fit = function(x, y, first, last, ...){
+                 library(randomForest)
+                 randomForest(x, y, importance = first, ...)
+                 },
+               pred = function(object, x)  predict(object, x),
+               rank = function(object, x, y) {
+                 vimp <- varImp(object)
+                 vimp <- vimp[order(vimp$Overall,decreasing = TRUE),,drop = FALSE]
+                 vimp$var <- rownames(vimp)                  
+                 vimp
+                 },
+               selectSize = pickSizeBest,
+               selectVar = pickVars)
 ```
@@
 ```{r}
-require(reshape2)
-multiVar %>% 
-  dplyr::select(size.length.um.used.for.conversions, particle.volume.um3, exposure.duration.d ,media.temp, dose.particles.mL.master) %>% 
-  melt() %>%  #convert wide to long
-   mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  ggplot(aes(x = value)) + 
-  stat_density() + 
-  facet_wrap(~variable, scales = "free")
-```
-
-<!-- # ```{r} -->
-<!-- # # cdplot(as.factor(effect) ~ particle.volume.um3, data = multiVar, -->
-<!-- #        main = "Conditional Density Plot of ", xlab = "Dose", ylab = "Effect") -->
-<!-- # ``` -->
-
-## PCA
-
-# UMAP
-
-```{r}
-require(uwot)
-require(Rtsne)
-require(vizier) #devtools::install_github("jlmelville/vizier")
-
-# For some functions we need to strip out non-numeric columns and convert data to matrix
-x2m <- function(X) {
-  if (!methods::is(X, "matrix")) {
-    m <- as.matrix(X[, which(vapply(X, is.numeric, logical(1)))])
-  }
-  else {m <- X} 
-  m}
-
-#choose values with most completeness
-multiVar2 <- multiVar %>% 
-  filter(!environment == "Terrestrial") %>% 
-  dplyr::select(size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  drop_na() #drop missing
-
-#convert discrete variables to numeric
-multiVar2[] <- data.matrix(multiVar2)
-
-# build umap for small dataset (<10,000 points)
-multiVar_map  <- umap(multiVar2, pca = 10)
-
-# Remove duplicates for t-SNE
-#multiVar2_noNa_dup <- multiVar2_noNa[-which(duplicated(x2m(multiVar2_noNa))), ]
-
-#build t-SNE
-#multiVar_tsne <- Rtsne::Rtsne(multiVar2_noNa_dup, perplexity = 15, initial_dims = 100, partical_pca = TRUE, exaggeration_factor = 4)
-
-# Non-numeric columns are ignored, so in a lot of cases you can pass a data
-# frame directly to umap
-#iris_umap <- umap(iris, n_neighbors = 50, learning_rate = 0.5, init = "random")
-
-#visualize umap
-embed_img <- function(X, Y, k = 15, ...) {
-  args <- list(...)
-  args$coords <- Y
-  args$x <- X
-
-  do.call(vizier::embed_plot, args)
-}
-
-#plot
-embed_img(multiVar2, multiVar_map, pc_axes = TRUE, equal_axes = TRUE, alpha_scale = 0.5, title = "Tox UMAP", cex = 1)
+predictors(rfProfile)
 ```
@@
 ```{r}
-#PCA
-pca <- stats::prcomp(multiVar2[,-4], retx = TRUE, rank. = 2)
-#build color pallete
-my_colors = colorRampPalette(c("red", "yellow", "green"))(nrow(multiVar2))
-#plot
-embed_plot(pca$x, multiVar2$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), #turbo, #rainbow, #my_colors, 
-           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE)
+head(rfProfile$resample)
 ```
@@
 ```{r}
-require(plotly)
-#PCA for discrete variables
-multiVar_discrete <- multiVar %>% 
-  select(size.length.um.used.for.conversions, polymer, dose.mg.L.master, exposure.duration.d, effect) %>% 
-  drop_na %>% 
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  mutate(effect_10 = case_when(
-    effect == "Y" ~ 1,
-    effect == "N" ~ 0
-  )) %>% 
-  select(-effect)
-
-#buildPCA
-pca_discrete <- stats::prcomp(multiVar_discrete[,-2], retx = TRUE, rank. = 5, scale. = TRUE)
-
-embed_plot(pca_discrete$x, multiVar_discrete$polymer)
-
-embed_plotly(pca_discrete$x, multiVar_discrete$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), 
-           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE,
-           tooltip = paste("Polymer:", multiVar_discrete$polymer))
-
-#embed_plotly also option
-```
-
-# Random Forest
-## Recursive Partitioning And Regression Trees
-
-The rpart algorithm works by splitting the dataset recursively, which means that the subsets that arise from a split are further split until a predetermined termination criterion is reached.
-At each step, the split is made based on the independent variable that results in the largest possible reduction in heterogeneity of the dependent (predicted) variable.
-
-It is important to note that the algorithm works by making the best possible choice at each particular stage, without any consideration of whether those choices remain optimal in future stages.
-That is, the algorithm makes a locally optimal decision at each stage.
-It is thus quite possible that such a choice at one stage turns out to be sub-optimal in the overall scheme of things.
-In other words,  the algorithm does not find a globally optimal tree.
-
-The size of the data set usually imposes a hard limit on how many features may be considered due to Curse of Dimensionality, i.e.
-your data becomes sparser and sparser as you increase the number of features considered, which usually leads to overfitting.
-While there is no rule of thumb relating to how many features vs.
-the number of observations you should use, I try to keep $e^Nf < No$ (Nf = number of features, No = number of observations) to minimise overfitting [this is not always possible and it does not ensure that we won’t overfit].
-In this case, our training set has 3200, which suggests that using 8 features (e^8 ~ 2980) is not entirely absurd.
-Obviously, this depends on your data, so we will cover some further overfitting checks later on.
-```{r}
-#trim data so effect is always known
-multiVar_sub <- multiVar %>% 
-  drop_na(effect_10)
-
-# Split data into training and test sets
-set.seed(42)
-multiVar_sub[,"train"] <- ifelse(runif(nrow(multiVar_sub)) < 0.8, 1, 0)
-# Separate trainig and test sets
-trainSet <- multiVar_sub[multiVar_sub$train==1,]
-testSet <- multiVar_sub[multiVar_sub$train==0,]
-#get column index of train flag
-trainColNum <- grep("train", names(trainSet))
-# Remove train flag column from train and test sets
-trainSet <- trainSet[, -trainColNum]
-testSet <- testSet[, -trainColNum]
-```
-Make a classification tree.
-
-```{r, echo = FALSE}
-#get column index of predicted variable in dataset
-typeColNum <- grep("effect",names(multiVar_sub))
-
-#build tree
-require(rpart)
-set.seed(15097)
-t1 <- rpart(effect ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,
-            method = "class", #classification because response is discrete
-            control = rpart.control(minbucket = 20, cp=0.008), #requires that there be at least 19 cases (responded + nonrespondents) in the final grouping of variable values of the terminal node of the tre..
-            data = trainSet)
-print(t1, digits=4)
-```
-
-Plot an interpretable tree.
-```{r, echo = FALSE}
-require(rpart.plot)
-cols <- ifelse(t1$frame$yval == 1, "gray50", "black")
-prp(t1, main="Tree for Effect",
-    extra=106, # display prob of survival and percent of obs
-    nn=TRUE, # display node numbers
-    fallen.leaves=TRUE, # put leaves on the bottom of page
-    branch=.5, # change angle of branch lines
-    faclen=0, # do not abbreviate factor levels
-    trace=1, # print automatically calculated cex
-    shadow.col="gray", # shadows under the leaves
-    branch.lty=1, # draw branches using solid lines
-    branch.type=5, # branch lines width = weight(frame$wt), no. of cases here
-    split.cex=1.2, # make split text larger than node text
-    split.prefix="is ", # put "is " before split text
-    split.suffix="?", # put "?" after split text
-    col=cols, border.col=cols, # cols[2] if survived
-    split.box.col="lightgray", # lightgray split boxes (default is white)
-    split.border.col="darkgray", # darkgray border on split boxes
-    split.round=0.5) # round the split box corners a tad
-```
-Next we see how good the model is by seeing how it fares against the test data.
-```{r}
-t1_predict <- predict(t1, newdata = testSet[,-typeColNum],
-                      type="class")
-mean(t1_predict==testSet$effect)
-# [1] 0.7094088
-#confusion matrix
-table(pred=t1_predict,true=testSet$effect)
-```
-```{r}
-par(mfrow=c(1,2)) # two plots on one page
-#plot approximate R-squared and relative error for different splits (2 plots). labels are only appropriate for the "anova" method.
-rsq.rpart(t1)
-```
-
-Next, we prune the tree using the cost complexity criterion.
-Basically, the intent is to see if a shallower subtree can give us comparable results.
-If so, we’d be better of choosing the shallower tree because it reduces the likelihood of overfitting.
-
-As described earlier, we choose the appropriate pruning parameter (aka cost-complexity parameter) \alpha by picking the value that results in the lowest prediction error.
-Note that all relevant computations have already been carried out by R when we built the original tree (the call to rpart in the code above).
-All that remains now is to pick the value of \alpha:
-
-### Pruning
-
-```{r}
-#cost-complexity pruning
-printcp(t1)
-```
-It is clear from the above, that the lowest cross-validation error (xerror in the table) occurs for \alpha =0.008 (this is CP in the table above).   One can find CP programatically like so:
-```{r}
-# get index of CP with lowest xerror
-opt <- which.min(t1$cptable[,"xerror"])
-#get its value
-cp <- t1$cptable[opt, "CP"]
-```
-Next, we prune the tree based on this value of CP:
-
-```{r}
-# prune the tree
-pt1 <- prune(t1,cp)
-#pt1<- prune(t1, cp= t1$cptable[which.min(t1$cptable[,"xerror"]),"CP"])
-
-# plot the pruned tree
-plot(pt1, uniform=TRUE,
-   main="Pruned Classification Tree");text(pt1, use.n=TRUE, all=TRUE, cex=.8)
-
-#post(pfit, file = "c:/ptree.ps",
- #  title = "Pruned Classification Tree for Kyphosis")
-```
-```{r}
-#find proportion of correct predictions using test set
-t1_pruned_predict <- predict(pt1, testSet, type="class")
-mean(t1_pruned_predict == testSet$effect)
-#
-```
-This is not an improvement over an unprunend tree..
-We need to check that this holds up for different training and test sets.
-This is easily done by creating multiple random partitions of the dataset and checking the efficacy of pruning for each.
-To do this efficiently, I’ll create a function that takes the training fraction, number of runs (partitions) and the name of the dataset as inputs and outputs the proportion of correct predictions for each run.
-It also optionally prunes the tree.
-
-```{r}
-#function to do multiple runs
-multiple_runs_classification <- function(train_fraction,n,dataset,prune_tree=FALSE){
-fraction_correct <- rep(NA,n)
-set.seed(42)
-for (i in 1:n){
-  dataset[,"train"] <- ifelse(runif(nrow(dataset))<0.8,1,0)
-  trainColNum <- grep("train",names(dataset))
-  typeColNum <- grep("effect",names(dataset))
-  trainSet <- dataset[dataset$train==1,-trainColNum]
-  testSet <- dataset[dataset$train==0,-trainColNum]
-  rpart_model <- rpart(effect~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,data = trainSet, method="class")
-if(prune_tree==FALSE) {
-  rpart_test_predict <- predict(rpart_model,testSet[,-typeColNum],type="class")
-  fraction_correct[i] <- mean(rpart_test_predict==testSet$effect)
-  }else{
-    opt <- which.min(rpart_model$cptable[,"xerror"])
-    cp <- rpart_model$cptable[opt, "CP"]
-    pruned_model <- prune(rpart_model,cp)
-    rpart_pruned_predict <- predict(pruned_model,testSet[,-typeColNum],type="class")
-    fraction_correct[i] <- mean(rpart_pruned_predict == testSet$effect)
-  }
-  }
-return(fraction_correct)
-}
-```
-
-Note that in the above,  I have set the default value of the prune_tree to FALSE, so the function will execute the first branch of the if statement unless the default is overridden.
-
-OK, so let’s do 50 runs with and without pruning, and check the mean and variance of the results for both sets of runs.
-```{r}
-#50 runs, no pruning
-unpruned_set <- multiple_runs_classification(0.8,50,multiVar_sub)
-mean(unpruned_set)
-#[1] 0.7261882
-sd(unpruned_set)
-#[1] 0.01554347
-#50 runs, with pruning
-pruned_set <- multiple_runs_classification(0.8,50,multiVar_sub,prune_tree=TRUE)
-mean(pruned_set)
-#[1] 0.7261875
-sd(pruned_set)
-#[1] 0.01552285
-```
-
-## CForest
-```{r run model}
-require(party)
-crf <- cforest(as.factor(effect_10) ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d +
-                 media.temp + dose.particles.mL.master,
-               controls = cforest_control(ntree = 500,
-                                          mincriterion = qnorm(0.8), 
-                                          trace = TRUE), # adds project bar because it's very slow
-               data = multiVar_sub)
-
-crf
-```
-
-```{r fit results}
-train2 <- multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
-                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master)
-train3 <-multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
-                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master,
-                                  effect_10)
-
-fitted <- predict(crf, train2, OOB = TRUE, type ="response")
-#rpart.prob <- predict(t1, newdata=imputedSmalls.requested.voluntary,type="prob")
-
-misClasificError <- mean(fitted != train3$effect_10)
-print(paste('Training Accuracy', 1 - misClasificError))
+trellis.par.set(caretTheme())
+plot(rfProfile, type = c("g", "o"))
 ```
@@
 ```{r}
-print(crf)
+sessionInfo()
 ```
-
-```{r}
-#Obtaining predicted probabilites for Test data
-tree.probs=predict(crf,
-                 newdata = train2,
-                 type="prob")
-
-#Calculate ROC curve
-rocCurve.tree <- roc(train3$effect_10,tree.probs[,'1'])
-#plot the ROC curve
-plot(rocCurve.tree,col=c(4))
-```
-
-Alternative ROC Curve
-```{r}
-require(pROC)
-predicted <- predict(crf, train2, OOB=TRUE, type= "response")
-auc(as.numeric(train3$effect_10), as.numeric(predicted))
-#Calculate ROC curve
-rocCurve.tree <- roc(train3$effect_10,as.numeric(predicted))
-#plot the ROC curve
-plot(rocCurve.tree,col=c(4))
-```
-```{r}
-#plot reciever operating curve
-perf <- performance(preds,"tpr","fpr")
-```
-
-```{r}
-# compute in-sample results
-caret::confusionMatrix(fitted,as.factor(train3$effect_10))
-```
-
-```{r}
-#plot feature importance
-cforestImpPlot <- function(x) {
-  cforest_importance <<- v <- varimp(x)
-  dotchart(v[order(v)])
-}
-
-importancePlot <- cforestImpPlot(crf)
-importancePlot
-```
-
-```{r}
-crf.prob <- matrix(unlist(crfsrvy.prob), ncol=2, byrow=TRUE)
-apply(crf.prob,2,mean)
-#[1] 0.2524514 0.7475486
-tab <- round(cbind(by(rpart.prob[,2], INDICES=t1$where, mean), by(crf.prob[,2], INDICES=t1$where, mean)), 4)
-colnames(tab) <- c("rpart", "cforest")
-kable(tab)
-```
-# Random Forest Package
-
-## Preparation
-```{r}
-library(quantregForest)
-library(caret)
-library(tidyverse)
-library(tidymodels)
-library(skimr)
-library(sf)
-library(ggspatial)
-library(nhdplusTools)
-library(patchwork)
-library(Metrics)
-library(gt)
-```
-
-```{r}
-multiVar <- multiVar %>% 
-  mutate_if(is.character, as.factor) %>% 
-  mutate(effect_10 = case_when( #convert ordinal to numeric
-     effect_f == "Yes" ~ 1,
-     effect_f == "No" ~ 0))
-
-#choose relevant predictors and log-transform
-multiVar_small_log <- multiVar %>% 
-  dplyr::select(ID, size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  drop_na() %>%  #drop missing
- mutate(effect_10 = case_when( #convert ordinal to numeric
-     effect_f == "Yes" ~ 1,
-     effect_f == "No" ~ 0
-   ))# %>% 
- #  select(-effect_f)
-
-#no log transform for comparison
-multiVar_small <- multiVar %>% 
-  dplyr::select(ID, size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, #effect_10 , 
-                exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>% 
-  drop_na() #%>%  #drop missing
-# mutate(effect_10 = case_when( #convert ordinal to numeric
-#     effect == "Y" ~ 1,
-#     effect == "N" ~ 0
-#   )) %>% 
-  # select(-effect)
-
-#ensure completeness
-skim(multiVar_small)
-```
-## Training Data
-
-```{r}
-# Create calibration and validation splits with tidymodels initial_split() function.
-set.seed(4)
-multiVar_small_split <- multiVar_small %>%
-  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
-# default is 3/4ths split (but 75% training, 25% testing).
-# Stratification (strata) = grouping training/testing sets by region, state, etc.
-# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
-
-# Create a training data set with the training() function
-# Pulls from training and testing sets created by initial_split()
-multiVar_small_train <- training(multiVar_small_split)
-multiVar_small_test <- testing(multiVar_small_split)
-# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
-
-# Create a separate dataset of available IDs that were not used in the training dataset.
-notTrain <- aoc %>% # all COMIDS from StreamCat data, sampled or not
-  filter(!ID %in% aoc$ID) # Removing sites used to train the model. n = 140,097
-```
-## Kitchen Sink Model
-
-```{r}
-# Step Three - Kitchen Sink model -----------------------------------------
-
-# Create finalized training dataset and include all possible variables. 
-rf_dat <- multiVar_small_train %>%
-  select(-ID)
-
-# Random forest -- 
-# a decision tree model, using predictors to answer dichotomous questions to create nested splits.
-# no pruning happens - rather, multiple trees are built (the forest) and then you are looking for consensus across trees
-# training data goes down the tree and ends up in a terminal node.
-# if testing data goes down the same route, then this upholds our conclusions. Or, if it goes awry, this allows us to look for patterns in how it goes awry.
-
-set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
-myrf <- randomForest(y = rf_dat$effect_f, # dependent variable
-  x = rf_dat %>%
-    select(-effect_f), # selecting all predictor variables
-  importance = T, # how useful is a predictor in predicting values (nothing causal)
-  proximity = T, 
-  ntrees = 100) # 500 trees default.
-
-myrf # examine the results.
-```
-~20% error rate. Let's compare this with the same model with log-transformed values.
-```{r}
-# Create calibration and validation splits with tidymodels initial_split() function.
-set.seed(4)
-multiVar_small_log_split <- multiVar_small_log %>%
-  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
-
-# Create a training data set with the training() function
-# Pulls from training and testing sets created by initial_split()
-multiVar_small_log_train <- training(multiVar_small_log_split)
-multiVar_small_log_test <- testing(multiVar_small_log_split)
-
-# Create finalized training dataset and include all possible variables. 
-rf_dat_log <- multiVar_small_log_train %>%
-  select(-ID, -effect_10)
-
-set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
-myrf_log <- randomForest(y = rf_dat_log$effect_f, # dependent variable
-  x = rf_dat_log %>%
-    select(-effect_f), # selecting all predictor variables
-  importance = T, # how useful is a predictor in predicting values (nothing causal)
-  proximity = T, 
-  ntrees = 100) # 500 trees default.
-
-myrf_log # examine the results.
-```
-No performance enhancement with log-transformed values.
-
-## Continuous Predictors
-Repeat with continous variables whenever possible, and max of 8 predictors.
-```{r}
-skim(multiVar)
-```
-```{r}
-acute <- multiVar %>% filter(acute.chronic_f == "Chronic") %>% 
-  filter(!environment == "Terrestrial") %>% 
-  filter(bio.org == "organism") %>% 
-  filter(lvl1_f == "Fitness") %>% 
-  filter(!exposure.route == "food") %>% 
-  select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d,bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master)) %>%
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  drop_na()
-skim(acute)
-```
-With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
-Let's determine if we have enough data for ML.
-```{r}
-exp(1)^6
-```
-$e^6$ is less than n (453), so we may proceed.
-
-```{r}
-plot(myrf)
-# model performance appears to improve most at ~75 trees
-```
-
-```{r}
-varImpPlot(myrf)
-# displays which variables are most important
-# helps to winnow down list of predictors
-# recommended to weigh left pane more
-# right pane also shows how evenly things split based on the list of predictors
-# values close to 0 can be dropped, but don't have to be
-```
-Dose, organism group, exposure duration are important.
-
-```{r}
-importance <- myrf$importance
-View(importance)
-# displays the data plotted in the plot above
-```
-
-```{r}
-# predict()
-# returns out of bag predictions for training data
-# in the bag: every time a tree is built, it uses ~80% of the original 75% we set aside from the original dataset used to create a tree to assure random data selection
-# out of bag: looking at the remaining 20% of the training data to predict, when you want to know what your model does at the training location sites
-
-# # Predict CRAM scores state-wide for all COMIDs.
-# notTrain_prediction <- notTrain %>% # taking all available IDs, that haven't been used in training
-#   na.omit() %>% # remove NAs
-#   mutate(effect_predicted = predict(myrf, newdata = notTrain %>% na.omit())) # using developed model (myrf), inputting predictor variables (notTrain - which contains IDs and associated  data) to predict output/dependent variable (effect_predicted a.k.a.
-effect).
-
-# rePredict CRAM scores for training data.
-multiVar_small_train$effect_predicted <- predict(myrf) # Add column of predicted CRAM values to training dataset.
-
-# Creates new dataset of bound rows for both ...
-# effect_predictions <- bind_rows(nottrain_prediction %>%
-#                             mutate(Set = "Non-training"), # statewide COMIDs (not used for training)
-#                             multiVar_smal_train %>%
-#                             mutate(Set = "Training")) # COMIDS from training dataset
-# This creates the dataset that will be plotted to create a state-wide plot of predicted CRAM scores.
-
-# Plot the data.
-multiVar_small_train %>% 
-  mutate(effect_10 = case_when( #convert ordinal to numeric
-     effect_predicted == "Yes" ~ 1,
-     effect_predicted == "No" ~ 0
-   )) %>% 
-  ggplot(aes(x = log10(dose.mg.L.master), y = effect_10)) +
-  geom_point(alpha = 0.7) +#color = "black") +
-  labs(x = "Density",
-    y = "Predicted CRAM Score") +
-  theme_dark() #+
-  #facet_wrap(.~polymer)
-```
-### Predictor Selector
-```{r}
-# Using caret to select the best predictors
-# What are the parameters you want to use to run recursive feature elimination (rfe)?
-my_ctrl <- rfeControl(functions = rfFuncs,
-                      method = "cv",
-                      verbose = FALSE,
-                      returnResamp = "all")
-
-# rfe = recursive feature elimination
-# THIS STEP TAKES FOR-EV-ER!!!
-set.seed(22)
-my_rfe <- rfe(y = rf_dat$effect_f, # set dependent variable
-              x = rf_dat %>% select(-effect_f), # set predictor variables
-              size = c(1:2, 4, 6, 8, 10, 13), # sets how many variables are in the overall model
-              # I have 13 total possible variables, so I've chosen increments of 3 to look at.
-              rfeControl = my_ctrl) # pull in control from above
-
-# can you make your model even simpler?
-# the following will pick a model with the smallest number of predictor variables based on the tolerance ("tol") that you specify (how much less than the best are you willing to tolerate?)
-my_size <- pickSizeTolerance(my_rfe$results, metric = "Accuracy", tol = 1, maximize = F)
-# higher tol (~10) gives you less variables
-# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
-pickVars(my_rfe$variables, size = my_size)
-```
-
-```{r}
-# Predict scores using the above 2 variables:
-
-# Create re-finalized training dataset and include all possible variables. 
-rf_dat2 <- multiVar_small_train %>%  select(dose.mg.L.master, organism.group, effect_f)
-
-set.seed(4) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
-myrf2 <- randomForest(y = rf_dat2$effect_f, # dependent variable
-  x = rf_dat2 %>%
-    select(-effect_f),
-  importance = T, 
-  proximity = T, 
-  ntrees = 500)
-
-myrf2 # examine the results. 
-```
-
-```{r}
-importance <- as.data.frame(as.table(myrf$importance))
-
-# Nicer ggplot variable importance plot.
-vip_plot_a <- importance %>%
- filter(Var2 == "MeanDecreaseAccuracy") %>%
-  mutate(Var1 = factor(Var1)) %>%
-  mutate(Var1_f = fct_reorder(Var1, Freq)) %>%
-  ggplot(aes(x = Freq, y = Var1_f)) +
-  geom_point(size = 3, alpha = 0.75, color = "black") +
-  labs(x = "Importance",
-    y = "Variables") +
-  theme_bw()
-
-# vip_plot_b <- importance %>%
-#   filter(Var2 == "MeanDecreaseGini") %>%
-#   mutate(Var1 = factor(Var1)) %>%
-#   mutate(Var1_f = fct_reorder(Var1, Freq)) %>%
-#   ggplot(aes(x = Freq, y = Var1_f)) +
-#   geom_point(size = 3, alpha = 0.75, color = "black") +
-#   labs(x = "Node Purity",
-#     y = "Variables") +
-#   theme_bw()
-
-vip_plot <- vip_plot_a #+ vip_plot_b
-
-vip_plot
-# ggsave("cram_vip_plot.png",
-#      path = "/Users/heilil/Desktop/R_figures",
-#      width = 25,
-#      height = 10,
-#      units = "cm"
-#    )
-```
-GINI importance measures the average gain of purity by splits of a given variable.
-If the variable is useful, it tends to split mixed labeled nodes into pure single class nodes.
-Splitting by a permuted variables tend neither to increase nor decrease node purities.
-Permuting a useful variable, tend to give relatively large decrease in mean gini-gain.
-GINI importance is closely related to the local decision function, that random forest uses to select the best available split.
-Therefore, it does not take much extra time to compute.
-On the other hand, mean gini-gain in local splits, is not necessarily what is most useful to measure, in contrary to change of overall model performance.
-Gini importance is overall inferior to (permutation based) variable importance as it is relatively more biased, more unstable and tend to answer a more indirect question.
-
-### Quantile Regression model
-```{r}
-# Quantile random forest regression mode, instead of looking at the mode of trees, can compare to 10th, 50th, 90th percentiles etc.
-
-# Need to make a new dataset taking the above results of pickVars into account.
-# Create finalized training dataset and include all possible variables. 
-qrf_dat <- multiVar_small_train  %>%
-  select(-effect_predicted, -ID)
-#   select(asci, RdCrsWs, PctAgWs, PctUrbWsRp100, PctOpWsRp100, PctOpWs, DamDensWs, RdDensWs, NABD_DensWs, PctUrbWs, PctUrbCatRp100, RdDensWsRp100, PctOpCat, PctUrbCat, RdDensCat, CBNFWs, PctOpCatRp100, PctAgWsRp100, TRIDensWs, AgKffactWs, FertWs)
-
- set.seed(20)
- myqrf <- quantregForest(y = qrf_dat$effect_f, # dependent variable
-              x = qrf_dat %>%
-                  select(-effect_f),
-              importance = T,
-              proximity = T,
-              keep.inbag=T,
-              ntrees = 500)
-
-#predict(myqrf) # automatically presents 10th %tile, median, and 90th %tile
-```
-
-```{r}
-#predict(myqrf, what=c(0.2, 0.3, 0.999)) # to print specific quantiles
-```
-
-```{r}
-plot(myqrf) # plots the results.
-```
-# Again appears to improve after ~100 trees.
-
-### Model Validation
-#### Categorical Response
-
-#### Continuous Response
-```{r}
-# Compare predicted vs. actual results, including by size.
-# Adding lines of slope=1 and linear models to each plot.
-val1 <- multiVar_small_train %>% 
-  ggplot(aes(x = effect_predicted, y = effect_f)) +
-  geom_point(color = "#2A3927", alpha = 0.5) +
-  geom_smooth(method = "lm", se = FALSE, color = "#2A3927") +
-  labs(x = "Predicted Effect",
-    y = "Effect measured",
-    title = "Training Data\nn=613") +
-  geom_abline(intercept = 0, slope = 1) +
-  theme_bw()
-
-val1
-
-lm1 <- lm(cram~cram_predicted, data = mydf2_train2)
-summary(lm1)
-
-val2 <- ggplot(mydf2_test2, aes(x = cram_predicted, y = cram)) +
-  geom_point(color = "#3793EC", alpha = 0.5) +
-  geom_smooth(method = "lm", se = FALSE, color = "#3793EC") +
-  scale_x_continuous(breaks = c(0.5, 0.7, 0.9)) +
-  labs(x = "CRAM predicted",
-    y = "CRAM measured",
-    title = "Testing Data\nn=202") +
-  geom_abline(intercept = 0, slope = 1) +
-  theme_bw()
-
-val2
-
-lm2 <- lm(cram~cram_predicted, data = mydf2_test2)
-summary(lm2)
-
-# Create the full testing + training dataset to plot together.
-mydf2_test2$set <- "Testing"
-mydf2_train2$set <- "Training"
-full_train_test <- bind_rows(mydf2_test2, mydf2_train2) %>%
-  mutate(set_f = factor(set, levels = c("Training", "Testing")))
-
-val3 <- ggplot(full_train_test, aes(x = cram_predicted, y = cram, color = set_f)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = FALSE) +
-  scale_color_manual(name = "Set", values = c("#2A3927", "#3793EC"), drop = FALSE) +
-  labs(x = "CRAM predicted",
-    y = "CRAM measured",
-    title = "All Data\nn=815") +
-  geom_abline(intercept = 0, slope = 1, color = "black") +
-  facet_wrap(~PSA6) +
-  theme_bw()
-
-val3
-
-val_fig <- (val1 + val2) /
-  (val3)
-
-val_fig + plot_annotation(
-  title = 'CRAM Random Forest Results',
-  subtitle = 'All modeling performed using StreamCAT datasets.',
-  caption = 'Linear models are colored according to dataset. Lines of slope = 1 are denoted in black.'
-)
-
-# Save figure.
-# ggsave("cram_rfmodel_validation.png",
-#      path = "/Users/heilil/Desktop/R_figures",
-#      width = 35,
-#      height = 25,
-#      units = "cm"
-#    )
-
-lm3 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Central_Valley") %>%
-    filter(set_f == "Training"))
-
-lm4 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Central_Valley") %>%
-    filter(set_f == "Testing"))
-
-lm5 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Chaparral") %>%
-    filter(set_f == "Training"))
-
-lm6 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Chaparral") %>%
-    filter(set_f == "Testing"))
-
-lm7 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Deserts_Modoc") %>%
-    filter(set_f == "Training"))
-
-lm8 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Deserts_Modoc") %>%
-    filter(set_f == "Testing"))
-
-lm9 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "North_Coast") %>%
-    filter(set_f == "Training"))
-
-lm10 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "North_Coast") %>%
-    filter(set_f == "Testing"))
-
-lm11 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Sierra") %>%
-    filter(set_f == "Training"))
-
-lm12 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Sierra") %>%
-    filter(set_f == "Testing"))
-
-lm13 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "South_Coast") %>%
-    filter(set_f == "Training"))
-
-lm14 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "South_Coast") %>%
-    filter(set_f == "Testing"))
-
-# Chose not to compute confusion matrix / accuracy score since this is more applicable to categorical ouputs from random forest models -
-# Instead, calculated Root Mean Squared Error (RMSE) of both training and test datasets.
-# If test RMSE values are much greater than training, then possible the model has been over fit.
-
-predtest <- predict(myrf2, mydf2_test2)
-rmse(mydf2_test2$cram,predtest)
-# 9.03
-
-predtrain <- predict(myrf2, mydf2_train2)
-rmse(mydf2_train2$cram,predtrain)
-# 4.20
-
-# Double checking using the original random forest dataset (rf_dat) with all 35 possible variables included to see where the error in number of predictors starts to increase dramatically (to help double check our decision to include 25 parameters).
-dc <- rfcv(rf_dat %>%
-    select(-cram), 
-  rf_dat$cram,
-  step = 0.7, # default is 0.5
-  scale="log")
-
-dc$error.cv
-#34        24        17        12         8         6         4         3         2 
-# 96.28062  97.29366  99.08743 102.93752 115.52580 120.63049 122.09563 123.74425 139.30181 
-# 1 
-# 206.58921
-
-# Appears between 34 and 24 variables, there is an insignificant increase in error.
-# However, this model is much larger than the CSCI (20) and ASCI (10) models, so we may decide to trim this down in the future.
-
-```
-
-# Works Cited
-
-# Non-linear Dose Response
-If we hold all paramaters constant except for dose, we should observe a non-linear dose-response behavior that can be best predicted by a 3- or 4-parameter Hill Equation.
-The most available information is likely for polystyrene spheres in crustaceans.There are 21 studies available.

### 2021-02-19 multivariate.Rmd
many more models!

@@
 library(randomForest)
+library(DALEX) #explanatory model analysis
+
 ```
@@
 # Load finalized dataset.
-aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) %>% 
-  mutate(ID = paste0("ID",row_number()))
+aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) #%>% 
+  #mutate(ID = paste0("ID",row_number()))
 
@@
                                    dose.particles.mL.master.AF.noec,
+                                   dose.um3.mL.master,
                                    max.size.ingest.mm, #maximum ingestible size range
@@
 <!-- # ``` -->
-
-# PCA
-
-# UMAP
-
-```{r}
-require(uwot)
-require(Rtsne)
-require(vizier) #devtools::install_github("jlmelville/vizier")
-
-# For some functions we need to strip out non-numeric columns and convert data to matrix
-x2m <- function(X) {
-  if (!methods::is(X, "matrix")) {
-    m <- as.matrix(X[, which(vapply(X, is.numeric, logical(1)))])
-  }
-  else {m <- X} 
-  m}
-
-#choose values with most completeness
-multiVar2 <- multiVar %>% 
-  filter(!environment == "Terrestrial") %>% 
-  dplyr::select(size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  drop_na() #drop missing
-
-#convert discrete variables to numeric
-multiVar2[] <- data.matrix(multiVar2)
-
-# build umap for small dataset (<10,000 points)
-multiVar_map  <- umap(multiVar2, pca = 10)
-
-# Remove duplicates for t-SNE
-#multiVar2_noNa_dup <- multiVar2_noNa[-which(duplicated(x2m(multiVar2_noNa))), ]
-
-#build t-SNE
-#multiVar_tsne <- Rtsne::Rtsne(multiVar2_noNa_dup, perplexity = 15, initial_dims = 100, partical_pca = TRUE, exaggeration_factor = 4)
-
-# Non-numeric columns are ignored, so in a lot of cases you can pass a data
-# frame directly to umap
-#iris_umap <- umap(iris, n_neighbors = 50, learning_rate = 0.5, init = "random")
-
-#visualize umap
-embed_img <- function(X, Y, k = 15, ...) {
-  args <- list(...)
-  args$coords <- Y
-  args$x <- X
-
-  do.call(vizier::embed_plot, args)
-}
-
-#plot
-embed_img(multiVar2, multiVar_map, pc_axes = TRUE, equal_axes = TRUE, alpha_scale = 0.5, title = "Tox UMAP", cex = 1)
-```
-
-```{r}
-#PCA
-pca <- stats::prcomp(multiVar2[,-5], retx = TRUE, rank. = 2)
-#build color pallete
-my_colors = colorRampPalette(c("red", "yellow", "green"))(nrow(multiVar2))
-#plot
-embed_plot(pca$x, multiVar2$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), #turbo, #rainbow, #my_colors, 
-           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE)
-```
-
-```{r}
-require(plotly)
-#PCA for discrete variables
-multiVar_discrete <- multiVar %>% 
-  select(size.length.um.used.for.conversions, polymer, dose.mg.L.master, exposure.duration.d) %>% 
-  drop_na %>% 
-  mutate_if(~is.numeric(.) && (.) > 0, log10)
-
-#buildPCA
-pca_discrete <- stats::prcomp(multiVar_discrete[,-2], retx = TRUE, rank. = 2)
-
-embed_plotly(pca_discrete$x, multiVar_discrete$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), 
-           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE,
-           tooltip = paste("Polymer:", multiVar_discrete$polymer))
-```
 
@@
 Alternative ROC Curve
-```{r}
+```{r eval=FALSE, include=FALSE}
 require(pROC)
@@
 ```
+# Multiple Models with DALEX
+
+```{r}
+#no log transform for comparison
+optimized <- aoc_z %>% 
+  dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, life.stage, bio.org, polymer, shape, size.length.um.used.for.conversions, treatments, effect)) %>% 
+  drop_na() %>%  #drop missing
+mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  mutate(effect_10 = case_when( #convert ordinal to numeric
+      effect == "Y" ~ 1,
+      effect == "N" ~ 0
+    )) %>%
+  mutate(effect_10 = factor(effect_10)) %>% 
+  dplyr::select(-(effect))
+
+#ensure completeness
+skim(optimized)
+```
+
+Alt Method
+https://raw.githack.com/pbiecek/DALEX_docs/master/vignettes/DALEX_caret.html#3_classification_use_case_-_wine_data
+```{r}
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+optimized_split <- optimized %>%
+  initial_split(prop = 0.75)
+# default is 3/4ths split (but 75% training, 25% testing).
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+optimized_train <- training(optimized_split)
+optimized_test <- testing(optimized_split)
+# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+
+classif_rf <- train(effect_10~., data = optimized_train, method = "rf", ntree = 100, tuneLength = 1)
+
+classif_glm <- train(effect_10~., data = optimized_train, method = "glm", family = "binomial")
+
+classif_svm <- train(effect_10~., data = optimized_train, method = "svmRadial", prob.model = TRUE, tuneLength = 1)
+
+# In this case we consider the differences between observed class and predicted probabilities to be residuals.
+So, we have to provide custom predict function which takes two arguments: model and newdata and returns a numeric vector with probabilities.
+
+# create custom predict function
+p_fun <- function(object, newdata){
+  predict(object, newdata = newdata, type="prob")[,11]}
+
+# make response variable numeric binary vector
+yTest <- as.numeric(as.character(optimized_test$effect_10))
+
+explainer_classif_rf <- DALEX::explain(classif_rf, label = "rf",
+                                       data = optimized_test, 
+                                       y = yTest,
+                                       predict_function = p_fun, 
+                                       verbose = FALSE)
+
+explainer_classif_glm <- DALEX::explain(classif_glm, label = "glm", 
+                                        data = optimized_test, y = yTest,
+                                        predict_function = p_fun,
+                                        verbose = FALSE)
+
+explainer_classif_svm <- DALEX::explain(classif_svm,  label = "svm", 
+                                        data = optimized_test, y = yTest,
+                                        predict_function = p_fun,
+                                        verbose = FALSE)
+
+```
+### Model Performance
+```{r}
+mp_classif_rf <- model_performance(explainer_classif_rf)
+mp_classif_glm <- model_performance(explainer_classif_glm)
+mp_classif_svm <- model_performance(explainer_classif_svm)
+
+plot(mp_classif_rf, mp_classif_glm, mp_classif_svm)
+```
+#### Boxplot
+```{r}
+plot(mp_classif_rf, mp_classif_glm, mp_classif_svm, geom = "boxplot")
+```
+### Variable Importance
+```{r}
+vi_classif_rf <- model_parts(explainer_classif_rf, loss_function = loss_root_mean_square)
+vi_classif_glm <- model_parts(explainer_classif_glm, loss_function = loss_root_mean_square)
+vi_classif_svm <- model_parts(explainer_classif_svm, loss_function = loss_root_mean_square)
+
+plot(vi_classif_rf, vi_classif_glm, vi_classif_svm)
+```
+#### Partial Dependence Plot
+```{r}
+pdp_classif_rf  <- model_profile(explainer_classif_rf, variable = "dose.um3.mL.master", type = "partial")
+pdp_classif_glm  <- model_profile(explainer_classif_glm, variable = "dose.um3.mL.master", type = "partial")
+pdp_classif_svm  <- model_profile(explainer_classif_svm, variable = "dose.um3.mL.master", type = "partial")
+
+plot(pdp_classif_rf, pdp_classif_glm, pdp_classif_svm)
+```
+#### Accumulated Local Effects Plot
+
+Accumulated local effects31 describe how features influence the prediction of a machine learning model on average. ALE plots are a faster and unbiased alternative to partial dependence plots (PDPs).
+
+```{r}
+ale_classif_rf  <- model_profile(explainer_classif_rf, variable = "alcohol", type = "accumulated")
+ale_classif_glm  <- model_profile(explainer_classif_glm, variable = "alcohol", type = "accumulated")
+ale_classif_svm  <- model_profile(explainer_classif_svm, variable = "alcohol", type = "accumulated")
+
+plot(ale_classif_rf, ale_classif_glm, ale_classif_svm)
+```
+
+```{r}
+sessionInfo()
+```
+
 # Interpretable Tree

### 2021-02-19 BMU/BMuCaretapp.R

@@ -1,0 +1,334 @@
+## Best Model Selection
+#https://datascienceplus.com/find-the-best-predictive-model-using-r-caret-package-modelgrid/
+library(shiny)
+library(shinythemes)
+library(modelgrid)
+library(magrittr)
+library(caret)
+library(foreach)
+library(recipes)
+library(plyr)
+library(purrr)
+library(dplyr)
+library(pROC)
+library(BradleyTerry2)
+library(corrplot)
+library(DataExplorer)
+library(randomForest)
+library(glmnet)
+library(xgboost)
+library(e1071)
+library(tidyverse)
+library(tigerstats)
+library(skimr)
+
+##### load data #####
+set.seed(9650)
+
+# Load finalized dataset.
+data_set <- read_csv("multiVar.csv", guess_max = 10000) 
+
+data_set <- data_set %>% 
+  dplyr::select(c(size.length.um.used.for.conversions, shape, polymer, organism.group, environment, bio.org, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master, dose.particles.mL.master, dose.um3.mL.master)) %>% 
+    mutate_if(is.character, as.factor) %>% 
+  drop_na()
+
+###### Pre-Processing the data  using recipe #####
+set.seed(9650)
+Attri_rec_2 <- 
+    recipe(effect_f ~ ., data = data_set) %>%        # Formula.
+    step_dummy(all_predictors())%>%               # convert nominal data into one or more numeric.
+    step_knnimpute(all_predictors())%>%           # impute missing data using nearest neighbors.
+    step_zv(all_predictors()) %>%                 # remove variables that are highly sparse and unbalanced.
+    step_corr(all_predictors()) %>%               # remove variables that have large absolute correlations with 
+    # other variables.
+    step_center(all_predictors()) %>%             # normalize numeric data to have a mean of zero.
+    step_scale(all_predictors()) %>%              # normalize numeric data to have a standard deviation of one.
+    prep(training = data_set, retain = TRUE)      # train the data recipe
+
+
+data_rec_2 <- as.data.frame(juice(Attri_rec_2))
+# add the response variable
+data_rec_2$effect_f <- data_set$effect_f
+# str(data_rec_2)
+# Create a Validation Dataset (training data 70% and validation data 30% of the original data)
+set.seed(9650)
+validation_index <- createDataPartition(data_rec_2$effect_f,times= 1,  p= 0.70, list=FALSE)
[…]

### 2021-02-23 DALEX.Rmd
more diagnostics and tuned model!

@@
 date: "2/19/2021"
-output: html_document
+output: 
+  html_document:
+    code_folding: hide
+    theme: sandstone
+    toc: yes
+    toc_float: yes
+    toc_depth: 5
+    number_sections: true
+  word_document:
+    toc: yes
 ---
+```{r global-options, include=FALSE}
+knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
+                      warning=FALSE, message=FALSE, echo = FALSE)
+```
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-https://uc-r.github.io/dalex#:~:text=DALEX%20is%20an%20R%20package,performance%20across%20multiple%20models%20convenient.
-```{r}
+```{r message=FALSE, warning=FALSE}
 # load required packages
@@
 library(caret)
+library(skimr)
+library(ggeffects)
+library(knitr)
+library(doParallel)
+library(modelplotr)
 ```
-# Data Import
+# Setup
+## Data Import
 ```{r}
@@
 df <- aoc_z %>% 
-  dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, dose.mg.L.master, dose.particles.mL.master, life.stage, bio.org, polymer, shape, size.length.um.used.for.conversions, treatments, effect)) %>%
+  dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, dose.mg.L.master, dose.particles.mL.master, life.stage, bio.org, polymer, shape, size.length.um.used.for.conversions, treatments, effect, effect.metric, particle.volume.um3, density.mg.um.3, environment, exposure.route, lvl2_f)) %>%
+  mutate(effect.metric = as.character(effect.metric)) %>% 
+  mutate(effect.metric = (case_when(
+    effect.metric == "NONEC" ~ "NOEC",
+    effect.metric == "HONEC" ~ "NOEC",
+    effect.metric == "LOEC" ~ "LOEC",
+    effect.metric == "LC50" ~ "LC50",
+    effect.metric == "EC50" ~"EC50",
+    effect.metric == "EC10" ~ "EC10"
+  ))) %>% 
+  mutate(effect.metric = replace_na(effect.metric,"not_available")) %>% 
+  mutate(effect.metric = as.factor(effect.metric)) %>% 
   drop_na() %>%  #drop missing
@@
 
-```{r message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
 # Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
@@
 
-classif_svm <- train(effect_10~., data = train, method = "svmRadial", prob.model = TRUE, tuneLength = 1,
-                     preProcess = c("pca","scale","center"))
+classif_svm <- train(effect_10~., data = train, method = "svmRadial", prob.model = TRUE, tuneLength = 1)#,
+                    # preProcess = c("pca","scale","center"))
 
@@
 ```
-### Model Performance
+# Modelling
+## Model Performance
 ```{r}
@@
 ```
+
+#### Lift Curves
+
+Lift curves describe a performance coefficient (lift) over the cumulative proportion of a population.
+Lift is calculated as the ratio of "yes's" on a certain sample point (for toxicity) divided by the ratio of "yes's" on the whole dataset.
+$Lift = Predicted Rate/Average Rate$.
+
+```{r eval=FALSE, include=FALSE}
+#### DO NOT RUN - MAKES WEIRD PLOT###
+
+# #recode class as character
+# df2 <- df %>% 
+#  mutate(effect = factor(case_when( #convert ordinal to numeric
+#       effect_10 == "1" ~ "Y",
+#       effect_10 == "0" ~ "N"
+#     )))
+# #split test/train
+# set.seed(4)
+# df2_split <- df2 %>%
+#   initial_split(prop = 0.75)
+# train2 <- training(df2_split)
+# test2 <- testing(df2_split)
+# 
+# #set controls
+# ctrl <- trainControl(method = "cv", classProbs = TRUE,
+#                      summaryFunction = twoClassSummary)
+# #create models for lift plots
+# lift_rf <- train(effect ~., data = train2, method = "rf", ntree = 100, tuneLength = 1,
+#                     trControl = ctrl)
+# 
+# lift_glm <- train(effect~., data = train2, method = "glm", family = "binomial", 
+#                      trControl = ctrl)
+# 
+# lift_svm <- train(effect~., data = train2, method = "svmRadial", prob.model = TRUE, 
+#                      tuneLength = 1, trControl = ctrl)#,
+#                     # preProcess = c("pca","scale","center"))
+# 
+# lift_decTree <- train(effect~., data = train2, method = "C5.0", 
+#                          preProcess =c("scale","center"), trControl = ctrl) 
+# 
+# lift_nnet <- train(effect~., data = train2, method = "nnet", preProcess=c("scale","center"),
+#                       trControl = ctrl)
+# 
+# lift_xgbTree <- train(effect~., data = train2, method = "xgbTree", trControl = ctrl)
+# 
+# 
+# ## Generate the test set results
+# lift_results <- data.frame(effect = test2$effect)
+# lift_results$rf <- predict(lift_rf, test, type = "prob")[,"Y"]
+# lift_results$glm <- predict(lift_glm, test, type = "prob")[,"Y"]
+# lift_results$svm <- predict(lift_svm, test, type = "prob")[,"Y"]
+# lift_results$decTree <- predict(lift_decTree, test, type = "prob")[,"Y"]
+# lift_results$nnet <- predict(lift_nnet, test, type = "prob")[,"Y"]
+# lift_results$xgbTree <- predict(lift_xgbTree, test, type = "prob")[,"Y"]
+# head(lift_results)
+# 
+# #plot results
+# trellis.par.set(caretTheme())
+# lift_obj <- lift(effect ~ rf + glm + svm + decTree + nnet + xgbTree, data = lift_results)
+# plot(lift_obj, values = 60, auto.key = list(columns = 3,
+#                                             lines = TRUE,
+#                                             points = FALSE))
+```
+##### ALT METHOD
+http://rstudio-pubs-static.s3.amazonaws.com/436131_3212dcf341cc422590f1a9f52830cfd6.html
+
+```{r}
+# transform datasets and model objects into scored data and calculate deciles 
+scores_and_ntiles <- prepare_scores_and_ntiles(datasets=list("train","test"),
+                                               dataset_labels = list("train data","test data"),
+                                               models = list("classif_rf", "classif_decTree",
+                                                             "classif_glm", "classif_nnet",
+                                                             "classif_svm", "classif_xgbTree"),
+                                               model_labels = list("random forest", "Decision
+                                                                   tree", "General linear model",
+                                                                   "Neural net", "Support Vector
+                                                                   machine", "eXtreme Gradient
+                                                                   Boosting Trees"),
+                                               target_column="effect_10",
+                                               ntiles = 100)
+
+# transform data generated with prepare_scores_and_deciles into aggregated data for chosen plotting scope 
+plot_input <- plotting_scope(prepared_input = scores_and_ntiles, scope = 'compare_models')
+plot_cumgains(data = plot_input, custom_line_colors = RColorBrewer::brewer.pal(2,'Accent'))
+```
+
+```{r}
+plot_cumlift(data = plot_input,custom_line_colors = RColorBrewer::brewer.pal(2,'Accent'))
+```
+```{r}
+plot_cumresponse(data = plot_input,highlight_ntile = 20, 
+                 custom_line_colors = RColorBrewer::brewer.pal(2,'Accent'))
+```
+```{r}
+plot_multiplot(data = plot_input,  custom_line_colors = RColorBrewer::brewer.pal(2,'Accent'))
+```
+
 #### Roc Curves
@@
 ### Variable Importance
-```{r fig.height=7, fig.width=5}
+```{r fig.height=7, fig.width=5, warning=FALSE}
 vi_classif_rf <- model_parts(explainer_classif_rf, loss_function = loss_root_mean_square)
@@
 
-plot(vi_classif_rf, vi_classif_glm, vi_classif_svm, vi_classif_decTree, vi_classif_nnet, vi_classif_xgbTree)#, #vi_classif_logicBag)
+plot(vi_classif_rf, vi_classif_glm, vi_classif_svm)#, #vi_classif_logicBag)
 ```
+```{r}
+plot(vi_classif_decTree, vi_classif_nnet, vi_classif_xgbTree)
+```
+
 #### Partial Dependence Plot
@@
 # filter for top 10 influential variables for each model and plot
-list(new_cust_glm, new_cust_rf, new_cust_svm, new_cust_decTree, new_cust_nnet, 
-     #new_cust_logicBag,
-     new_cust_xgbTree) %>%
+list(new_cust_glm, new_cust_rf) %>%
   purrr::map(~ top_n(., 11, wt = abs(contribution))) %>%
@@
 ```
+
+```{r}
+library(ggplot2)
+
+# filter for top 10 influential variables for each model and plot
+list(new_cust_svm, new_cust_decTree) %>%
+  purrr::map(~ top_n(., 11, wt = abs(contribution))) %>%
+  do.call(rbind, .) %>%
+  mutate(variable = paste0(variable, " (", label, ")")) %>%
+  ggplot(aes(contribution, reorder(variable, contribution))) +
+  geom_point() +
+  geom_vline(xintercept = 0, size = 3, color = "white") +
+  facet_wrap(~ label, scales = "free_y", ncol = 1) +
+  ylab(NULL)
+```
+
+```{r}
+library(ggplot2)
+
+# filter for top 10 influential variables for each model and plot
+list(new_cust_nnet,new_cust_xgbTree) %>%
+  purrr::map(~ top_n(., 11, wt = abs(contribution))) %>%
+  do.call(rbind, .) %>%
+  mutate(variable = paste0(variable, " (", label, ")")) %>%
+  ggplot(aes(contribution, reorder(variable, contribution))) +
+  geom_point() +
+  geom_vline(xintercept = 0, size = 3, color = "white") +
+  facet_wrap(~ label, scales = "free_y", ncol = 1) +
+  ylab(NULL)
+```
+
 # Final Model
-All in all random forest is my final model of choice: it appears the more balanced and is the most accurate overall.
+All in all random forest is my final model of choice: it appears the more balanced and is the most accurate overall. This model will now be tuned and refined for maximum performance.
+
+## Tuning
+http://rstudio-pubs-static.s3.amazonaws.com/480890_237ad52b09b6440e9c849a3c07a04d2f.html
+```{r include=FALSE}
+cache = TRUE
+set.seed(1000)
+train_control <- trainControl(method = "repeatedcv", 
+                              number = 10,
+                              repeats = 5,
+                              verboseIter = TRUE,
+                              allowParallel = TRUE,
+                              summaryFunction = multiClassSummary)
+cache = TRUE
+set.seed(1000)
+
+start_time <- Sys.time() # Start timer
+
+my_grid1 <- expand.grid(mtry = 1:17)
+
+rf1 <- train(effect_10 ~ .,
+             data = train,
+             method = "rf",
+             metric = "Accuracy",
+             tuneGrid = my_grid1,
+             trControl = train_control)
+rf1
+```
+
+Plot tuning.
+
+```{r}
+cache = TRUE
+
+my_plot <- function(model) {
+    theme_set(theme_minimal())
+    u <- model$results %>%
+        select(mtry, Accuracy, Kappa, F1, Sensitivity,
+              Specificity,Pos_Pred_Value, Neg_Pred_Value, 
+               Precision, Recall, Detection_Rate) %>%
+        gather(a, b, -mtry)
+
+    u %>% ggplot(aes(mtry, b)) + geom_line() + geom_point() + 
+        facet_wrap(~ a, scales = "free") + 
+        labs(x = "Number of mtry", y = NULL, 
+             title = "The Relationship between Model Performance and mtry")
+}
+
+rf1 %>% my_plot()
+```
+
 I will further refine this model using recursive feature elimination and compare accuracy with the other models.
@@
 
-my_ctrl <- rfeControl(functions = rfFuncs,
+my_ctrl <- rfeControl(functions = rfFuncs, #random forests
                       method = "repeatedcv",
@@
 ```
+
+## Predictor Selection
+The following variables are those that were picked in the final (most accurate) model. 
 ```{r}
-y <- "effect_10"
-x <- setdiff(names(df), y)
-
-#alt method
-rfRFE <-  list(summary = defaultSummary,
-               fit = function(x, y, first, last, ...){
-                 library(randomForest)
-                 randomForest(x, y, importance = first, ...)
-                 },
-               pred = function(object, x)  predict(object, x),
-               rank = function(object, x, y) {
-                 vimp <- varImp(object)
-                 vimp <- vimp[order(vimp$Overall,decreasing = TRUE),,drop = FALSE]
-                 vimp$var <- rownames(vimp)                  
-                 vimp
-                 },
-               selectSize = pickSizeBest,
-               selectVar = pickVars)
-```
-
-```{r}
+#get variable names picked in the final model
 predictors(rfProfile)
 ```
-
+The first 6 models are shown below with their corresponding accuracy and kappa values. 
 ```{r}
@@
 trellis.par.set(caretTheme())
-plot(rfProfile, type = c("g", "o"))
+#plot(rfProfile, type = c("g", "o"), metric = "Accuracy")
+ggplot(rfProfile) + theme_bw()
+```
+```{r}
+plot1 <- xyplot(rfProfile, type = c("g", "p", "smooth"), main = "Accuracy of individual resampling results")
+plot2 <- densityplot(rfProfile, 
+                     subset = Variables < 5, 
+                     adjust = 1.25, 
+                     as.table = TRUE, 
+                     xlab = "Accuracy Estimates", 
+                     pch = "|")
+print(plot1, split=c(1,1,1,2), more=TRUE)
+print(plot2, split=c(1,2,1,2))
+```
+
+### Final Predictor Selection
+```{r}
+my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 1, maximize = TRUE)
+# higher tol (~10) gives you less variables
+# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
+accuracy1 <- pickVars(rfProfile$variables, size = my_size)
+accuracy1
+```
+A random forest model with the above four predictors is within 1% of the best model.
+If a more accurate model is desired, and data is available, additional factors can be included in the model.
+Shown below are factors that should be included for a model that is within 0.5% accuracy of the best model.
+
+```{r}
+my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 0.5, maximize = TRUE)
+# higher tol (~10) gives you less variables
+# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
+accuracy0.5 <- pickVars(rfProfile$variables, size = my_size)
+accuracy0.5
+```
+Below is a table showing which variables are included in models of varying accuracies.
+
+```{r}
+my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 0.1, maximize = TRUE)
+accuracy0.1 <- pickVars(rfProfile$variables, size = my_size)
+my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 0.3, maximize = TRUE)
+accuracy0.3 <- pickVars(rfProfile$variables, size = my_size)
+my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 5, maximize = TRUE)
+accuracy5 <- pickVars(rfProfile$variables, size = my_size)
+my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 10, maximize = TRUE)
+accuracy10 <- pickVars(rfProfile$variables, size = my_size)
+
+varsTable = list('0.1%' =accuracy0.1, '1%' = accuracy1, '5%' = accuracy5, '10%' = accuracy10)
+
+#$make padded dataframe
+na.pad <- function(x,len){
+    x[1:len]
+}
+
+makePaddedDataFrame <- function(l,...){
+    maxlen <- max(sapply(l,length))
+    data.frame(lapply(l,na.pad,len=maxlen),...)
+}
+#fancy print
+kable(makePaddedDataFrame(
+  list('Accuracy (0.1)' = accuracy0.1, 'Accuracy (0.3)' = accuracy0.3, "Accuracy (1)" = accuracy1, 'Accuracy (5)' = accuracy5, 'Accuracy (10)' = accuracy10)
+  ),
+      caption = "Variables to include in Random Forest to achieve Model Accuracy (within x% of best model)",
+    footnote = "Random Forest Recursive Feature Elimination")
+```
+### Final Model Tuning
+While the absolute best model contains 18 variables with an estimated accuracy of 93.89%, it is possible to achieve ~93% accuracy with just eight variables, as displayed in the above table.
+The advantage of including fewer variables is that less information is needed to accurately predict toxicity.
+The final model with these eight variables will be further refined through an iterative tuning process to determine the optimal number of variables to be randomly collected for sampling at each split (mtry), and the number of branches to grow after each split.Code for the following section is inspired from https://rpubs.com/phamdinhkhanh/389752.
+```{r}
+final_df <-df %>% 
+  dplyr::select(c(effect.metric, dose.mg.L.master, lvl2_f, dose.particles.mL.master, effect_10, dose.um3.mL.master, organism.group, exposure.duration.d, treatments)) %>% 
+  droplevels()
+
+#ensure completeness
+skim(final_df)
 ```
@@
 ```{r}
+# create train, validation, and test splits
+# Create calibration and validation splits with tidymodels initial_split() function.
+set.seed(4)
+df_final_split <- final_df %>%
+  initial_split(prop = 0.75)
+# default is 3/4ths split (but 75% training, 25% testing).
+
+# Create a training data set with the training() function
+# Pulls from training and testing sets created by initial_split()
+train_final <- training(df_final_split)
+test_final <- testing(df_final_split)
+
+# variable names for resonse & features
+y <- "effect_10"
+x <- setdiff(names(final_df), y) 
+```
+
+The mtry parameter will be optimized below:
+
+```{r}
+tunegrid <- expand.grid(.mtry=(1:7))
+                       # , .ntree=c(500, 1000, 1500, 2000, 2500)) #would like to optimize but can't figure out how!
+
+fitControl <- trainControl(## 10-fold CV
+                           method = "repeatedcv",
+                           number = 10,
+                           ## repeat 3 times
+                           repeats = 3)
+
+nrow(tunegrid)
+
+set.seed(825)
+tuneFit <- train(effect_10 ~ ., data = train_final, 
+                 method = "rf", 
+                 trControl = fitControl, 
+                 verbose = FALSE, 
+                 ## Now specify the exact models 
+                 ## to evaluate:
+                 metric = 'Accuracy',
+                 tuneGrid = tunegrid)
+tuneFit
+```
+
+```{r}
+plot(tuneFit)
+```
+Now that mtry is optimized, the number of trees will be optimized, holding mtry constant.
+
+```{r}
+# Manual Search
+control <- trainControl(method="repeatedcv", number=10, repeats=3, search="grid")
+tunegrid <- expand.grid(.mtry=c(7)) #replace with optimal value from previous chunk
+modellist <- list()
+for (ntree in c(1000, 1500, 2000, 2500)) {
+	set.seed(123)
+	fit <- train(effect_10~., data=train_final, method="rf", metric='Accuracy', 
+	             tuneGrid=tunegrid, trControl=control, ntree=ntree)
+	key <- toString(ntree)
+	modellist[[key]] <- fit
+}
+# compare results
+results <- resamples(modellist)
+summary(results)
+```
+
+```{r}
+dotplot(results)
+```
+
+### Final Model
+Now that we have optimized the variables for our final model, we will build and save the final simplified model that has the highest accuracy.
+```{r}
+tunegrid <- expand.grid(.mtry=c(7))
+
+final_model <- train(effect_10~., data = train_final, method = "rf", ntree = 2500, tuneLength = 5)
+
+# explain
+yTest <- as.numeric(as.character(test_final$effect_10))
+
+explainer_final_model <- DALEX::explain(final_model, label = "rf",
+                                       data = test_final, 
+                                       y = yTest)
+
+mp_classif_final_model <- model_performance(explainer_final_model)
+mp_classif_final_model
+```
+#### Performance
+Performance metrics for the final model are below.
+
+##### ROC Curve
+
+```{r}
+plot(mp_classif_final_model,
+     geom = "roc") +
+  ggtitle("ROC Curve - Final Model",  
+          paste("AUC = ",round(mp_classif_final_model$measures$auc,2),
+                paste("Accuracy = ", 100 * round(mp_classif_final_model$measures$accuracy,3), "%")
+          )) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.subtitle = element_text(hjust = 0.5),
+        legend.position = "none")
+```
+
+##### Confusion Matrix
+The tuned, final, simplified model is now validated using test set data.
+```{r}
+test_pred <- predict(final_model, newdata = test_final)
+confusionMatrix(table(test_pred, test_final$effect_10))
+```
+This confusion matrix is plotted below.
+```{r}
+table <- data.frame(confusionMatrix(test_pred, test_final$effect_10)$table)
+
+plotTable <- table %>%
+  mutate(goodbad = ifelse(table$Prediction == table$Reference, "good", "bad")) %>%
+  group_by(Reference) %>%
+  mutate(prop = Freq/sum(Freq))
+
+# fill alpha relative to sensitivity/specificity by proportional outcomes within reference groups (see dplyr code above as well as original confusion matrix for comparison)
+ggplot(data = plotTable, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
+  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
+  scale_x_discrete(labels = c("No Effect", "Effect")) +
+  scale_y_discrete(labels = c("No Effect", "Effect"))
+  theme_bw() +
+  xlim(rev(levels(table$Reference)))
+```
+
+# Machine Settings
+
+Machine Info:
+```{r}
+Sys.info()[c(1:3,5)]
+```
+
+Session Info:
+```{r}
 sessionInfo()

### 2021-02-23 BMU/BMuCaretapp.R

@@ -31,7 +31,7 @@ data_set <- read_csv("multiVar.csv", guess_max = 10000)
 data_set <- data_set %>% 
   dplyr::select(c(size.length.um.used.for.conversions, shape, polymer, organism.group, environment, bio.org, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master, dose.particles.mL.master, dose.um3.mL.master)) %>% 
     mutate_if(is.character, as.factor) %>% 
-  drop_na()
+  drop_na(effect_f)
 
 ###### Pre-Processing the data  using recipe #####
 set.seed(9650)

### 2021-02-23 SVM,kNN.Rmd

--- 2021-02-04 multivariate.Rmd
@@
 ---
-title: "Multivariate Analysis"
+title: "SVM, K-NN"
 author: "Scott Coffin"
-date: "1/19/2021"
-output: html_document
+date: "2/18/2021"
+output: 
+ html_document:
+    code_folding: hide
+    theme: sandstone
+    toc: yes
+    toc_float: yes
+    toc_depth: 4
+    number_sections: true
+  word_document:
+    toc: yes
 ---
+```{r global-options, include=FALSE}
+knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
+                      warning=FALSE, message=FALSE, echo = FALSE)
+```
+---
+
 # Setup
@@
 library(ISLR)  #for the Carseat Data
+library(lme4) #general linearized mixed model GLMM
 library(quantregForest)
-library(caret)
+library(caret) #support vector machines
 library(tidyverse)
@@
 library(gt)
+library(randomForest)
+
+# Attach Packages
+library(kernlab)      # SVM methodology
+library(e1071)        # SVM methodology
 ```
@@
 # Load finalized dataset.
-aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
-
-#assign row ID
-aoc <- rowid_to_column(aoc, "ID")
-aoc <- aoc %>% 
-  mutate(ID = paste0(ID, "ID"))
+aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) %>% 
+  mutate(ID = paste0("ID",row_number()))
 
@@
 # subset data to selected variables
+
 multiVar <- aoc_z %>% dplyr::select(#doi, size.category, 
-                                    ID,
                                     size_f,
@@
                                     treatments, #number of doses (no including control)
-                                    effect_f, #yes no
+                                    effect, #yes no
+                                    effect_f,
+                                    effect_10,
+                                    size_f,
                                     exposure.duration.d, 
@@
                                    dose.particles.mL.master.AF.noec,
+                                   max.size.ingest.mm, #maximum ingestible size range
                                    effect.score) %>%  #1 = minor, 2 = photosynthesis, feeding, 3 = growth, chlorophyll content, 4 = reproduction, 5  = population growth, 6 = survival
   filter(!size_f == "Not Reported") %>%   #take out not reported 
-  mutate_if(is.character, as.factor) %>% mutate(effect_10 = case_when(
-     effect_f == "Yes" ~ 1,
-     effect_f == "No" ~ 0))
-```
-
-# Analysis
-## Data Exploration
-### Completeness vby Size
-```{r completeness heatmap}
-CompletenessSummary_size <- multiVar %>%
-  group_by(size_f) %>% 
-     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
-  mutate(across(is.numeric,~round(., 2))) %>% 
-  mutate(across(is.numeric, ~100 *(1 -.)))
-CompletenessSummary_size
-```
-
-```{r}
-require()
-#convert to matrix and transpose
-transposed_size <- as.data.frame(t(as.matrix(CompletenessSummary_size[2:28]))) %>% #1 is category, 2-6 are variables
-  arrange('1nm < 100nm')
-#reassign column names
-colnames(transposed_size) <- c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm")
-#transposedTag
-#format as matrix
-MissingMatrix_size <- data.matrix(transposed_size)
-#build heatmap
-pheatmap(MissingMatrix_size,
-                           main = "Data Completeness by Size Bin", #title
-                           fontsize = 12,
-                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
-                           display_numbers = TRUE,
-                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
-                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)
-
-```
-### Completeness by Environment
-```{r completeness heatmap}
-CompletenessSummary_environment <- multiVar %>%
-  filter(!environment == "NA") %>% 
-  group_by(environment) %>% 
-     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
-  mutate(across(is.numeric,~round(., 2))) %>% 
-  mutate(across(is.numeric, ~100 *(1 -.)))
-CompletenessSummary_environment
-```
-
-```{r}
-require()
-#convert to matrix and transpose
-transposed_environment <- as.data.frame(t(as.matrix(CompletenessSummary_environment[2:28])))
-#reassign column names
-colnames(transposed_environment) <- c("Freshwater", "Marine", "Terrestrial")
-#transposedTag
-#format as matrix
-MissingMatrix_environment <- data.matrix(transposed_environment)
-#build heatmap
-pheatmap(MissingMatrix_environment,
-                           main = "Data Completeness by environment", #title
-                           fontsize = 12,
-                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
-                           display_numbers = TRUE,
-                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
-                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)
-
-```
-
-## Stepwise Regression
-
-```{r}
-## Estimate an OLS Regression
-fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master +mouth, 
-             na.action = na.omit, 
-             data = multiVar)
-summary(fitols)
-```
-
-```{r}
-require(reshape2)
-multiVar %>% 
-  dplyr::select(size.length.um.used.for.conversions, particle.volume.um3, exposure.duration.d ,media.temp, dose.particles.mL.master) %>% 
-  melt() %>%  #convert wide to long
-   mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  ggplot(aes(x = value)) + 
-  stat_density() + 
-  facet_wrap(~variable, scales = "free")
-```
-
-<!-- # ```{r} -->
-<!-- # # cdplot(as.factor(effect) ~ particle.volume.um3, data = multiVar, -->
-<!-- #        main = "Conditional Density Plot of ", xlab = "Dose", ylab = "Effect") -->
-<!-- # ``` -->
-
-## PCA
-
-# UMAP
-
-```{r}
-require(uwot)
-require(Rtsne)
-require(vizier) #devtools::install_github("jlmelville/vizier")
-
-# For some functions we need to strip out non-numeric columns and convert data to matrix
-x2m <- function(X) {
-  if (!methods::is(X, "matrix")) {
-    m <- as.matrix(X[, which(vapply(X, is.numeric, logical(1)))])
-  }
-  else {m <- X} 
-  m}
-
-#choose values with most completeness
-multiVar2 <- multiVar %>% 
-  filter(!environment == "Terrestrial") %>% 
-  dplyr::select(size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  drop_na() #drop missing
-
-#convert discrete variables to numeric
-multiVar2[] <- data.matrix(multiVar2)
-
-# build umap for small dataset (<10,000 points)
-multiVar_map  <- umap(multiVar2, pca = 10)
-
-# Remove duplicates for t-SNE
-#multiVar2_noNa_dup <- multiVar2_noNa[-which(duplicated(x2m(multiVar2_noNa))), ]
-
-#build t-SNE
-#multiVar_tsne <- Rtsne::Rtsne(multiVar2_noNa_dup, perplexity = 15, initial_dims = 100, partical_pca = TRUE, exaggeration_factor = 4)
-
-# Non-numeric columns are ignored, so in a lot of cases you can pass a data
-# frame directly to umap
-#iris_umap <- umap(iris, n_neighbors = 50, learning_rate = 0.5, init = "random")
-
-#visualize umap
-embed_img <- function(X, Y, k = 15, ...) {
-  args <- list(...)
-  args$coords <- Y
-  args$x <- X
-
-  do.call(vizier::embed_plot, args)
-}
-
-#plot
-embed_img(multiVar2, multiVar_map, pc_axes = TRUE, equal_axes = TRUE, alpha_scale = 0.5, title = "Tox UMAP", cex = 1)
-```
-
-```{r}
-#PCA
-pca <- stats::prcomp(multiVar2[,-4], retx = TRUE, rank. = 2)
-#build color pallete
-my_colors = colorRampPalette(c("red", "yellow", "green"))(nrow(multiVar2))
-#plot
-embed_plot(pca$x, multiVar2$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), #turbo, #rainbow, #my_colors, 
-           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE)
-```
-
-```{r}
-require(plotly)
-#PCA for discrete variables
-multiVar_discrete <- multiVar %>% 
-  select(size.length.um.used.for.conversions, polymer, dose.mg.L.master, exposure.duration.d, effect) %>% 
-  drop_na %>% 
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  mutate(effect_10 = case_when(
-    effect == "Y" ~ 1,
-    effect == "N" ~ 0
-  )) %>% 
-  select(-effect)
-
-#buildPCA
-pca_discrete <- stats::prcomp(multiVar_discrete[,-2], retx = TRUE, rank. = 5, scale. = TRUE)
-
-embed_plot(pca_discrete$x, multiVar_discrete$polymer)
-
-embed_plotly(pca_discrete$x, multiVar_discrete$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), 
-           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE,
-           tooltip = paste("Polymer:", multiVar_discrete$polymer))
-
-#embed_plotly also option
-```
-
-# Random Forest
-## Recursive Partitioning And Regression Trees
-
-The rpart algorithm works by splitting the dataset recursively, which means that the subsets that arise from a split are further split until a predetermined termination criterion is reached.
-At each step, the split is made based on the independent variable that results in the largest possible reduction in heterogeneity of the dependent (predicted) variable.
-
-It is important to note that the algorithm works by making the best possible choice at each particular stage, without any consideration of whether those choices remain optimal in future stages.
-That is, the algorithm makes a locally optimal decision at each stage.
-It is thus quite possible that such a choice at one stage turns out to be sub-optimal in the overall scheme of things.
-In other words,  the algorithm does not find a globally optimal tree.
-
-The size of the data set usually imposes a hard limit on how many features may be considered due to Curse of Dimensionality, i.e.
-your data becomes sparser and sparser as you increase the number of features considered, which usually leads to overfitting.
-While there is no rule of thumb relating to how many features vs.
-the number of observations you should use, I try to keep $e^Nf < No$ (Nf = number of features, No = number of observations) to minimise overfitting [this is not always possible and it does not ensure that we won’t overfit].
-In this case, our training set has 3200, which suggests that using 8 features (e^8 ~ 2980) is not entirely absurd.
-Obviously, this depends on your data, so we will cover some further overfitting checks later on.
-```{r}
-#trim data so effect is always known
-multiVar_sub <- multiVar %>% 
-  drop_na(effect_10)
-
-# Split data into training and test sets
-set.seed(42)
-multiVar_sub[,"train"] <- ifelse(runif(nrow(multiVar_sub)) < 0.8, 1, 0)
-# Separate trainig and test sets
-trainSet <- multiVar_sub[multiVar_sub$train==1,]
-testSet <- multiVar_sub[multiVar_sub$train==0,]
-#get column index of train flag
-trainColNum <- grep("train", names(trainSet))
-# Remove train flag column from train and test sets
-trainSet <- trainSet[, -trainColNum]
-testSet <- testSet[, -trainColNum]
-```
-Make a classification tree.
-
-```{r, echo = FALSE}
-#get column index of predicted variable in dataset
-typeColNum <- grep("effect",names(multiVar_sub))
-
-#build tree
-require(rpart)
-set.seed(15097)
-t1 <- rpart(effect ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,
-            method = "class", #classification because response is discrete
-            control = rpart.control(minbucket = 20, cp=0.008), #requires that there be at least 19 cases (responded + nonrespondents) in the final grouping of variable values of the terminal node of the tre..
-            data = trainSet)
-print(t1, digits=4)
-```
-
-Plot an interpretable tree.
-```{r, echo = FALSE}
-require(rpart.plot)
-cols <- ifelse(t1$frame$yval == 1, "gray50", "black")
-prp(t1, main="Tree for Effect",
-    extra=106, # display prob of survival and percent of obs
-    nn=TRUE, # display node numbers
-    fallen.leaves=TRUE, # put leaves on the bottom of page
-    branch=.5, # change angle of branch lines
-    faclen=0, # do not abbreviate factor levels
-    trace=1, # print automatically calculated cex
-    shadow.col="gray", # shadows under the leaves
-    branch.lty=1, # draw branches using solid lines
-    branch.type=5, # branch lines width = weight(frame$wt), no. of cases here
-    split.cex=1.2, # make split text larger than node text
-    split.prefix="is ", # put "is " before split text
-    split.suffix="?", # put "?" after split text
-    col=cols, border.col=cols, # cols[2] if survived
-    split.box.col="lightgray", # lightgray split boxes (default is white)
-    split.border.col="darkgray", # darkgray border on split boxes
-    split.round=0.5) # round the split box corners a tad
-```
-Next we see how good the model is by seeing how it fares against the test data.
-```{r}
-t1_predict <- predict(t1, newdata = testSet[,-typeColNum],
-                      type="class")
-mean(t1_predict==testSet$effect)
-# [1] 0.7094088
-#confusion matrix
-table(pred=t1_predict,true=testSet$effect)
-```
-```{r}
-par(mfrow=c(1,2)) # two plots on one page
-#plot approximate R-squared and relative error for different splits (2 plots). labels are only appropriate for the "anova" method.
-rsq.rpart(t1)
-```
-
-Next, we prune the tree using the cost complexity criterion.
-Basically, the intent is to see if a shallower subtree can give us comparable results.
-If so, we’d be better of choosing the shallower tree because it reduces the likelihood of overfitting.
-
-As described earlier, we choose the appropriate pruning parameter (aka cost-complexity parameter) \alpha by picking the value that results in the lowest prediction error.
-Note that all relevant computations have already been carried out by R when we built the original tree (the call to rpart in the code above).
-All that remains now is to pick the value of \alpha:
-
-### Pruning
-
-```{r}
-#cost-complexity pruning
-printcp(t1)
-```
-It is clear from the above, that the lowest cross-validation error (xerror in the table) occurs for \alpha =0.008 (this is CP in the table above).   One can find CP programatically like so:
-```{r}
-# get index of CP with lowest xerror
-opt <- which.min(t1$cptable[,"xerror"])
-#get its value
-cp <- t1$cptable[opt, "CP"]
-```
-Next, we prune the tree based on this value of CP:
-
-```{r}
-# prune the tree
-pt1 <- prune(t1,cp)
-#pt1<- prune(t1, cp= t1$cptable[which.min(t1$cptable[,"xerror"]),"CP"])
-
-# plot the pruned tree
-plot(pt1, uniform=TRUE,
-   main="Pruned Classification Tree");text(pt1, use.n=TRUE, all=TRUE, cex=.8)
-
-#post(pfit, file = "c:/ptree.ps",
- #  title = "Pruned Classification Tree for Kyphosis")
-```
-```{r}
-#find proportion of correct predictions using test set
-t1_pruned_predict <- predict(pt1, testSet, type="class")
-mean(t1_pruned_predict == testSet$effect)
-#
-```
-This is not an improvement over an unprunend tree..
-We need to check that this holds up for different training and test sets.
-This is easily done by creating multiple random partitions of the dataset and checking the efficacy of pruning for each.
-To do this efficiently, I’ll create a function that takes the training fraction, number of runs (partitions) and the name of the dataset as inputs and outputs the proportion of correct predictions for each run.
-It also optionally prunes the tree.
-
-```{r}
-#function to do multiple runs
-multiple_runs_classification <- function(train_fraction,n,dataset,prune_tree=FALSE){
-fraction_correct <- rep(NA,n)
-set.seed(42)
-for (i in 1:n){
-  dataset[,"train"] <- ifelse(runif(nrow(dataset))<0.8,1,0)
-  trainColNum <- grep("train",names(dataset))
-  typeColNum <- grep("effect",names(dataset))
-  trainSet <- dataset[dataset$train==1,-trainColNum]
-  testSet <- dataset[dataset$train==0,-trainColNum]
-  rpart_model <- rpart(effect~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,data = trainSet, method="class")
-if(prune_tree==FALSE) {
-  rpart_test_predict <- predict(rpart_model,testSet[,-typeColNum],type="class")
-  fraction_correct[i] <- mean(rpart_test_predict==testSet$effect)
-  }else{
-    opt <- which.min(rpart_model$cptable[,"xerror"])
-    cp <- rpart_model$cptable[opt, "CP"]
-    pruned_model <- prune(rpart_model,cp)
-    rpart_pruned_predict <- predict(pruned_model,testSet[,-typeColNum],type="class")
-    fraction_correct[i] <- mean(rpart_pruned_predict == testSet$effect)
-  }
-  }
-return(fraction_correct)
-}
-```
-
-Note that in the above,  I have set the default value of the prune_tree to FALSE, so the function will execute the first branch of the if statement unless the default is overridden.
-
-OK, so let’s do 50 runs with and without pruning, and check the mean and variance of the results for both sets of runs.
-```{r}
-#50 runs, no pruning
-unpruned_set <- multiple_runs_classification(0.8,50,multiVar_sub)
-mean(unpruned_set)
-#[1] 0.7261882
-sd(unpruned_set)
-#[1] 0.01554347
-#50 runs, with pruning
-pruned_set <- multiple_runs_classification(0.8,50,multiVar_sub,prune_tree=TRUE)
-mean(pruned_set)
-#[1] 0.7261875
-sd(pruned_set)
-#[1] 0.01552285
-```
-
-## CForest
-```{r run model}
-require(party)
-crf <- cforest(as.factor(effect_10) ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d +
-                 media.temp + dose.particles.mL.master,
-               controls = cforest_control(ntree = 500,
-                                          mincriterion = qnorm(0.8), 
-                                          trace = TRUE), # adds project bar because it's very slow
-               data = multiVar_sub)
-
-crf
-```
-
-```{r fit results}
-train2 <- multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
-                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master)
-train3 <-multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
-                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master,
-                                  effect_10)
-
-fitted <- predict(crf, train2, OOB = TRUE, type ="response")
-#rpart.prob <- predict(t1, newdata=imputedSmalls.requested.voluntary,type="prob")
-
-misClasificError <- mean(fitted != train3$effect_10)
-print(paste('Training Accuracy', 1 - misClasificError))
-```
-
-```{r}
-print(crf)
-```
-
-```{r}
-#Obtaining predicted probabilites for Test data
-tree.probs=predict(crf,
-                 newdata = train2,
-                 type="prob")
-
-#Calculate ROC curve
-rocCurve.tree <- roc(train3$effect_10,tree.probs[,'1'])
-#plot the ROC curve
-plot(rocCurve.tree,col=c(4))
-```
-
-Alternative ROC Curve
-```{r}
-require(pROC)
-predicted <- predict(crf, train2, OOB=TRUE, type= "response")
-auc(as.numeric(train3$effect_10), as.numeric(predicted))
-#Calculate ROC curve
-rocCurve.tree <- roc(train3$effect_10,as.numeric(predicted))
-#plot the ROC curve
-plot(rocCurve.tree,col=c(4))
-```
-```{r}
-#plot reciever operating curve
-perf <- performance(preds,"tpr","fpr")
-```
-
-```{r}
-# compute in-sample results
-caret::confusionMatrix(fitted,as.factor(train3$effect_10))
-```
-
-```{r}
-#plot feature importance
-cforestImpPlot <- function(x) {
-  cforest_importance <<- v <- varimp(x)
-  dotchart(v[order(v)])
-}
-
-importancePlot <- cforestImpPlot(crf)
-importancePlot
-```
-
-```{r}
-crf.prob <- matrix(unlist(crfsrvy.prob), ncol=2, byrow=TRUE)
-apply(crf.prob,2,mean)
-#[1] 0.2524514 0.7475486
-tab <- round(cbind(by(rpart.prob[,2], INDICES=t1$where, mean), by(crf.prob[,2], INDICES=t1$where, mean)), 4)
-colnames(tab) <- c("rpart", "cforest")
-kable(tab)
-```
-# Random Forest Package
-
-## Preparation
-```{r}
-library(quantregForest)
-library(caret)
-library(tidyverse)
-library(tidymodels)
-library(skimr)
-library(sf)
-library(ggspatial)
-library(nhdplusTools)
-library(patchwork)
-library(Metrics)
-library(gt)
-```
-
-```{r}
-multiVar <- multiVar %>% 
-  mutate_if(is.character, as.factor) %>% 
-  mutate(effect_10 = case_when( #convert ordinal to numeric
-     effect_f == "Yes" ~ 1,
-     effect_f == "No" ~ 0))
-
-#choose relevant predictors and log-transform
-multiVar_small_log <- multiVar %>% 
-  dplyr::select(ID, size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  drop_na() %>%  #drop missing
- mutate(effect_10 = case_when( #convert ordinal to numeric
-     effect_f == "Yes" ~ 1,
-     effect_f == "No" ~ 0
-   ))# %>% 
- #  select(-effect_f)
-
-#no log transform for comparison
-multiVar_small <- multiVar %>% 
-  dplyr::select(ID, size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, #effect_10 , 
-                exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>% 
-  drop_na() #%>%  #drop missing
-# mutate(effect_10 = case_when( #convert ordinal to numeric
+                                 #  max.size.ingest.mm) %>%  #max ingestible size
+  filter(!size_f == "Not Reported")  #take out not reported
+
+#recode variables
+# multiVar <- multiVar %>% mutate(effect_10 = case_when(
 #     effect == "Y" ~ 1,
 #     effect == "N" ~ 0
-#   )) %>% 
-  # select(-effect)
-
-#ensure completeness
-skim(multiVar_small)
-```
-## Training Data
-
+#   ))# %>% 
+#   #mutate_all(is.character, ~as.factor())
+
+```
+
+# Support Vector Machine
+## Kitchen Sink
+### Visualization
+
+#### Support Vector Machine - Non-linear
+
+Support Vector Classifiers are a subset of the group of classification structures known as Support Vector Machines.
+Support Vector Machines can construct classification boundaries that are nonlinear in shape.
+The options for classification structures using the svm() command from the e1071 package are linear, polynomial, radial, and sigmoid.
+Toobserve relationships of classifiers, we plot the data in two dimensions.
+
+```{r}
+# Plot data
+multiVar %>% 
+ggplot(aes(x = log(size.length.um.used.for.conversions), y = log(dose.mg.L.master), color = effect_f, shape = effect_f)) + 
+  geom_point(size = 2) +
+  scale_color_manual(values=c("#000000", "#FF0000")) +
+  theme(legend.position = "none")
+```
+Notice that the data is not linearly separable, and furthermore, isn’t all clustered together in a single group.
+There are two sections of class 1 observations with a cluster of class 2 observations in between.
+To demonstrate the power of SVMs, we’ll take 100 random observations from the set and use them to construct our boundary.
+We set kernel = "radial" based on the shape of our data and plot the results.
+
+```{r}
+# set pseudorandom number generator
+set.seed(123)
+# sample training data
+set.seed(4)
+kitchenSink_split <- kitchenSink %>%
+  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
+kitchenSink_train <- training(kitchenSink_split)
+kitchenSink_test <- testing(kitchenSink_split)
+# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
+count_kitchenSink <- paste0('n = ',nrow(kitchenSink))
+skim(kitchenSink)
+
+subTrain <- kitchenSink_test %>% 
+  dplyr::select(c(effect_f, dose.mg.L.master, size.length.um.used.for.conversions)) %>% 
+  mutate_if(~is.numeric(.) && (.) > 0, log10)
+
+#fit model
+svmfit_kitchenSink <- svm(effect_f ~ ., 
+                          data = subTrain, kernel = "radial", gamma = 1, cost = 1)
+summary(svmfit_kitchenSink)
+# plot classifier
+plot(svmfit_kitchenSink, data = subTrain)
+```
+The same procedure can be run using the kernlab package, which has far more kernel options than the corresponding function in e1071.
+In addition to the four choices in e1071, this package allows use of a hyperbolic tangent, Laplacian, Bessel, Spline, String, or ANOVA RBF kernel.
+To fit this data, we set the cost to be the same as it was before, 1.
+```{r}
+# Fit radial-based SVM in kernlab
+kernfit <- ksvm(effect_f ~ ., data = subTrain, type = "C-svc", kernel = 'rbfdot', C = 1, scaled = c())
+# Plot training data
+plot(kernfit, data = subTrain)
+```
+
+### Data Split
+```{r}
+#choose relevant predictors and log-transform
+kitchenSink <- aoc_z %>% 
+  mutate_if(is.character, as.factor) %>% 
+  mutate(doi = as.character(doi)) %>% 
+  dplyr::select(size_f,
+                #doi, #need to split studies
+                size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
+  # mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+ # mutate(effect_10 = case_when( #convert ordinal to numeric
+ #     effect_f == "Y" ~ 1,
+ #     effect_f == "N" ~ 0
+ #   ))
+drop_na() #drop missing
+
+skim(kitchenSink)
+```
+#### Drop NA
 ```{r}
@@
 set.seed(4)
-multiVar_small_split <- multiVar_small %>%
+kitchenSink_split <- kitchenSink %>%
   initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
@@
 # Pulls from training and testing sets created by initial_split()
-multiVar_small_train <- training(multiVar_small_split)
-multiVar_small_test <- testing(multiVar_small_split)
+kitchenSink_train <- training(kitchenSink_split)
+kitchenSink_test <- testing(kitchenSink_split)
 # Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
 
-# Create a separate dataset of available IDs that were not used in the training dataset.
-notTrain <- aoc %>% # all COMIDS from StreamCat data, sampled or not
-  filter(!ID %in% aoc$ID) # Removing sites used to train the model. n = 140,097
-```
-## Kitchen Sink Model
-
-```{r}
-# Step Three - Kitchen Sink model -----------------------------------------
-
-# Create finalized training dataset and include all possible variables. 
-rf_dat <- multiVar_small_train %>%
-  select(-ID)
-
-# Random forest -- 
-# a decision tree model, using predictors to answer dichotomous questions to create nested splits.
-# no pruning happens - rather, multiple trees are built (the forest) and then you are looking for consensus across trees
-# training data goes down the tree and ends up in a terminal node.
-# if testing data goes down the same route, then this upholds our conclusions. Or, if it goes awry, this allows us to look for patterns in how it goes awry.
-
-set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
-myrf <- randomForest(y = rf_dat$effect_f, # dependent variable
-  x = rf_dat %>%
-    select(-effect_f), # selecting all predictor variables
-  importance = T, # how useful is a predictor in predicting values (nothing causal)
-  proximity = T, 
-  ntrees = 100) # 500 trees default.
-
-myrf # examine the results.
-```
-~20% error rate. Let's compare this with the same model with log-transformed values.
-```{r}
+count_kitchenSink <- paste0('n = ',nrow(kitchenSink))
+
+skim(kitchenSink)
+```
+
+#### Rough Fix NA
+```{r}
+#rough fix
+kitchenSink_roughfix <- na.roughfix(kitchenSink)
+
 # Create calibration and validation splits with tidymodels initial_split() function.
 set.seed(4)
-multiVar_small_log_split <- multiVar_small_log %>%
-  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
-
-# Create a training data set with the training() function
-# Pulls from training and testing sets created by initial_split()
-multiVar_small_log_train <- training(multiVar_small_log_split)
-multiVar_small_log_test <- testing(multiVar_small_log_split)
-
-# Create finalized training dataset and include all possible variables. 
-rf_dat_log <- multiVar_small_log_train %>%
-  select(-ID, -effect_10)
-
-set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
-myrf_log <- randomForest(y = rf_dat_log$effect_f, # dependent variable
-  x = rf_dat_log %>%
-    select(-effect_f), # selecting all predictor variables
-  importance = T, # how useful is a predictor in predicting values (nothing causal)
-  proximity = T, 
-  ntrees = 100) # 500 trees default.
-
-myrf_log # examine the results.
-```
-No performance enhancement with log-transformed values.
-
-## Continuous Predictors
-Repeat with continous variables whenever possible, and max of 8 predictors.
-```{r}
-skim(multiVar)
-```
-```{r}
-acute <- multiVar %>% 
-  filter(acute.chronic_f == "Acute") %>% 
-  filter(!environment == "Terrestrial") %>% 
-  filter(bio.org == "organism") %>% 
-  filter(lvl1_f == "Fitness") %>% 
-  filter(!exposure.route == "food") %>% 
-  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  drop_na()
-skim(acute)
-```
-With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
-Let's determine if we have enough data for ML.
-```{r}
-exp(1)^6
-```
-$e^6$ is less than n (453), so we may proceed.
-
-```{r}
-# Create calibration and validation splits with tidymodels initial_split() function.
-set.seed(4)
-acute_split <- acute %>%
+kitchenSink_split_roughfix <- kitchenSink_roughfix %>%
   initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
@@
 # Pulls from training and testing sets created by initial_split()
-acute_train <- training(acute_split)
-acute_test <- testing(acute_split)
+kitchenSink_train_roughfix <- training(kitchenSink_split_roughfix)
+kitchenSink_test_roughfix <- testing(kitchenSink_split_roughfix)
 # Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
 
-# Random forest -- 
-set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
-acuterf <- randomForest(y = acute_train$effect_f, # dependent variable
-  x = acute_train %>%
-    dplyr::select(-effect_f), # selecting all predictor variables
-  importance = T, # how useful is a predictor in predicting values (nothing causal)
-  proximity = T, 
-  ntrees = 100) # 500 trees default.
-
-acuterf # examine the results.
-```
-
-```{r}
-plot(acuterf)
-# model performance appears to improve most at ~75 trees
-```
-
-```{r}
-varImpPlot(acuterf)
-# displays which variables are most important
-# helps to winnow down list of predictors
-# recommended to weigh left pane more
-# right pane also shows how evenly things split based on the list of predictors
-# values close to 0 can be dropped, but don't have to be
-```
-Dose, organism group, exposure duration are important.
-
-```{r}
-importance <- acuterf$importance
-View(importance)
-# displays the data plotted in the plot above
-```
-
-###Validation
-
-```{r fit results}
-fitted <- predict(acuterf, acute_test[,-6], OOB = TRUE, type ="response")
-misClasificError <- mean(fitted != acute_test$effect_f)
-print(paste('Training Accuracy', 1 - misClasificError))
-```
-
-Alternative ROC Curve
-```{r}
-require(pROC)
-predicted <- predict(acuterf, acute_test %>%  dplyr::select(-effect_f),
-                       OOB=TRUE, type= "response")
-#Calculate ROC curve
-rocCurve.tree <- roc(as.numeric(acute_test$effect_f),as.numeric(predicted))
-
-##gplot
-# rocks <- roc()
-
-#plot the ROC curve
-plot(rocCurve.tree,col=c(4))
-```
-##### Chronic
-```{r}
-chronic <- multiVar %>% 
-  filter(acute.chronic_f == "Chronic") %>% 
-  filter(!environment == "Terrestrial") %>% 
-  filter(bio.org == "organism") %>% 
-  filter(lvl1_f == "Fitness") %>% 
-  filter(!exposure.route == "food") %>% 
-  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  drop_na()
-skim(chronic)
-```
-With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
-Let's determine if we have enough data for ML.
-```{r}
-exp(1)^6
-```
-$e^6$ is less than n (453), so we may proceed.
-
-```{r}
+count_roughfix <- paste0('n = ',nrow(kitchenSink_roughfix))
+
+skim(kitchenSink_roughfix)
+```
+
+#### Multiple Imputation NA
+
+The algorithm starts by imputing NAs using na.roughfix.
+Then randomForest is called with the completed data.
+The proximity matrix from the randomForest is used to update the imputation of the NAs.
+For continuous predictors, the imputed value is the weighted average of the non-missing obervations, where the weights are the proximities.
+For categorical predictors, the imputed value is the category with the largest average proximity.
+This process is iterated iter times.
+
+Note: Imputation has not (yet) been implemented for the unsupervised case.
+Also, Breiman (2003) notes that the OOB estimate of error from randomForest tend to be optimistic when run on the data matrix with imputed values.
+```{r}
+# impute values
+#drop NA's in response
+kitchenSink_noNa <- kitchenSink %>% drop_na(effect_f)
+
+#impute
+set.seed(111)
+kitchenSink_rfImpute <- rfImpute(data = kitchenSink_noNa, 
+                                    effect_f ~., #response value, cannot contain NA's
+                                    iter = 4,
+                                    ntree =75)
+
 # Create calibration and validation splits with tidymodels initial_split() function.
 set.seed(4)
-chronic_split <- chronic %>%
+kitchenSink_split_imputed <- kitchenSink_rfImpute %>%
   initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
@@
 # Pulls from training and testing sets created by initial_split()
-chronic_train <- training(chronic_split)
-chronic_test <- testing(chronic_split)
+kitchenSink_train_imputed <- training(kitchenSink_split_imputed)
+kitchenSink_test_imputed <- testing(kitchenSink_split_imputed)
 # Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
 
-# Random forest -- 
-set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
-chronicrf <- randomForest(y = chronic_train$effect_f, # dependent variable
-  x = chronic_train %>%
-    dplyr::select(-effect_f), # selecting all predictor variables
-  importance = T, # how useful is a predictor in predicting values (nothing causal)
-  proximity = T, 
-  ntrees = 100) # 500 trees default.
-
-chronicrf # examine the results.
-```
-
-```{r}
-plot(chronicrf)
-# model performance appears to improve most at ~75 trees
-```
-
-```{r}
-varImpPlot(chronicrf)
-# displays which variables are most important
-# helps to winnow down list of predictors
-# recommended to weigh left pane more
-# right pane also shows how evenly things split based on the list of predictors
-# values close to 0 can be dropped, but don't have to be
-```
-Dose, organism group, exposure duration are important.
-
-```{r}
-importance <- chronicrf$importance
-View(importance)
-# displays the data plotted in the plot above
-```
-
-```{r fit results}
-fitted <- predict(chronicrf, chronic_test[,-6], OOB = TRUE, type ="response")
-misClasificError <- mean(fitted != chronic_test$effect_f)
-print(paste('Training Accuracy', 1 - misClasificError))
-```
-
-#### All Data
-```{r}
-all <- multiVar %>% 
-  #filter(acute.chronic_f == "Chronic") %>% 
-  filter(!environment == "Terrestrial") %>% 
-  filter(bio.org == "organism") %>% 
-  filter(lvl1_f == "Fitness") %>% 
-  filter(!exposure.route == "food") %>% 
-  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  drop_na()
-skim(all)
-```
-With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
-Let's determine if we have enough data for ML.
-```{r}
-exp(1)^6
-```
-$e^6$ is less than n (453), so we may proceed.
-
-```{r}
-# Create calibration and validation splits with tidymodels initial_split() function.
-set.seed(4)
-all_split <- all %>%
-  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
-# default is 3/4ths split (but 75% training, 25% testing).
-# Stratification (strata) = grouping training/testing sets by region, state, etc.
-# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
-
-# Create a training data set with the training() function
-# Pulls from training and testing sets created by initial_split()
-all_train <- training(all_split)
-all_test <- testing(all_split)
-# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
-
-# Random forest -- 
-set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
-allrf <- randomForest(y = all_train$effect_f, # dependent variable
-  x = all_train %>%
-    dplyr::select(-effect_f), # selecting all predictor variables
-  importance = T, # how useful is a predictor in predicting values (nothing causal)
-  proximity = T, 
-  ntrees = 100) # 500 trees default.
-
-allrf # examine the results.
-```
-
-```{r}
-plot(allrf)
-# model performance appears to improve most at ~75 trees
-```
-
-```{r}
-varImpPlot(allrf)
-# displays which variables are most important
-# helps to winnow down list of predictors
-# recommended to weigh left pane more
-# right pane also shows how evenly things split based on the list of predictors
-# values close to 0 can be dropped, but don't have to be
-```
-Dose, organism group, exposure duration are important.
-
-```{r}
-importance <- allrf$importance
-View(importance)
-# displays the data plotted in the plot above
-```
-
-```{r fit results}
-fitted <- predict(allrf, all_test[,-6], OOB = TRUE, type ="response")
-misClasificError <- mean(fitted != all_test$effect_f)
-print(paste('Training Accuracy', 1 - misClasificError))
-```
-
-#### NO FILTERS
-```{r}
-nofilter <- multiVar %>% 
-  #filter(acute.chronic_f == "Chronic") %>% 
-  ##filter(!environment == "Terrestrial") %>% 
-  #filter(bio.org == "organism") %>% 
-  #filter(lvl1_f == "Fitness") %>% 
-  #filter(!exposure.route == "food") %>% 
-  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
-  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
-  drop_na()
-skim(nofilter)
-```
-With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID.
-Let's determine if we have enough data for ML.
-```{r}
-exp(1)^6
-```
-$e^6$ is less than n (453), so we may proceed.
-
-```{r}
-# Create calibration and validation splits with tidymodels initial_split() function.
-set.seed(4)
-nofilter_split <- nofilter %>%
-  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
-# default is 3/4ths split (but 75% training, 25% testing).
-# Stratification (strata) = grouping training/testing sets by region, state, etc.
-# Using the "strata" cnofilter ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
-
-# Create a training data set with the training() function
-# Pulls from training and testing sets created by initial_split()
-nofilter_train <- training(nofilter_split)
-nofilter_test <- testing(nofilter_split)
-# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
-
-# Random forest -- 
-set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
-nofilterrf <- randomForest(y = nofilter_train$effect_f, # dependent variable
-  x = nofilter_train %>%
-    dplyr::select(-effect_f), # selecting nofilter predictor variables
-  importance = T, # how useful is a predictor in predicting values (nothing causal)
-  proximity = T, 
-  ntrees = 100) # 500 trees default.
-
-nofilterrf # examine the results.
-```
-
-```{r}
-plot(nofilterrf)
-# model performance appears to improve most at ~75 trees
-```
-
-```{r}
-varImpPlot(nofilterrf)
-# displays which variables are most important
-# helps to winnow down list of predictors
-# recommended to weigh left pane more
-# right pane also shows how evenly things split based on the list of predictors
-# values close to 0 can be dropped, but don't have to be
-```
-Dose, organism group, exposure duration are important.
-
-```{r}
-importance <- nofilterrf$importance
-View(importance)
-# displays the data plotted in the plot above
-```
-
-```{r fit results}
-fitted <- predict(nofilterrf, nofilter_test[,-6], OOB = TRUE, type ="response")
-misClasificError <- mean(fitted != nofilter_test$effect_f)
-print(paste('Training Accuracy', 1 - misClasificError))
-```
-
-#### ROC Curve
-yet anothyer try
-```{r}
+count_imputed <- paste0('n = ',nrow(kitchenSink_rfImpute))
+
+skim(kitchenSink_rfImpute)
+```
+
+### Train
+#### Drop Na
+```{r}
+#examine train and test sets
+dim(kitchenSink_train); dim(kitchenSink_test)
+```
+
+Before we train our model, we’ll first implement the trainControl() method.
+This will control all the computational overheads so that we can use the train() function provided by the caret package.
+The training method will train our data on different algorithms.
+
+First, let’s focus on the traincontrol() method:
+
+```{r}
+trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
+# The trainControl() method here, is taking 3 parameters.
+# The “method” parameter defines the resampling method, in this demo we’ll be using the repeatedcv or the repeated cross-validation method.
+# The next parameter is the “number”, this basically holds the number of resampling iterations.
+# The “repeats ” parameter contains the sets to compute for our repeated cross-validation. We are using setting number =10 and repeats =3
+# This trainControl() method returns a list. We are going to pass this on our train() method.
+
+svm_Linear <- train(effect_f ~., 
+                    data = kitchenSink_train, 
+                    method = "bayesglm", #Bayesian Model
+                    trControl = trctrl, 
+                    preProcess = c("center", "scale"), 
+                    tuneLength = 10)
+#The “preProcess” parameter is for preprocessing our training data.
+# We are passing 2 values in our “pre-process” parameter “center” & “scale”. These two help for centering and scaling the data.
+# After pre-processing, these convert our training data with mean value as approximately “0” and standard deviation as “1”.
+The “tuneLength” parameter holds an integer value.
+This is for tuning our algorithm.
+
+svm_Linear
+```
+It’s a linear model therefore, it just tested at value “C” = 1.
+
+Now, our model is trained with C value as 1. We are ready to predict classes for our test set. We can use predict() method.
+
+The caret package provides predict() method for predicting results.
+We are passing 2 arguments.
+Its first parameter is our trained model and second parameter “newdata” holds our testing data frame.
+The predict() method returns a list, we are saving it in a test_pred variable.
+
+```{r}
+test_pred <- predict(svm_Linear, newdata = kitchenSink_test)
+confusionMatrix(table(test_pred, kitchenSink_test$effect_f))
+```
+
+The output shows that our model accuracy for test set is %
+
+By following the above procedure, we can build our svmLinear classifier.
+
+We can also do some customization for selecting C value(Cost) in Linear classifier. This can be done by inputting values in grid search.
+
+The next code snippet will show you, building & tuning of an SVM classifier with different values of C.
+
+We are going to put some values of C using expand.grid() into “grid” dataframe.
+Next step is to use this dataframe for testing our classifier at specific C values.
+It needs to be put in train() method with tuneGrid parameter.
+
+```{r}
+grid <- expand.grid(C = c(0,0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2,5))
+svm_Linear_Grid <- train(effect_f ~., data = kitchenSink_train, method = "svmLinear",
+                         trControl = trctrl,
+                         preProcess = c("center", "scale"),
+                         tuneGrid = grid,
+                         tuneLength = 10)
+svm_Linear_Grid
+plot(svm_Linear_Grid)
+```
+The above plot is showing that our classifier is giving best accuracy on C = 1.75. Let’s try to make predictions using this model for our test set.
+
+```{r}
+test_pred_grid <- predict(svm_Linear_Grid, newdata = kitchenSink_train)
+confusionMatrix(table(test_pred_grid, kitchenSink_train$effect_f))
+```
+
+# ROC Curves
+
+Prep for plotting.
+```{r}
+
+#### kitchenSink
+kitchenSinkpredictions <- as.data.frame(predict(kitchenSinkrf, kitchenSink_test%>% dplyr::select(-effect_f), type = "prob"))
+# predict class and then attach test class
+kitchenSinkpredictions$predict <- names(kitchenSinkpredictions)[1:2][apply(kitchenSinkpredictions[,1:2], 1, which.max)]
+kitchenSinkpredictions$observed <- kitchenSink_test$effect_f
+head(kitchenSinkpredictions)
+
+#### kitchenSink_roughfix_roughfix
+kitchenSink_roughfixpredictions <- as.data.frame(predict(kitchenSink_roughfixrf, kitchenSink_roughfix_test%>% dplyr::select(-effect_f), type = "prob"))
+# predict class and then attach test class
+kitchenSink_roughfixpredictions$predict <- names(kitchenSink_roughfixpredictions)[1:2][apply(kitchenSink_roughfixpredictions[,1:2], 1, which.max)]
+kitchenSinkpredictions$observed <- kitchenSink_test$effect_f
+head(kitchenSinkpredictions)
+
 ###CHRONIC
-chronicpredictions <- as.data.frame(predict(chronicrf, chronic_test[,-6], type = "prob"))
+chronicpredictions <- as.data.frame(predict(chronicrf, chronic_test %>% dplyr::select(-effect_f), type = "prob"))
 # predict class and then attach test class
@@
 ####ACUTE
-predictions <- as.data.frame(predict(acuterf, acute_test[,-6], type = "prob"))
+predictions <- as.data.frame(predict(acuterf, acute_test%>% dplyr::select(-effect_f), type = "prob"))
 # predict class and then attach test class
@@
 
-#### all
-allpredictions <- as.data.frame(predict(allrf, all_test[,-6], type = "prob"))
-# predict class and then attach test class
-allpredictions$predict <- names(allpredictions)[1:2][apply(allpredictions[,1:2], 1, which.max)]
-allpredictions$observed <- all_test$effect_f
-head(allpredictions)
-
 ###nofilter
-nofilterpredictions <- as.data.frame(predict(nofilterrf, nofilter_test[,-6], type = "prob"))
+nofilterpredictions <- as.data.frame(predict(nofilterrf, nofilter_test%>% dplyr::select(-effect_f), type = "prob"))
 # predict class and then attach test class
@@
 nofilterpredictions$observed <- nofilter_test$effect_f
-```
-```{r}
+
+###nofilterOptimized
+nofilter.optimizedpredictions <- as.data.frame(predict(myrf_optimized, multiVar_smkitchenSink_test %>%  dplyr::select(-effect_f), type = "prob"))
+# predict class and then attach test class
+nofilter.optimizedpredictions$predict <- names(nofilter.optimizedpredictions)[1:2][apply(nofilter.optimizedpredictions[,1:2], 1, which.max)]
+nofilter.optimizedpredictions$observed <- multiVar_smkitchenSink_test$effect_f
+
+###nofilterOptimizedImputed
+nofilter.optimized.imputedpredictions <- as.data.frame(predict(myrf_optimized_imputed, multiVar_smkitchenSink_test_imputed %>%  dplyr::select(-effect_f), type = "prob"))
+# predict class and then attach test class
+nofilter.optimized.imputedpredictions$predict <- names(nofilter.optimized.imputedpredictions)[1:2][apply(nofilter.optimized.imputedpredictions[,1:2], 1, which.max)]
+nofilter.optimized.imputedpredictions$observed <- multiVar_smkitchenSink_test_imputed$effect_f
+```
+Plot.
+```{r}
+require(ggdark)
 # 1 ROC curve, yes vs no for acute
-roc.acute <- roc(ifelse(predictions$observed=="Yes", "Yes", "No"), as.numeric(predictions$Yes))
+roc.acute <- roc(ifelse(predictions$observed=="Y", "Y", "N"), as.numeric(predictions$Y))
 
 #chronic
-roc.chronic <- roc(ifelse(chronicpredictions$observed=="Yes", "Yes", "No"), as.numeric(chronicpredictions$Yes))
-#all
-roc.all <- roc(ifelse(allpredictions$observed=="Yes", "Yes", "No"), as.numeric(allpredictions$Yes))
+roc.chronic <- roc(ifelse(chronicpredictions$observed=="Y", "Y", "N"), as.numeric(chronicpredictions$Y))
+#kitchenSink
+roc.kitchenSink <- roc(ifelse(kitchenSinkpredictions$observed=="Y", "Y", "N"), as.numeric(kitchenSinkpredictions$Y))
 
 #nofilter
-roc.nofilter <- roc(ifelse(nofilterpredictions$observed=="Yes", "Yes", "No"), as.numeric(nofilterpredictions$Yes))
-
-all <- ggroc(roc.all) + labs(title = "Chronic and Acute",
-                             subtitle = "Accuracy = 81.7%, n = 1510") +
-  theme_minimal()
-acute <- ggroc(roc.acute, col = "red") + labs(title = "Acute",
-                                              subtitle = "Accuracy = 84.8%, n = 883") + theme_minimal()
-chronic <- ggroc(roc.chronic, col = "blue") + labs(title = "Chronic",
-                                                   subtitle = "Accuracy = 67.3%, n = 452") + theme_minimal()
-
-nofilter<- ggroc(roc.nofilter, col = "green") + labs(title = "Entire Dataset",
-                                                   subtitle = "Accuracy = 79.9%, n = 4615") + theme_minimal()
+roc.nofilter <- roc(ifelse(nofilterpredictions$observed=="Y", "Y", "N"), as.numeric(nofilterpredictions$Y))
+
+#no filter (optimized)
+roc.nofilter.optimized <- roc(ifelse(nofilter.optimizedpredictions$observed=="Y", "Y", "N"), as.numeric(nofilter.optimizedpredictions$Y))
+
+#no filter (optimized; imputed)
+roc.nofilter.optimized.imputed <- roc(ifelse(nofilter.optimized.imputedpredictions$observed=="Y", "Y", "N"), as.numeric(nofilter.optimized.imputedpredictions$Y))
+
+##make ROC curves
+
+#kitchenSink
+kitchenSinkROC <- ggroc(roc.kitchenSink, col = "yellow") + 
+  labs(title = "Chronic and Acute",
+       subtitle = paste0(accuracy_kitchenSink,', ',count_kitchenSink)) + 
+  dark_theme_bw()
+
+#acute
+acuteROC <- ggroc(roc.acute, col = "green") + 
+  labs(title = "Acute",
+       subtitle = paste0(accuracy_acute,', ',count_acute)) + 
+  dark_theme_bw()
+
+#chronic
+chronicROC <- ggroc(roc.chronic, col = "blue") + 
+  labs(title = "Chronic",
+       subtitle = paste0(accuracy_chronic,', ',count_chronic)) + #auto label
+  dark_theme_bw()
+
+#no filter
+nofilterROC <- ggroc(roc.nofilter, col = "red3") + 
+  labs(title = "Entire Dataset",
+       subtitle = paste0(accuracy_nofilter,', ',count_nofilter)) + 
+  dark_theme_bw()
+
+#optimized (rough fix)
+nofilteroptimizedROC <- ggroc(roc.nofilter.optimized, col = "orangered") + 
+  labs(title = "Entire Dataset (optimized)",
+       subtitle = paste0(accuracy_optimized,', ',count_optimized)) + 
+  dark_theme_bw()
+
+#optimized (multiple imputation)
+nofilteroptimizedimputedROC <- ggroc(roc.nofilter.optimized.imputed, col = "orange") + 
+  labs(title = "Entire Dataset (optimized; imputed)",
+       subtitle = paste0(accuracy_optimized_imputed,', ',count_optimized_imputed)) + 
+  dark_theme_bw()
+
+#arrange together and print
 require(gridExtra)
-grid.arrange(all, acute, chronic, nofilter,
-             ncol = 4)
-```
-
-```{r}
-#plot reciever operating curve
-perf <- performance(preds,"tpr","fpr")
-```
-
-```{r}
-# compute in-sample results
-caret::confusionMatrix(fitted,as.factor(acute_test$effect_f))
-```
-
-```{r}
-# Validation set assessment #1: looking at confusion matrix
-prediction_for_table <- predict(acuterf, acute_test[,-6])
-table(observed = acute_test[,6], predicted = prediction_for_table)
-```
-
-```{r}
-# Validation set assessment #2: ROC curves and AUC
-# Needs to import ROCR package for ROC curve plotting:
-library(ROCR)
-# Calculate the probability of new observations belonging to each class
-# prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
-prediction_for_roc_curve <- predict(acuterf ,acute_test[,-6],type="response")
-# Use pretty colours:
-pretty_colours <- c("#F8766D","#00BA38","#619CFF")
-# Specify the different classes 
-classes <- levels(acute_test$organism.group)
-# For each class
-for (i in 1:13)
-{
- # Define which observations belong to class[i]
- true_values <- ifelse(acute_test[,6]==classes[i],"Yes","No")
- # Assess the performance of classifier for class[i]
- pred <- prediction(prediction_for_roc_curve[,i],true_values)
- perf <- performance(pred, "tpr", "fpr")
- if (i==1)
- {
-     plot(perf,main="ROC Curve",col=pretty_colours[i]) 
- }
- else
- {
-     plot(perf,main="ROC Curve",col=pretty_colours[i],add=TRUE) 
- }
- # Calculate the AUC and print it to screen
- auc.perf <- performance(pred, measure = "auc")
- print(auc.perf@y.values)
-}
-```
-
-```{r}
-# predict()
-# returns out of bag predictions for training data
-# in the bag: every time a tree is built, it uses ~80% of the original 75% we set aside from the original dataset used to create a tree to assure random data selection
-# out of bag: looking at the remaining 20% of the training data to predict, when you want to know what your model does at the training location sites
-
-# # Predict CRAM scores state-wide for all COMIDs.
-# notTrain_prediction <- notTrain %>% # taking all available IDs, that haven't been used in training
-#   na.omit() %>% # remove NAs
-#   mutate(effect_predicted = predict(myrf, newdata = notTrain %>% na.omit())) # using developed model (myrf), inputting predictor variables (notTrain - which contains IDs and associated  data) to predict output/dependent variable (effect_predicted a.k.a.
-effect).
-
-# rePredict CRAM scores for training data.
-multiVar_small_train$effect_predicted <- predict(myrf) # Add column of predicted CRAM values to training dataset.
-
-# Creates new dataset of bound rows for both ...
-# effect_predictions <- bind_rows(nottrain_prediction %>%
-#                             mutate(Set = "Non-training"), # statewide COMIDs (not used for training)
-#                             multiVar_smal_train %>%
-#                             mutate(Set = "Training")) # COMIDS from training dataset
-# This creates the dataset that will be plotted to create a state-wide plot of predicted CRAM scores.
-
-# Plot the data.
-multiVar_small_train %>% 
-  mutate(effect_10 = case_when( #convert ordinal to numeric
-     effect_predicted == "Yes" ~ 1,
-     effect_predicted == "No" ~ 0
-   )) %>% 
-  ggplot(aes(x = log10(dose.mg.L.master), y = effect_10)) +
-  geom_point(alpha = 0.7) +#color = "black") +
-  labs(x = "Density",
-    y = "Predicted CRAM Score") +
-  theme_dark() #+
-  #facet_wrap(.~polymer)
-```
-### Predictor Selector
-```{r}
-# Using caret to select the best predictors
-# What are the parameters you want to use to run recursive feature elimination (rfe)?
-my_ctrl <- rfeControl(functions = rfFuncs,
-                      method = "cv",
-                      verbose = FALSE,
-                      returnResamp = "all")
-
-# rfe = recursive feature elimination
-# THIS STEP TAKES FOR-EV-ER!!!
-set.seed(22)
-my_rfe <- rfe(y = rf_dat$effect_f, # set dependent variable
-              x = rf_dat %>% select(-effect_f), # set predictor variables
-              size = c(1:2, 4, 6, 8, 10, 13), # sets how many variables are in the overall model
-              # I have 13 total possible variables, so I've chosen increments of 3 to look at.
-              rfeControl = my_ctrl) # pull in control from above
-
-# can you make your model even simpler?
-# the following will pick a model with the smallest number of predictor variables based on the tolerance ("tol") that you specify (how much less than the best are you willing to tolerate?)
-my_size <- pickSizeTolerance(my_rfe$results, metric = "Accuracy", tol = 1, maximize = F)
-# higher tol (~10) gives you less variables
-# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
-pickVars(my_rfe$variables, size = my_size)
-```
-
-```{r}
-# Predict scores using the above 2 variables:
-
-# Create re-finalized training dataset and include all possible variables. 
-rf_dat2 <- multiVar_small_train %>%  select(dose.mg.L.master, organism.group, effect_f)
-
-set.seed(4) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
-myrf2 <- randomForest(y = rf_dat2$effect_f, # dependent variable
-  x = rf_dat2 %>%
-    select(-effect_f),
-  importance = T, 
-  proximity = T, 
-  ntrees = 500)
-
-myrf2 # examine the results. 
-```
-
-```{r}
-importance <- as.data.frame(as.table(myrf$importance))
-
-# Nicer ggplot variable importance plot.
-vip_plot_a <- importance %>%
- filter(Var2 == "MeanDecreaseAccuracy") %>%
-  mutate(Var1 = factor(Var1)) %>%
-  mutate(Var1_f = fct_reorder(Var1, Freq)) %>%
-  ggplot(aes(x = Freq, y = Var1_f)) +
-  geom_point(size = 3, alpha = 0.75, color = "black") +
-  labs(x = "Importance",
-    y = "Variables") +
-  theme_bw()
-
-# vip_plot_b <- importance %>%
-#   filter(Var2 == "MeanDecreaseGini") %>%
-#   mutate(Var1 = factor(Var1)) %>%
-#   mutate(Var1_f = fct_reorder(Var1, Freq)) %>%
-#   ggplot(aes(x = Freq, y = Var1_f)) +
-#   geom_point(size = 3, alpha = 0.75, color = "black") +
-#   labs(x = "Node Purity",
-#     y = "Variables") +
-#   theme_bw()
-
-vip_plot <- vip_plot_a #+ vip_plot_b
-
-vip_plot
-# ggsave("cram_vip_plot.png",
-#      path = "/Users/heilil/Desktop/R_figures",
-#      width = 25,
-#      height = 10,
-#      units = "cm"
-#    )
-```
-GINI importance measures the average gain of purity by splits of a given variable.
-If the variable is useful, it tends to split mixed labeled nodes into pure single class nodes.
-Splitting by a permuted variables tend neither to increase nor decrease node purities.
-Permuting a useful variable, tend to give relatively large decrease in mean gini-gain.
-GINI importance is closely related to the local decision function, that random forest uses to select the best available split.
-Therefore, it does not take much extra time to compute.
-On the other hand, mean gini-gain in local splits, is not necessarily what is most useful to measure, in contrary to change of overall model performance.
-Gini importance is overall inferior to (permutation based) variable importance as it is relatively more biased, more unstable and tend to answer a more indirect question.
-
-### Quantile Regression model
-```{r}
-# Quantile random forest regression mode, instead of looking at the mode of trees, can compare to 10th, 50th, 90th percentiles etc.
-
-# Need to make a new dataset taking the above results of pickVars into account.
-# Create finalized training dataset and include all possible variables. 
-qrf_dat <- multiVar_small_train  %>%
-  select(-effect_predicted, -ID)
-#   select(asci, RdCrsWs, PctAgWs, PctUrbWsRp100, PctOpWsRp100, PctOpWs, DamDensWs, RdDensWs, NABD_DensWs, PctUrbWs, PctUrbCatRp100, RdDensWsRp100, PctOpCat, PctUrbCat, RdDensCat, CBNFWs, PctOpCatRp100, PctAgWsRp100, TRIDensWs, AgKffactWs, FertWs)
-
- set.seed(20)
- myqrf <- quantregForest(y = qrf_dat$effect_f, # dependent variable
-              x = qrf_dat %>%
-                  select(-effect_f),
-              importance = T,
-              proximity = T,
-              keep.inbag=T,
-              ntrees = 500)
-
-#predict(myqrf) # automatically presents 10th %tile, median, and 90th %tile
-```
-
-```{r}
-#predict(myqrf, what=c(0.2, 0.3, 0.999)) # to print specific quantiles
-```
-
-```{r}
-plot(myqrf) # plots the results.
-```
-# Again appears to improve after ~100 trees.
-
-### Model Validation
-#### Categorical Response
-
-#### Continuous Response
-```{r}
-# Compare predicted vs. actual results, including by size.
-# Adding lines of slope=1 and linear models to each plot.
-val1 <- multiVar_small_train %>% 
-  ggplot(aes(x = effect_predicted, y = effect_f)) +
-  geom_point(color = "#2A3927", alpha = 0.5) +
-  geom_smooth(method = "lm", se = FALSE, color = "#2A3927") +
-  labs(x = "Predicted Effect",
-    y = "Effect measured",
-    title = "Training Data\nn=613") +
-  geom_abline(intercept = 0, slope = 1) +
-  theme_bw()
-
-val1
-
-lm1 <- lm(cram~cram_predicted, data = mydf2_train2)
-summary(lm1)
-
-val2 <- ggplot(mydf2_test2, aes(x = cram_predicted, y = cram)) +
-  geom_point(color = "#3793EC", alpha = 0.5) +
-  geom_smooth(method = "lm", se = FALSE, color = "#3793EC") +
-  scale_x_continuous(breaks = c(0.5, 0.7, 0.9)) +
-  labs(x = "CRAM predicted",
-    y = "CRAM measured",
-    title = "Testing Data\nn=202") +
-  geom_abline(intercept = 0, slope = 1) +
-  theme_bw()
-
-val2
-
-lm2 <- lm(cram~cram_predicted, data = mydf2_test2)
-summary(lm2)
-
-# Create the full testing + training dataset to plot together.
-mydf2_test2$set <- "Testing"
-mydf2_train2$set <- "Training"
-full_train_test <- bind_rows(mydf2_test2, mydf2_train2) %>%
-  mutate(set_f = factor(set, levels = c("Training", "Testing")))
-
-val3 <- ggplot(full_train_test, aes(x = cram_predicted, y = cram, color = set_f)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = FALSE) +
-  scale_color_manual(name = "Set", values = c("#2A3927", "#3793EC"), drop = FALSE) +
-  labs(x = "CRAM predicted",
-    y = "CRAM measured",
-    title = "All Data\nn=815") +
-  geom_abline(intercept = 0, slope = 1, color = "black") +
-  facet_wrap(~PSA6) +
-  theme_bw()
-
-val3
-
-val_fig <- (val1 + val2) /
-  (val3)
-
-val_fig + plot_annotation(
-  title = 'CRAM Random Forest Results',
-  subtitle = 'All modeling performed using StreamCAT datasets.',
-  caption = 'Linear models are colored according to dataset. Lines of slope = 1 are denoted in black.'
-)
-
-# Save figure.
-# ggsave("cram_rfmodel_validation.png",
-#      path = "/Users/heilil/Desktop/R_figures",
-#      width = 35,
-#      height = 25,
-#      units = "cm"
-#    )
-
-lm3 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Central_Valley") %>%
-    filter(set_f == "Training"))
-
-lm4 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Central_Valley") %>%
-    filter(set_f == "Testing"))
-
-lm5 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Chaparral") %>%
-    filter(set_f == "Training"))
-
-lm6 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Chaparral") %>%
-    filter(set_f == "Testing"))
-
-lm7 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Deserts_Modoc") %>%
-    filter(set_f == "Training"))
-
-lm8 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Deserts_Modoc") %>%
-    filter(set_f == "Testing"))
-
-lm9 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "North_Coast") %>%
-    filter(set_f == "Training"))
-
-lm10 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "North_Coast") %>%
-    filter(set_f == "Testing"))
-
-lm11 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Sierra") %>%
-    filter(set_f == "Training"))
-
-lm12 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "Sierra") %>%
-    filter(set_f == "Testing"))
-
-lm13 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "South_Coast") %>%
-    filter(set_f == "Training"))
-
-lm14 <- lm(cram~cram_predicted, 
-  data = full_train_test %>%
-    filter(PSA6 == "South_Coast") %>%
-    filter(set_f == "Testing"))
-
-# Chose not to compute confusion matrix / accuracy score since this is more applicable to categorical ouputs from random forest models -
-# Instead, calculated Root Mean Squared Error (RMSE) of both training and test datasets.
-# If test RMSE values are much greater than training, then possible the model has been over fit.
-
-predtest <- predict(myrf2, mydf2_test2)
-rmse(mydf2_test2$cram,predtest)
-# 9.03
-
-predtrain <- predict(myrf2, mydf2_train2)
-rmse(mydf2_train2$cram,predtrain)
-# 4.20
-
-# Double checking using the original random forest dataset (rf_dat) with all 35 possible variables included to see where the error in number of predictors starts to increase dramatically (to help double check our decision to include 25 parameters).
-dc <- rfcv(rf_dat %>%
-    select(-cram), 
-  rf_dat$cram,
-  step = 0.7, # default is 0.5
-  scale="log")
-
-dc$error.cv
-#34        24        17        12         8         6         4         3         2 
-# 96.28062  97.29366  99.08743 102.93752 115.52580 120.63049 122.09563 123.74425 139.30181 
-# 1 
-# 206.58921
-
-# Appears between 34 and 24 variables, there is an insignificant increase in error.
-# However, this model is much larger than the CSCI (20) and ASCI (10) models, so we may decide to trim this down in the future.
-
-```
-
-# Works Cited
-
-# Non-linear Dose Response
-If we hold all paramaters constant except for dose, we should observe a non-linear dose-response behavior that can be best predicted by a 3- or 4-parameter Hill Equation.
-The most available information is likely for polystyrene spheres in crustaceans.There are 21 studies available.
+grid.arrange(nofilterROC, nofilteroptimizedROC, nofilteroptimizedimputedROC,kitchenSinkROC, acuteROC, chronicROC,
+             ncol = 3)
+```
+ROC curves may also be visualized together
+```{r}
+require(pROC)
+require(tidyverse)
+require(ggdark)
+require(ggsci)
+ggroc(list(kitchenSink = roc.nofilter, optimized = roc.nofilter.optimized, optimized.imputed = roc.nofilter.optimized.imputed, organisms = roc.kitchenSink, acute = roc.acute, chronic = roc.chronic), aes = c("linetype", "color")) +
+  labs(title = "ROC Curves for Aquatic Toxicity Random Forest",
+       subtitle = "n = 4615",
+       color = "Dataset",
+       linetype = "Dataset") +
+   scale_color_tron() +
+  # theme_bw(base_size = 20)
+ dark_theme_bw(base_size = 20)# +
+ theme(plot.title.position = element_text(hjust = 0.5),
+     plot.subtitle.position = element_text(hjust = 0.5))
+```

### 2021-02-23 Stepwise Multiple Regressions.Rmd

@@
   html_document: 
-    toc: true
+    code_folding: hide
+    theme: sandstone
+    toc: yes
+    toc_float: yes
     toc_depth: 4
     number_sections: true
-    toc_float: true
+  word_document:
+    toc: yes
+---
+```{r global-options, include=FALSE}
+knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
+                      warning=FALSE, message=FALSE, echo = FALSE)
+```
 ---
@@
 library(ggiraphExtra) #For plotting
+library(ggeffects)
+library(ggdark)
+library(ggsci)
+library(viridis)
 
@@
 
-ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
+ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE, interactive = TRUE)
 ```
@@
 point = FALSE,
-          interactive = FALSE, 
+          interactive = TRUE, 
 show.summary = TRUE,
@@
 ```{r}
-ggPredict(m1_crust_model_particles, colorn=1000)
+ggPredict(m1_crust_model_particles, colorn=1000, interactive = TRUE)
 ```
@@
 ```{r}
-ggPredict(m1_crust_model_volume, colorn=500)
+ggPredict(m1_crust_model_volume, colorn=500, interactive = TRUE)
 ```
@@
 ```{r}
-ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
+ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE, interactive = TRUE)
 ```
@@
 
-ggPredict(m1_crust_model_dose,colorn=100,jitter=TRUE, interactive = FALSE, digits = 3, se = FALSE)
+ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE, interactive = TRUE)
 
@@
 ```{r}
+require(ggiraphExtra)
 #Plots
-ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
+ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE, interactive = TRUE)
 ```
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 #alternative method using ggpredict
@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 #alt method
@@
 ```
-Plot.
+Plotted below.
 ```{r}
 discrete <- ggpredict(m1_crust_model_discrete,terms = c("logdose.mg.L.master", "size_f"))
-plot(discrete, facet = TRUE, add.data = TRUE)
+plot(discrete, facet = TRUE, add.data = TRUE, colors = "quadro", alpha = 0.2, dodge = 0.2)
+     #title = "All Aquatic Organisms")
 ```
-### Survival package
+#### Acute only crustacea
+```{r, echo = FALSE}
+require(arm) #bayesglm()
+#Select the columns that we want to feed into the model for simplicity
+m1_crust_acute <- aoc_setup %>% 
+  filter(!environment == "Terrestrial") %>% 
+  filter(bio.org == "organism") %>% 
+  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
+  filter(organism.group == "Crustacea") %>% 
+  filter(!acute.chronic_f == "Unavailable") %>% 
+  filter(lvl1_f == "Fitness") %>% 
+  filter(!size_f == "Not Reported") %>% 
+  filter(!effect_f == "NA") %>% 
+  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>% 
+  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
+ # filter(acute.chronic_f == "Acute") %>% 
+  dplyr::select(c(effect_10, 
+                  logdose.mg.L.master, 
+                  logdose.particles.mL.master, 
+                  size_f)) %>% 
+  drop_na()
+
+#Fit the full model
+m1_crust_acute_model_discrete <- glm(effect_10 ~ 
+                                      logdose.mg.L.master * size_f  + 
+                                       logdose.particles.mL.master * size_f, 
+                                     data = m1_crust_acute, 
+                                     na.action = "na.exclude", 
+                                     family = "binomial"#,
+                                     #glm.control(maxit = 50)
+                                     )
+
+summary(m1_crust_acute_model_discrete)
+```
+The above glm is visualized below.
+```{r}
+acute_discrete <- ggpredict(m1_crust_acute_model_discrete,terms = c(#"logdose.mg.L.master", 
+                                                                    "logdose.particles.mL.master",
+                                                                    "size_f")) #ensure discrete factor is second
+plot(acute_discrete)#, facet = TRUE, add.data = TRUE)
+```
+
+#### Alt Method using GG Plot
+This approach achieves a similiar product as using the ggPredict() function, except it relies on ggplot(), which is more malleable and transparent.
+The general steps are to first create a new dataframe over 1000 values of size using expand.grid() then use predict() and plot() with geom_line() and colour=size.
+```{r}
+require(viridis)
+#filtered dataset
+#m1_crust
+#model
+#summary(m1_crust_model_dose )
+#generate distribution of data
+mockData <- expand.grid(size.length.um.used.for.conversions = seq(0.034, 5000, 1),
+            logdose.particles.mL.master = seq(-4.195, 12.650,0.1))
+
+mockData$effect_10 <- predict(m1_crust_model_dose,
+                     mockData,
+                     type = "response")
+#plot
+mockData %>% 
+  ggplot(aes(x = logdose.particles.mL.master, 
+             y = effect_10, 
+             color = size.length.um.used.for.conversions)) +
+  geom_line() +
+  scale_color_gradientn(colors = topo.colors(7)) +
+  #scale_color_viridis_c(option = "A") +
+  dark_theme_bw()
+```
+
+## Survival package
+### Cumulative Hazard by Time
 ```{r}
@@
 ```
+### Probability of Survival by Time
+
 ```{r}

### 2021-02-23 app.R
Added red criteria to setup

@@ -331,8 +331,11 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
   mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
   mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
                                             af.time_noNA == 1 ~ "Chronic",
-                                            af.time_noNA == "Unavailable" ~ "Unavailable")))   #factorize assesment factor time into chronic/acute
-
+                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>% #factorize assesment factor time into chronic/acute
+  mutate(tier_zero_tech_f = factor(case_when(tech.tier.zero == "Y" ~ "Red Criteria Failed",
+                                             tech.tier.zero == "N" ~ "Red Criteria Passed"))) %>% 
+  mutate(tier_zero_risk_f = factor(case_when(risk.tier.zero == "Y" ~ "Red Criteria Failed",
+                                             risk.tier.zero == "N" ~ "Red Criteria Passed")))
 
 #### SSD AO Setup ####
 

### 2021-02-23 app.R
Added red criteria widgets to exploration page and rearranged all widgets

@@ -636,28 +636,70 @@ column(width = 12,
                       column(width = 3,offset = 9,  
                              p("*Warning: Exposure duration selections will limit data to the following organism groups: Fish, Molluscs, Crustacea and Algae."))),
 
-                    radioButtons(inputId = "dose_check", # dosing units
-                                 label = "Particles/mL, mg/L, or um3/mL:",
-                                 choices = c("Particles/mL", "mg/L", "um3/mL"),
-                                 selected = "mg/L"),
+                    # second row of widget headers
+                    column(width=12,
+                           
+                           column(width = 3,
+                                  h4("Quality Criteria")),
+                           
+                           column(width = 3,
+                                  h4("Exposure Concentrations")),
+                           
+                           column(width = 3,
+                                  h4("Aesthetics"))),
                     
-                    p("Concentrations may be reported in mass/volume or particle #/volume (or sometimes both). Using methods described in", a(href ="https://pubs.acs.org/doi/10.1021/acs.est.0c02982", "Koelmans et. al (2020)"), " units have been converted."),
+                    #New row of widgets
                     
-                    radioButtons(inputId = "Rep_Con_rad",
-                      label = "Do you want to use just the reported, just the converted, or all exposure concentrations?",
-                      choices = c("reported", "converted", "all"),
-                      selected = "all"),
+                    column(width = 12,
+                           
+                           column(width = 3,
+                                  pickerInput(inputId = "tech_tier_zero_check", # chronic/acute checklist
+                                              label = "Technical Quality:", 
+                                              choices = levels(aoc_setup$tier_zero_tech_f),
+                                              selected = levels(aoc_setup$tier_zero_tech_f),
+                                              options = list(`actions-box` = TRUE), 
+                                              multiple = TRUE)),
+                           
+                           column(width = 3, 
+                                  radioButtons(inputId = "dose_check", # dosing units
+                                               label = "Particles/mL, mg/L, or um3/mL:",
+                                               choices = c("Particles/mL", "mg/L", "um3/mL"),
+                                               selected = "mg/L")),
+                           
+                           column(width = 3,
+                                  selectInput(inputId = "plot.type", "Plot Type:", 
+                                       list(boxplot = "boxplot", violin = "violin", beeswarm = "beeswarm")),
+                                  checkboxInput(inputId = "show.points", "Show All Points", FALSE))),
+                           
+                  #New row of widgets
                     
-                    #aesthethics
[…]

### 2021-02-23 app.R
Connected widgets to data for exploration tab

@@ -1261,6 +1261,8 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
     shape_c <- input$shape_check # assign values to "shape_c" 
     size_c <- input$size_check # assign values to "size_c"
     species_c <- input$species_check #assign values to "species_c"
+    tech_tier_zero_c<-input$tech_tier_zero_check #assign values to "design_tier_zero_c"
+    risk_tier_zero_c<-input$risk_tier_zero_check #assign values to "risk_tier_zero_c"
     range_n <- input$range # assign values to "range_n"
     dose_check <- input$dose_check #renames selection from radio button
     Rep_Con_rad <- input$Rep_Con_rad #use nominal or calculated exposure concentrations. Options are TRUE (calculated) or FALSE (reported)
@@ -1324,7 +1326,9 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
       filter(shape_f %in% shape_c) %>% #filter by shape
       filter(species_f %in% species_c) %>%  #filter by species
       filter(env_f %in% env_c) %>% #filter by environment
-      filter(acute.chronic_f %in% acute.chronic.c) #acute/chronic
+      filter(acute.chronic_f %in% acute.chronic.c) %>%  #acute/chronic
+      filter(tier_zero_tech_f %in% tech_tier_zero_c) %>% #technical quality
+      filter(tier_zero_risk_f %in% risk_tier_zero_c) #risk assessment quality
       #filter(size.length.um.used.for.conversions <= range_n) #For size slider widget - currently commented out
 
   })
@@ -1861,6 +1865,8 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
     shinyjs::reset("life_check")
     shinyjs::reset("bio_check")
     shinyjs::reset("species_check")
+    shinyjs::reset("tech_tier_zero_check")
+    shinyjs::reset("risk_tier_zero_check")
   }) #If we add more widgets, make sure they get added here. 
 
 #### SSD AO S ####

### 2021-02-23 DALEX.Rmd
slight changes

@@
 author: "Scott Coffin"
-date: "2/19/2021"
+date: "2/22/2021"
 output: 
@@
     toc_float: yes
-    toc_depth: 5
+    toc_depth: 6
     number_sections: true
@@
   scale_x_discrete(labels = c("No Effect", "Effect")) +
-  scale_y_discrete(labels = c("No Effect", "Effect"))
-  theme_bw() +
-  xlim(rev(levels(table$Reference)))
+  scale_y_discrete(labels = c("No Effect", "Effect")) +
+  theme_bw()
 ```

### 2021-02-23 Simplifying via continuous probability distributions.R

@@ -54,4 +54,50 @@ library(ggplot2)
 tbl <- tibble(df)
 
 ggplot(tbl) + 
+  
+# From Merel Kooi
+rm(list=ls())
+
+## Parameters
+n = 1E5     #number of values to generate
+xmin = 1   #minimum size (um)
+xmax = 5000 #maximum size (um)
+alpha = 1.6 #exponent of power law distribution
+
+## Generate random data given parameters of the power law distribution
+## Based on PDF
+
+## First approach 
+a = -alpha     # It has to be negative
+y = runif(n)   # Number of samples
+Size1 = ((xmax^(a+1) - xmin^(a+1))*y + xmin^(a+1))^(1/(a+1))
+
+## Second approach
+## Function that will randomly generate 1 value from the distribution
+X.func <- function (X){
+  success <- FALSE
+  while (!success){
+    U = runif(1, 0, 1)
+    X = xmin*(1-U)^(1/(1-alpha))
+    success <- X < xmax }
+  return(X)
+}
+
+## df to store your generated results in
+Size2 <- data.frame(Size = numeric(0))
+
+## loop through function n times (can probably do this in a function too)
+for(i in 1:n){
+  X <- X.func()
+  Size2 <- rbind(Size2, X)
+}
+
+colnames(Size2) <- "Size2"
+
+## check results - visual comparison
+Size <- cbind(Size1, Size2)
+
+par(mfrow = c(1,2))
+hist(Size$Size1)
+hist(Size$Size2)
   geom_density(aes(y))

### 2021-02-23 app.R
Added quality criteria widgets to ssd tab

@@ -779,8 +779,17 @@ column(width = 12,
                     br(),
                     p("Use the options below to filter the toxicity thresholds dataset. Once complete, hit the 'submit' button"),
                     
-                    # widget 1
+                    # Row 1 of Headers
+                    column(width=12,
+                           
+                           column(width = 3,
+                                  h4("Environment & Organism Type")),
+                    ), #closes out column
+                    
+                    # Row 1 of Widgets
                     column(width = 12,
+                           
+                           #Environment widget
                            column(width = 3,
                                   # alternative to fully listed checklists
                                   # requires shinyWidgets package
@@ -790,40 +799,15 @@ column(width = 12,
                                               selected = levels(aoc_z$env_f),   
                                               options = list(`actions-box` = TRUE), # option to de/select all
                                               multiple = TRUE)), # allows for multiple inputs
-                           # Organism widget
+                           
+                           #Organism group widget - reactive to environment
                            column(width = 3,
                                   htmlOutput("GroupSelection")), # organism checklist
+                          
+                           #Species widget - reactive to environment 
                            column(width = 3,
-                                  htmlOutput("SpeciesSelection"))), # dependent Species checklist
-                           br(),
-                           p("Advanced options. Suggest using defaults."),
-                           br(),
-                    
-                    column(width = 12,
-                           #Size widget
-                           column(width = 3,
-                                  pickerInput(inputId = "size_check_ssd", # organism checklist
-                                              label = "Sizes:",
-                                              choices = levels(aoc_z$size_f),
-                                              selected = levels(aoc_z$size_f),
-                                              options = list(`actions-box` = TRUE), # option to de/select all
-                                              multiple = TRUE)), # allows for multiple inputs
-                           #Morphology widget
-                           column(width = 3,
-                                  htmlOutput("shapeSelection")), # dependent on sizes
+                                  htmlOutput("SpeciesSelection")),
                            
-                            #Endpoint widget
-                           column(width = 3,
-                                  htmlOutput("lvl1Selection")), # allows for multiple inputs
-                           column(width = 3,
-                                  htmlOutput("lvl2Selection")), #specific endpoint based on previous checkbox
-                    
-                            ), #closes out column
-                       
-                    column(width = 12,    
-                           #Polymer widget
-                           column(width = 3,
-                                  htmlOutput("polySelection")),# polymer selection based on other inputs
                            #acute/chronic widget
                            column(width = 3,
[…]

### 2021-02-23 app.R
Added reference document for quality criteria

@@ -1111,6 +1111,8 @@ tabPanel("5: Resources",
          br(),     
          h3(align = "center", a(href = "https://sccwrp-my.sharepoint.com/:b:/g/personal/leahth_sccwrp_org/EYUFX1dOfSdGuHSfrUDcnewBxgttfTCOwom90hrt5nx1FA?e=jFXEyQ", 'Data Category Descriptions')),
          br(),
+         h3(align = "center", a(href = "https://sccwrp-my.sharepoint.com/:b:/g/personal/leahth_sccwrp_org/EXOluRMsb_RPpjsqTjhmuaUBNz3Pd9vkl7Hl09lKFxaxEA?e=nlPaGl", 'Quality Screening: Red Criteria')),
+         br(),
          h3(align = "center", a(href = "https://sccwrp-my.sharepoint.com/:b:/g/personal/leahth_sccwrp_org/ETy8vDCXe_pAq88Ky0Xob1gBmCdAXYCsEwDFqCfDTL-DNA?e=e7Ic21", 'Aquatic Organisms Study List')),
          br(),
          h3(align = "center", a(href = "https://sccwrp-my.sharepoint.com/:b:/g/personal/leahth_sccwrp_org/EXf0crCKDPVHo5xBEdw4PQwBxA8cnu0x4WY477CuEzZcPw?e=qs00V3", 'Dose Conversion Methods')),

### 2021-02-24 DALEX.Rmd
more plots!

@@
 ---
-title: "DALEX"
+title: "DALEX_TechQualityControl"
 author: "Scott Coffin"
-date: "2/22/2021"
+date: "2/23/2021"
 output: 
@@
 library(modelplotr)
+library(gridExtra)
 ```
@@
 df <- aoc_z %>% 
+  filter(tech.tier.zero == "N") %>% #gives studies that pass technical quality criteria
   dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, dose.mg.L.master, dose.particles.mL.master, life.stage, bio.org, polymer, shape, size.length.um.used.for.conversions, treatments, effect, effect.metric, particle.volume.um3, density.mg.um.3, environment, exposure.route, lvl2_f)) %>%
@@
 
-```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+```{r All Model Training, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
 # Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
@@
 ## Model Performance
-```{r}
+```{r model performance, include=FALSE}
 mp_classif_rf <- model_performance(explainer_classif_rf)
@@
 #mp_classif_logicBag <- model_performance(explainer_classif_logicBag)
-
+```
+
+### Histogram of Residuals
+```{r}
+plot(mp_classif_rf, mp_classif_glm, geom ="histogram")
+```
+```{r}
+plot(mp_classif_svm, mp_classif_decTree, geom ="histogram")
+```
+
+```{r}
+plot(mp_classif_nnet, mp_classif_xgbTree, geom ="histogram")
+```
+### Precision Recall Curves
+```{r}
+plot(mp_classif_rf, mp_classif_glm, mp_classif_svm, mp_classif_decTree, mp_classif_nnet, mp_classif_xgbTree, geom ="prc")
+```
+
+### Reverse Cumulative Distribution of Residuals
+```{r}
 # compare residuals plots
@@
 
-#### Lift Curves
+### Lift Curves
 
@@
 
-#### Roc Curves
+### Roc Curves
 ```{r}
@@
   ggtitle("ROC Curves - All Models",  
-          paste("AUC_rf = ",round(mp_classif_rf$measures$auc,2), 
-                paste("AUC_svm = ",round(mp_classif_svm$measures$auc,2)),
-          paste("AUC_nnet = ",round(mp_classif_nnet$measures$auc,2))
+          paste("AUC_rf = ",round(mp_classif_rf$measures$auc,3), 
+                paste("AUC_svm = ",round(mp_classif_svm$measures$auc,3)),
+          paste("AUC_nnet = ",round(mp_classif_nnet$measures$auc,3))
           )) +
@@
 
-#### Partial Dependence Plot
+### Partial Dependence Plot
 ```{r}
@@
 
-#### Breakdown
+##### Grouped Partial Dependence Plots
+```{r include=FALSE}
+plot(pdp_classif_rf)
+#Partial Dependence by dose.particles.mL.master and effect.metric
+pdp_classif_rf  <- model_profile(explainer_classif_rf, variable = "dose.particles.mL.master", 
+                                 groups = "effect.metric")
+pdp_classif_glm  <- model_profile(explainer_classif_glm, variable = "dose.particles.mL.master", 
+                                 groups = "effect.metric")
+pdp_classif_svm  <- model_profile(explainer_classif_svm, variable = "dose.particles.mL.master", 
+                                 groups = "effect.metric")
+pdp_classif_decTree  <- model_profile(explainer_classif_decTree, variable = "dose.particles.mL.master", 
+                                 groups = "effect.metric")
+pdp_classif_nnet  <- model_profile(explainer_classif_nnet, variable = "dose.particles.mL.master", 
+                                 groups = "effect.metric")
+pdp_classif_xgbTree  <- model_profile(explainer_classif_xgbTree, variable = "dose.particles.mL.master", 
+                                 groups = "effect.metric")
+```
+
+```{r}
+#plot neural net and glm
+glm <- plot(pdp_classif_glm$agr_profiles) +
+  ggtitle("GLM: Partial Dependence Profiles by Dose and Effect Metric", "") +
+  xlab("log10(dose particles/mL)") +
+  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
+  scale_color_discrete(name = "Model x Effect Metric") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+nnet <- plot(pdp_classif_nnet$agr_profiles) +
+  ggtitle("Neural Net: Partial Dependence Profiles by Dose and Effect Metric", "") +
+  xlab("log10(dose particles/mL)") +
+  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
+  scale_color_discrete(name = "Model x Effect Metric") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+grid.arrange(glm, nnet)
+```
+
+```{r}
+#plot rf and decision tree
+rf <- plot(pdp_classif_rf$agr_profiles) +
+  ggtitle("Random Forest:Partial Dependence Profiles by Dose and Effect Metric", "") +
+  xlab("log10(dose particles/mL)") +
+  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
+  scale_color_discrete(name = "Model x Effect Metric") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+decTree <- plot(pdp_classif_decTree$agr_profiles) +
+  ggtitle("Decision Tree: Partial Dependence Profiles by Dose and Effect Metric", "") +
+  xlab("log10(dose particles/mL)") +
+  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
+  scale_color_discrete(name = "Model x Effect Metric") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+grid.arrange(rf, decTree)
+```
+
+```{r}
+#plot rf and decision tree
+svm <- plot(pdp_classif_svm$agr_profiles) +
+  ggtitle("Support Vector Machine:Partial Dependence Profiles by Dose and Effect Metric", "") +
+  xlab("log10(dose particles/mL)") +
+  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
+  scale_color_discrete(name = "Model x Effect Metric") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+xgbTree <- plot(pdp_classif_xgbTree$agr_profiles) +
+  ggtitle("eXtreme Gradient Boosing Trees: Partial Dependence Profiles by Dose and Effect Metric", "") +
+  xlab("log10(dose particles/mL)") +
+  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
+  scale_color_discrete(name = "Model x Effect Metric") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+grid.arrange(svm, xgbTree)
+```
+
+### Confusion Matrices
+
+```{r include=FALSE}
+#predict test data
+pred_rf <- predict(classif_rf, newdata = test)
+pred_glm <- predict(classif_glm, newdata = test)
+pred_svm <- predict(classif_svm, newdata = test)
+pred_decTree <- predict(classif_decTree, newdata = test)
+pred_nnet <- predict(classif_nnet, newdata = test)
+pred_xgbTree <- predict(classif_xgbTree, newdata = test)
+
+#build confusion matrices
+table_rf <- data.frame(confusionMatrix(pred_rf, test$effect_10)$table)
+table_glm <- data.frame(confusionMatrix(pred_glm, test$effect_10)$table)
+table_svm <- data.frame(confusionMatrix(pred_svm, test$effect_10)$table)
+table_decTree <- data.frame(confusionMatrix(pred_decTree, test$effect_10)$table)
+table_nnet <- data.frame(confusionMatrix(pred_nnet, test$effect_10)$table)
+table_xgbTree <- data.frame(confusionMatrix(pred_xgbTree, test$effect_10)$table)
+
+#build plots
+plotTable_rf <- table_rf %>% mutate(goodbad = ifelse(table_rf$Prediction == table_rf$Reference, "good", "bad")) %>%
+  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))
+
+plotTable_glm <- table_glm %>% mutate(goodbad = ifelse(table_glm$Prediction == table_glm$Reference, "good", "bad")) %>%
+  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))
+
+plotTable_svm <- table_svm %>% mutate(goodbad = ifelse(table_svm$Prediction == table_svm$Reference, "good", "bad")) %>%
+  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))
+
+plotTable_decTree <- table_decTree %>% mutate(goodbad = ifelse(table_decTree$Prediction == table_decTree$Reference, "good", "bad")) %>%
+  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))
+
+plotTable_nnet <- table_nnet %>% mutate(goodbad = ifelse(table_nnet$Prediction == table_nnet$Reference, "good", "bad")) %>%
+  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))
+
+plotTable_xgbTree <- table_xgbTree %>% mutate(goodbad = ifelse(table_xgbTree$Prediction == table_xgbTree$Reference, "good", "bad")) %>%
+  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))
+```
+
+Confusion Matrices for the six tested models are below.
+```{r}
+# fill alpha relative to sensitivity/specificity by proportional outcomes within reference groups (see dplyr code above as well as original confusion matrix for comparison)
+CM_rf <- ggplot(data = plotTable_rf, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
+  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
+  scale_x_discrete(labels = c("No Effect", "Effect")) +
+  scale_y_discrete(labels = c("No Effect", "Effect")) +
+  ggtitle("Random Forest",
+          paste("Accuracy = ", 100 * round(mp_classif_rf$measures$accuracy,3), "%")) +
+  theme_bw() +
+    theme(plot.title = element_text(hjust = 0.5),
+          legend.position = "none")
+
+CM_glm <- ggplot(data = plotTable_glm, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
+  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
+  scale_x_discrete(labels = c("No Effect", "Effect")) +
+  scale_y_discrete(labels = c("No Effect", "Effect")) +
+  ggtitle("General Linear Model",
+          paste("Accuracy = ", 100 * round(mp_classif_glm$measures$accuracy,3), "%")) +
+  theme_bw() +
+    theme(plot.title = element_text(hjust = 0.5),
+          legend.position = "none")
+
+CM_svm <- ggplot(data = plotTable_svm, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
+  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
+  scale_x_discrete(labels = c("No Effect", "Effect")) +
+  scale_y_discrete(labels = c("No Effect", "Effect")) +
+  ggtitle("Support Vector Machine",
+          paste("Accuracy = ", 100 * round(mp_classif_svm$measures$accuracy,3), "%")) +
+  theme_bw() +
+ theme(plot.title = element_text(hjust = 0.5),
+          legend.position = "none")
+
+CM_decTree <- ggplot(data = plotTable_decTree, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
+  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
+  scale_x_discrete(labels = c("No Effect", "Effect")) +
+  scale_y_discrete(labels = c("No Effect", "Effect")) +
+  ggtitle("Decision Tree",
+          paste("Accuracy = ", 100 * round(mp_classif_decTree$measures$accuracy,3), "%")) +
+  theme_bw() +
+  theme(plot.title = element_text(hjust = 0.5),
+          legend.position = "none")
+
+CM_nnet <- ggplot(data = plotTable_nnet, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
+  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
+  scale_x_discrete(labels = c("No Effect", "Effect")) +
+  scale_y_discrete(labels = c("No Effect", "Effect")) +
+  ggtitle("Neural Net",
+          paste("Accuracy = ", 100 * round(mp_classif_nnet$measures$accuracy,3), "%")) +
+  theme_bw() +
+    theme(plot.title = element_text(hjust = 0.5),
+          legend.position = "none")
+
+CM_xgbTree <- ggplot(data = plotTable_xgbTree, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
+  geom_tile() +
+  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
+  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
+  scale_x_discrete(labels = c("No Effect", "Effect")) +
+  scale_y_discrete(labels = c("No Effect", "Effect")) +
+  ggtitle("eXtreme Gradient Boosting Trees",
+          paste("Accuracy = ", 100 * round(mp_classif_xgbTree$measures$accuracy,3), "%")) +
+  theme_bw() +
+    theme(plot.title = element_text(hjust = 0.5),
+          legend.position = "none")
+
+grid.arrange(CM_rf, CM_glm, CM_svm, CM_decTree, CM_nnet, CM_xgbTree,
+             ncol = 2, top = "Confusion Matrices for ML Models")
+```
+
+### Breakdown
 
@@
 http://rstudio-pubs-static.s3.amazonaws.com/480890_237ad52b09b6440e9c849a3c07a04d2f.html
-```{r include=FALSE}
+```{r Tuning, include=FALSE}
 cache = TRUE
@@
 
-```{r}
+```{r Recursive Feature Elimination}
 
@@
 Now that we have optimized the variables for our final model, we will build and save the final simplified model that has the highest accuracy.
-```{r}
+```{r Final Model}
 tunegrid <- expand.grid(.mtry=c(7))

### 2021-02-24 DALEX.Rmd
more plots for DALEx and default preferences updates for app.R

@@
   filter(tech.tier.zero == "N") %>% #gives studies that pass technical quality criteria
+  filter(risk.tier.zero == "N") %>%  #only studies applicable for risk assessment. VERY RESTRICTIVE 
   dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, dose.mg.L.master, dose.particles.mL.master, life.stage, bio.org, polymer, shape, size.length.um.used.for.conversions, treatments, effect, effect.metric, particle.volume.um3, density.mg.um.3, environment, exposure.route, lvl2_f)) %>%
@@
 
-#classif_logicBag <- train(effect_10~., data = train, method = "logicBag", ntree = 100)
-
 # explain
@@
                                         data = test, y = yTest)
-
-#explainer_classif_logicBag <- DALEX::explain(classif_svm,  label = "Bagged Logic Regression", 
-                                       # data = test, y = yTest)
-
 ```
@@
 mp_classif_xgbTree <- model_performance(explainer_classif_xgbTree)
-#mp_classif_logicBag <- model_performance(explainer_classif_logicBag)
 ```

### 2021-02-24 app.R

@@ -951,7 +951,7 @@ column(width = 12,
                                   pickerInput(inputId = "effect.metric_rad_ssd", # effect metric checklist
                                               label = "Effect Metric:",
                                               choices = levels(aoc_z$effect.metric),
-                                              selected = "HONEC",
+                                              selected = c("HONEC", "NOEC"),
                                               options = list(`actions-box` = TRUE), 
                                               multiple = TRUE)), # allows for multiple inputs
                           
@@ -975,7 +975,7 @@ column(width = 12,
                                     inputId = "conc.select.rad",
                                     label = "What should be considered the 'sensitive' concentration for each species?",
                                     choices = list("minimum", "lower 95% CI", "median", "mean"),
-                                    selected = "minimum")),
+                                    selected = "mean")),
                            
                     ),#close out column
                     

### 2021-02-24 app.R
Changed y/n to pass/fail for quality criteria

@@ -332,10 +332,10 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
   mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
                                             af.time_noNA == 1 ~ "Chronic",
                                             af.time_noNA == "Unavailable" ~ "Unavailable"))) %>% #factorize assesment factor time into chronic/acute
-  mutate(tier_zero_tech_f = factor(case_when(tech.tier.zero == "Y" ~ "Red Criteria Failed",
-                                             tech.tier.zero == "N" ~ "Red Criteria Passed"))) %>% 
-  mutate(tier_zero_risk_f = factor(case_when(risk.tier.zero == "Y" ~ "Red Criteria Failed",
-                                             risk.tier.zero == "N" ~ "Red Criteria Passed")))
+  mutate(tier_zero_tech_f = factor(case_when(tech.tier.zero == "Fail" ~ "Red Criteria Failed",
+                                             tech.tier.zero == "Pass" ~ "Red Criteria Passed"))) %>% 
+  mutate(tier_zero_risk_f = factor(case_when(risk.tier.zero == "Fail" ~ "Red Criteria Failed",
+                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed")))
 
 #### SSD AO Setup ####
 

### 2021-02-24 app.R
Made plots larger and made study labels slightly less ugly

@@ -736,37 +736,37 @@ column(width = 12,
                     hr(), # adds divider
                     
                     column(width = 12,
-                    plotOutput(outputId = "organism_plot_react"),
+                    plotOutput(outputId = "organism_plot_react", height = "600px"),
                     br())), 
                     
                     column(width = 12,
                   
                     column(width = 12,
-                    plotOutput(outputId = "lvl_plot_react"),
+                    plotOutput(outputId = "lvl_plot_react", height = "600px"),
                     br())), 
 
                     column(width = 12,
                     
                     column(width = 12,
-                    plotOutput(outputId = "lvl2_plot_react"),
+                    plotOutput(outputId = "lvl2_plot_react", height = "600px"),
                     br())), 
                     
                     column(width = 12,
                            
                     column(width = 12,
-                    plotOutput(outputId = "size_plot_react"),
+                    plotOutput(outputId = "size_plot_react", height = "600px"),
                     br())), 
                     
                     column(width = 12,
                   
                     column(width = 12,
-                    plotOutput(outputId = "shape_plot_react"),
+                    plotOutput(outputId = "shape_plot_react", height = "600px"),
                     br())), 
                     
                     column(width = 12,
                 
                     column(width = 12,
-                    plotOutput(outputId = "poly_plot_react"),
+                    plotOutput(outputId = "poly_plot_react", height = "600px"),
                     br()))), 
 
 #### SSD AO UI ####
@@ -1474,7 +1474,7 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
       group_by(org_f, effect_f) %>% # need to include so there's a recognized "y"
       summarize(dose_new = quantile(dose_new, .1), # need for recognized "x"
                 measurements = n(),
-                studies = n_distinct(article))
+                studies = n_distinct(article)) 
    
     p <- ggplot(aoc_filter(), aes(x = dose_new, y = org_f, fill = effect_f)) +
       plot.type + #adds user-defined geom()
@@ -1488,9 +1488,10 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
       # scale_fill_manual(values = c("#FD8D3C", "#7F2704")) +
       geom_label_repel(data = aoc_org1,
                       aes(label = paste("(",measurements,",",studies,")")),
[…]

### 2021-02-24 app.R
Darkened bar plots

@@ -1438,9 +1438,9 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
     
     #plot types
     plot.type<-switch(input$plot.type,
-                      "boxplot" 	= geom_boxplot(alpha = 0.7, aes(color = effect_f)),
-                      "violin" = geom_violin(alpha = 0.7, aes(color = effect_f)),
-                      "beeswarm" = geom_quasirandom(alpha = 0.7, aes(color = effect_f), 
+                      "boxplot" 	= geom_boxplot(alpha = 0.8, aes(color = effect_f)),
+                      "violin" = geom_violin(alpha = 0.8, aes(color = effect_f)),
+                      "beeswarm" = geom_quasirandom(alpha = 0.8, aes(color = effect_f), 
                                                     method = "smiley", groupOnX = FALSE, cex = 2)) #groupOnX specifies groups on y axis
     
     #Theme type
@@ -1522,9 +1522,9 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
     
     #plot types
     plot.type<-switch(input$plot.type,
-                      "boxplot" 	= geom_boxplot(alpha = 0.7, aes(color = effect_f)),
-                      "violin" = geom_violin(alpha = 0.7, aes(color = effect_f)),
-                      "beeswarm" = geom_quasirandom(alpha = 0.7, aes(color = effect_f), 
+                      "boxplot" 	= geom_boxplot(alpha = 0.8, aes(color = effect_f)),
+                      "violin" = geom_violin(alpha = 0.8, aes(color = effect_f)),
+                      "beeswarm" = geom_quasirandom(alpha = 0.8, aes(color = effect_f), 
                                                     method = "smiley", groupOnX = FALSE, cex = 2)) #groupOnX specifies groups on y axis
     
     #Theme type
@@ -1603,9 +1603,9 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
     
     #plot types
     plot.type<-switch(input$plot.type,
-                      "boxplot" 	= geom_boxplot(alpha = 0.7, aes(color = effect_f)),
-                      "violin" = geom_violin(alpha = 0.7, aes(color = effect_f)),
-                      "beeswarm" = geom_quasirandom(alpha = 0.7, aes(color = effect_f), 
+                      "boxplot" 	= geom_boxplot(alpha = 0.8, aes(color = effect_f)),
+                      "violin" = geom_violin(alpha = 0.8, aes(color = effect_f)),
[…]

### 2021-02-24 app.R
Small edits trying to get YAML to work

@@ -8,8 +8,8 @@
 
 # Load packages
 library(tidyverse) #General everything
-library(RColorBrewer)
-library(ggplot2) #General plotting
+library(RColorBrewer) #color palette
+library(ggplot2) #plotting
 library(ggrepel) #For adding text labels that repel away from data points
 library(calecopal) #Color palette
 library(shiny) #Runs shiny
@@ -30,8 +30,6 @@ library(ggdark) #dark mode ggplot
 library(ggsci) #color palettes
 #library(bslib) #better themes
 #library(thematic) #complete control over themes (including plots)
-
-
 
 # Load finalized dataset.
 aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
@@ -504,7 +502,7 @@ column(width = 12,
 
 
 #### Exploration AO UI ####
-                  tabPanel("3: Exploration",
+                tabPanel("3: Exploration",
                     shinyjs::useShinyjs(), # requires package for "reset" button, DO NOT DELETE - make sure to add any new widget to the reset_input in the server
                     id = "heili-tab", # adds ID for resetting Heili's tab's filters
                     
@@ -542,7 +540,7 @@ column(width = 12,
                     column(width = 12,
                       
                       column(width = 3,
-                      pickerInput(inputId = "lvl1_check", # endpoint checklist
+                        pickerInput(inputId = "lvl1_check", # endpoint checklist
                         label = "Broad Endpoint Category:", 
                         choices = levels(aoc_setup$lvl1_f),
                         selected = levels(aoc_setup$lvl1_f),
@@ -550,7 +548,7 @@ column(width = 12,
                         multiple = TRUE)), # allows for multiple inputs
                       
                       column(width = 3,
-                      pickerInput(inputId = "poly_check", # polymer checklist
+                        pickerInput(inputId = "poly_check", # polymer checklist
                         label = "Polymer:", 
                         choices = levels(aoc_setup$poly_f),
                         selected = levels(aoc_setup$poly_f),
@@ -558,7 +556,7 @@ column(width = 12,
                         multiple = TRUE)),
                       
                       column(width = 3,
-                      pickerInput(inputId = "organism_check", # organismal group checklist
+                        pickerInput(inputId = "organism_check", # organismal group checklist
                         label = "Organisms:", 
                         choices = levels(aoc_setup$org_f),
                         selected = levels(aoc_setup$org_f),
@@ -566,7 +564,7 @@ column(width = 12,
                         multiple = TRUE)), 
                       
[…]

### 2021-02-24 app.R
removed link

@@ -412,7 +412,7 @@ ui <- fluidPage(theme = shinytheme("flatly"), #light,
                  
                     p(align = "center", a(href = "https://www.sccwrp.org/about/staff/leah-thornton-hampton/", 'Dr. Leah Thornton Hampton'),", Southern California Coastal Water Research Project ", 
                       tags$a(href="https://twitter.com/DrLeahTH", tags$img(src="twitter.png", width="2%", height="2%")), tags$a(href="https://github.com/leahth", tags$img(src="github.png", width="2%", height="2%"))),
-                    p(align = "center", a(href = "https://www.sccwrp.org/about/staff/heili-lowman/", 'Dr. Heili Lowman'),", Southern California Coastal Water Research Project ",
+                    p(align = "center", "Heili Lowman, Southern California Coastal Water Research Project ",
                       tags$a(href="https://twitter.com/heili_lowman", tags$img(src="twitter.png", width="2%", height="2%")), tags$a(href="https://github.com/hlowman", tags$img(src="github.png", width="2%", height="2%"))), 
                     p(align = "center", a(href = "https://agency.calepa.ca.gov/staffdirectory/detail.asp?UID=69294&BDO=7&VW=DET&SL=S", 'Dr. Scott Coffin'),", California State Water Resources Control Board", 
                       tags$a(href="https://twitter.com/DrSCoffin", tags$img(src="twitter.png", width="2%", height="2%")), tags$a(href="https://github.com/ScottCoffin", tags$img(src="github.png", width="2%", height="2%"))),

### 2021-02-27 Modelling/simulated distributions.Rmd
simulated distributions

---
title: "Simulated Probabilistic Distributions"
author: "Scott Coffin"
date: "2/24/2021"
output:   
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
    includes:
     # after_body: footer.html
  word_document:
    toc: yes
---
#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE,time_it = TRUE) #report
```

```{r library}
# Bimodal universal shape distribution for environmental microplastic from Kooi and Koelmans (2019)
library(truncnorm)
library(tidyverse)
library(MonteCarlo)# Monte Carlo Simulation 
library(gridExtra)
library(msm) ## rtnorm - get upper and lower limit of shape distribution
library(GeneralizedHyperbolic) ## normal-inverse Gaussian
library(skimr)
```
# Resources: 
https://pubs.acs.org/doi/suppl/10.1021/acs.est.0c02982/suppl_file/es0c02982_si_001.pdf https://pubs.acs.org/doi/suppl/10.1021/acs.estlett.9b00379/suppl_file/ez9b00379_si_001.pdf
https://pubs.acs.org/doi/10.1021/acs.estlett.9b00379

<!-- ## Parameters -->
<!-- ```{r} -->
<!-- #average parameters from distribution -->
<!-- f1 = 0.06 -->
<!-- f2 = 0.94 -->
<!-- delta1 = 0.03 -->
<!-- delta2 = 0.19 -->
<!-- mu1 = 0.08 -->
<!-- mu2 = 0.44 -->
<!-- x = runif(n=100) -->

<!-- #define list from above -->
<!-- param_list = list("f1" = f1, "f2" = f2, "delta1" = delta1, "delta2" = delta2, "mu1" = mu1, "mu2" = mu2, "x" = x) -->

<!-- #permanently define euler's number -->
<!-- e <- exp(1) -->
<!-- lockBinding("e", globalenv()) -->

<!-- #starting conditions for x -->
<!-- x = runif(n = 100) #particle size in micrometers -->

<!-- #equation 4 -->
<!-- y = f1 * (1/sqrt(2 * pi * delta1 ^ 2)) * (e^-(x - mu1)^2 /2 *delta1 ^ 2)+ f2 * (1/sqrt(2 * pi * delta2 ^ 2)) * (e^-(x - mu2)^2 /2 *delta2 ^ 2) -->

<!-- #define equation 4 -->
<!-- binom <- function(x, f1, f2, delta1, delta2, mu1, mu2){ -->
<!--   #generate sample -->
<!--   #x = runif(n=1000) -->
<!--   #calculate CSF -->
<!--   y <- f1 * (1/sqrt(2 * pi * delta1 ^ 2)) * (e^-(x - mu1)^2 /2 *delta1 ^ 2)+ f2 * (1/sqrt(2 * pi * delta2 ^ 2)) * (e^-(x - mu2)^2 /2 *delta2 ^ 2) -->
<!--   #return result -->
<!--   return(list("y" = y)) -->
<!-- } -->

<!-- MC_result <- MonteCarlo(func=binom, nrep=1, param_list=param_list) -->

<!-- df<-MakeFrame(MC_result) -->
<!-- head(df) -->
<!-- tbl <- tibble(df) -->

<!-- ``` -->
<!-- #### Size -->

<!-- ```{r} -->
<!-- # From Merel Kooi -->
<!-- #rm(list=ls()) -->

<!-- ## Parameters -->
<!-- n = 1E5     #number of values to generate -->
<!-- xmin = 1   #minimum size (um) -->
<!-- xmax = 5000 #maximum size (um) -->
<!-- alpha = 1.6 #exponent of power law distribution -->

<!-- ## Generate random data given parameters of the power law distribution -->
<!-- ## Based on PDF -->

<!-- ## First approach  -->
<!-- a = -alpha     # It has to be negative -->
<!-- y = runif(n)   # Number of samples -->
<!-- Size1 = ((xmax^(a+1) - xmin^(a+1))*y + xmin^(a+1))^(1/(a+1)) -->

<!-- ## Second approach -->
<!-- ## Function that will randomly generate 1 value from the distribution -->
<!-- size.func <- function (X){ -->
<!--   success <- FALSE -->
<!--   while (!success){ -->
<!--     U = runif(1, 0, 1) -->
<!--     X = xmin*(1-U)^(1/(1-alpha)) -->
<!--     success <- X < xmax } -->
<!--   return(X) -->
<!-- } -->

<!-- ## df to store your generated results in -->
<!-- Size2 <- data.frame(Size = numeric(0)) -->

<!-- ## loop through function n times (can probably do this in a function too) -->
<!-- for(i in 1:n){ -->
<!--   X <- size.func() -->
<!--   Size2 <- rbind(Size2, X) -->
<!-- } -->

<!-- colnames(Size2) <- "Size2" -->

<!-- ## check results - visual comparison -->
<!-- Size <- cbind(Size1, Size2) -->

<!-- p1 <- ggplot(Size, aes(x = Size1)) + geom_histogram() + scale_x_log10() -->
<!-- p2 <- ggplot(Size, aes(x = Size2)) + geom_histogram() + scale_x_log10() -->
<!-- grid.arrange(p1, p2, -->
<!--              top = "Comparison of Two Methods to Generate Distributions") -->
<!-- ``` -->

<!-- #### Shape -->

<!-- The most abundant shape category of microplastic in water and sediment is fibers (48.5%), followed by fragments (31%), beads (6.5%), films (5.5%), and foam (3.5%)(3). Combining these abundance data with the triangular shape distributions (Table S2 and Figure S2) resulted in a continuous bimodal microplastic shape distribution (Figure 2A). The fitted parameter values for eq 4 were as follows: f1 = 0.06, f2 = 0.94, σ1 = 0.03, σ2 = 0.19, μ1 = 0.08, and μ2 = 0.44 (see Table S5 for standard errors). A Pearson χ2 test indicated that the optimized distribution fits the data well (the fitted model did not differ significantly from the data; p = 0.231 > 0.05). The distribution is dominated by fibers and fragments (CSF = 0.25–0.75) but also has a distinct second peak at a CSF of 0.07, which is mainly attributed to sheets (Figure 2A). The distribution captures the main features of shapes encountered in the environment as well as the relative abundances of these (now continuous) shapes in one go. Most illustrative though is the continuous character of microplastic shape. -->

<!-- ```{r} -->
<!-- # equation for shape -->

<!-- ## Parameters -->
<!-- n = 1E5     #number of values to generate -->
<!-- f1 = 0.059    -->
<!-- sigma1 = 0.076  -->
<!-- mu1 =  0.030 -->
<!-- f2 = 0.941 -->
<!-- sigma2 = 0.441 -->
<!-- mu2 =  0.189 -->

<!-- y=runif(n) #number of samples -->

<!-- ## Generate random data given parameters of the bimodal normal distributions Based on PDF -->

<!-- ## First approach  -->
<!-- CSF1 = (f1 * (1 / sqrt(2 * pi() * sigma1 ^ 2) * e ^ -(x - mu1) ^ 2 / (2 * sigma1 ^ 2) + f2 * (1 / sqrt(2 * pi() * sigma2 ^ 2) * e ^ -(x - mu2) ^ 2 / (2 * sigma2 ^ 2)))) -->

<!-- #a = -alpha     # Ithas to be negative -->
<!-- #y = runif(n)   # Number of samples -->
<!-- #Size1 = ((xmax^(a+1) - xmin^(a+1))*y + xmin^(a+1))^(1/(a+1)) -->

<!-- ``` -->

<!-- #### Density -->

# Generate Particles

Shape is defined as:
Simplifed equation for corey shape factor:

$CSF = H/sqrt(LW)$

Equation for abundance of particles based on density (two distributions). (Note that this is *equation 4* in Kooi and Koelmans (2019).)

$y = f1(\frac{1}{sqrt(2pi()\sigma_1^2})e^{-(x-\mu_1)^2/2\sigma_1^2} + f2(\frac{1}{sqrt(2pi()\sigma_2^2})e^{-(x-\mu_2)^2/2\sigma_2^2}$

For polymer abundance, for the K&K2019 paper I used the fixed values from Burns & Boxall review, and went straight for density. Given these fixed relative abundances, you could calculate polymer type randomly for N particles via:

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
n = 1000
sample(pols, size = n, prob = rel.ab, replace = TRUE)

Particle volume is determined according to shape as follows:

$V = \frac{\pi}{6}L^3CSF^2$

Where CSF = corey shape factor and L = length.

Particle mass is estimated as follows:

$m = pV*\frac{1}{1e12}*1000$

Where *m* is the mass (mg), *p* is density (g/cm^3), *V* is volume (um^3) - which is calculated by the cube of the length (um) of each particle, and additional conversion factors for mg to g (x1000) and cm^3 to um^3 (1e-12).
```{r generate size, include=FALSE}
#rm(list=ls())
n = 10000
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)
```

```{r}
##++++++++++++++++++++++++++++++++++++++++
## Calculate Volume/Volume and estimate mass/volume
##++++++++++++++++++++++++++++++++++++++++
Data_final <- Data %>% 
  mutate(vol.L = 1000) %>% 
  mutate(vol.mL = vol.L * 1000) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L)
skim(Data_final)
```

Parameters:
Shape, density, size, 
why do 3 polymers? Leave polymer out for now. 
For mechanistical point of view- polymer does not matter.

```{r}
##++++++++++++++++++++++++++++++++++++++++
## Quick check results
##++++++++++++++++++++++++++++++++++++++++
#p1 <- ggplot(Data_final, aes(x = Size, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7) + scale_x_log10()
p2 <- ggplot(Data_final, aes(x = Shape, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)
p3 <- ggplot(Data_final, aes(x = Density, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)
p4 <- ggplot(Data_final, aes(x = polymer, fill = polymer)) + geom_histogram(stat = "count")
#p5 <- ggplot(Data_final, aes(x = mass.mg, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7 )+ scale_x_log10()
#p6 <- ggplot(Data_final, aes(x = um3, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
p7 <- ggplot(Data_final, aes(x = dose.um3.mL.master, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
p8 <- ggplot(Data_final, aes(x = dose.particles.mL.master, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
p9 <- ggplot(Data_final, aes(x = dose.mg.L.master, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
grid.arrange(#p1,
             p2,
             p3,
             p4,
             #p5, 
             #p6,
             p7,
             p8,
             p9,
             top = "Simulated Microplastics Dataset (n = 1e5)")
```

# Predict Toxicities
```{r}
#pull in model from DALEX
final_model <- readRDS("final_rf_model.rds")

### parameters ###
df_1 <- Data_final %>% 
  mutate(treatments = 1) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 28) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Juvenile") %>% 
  mutate(bio.org = "organism.group") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality")

df_1_pred <- predict(final_model, newdata = df_1, type = "prob")
df_1_pred_final <- cbind(df_1, df_1_pred) %>% 
  rename("ToxicityProbability" = `1`)

df_1_pred_final %>% ggplot(aes(x = Size, y = ToxicityProbability)) + 
  stat_ecdf(geom = "point", size = 2, color = "blue")+
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.6, size = 1.5) +
  scale_x_log10() +
  scale_y_continuous(labels = scales::percent)+
  labs(title = "Cumulative Toxicity Probability",
       subtitle = "n = 1000 particles/L; Crustacea acute mortality",
       caption = "Predicted Toxicity from Random Forest")
```
HIGHLIGHT AREA UNDER THE CURVE
If we go down to lower number, we would not adequately mimic distribution. Stick to creating 1E5 particles, then change volume. 
Could capture variability for small bumbers by simulating small number of particles

How to calculate uncertainties? 
  Replace effect parameters by median +- SD, then predict tox for each distribution, plot, then if it is minor, FORGET about it! REsiduals in Merel's models were not randomly distributed -  so imperfect models.

What sizes should be used?
  1-5,000 um.

Wait for Merel to update interactions between shape/size/density/count.

What about max size ingest?
  Tox studies typically do not exposure animals to particles bigger than their mouth size opening, so model may not understand that relationship. 
  Could impose bioavailable fraction AFTER model. JUST do predictions for species that have bioavailable fraction data. Could estimate mouth size from prey to predator fraction. JAMS et al has mouth-size opening parameters.

Non-linear regression to get dose-response.

Mattson, Hasselof, Frontiers in Science - measures size distribution.

Final figure: SINGLE toxicity prediction for each dose. Use environmental data. Make SSD. Risk Characterization. The model's SSD is fundamentally different from the SSD made from monodisperse particles, so it is MORE reliable. 
Big assumption: all particle toxicities are ADDITIVE! Depends on effect mechanism - food dilution.

## Dose-response curves

## SSD

## Sensitivity Analysis
it would be great to use the full approach to find out which model parameters and MP subclasses the HC5 is most sensitive to.And also if that HC5 is sensitive to which of Merel’s compartment-specific MP parameterisations is used.

## Shiny App Tabs
1) ML predictor
2) Chemical data
  Bioaccumulation model, partitioning with chemicals/particles. Would add toxicity from chemicals to toxicity from particles! Use same model from Worm paper. Would show how it would compare with toxicities from particles. We have criticial body burden concept (lipids)

### 2021-02-27 Stepwise Multiple Regressions.Rmd
simulated distributions

@@
 knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
-                      warning=FALSE, message=FALSE, echo = FALSE)
+                      warning=FALSE, message=FALSE)
 ```

### 2021-03-09 app.R
Added lvl3 tidying and removed burrowing from lvl2 (was collapsed into locomotion)

@@ -271,7 +271,6 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
     lvl2 == "body.condition"~"Body Condition",
     lvl2 == "boldness"~"Boldness",
     lvl2 == "brain.histo"~"Brain Histological Abnormalities",
-    lvl2 == "burrowing"~"Burrowing",
     lvl2 == "carb.metabolism"~"Carb Metabolism",
     lvl2 == "chemokines.cytokines"~"Chemokines",
     lvl2 == "circulatory"~"Circulatory",
@@ -307,6 +306,530 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
     lvl2 == "shoaling"~"Shoaling",
     lvl2 == "stress"~"Stress",
     lvl2 == "vision.system"~"Vision System"))) %>% #Renames for widget
+  #Level 3 endpoint tidying
+  mutate(lvl3_f = factor(case_when(lvl3 == "11bhsd.mrnaexpression"~"11β HSD mRNA expression",
+                                   lvl3 == "17bhsd.mrnaexpression"~"17β HSD mRNA expression",
+                                   lvl3 == "1433e.mrnaexpression"~"1433e mRNA expression",
+                                   lvl3 == "3bhsd.mrnaexpression"~"3β HSD mRNA expression",
+                                   lvl3 == "5ht1.mrnaexpression"~"5ht1 mRNA expression",
+                                   lvl3 == "aadd45a.mrnaexpression"~"aadd45a mRNA expression",
+                                   lvl3 == "abcb1.mrnaexpression"~"abcb1 mRNA expression",
+                                   lvl3 == "abcc.mrnaexpression"~"abcc mRNA expression",
+                                   lvl3 == "abcc5.mrnaexpression"~"abcc5 mRNA expression",
+                                   lvl3 == "abcd.mrnaexpression"~"abcd mRNA expression",
+                                   lvl3 == "abnormality.occurrence"~"Abnormality Occurrence",
+                                   lvl3 == "abs.rc"~"Absorption Flux/Reaction Center",
+                                   lvl3 == "absorption.efficiency"~"Absorption Efficiency",
+                                   lvl3 == "abundance"~"Abundance",
+                                   lvl3 == "acc1.mrnaexpression"~"acc1 mRNA expression",
+                                   lvl3 == "ace.diversity.index"~"ACE Diversity Index",
+                                   lvl3 == "ache.activity"~"AcHE Activity",
[…]

### 2021-03-09 app.R
Put burrowing back in...definitely needed it and add collapsible tree plot

@@ -28,6 +28,7 @@ library(ggbeeswarm) #plot all points
 library(fitdistrplus) #alt SSD 
 library(ggdark) #dark mode ggplot
 library(ggsci) #color palettes
+library(collapsibleTree) #plot type for endpoint category tree
 #library(bslib) #better themes
 #library(thematic) #complete control over themes (including plots)
 
@@ -271,6 +272,7 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
     lvl2 == "body.condition"~"Body Condition",
     lvl2 == "boldness"~"Boldness",
     lvl2 == "brain.histo"~"Brain Histological Abnormalities",
+    lvl2 == "burrowing"~"Burrowing",
     lvl2 == "carb.metabolism"~"Carb Metabolism",
     lvl2 == "chemokines.cytokines"~"Chemokines",
     lvl2 == "circulatory"~"Circulatory",
@@ -872,6 +874,12 @@ aoc_z <- aoc_setup %>% # start with Heili's altered dataset (no filtration for t
 aoc_z$Species <- as.factor(paste(aoc_setup$genus,aoc_setup$species)) #must make value 'Species" (uppercase)
 aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
 aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
+
+#### Endpoint Category Setup ####
+
+aoc_endpoint <- aoc_setup %>% 
+  group_by(lvl1_f,lvl2_f,lvl3_f,bio_f) %>% 
+  summarise()
 
 #### User Interface ####
 #custom themes (EXPERIMENTAL, COMMENT OUT UNTIL SHINY UPDATES BUGS)
@@ -1632,9 +1640,19 @@ column(width = 12,
                           ) #closes out scott's main panel
                     ), #closes out Scott's tab panel
 
+#### Endpoint Category UI ####
+
+tabPanel("5: Endpoint Categorization", 
+         
+         
+         collapsibleTreeOutput("plot", height = "800px"),
+         
+         
+         ), #closes out tab
+
 #### Resources UI ####
 
-tabPanel("5: Resources", 
+tabPanel("6: Resources", 
          br(),     
[…]

### 2021-03-09 app.R
Adding bio org widget and aesthetic mods

@@ -1643,10 +1643,33 @@ column(width = 12,
 #### Endpoint Category UI ####
 
 tabPanel("5: Endpoint Categorization", 
+         br(),
+         p("This plot displays the categorization of measured endpoints in the database. Nodes correspond to the Broad Endpoint Category (blue), 
+         the Specific Endpoint Category (green), Endpoints (pink) and the level of biological organization (purple). Alternatively, the widget below may be used to select
+         endpoints at various Biological Levels of Organization. Click nodes to expand and collapse the plot."),
+         br(),
          
+         column(width = 12,
          
+         column(width = 3, 
+                pickerInput(inputId = "bio_check_endpoint", # bio org checklist
+                            label = "Level of Biological Organization", 
+                            choices = levels(aoc_endpoint$bio_f),
+                            selected = levels(aoc_endpoint$bio_f),
+                            options = list(`actions-box` = TRUE),
+                            multiple = TRUE)),
+         ), #closes out button column
+         
+         column(width = 12,
+         
+         #Go button
+         column(width = 3,
+                actionButton("go_endpoint", "Update Filters", class = "btn-success")), # adds update action button
+         
+         ), #closes out button column
+         
+         #collapsible tree plot
          collapsibleTreeOutput("plot", height = "800px"),
-         
          
          ), #closes out tab
 
@@ -3343,25 +3366,34 @@ output$downloadSsdPlot <- downloadHandler(
 
   #### Endpoint Category S ####
   
+  aoc_filter_endpoint <- eventReactive(list(input$go_endpoint),{
+  
+  # biological organization widget
+  bio_c_endpoint <- input$bio_check_endpoint # assign bio values to "bio_c"
+  
+  aoc_endpoint %>% # take original dataset
+        filter(bio_f %in% bio_c_endpoint) #filter by bio organization
+  
+  })
+  
   output$plot <- renderCollapsibleTree({
     
-    collapsibleTree(aoc_endpoint, root = "Aquatic Organisms Database", hierarchy = c("lvl1_f", "lvl2_f", "lvl3_f", "bio_f"),
-                    fontSize = 12,    
[…]

### 2021-03-09 app.R
Made exp duration button yellow on exploration tab to match ssd

@@ -1160,12 +1160,12 @@ column(width = 12,
                                          label = "Exposure Duration*:", 
                                          choices = levels(aoc_setup$acute.chronic_f),
                                          selected = levels(aoc_setup$acute.chronic_f),
-                                         options = list(`actions-box` = TRUE), 
+                                         options = list(`actions-box` = TRUE, style = "btn-warning"), 
                                          multiple = TRUE))),
                       column(width = 12,
                              
                       column(width = 3,offset = 9,  
-                             p("*Warning: Exposure duration selections will limit data to the following organism groups: Fish, Molluscs, Crustacea and Algae."))),
+                             p(strong("Warning: Using values other than defaults for yellow widgets will limit data to the following organism groups: Fish, Molluscs, Crustacea and Algae.")))),
 
                     # second row of widget headers
                     column(width=12,

### 2021-03-10 app.R
comments

@@ -28,8 +28,8 @@ library(ggbeeswarm) #plot all points
 library(fitdistrplus) #alt SSD 
 library(ggdark) #dark mode ggplot
 library(ggsci) #color palettes
-#library(bslib) #better themes
-#library(thematic) #complete control over themes (including plots)
+# library(bslib) #better themes. required for dark mode
+# library(thematic) #complete control over themes (including plots) required for dark mode
 
 # Load finalized dataset.
 aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
@@ -351,17 +351,21 @@ aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
 aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
 
 #### User Interface ####
+
 #custom themes (EXPERIMENTAL, COMMENT OUT UNTIL SHINY UPDATES BUGS)
-# light <- bs_theme()
-# dark <- bs_theme(bg = "black", fg = "white", primary = "purple")
+# light <- bs_theme() # DARK MODE SWITCH
+# dark <- bs_theme(bg = "black", fg = "white", primary = "purple") # DARK MODE SWITCH
+
 # build UI
 ui <- fluidPage(theme = shinytheme("flatly"), #light, 
-                
-                  # div(class = "custom-control custom-switch", 
-                  #   tags$input(id = "dark_mode", type = "checkbox", class = "custom-control-input",
-                  #     onclick = HTML("Shiny.setInputValue('dark_mode', document.getElementById('dark_mode').value);")),
-                  #   tags$label("Dark mode", `for` = "dark_mode", class = "custom-control-label")
-                  #   ),  
+
+                # #### DARK MODE SWITCH ####
+                #   div(class = "custom-control custom-switch",
+                #     tags$input(id = "dark_mode", type = "checkbox", class = "custom-control-input",
+                #       onclick = HTML("Shiny.setInputValue('dark_mode', document.getElementById('dark_mode').value);")),
+                #     tags$label("Dark mode", `for` = "dark_mode", class = "custom-control-label")
+                #     ),
+                # ####
   
   # App title
   titlePanel(title=div(img(src = "main_logo.png", width = "10%", height = "10%"), "Toxicity of Microplastics Explorer: Aquatic Organisms")),

### 2021-03-11 Concentration data/Concentration data.R
initial commit for risk characterization rmd

@@ -21,10 +21,11 @@ SFEI <- read.csv("Concentration data/SFEI.csv", stringsAsFactors = TRUE) %>%
 # Adam et al (2019) data needs correcting
 adam <- read.csv("Concentration data/adam2019.csv", na.strings = "N.A.") %>% 
   mutate(x1M = Min.particle.size.um,
-         x2M = Max.particle.size.um)
+         x2M = Max.particle.size.um) %>% 
+  mutate(Sample.Type = "sample")
 
 #function to derive correction factor (CF) from Koelmans et al (equation 2)
-CFfnx = function(a = 1.6, #default alpha from Koelmans et al (2020)
+CFfnx = function(a = 2.5, #default alpha from Koelmans et al (2020)
                  x2D = 5000, #set detault values to convert ranges to (1-5,000 um) #5mm is upper defuault 
                  x1D = 1, #1 um is lower default size
                  x2M, x1M){
@@ -48,7 +49,7 @@ adam <- adam %>%
   mutate(System = factor(System))
 
 #write.csv(adam, "Concentration data/adam.csv") #write file so that I can manually add a column that says "sample"
-adam<- read.csv("Concentration data/adam.csv", stringsAsFactors = TRUE)
+#adam<- read.csv("Concentration data/adam.csv", stringsAsFactors = TRUE)
 
 ##### SSD prep ####
 # Guess_max ensures columns with lots of NAs are not imported as logical vectors, but as numeric/double.
@@ -694,14 +695,62 @@ samplesADAM <- adam %>%
 
 #make new dataframe to plot both histograms together
 sampleSimpleADAM <- samplesADAM %>%
-  select(Conc, Sample.Type) %>% 
+  dplyr::select(Conc, Sample.Type) %>% 
   droplevels()
 
 #make new dataframe to plot both histograms together
 dfADAM <- rbind(sampleSimpleADAM,food.dilution.simple)
 
+##### Environmental Occurence ##### 
+HC5 = 75.6 #Bart's number
 
-#histogram of both tox and concentrations 
+#calculate exceedance
+dfADAM_exceedance <- dfADAM %>% 
+  mutate(aboveHC5 = factor(case_when(
+    Conc >= HC5 ~ "above HC5",
+    Conc < HC5 ~ "below HC5",
+  )))
+
+#give summary stat for exceedance
+exceedance <- dfADAM_exceedance  %>%
[…]

### 2021-03-11 DALEX_tech_QAQC.Rmd
multiple thresholds rmd initial commit

--- 2021-02-24 DALEX.Rmd
@@
     code_folding: hide
-    theme: sandstone
+    theme: journal
     toc: yes
@@
     number_sections: true
+    includes:
+      after_body: footer.html
   word_document:
@@
 knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
-                      warning=FALSE, message=FALSE, echo = FALSE)
-```
-
-```{r message=FALSE, warning=FALSE}
+                      warning=FALSE, message=FALSE,time_it = TRUE) #report time to knit for all chunks
+
+#knit time reporter
+all_times <- list()  # store the time for all chunks in a list
+knitr::knit_hooks$set(time_it = local({
+  now <- NULL
+  function(before, options) {
+    if (before) {
+      now <<- Sys.time()
+    } else {
+      res <- difftime(Sys.time(), now)
+      all_times[[options$label]] <<- res
+    }
+  }
+}))
+```
+
+```{r message=FALSE, warning=FALSE, include=FALSE}
 # load required packages
@@
 ```
-# Setup
-## Data Import
-```{r}
+
+```{r include=FALSE}
 # Load finalized dataset.
@@
 aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
-```
-## Data Selection
+
+# save data object
+#saveRDS(aoc_z, file = "aoc_z.Rda")
+```
+# Data Selection
+
+**Important**: this dataset has ONLY been filtered for 'red' criteria for technical quality.
+
 ```{r}
@@
 ```
+
+```{r final model save}
+#save model
+saveRDS(final_model, file="final_rf_model.rds")
+#load model
+# final_model <- readRDS("final_rf_model.rds")
+```
+
 #### Performance
@@
 
+##### What parameters bias the model?
+Test residuals for size, dose, etc.
+
 # Machine Settings
+Time to knit:
+```{r}
+all_times
+```
 

### 2021-03-13 app.R
Updated size categories to have 1-10 and 10-100 rather than 1-100

@@ -221,11 +221,12 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
   mutate(size_f = factor(case_when(
     size.category == 1 ~ "1nm < 100nm",
     size.category == 2 ~ "100nm < 1µm",
-    size.category == 3 ~ "1µm < 100µm",
+    size.category == 3.1 ~ "1µm < 10µm",
+    size.category == 3.2 ~ "10µm < 100µm",
     size.category == 4 ~ "100µm < 1mm",
     size.category == 5 ~ "1mm < 5mm",
     size.category == 0 ~ "Not Reported"),
-    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
+    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
   # shape category data tidying.
   mutate(shape_f = factor(case_when(
     shape == "fiber" ~ "Fiber",

### 2021-03-17 Scott_Modeling/DALEX.Rmd
Reorganizing things into folders

--- 2021-02-24 DALEX.Rmd
@@
 knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
-                      warning=FALSE, message=FALSE, echo = FALSE)
+                      warning=FALSE, message=FALSE, echo = FALSE,
+                      time_it = TRUE) #report time to knit for all chunks
+
+#knit time reporter
+all_times <- list()  # store the time for all chunks in a list
+knitr::knit_hooks$set(time_it = local({
+  now <- NULL
+  function(before, options) {
+    if (before) {
+      now <<- Sys.time()
+    } else {
+      res <- difftime(Sys.time(), now)
+      all_times[[options$label]] <<- res
+    }
+  }
+}))
 ```
@@
 ## Data Selection
+
+**Important**: this dataset has ONLY been filtered for 'red' criteria for technical quality.
+
 ```{r}
@@
 # Machine Settings
+Time to knit:
+```{r}
+all_times
+```
 

### 2021-03-17 Scott_Modeling/Modelling/simulated distributions.Rmd
Reorganizing things into folders

---
title: "Simulated Probabilistic Distributions"
author: "Scott Coffin"
date: "2/24/2021"
output:   
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
    includes:
     # after_body: footer.html
  word_document:
    toc: yes
---
#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE,time_it = TRUE) #report
```

```{r library}
# Bimodal universal shape distribution for environmental microplastic from Kooi and Koelmans (2019)
library(truncnorm)
library(tidyverse)
library(MonteCarlo)# Monte Carlo Simulation 
library(gridExtra)
library(msm) ## rtnorm - get upper and lower limit of shape distribution
library(GeneralizedHyperbolic) ## normal-inverse Gaussian
library(skimr)
library(drc) #dose-response curves
```

```{r}
# load tox data from RShiny app
aoc_z <- readRDS(file = "aoc_z.Rda")
```

# Resources: 
https://pubs.acs.org/doi/suppl/10.1021/acs.est.0c02982/suppl_file/es0c02982_si_001.pdf https://pubs.acs.org/doi/suppl/10.1021/acs.estlett.9b00379/suppl_file/ez9b00379_si_001.pdf
https://pubs.acs.org/doi/10.1021/acs.estlett.9b00379

# Particles constant, volume different
## Generate Particles

Shape is defined as:
Simplifed equation for corey shape factor:

$CSF = H/sqrt(LW)$

Equation for abundance of particles based on density (two distributions). (Note that this is *equation 4* in Kooi and Koelmans (2019).)

The most abundant shape category of microplastic in water and sediment is fibers (48.5%), followed by fragments (31%), beads (6.5%), films (5.5%), and foam (3.5%)(3). Combining these abundance data with the triangular shape distributions (Table S2 and Figure S2) resulted in a continuous bimodal microplastic shape distribution (Figure 2A). The fitted parameter values for eq 4 were as follows: f1 = 0.06, f2 = 0.94, σ1 = 0.03, σ2 = 0.19, μ1 = 0.08, and μ2 = 0.44 (see Table S5 for standard errors). A Pearson χ2 test indicated that the optimized distribution fits the data well (the fitted model did not differ significantly from the data; p = 0.231 > 0.05). The distribution is dominated by fibers and fragments (CSF = 0.25–0.75) but also has a distinct second peak at a CSF of 0.07, which is mainly attributed to sheets (Figure 2A). The distribution captures the main features of shapes encountered in the environment as well as the relative abundances of these (now continuous) shapes in one go. Most illustrative though is the continuous character of microplastic shape.

$y = f1(\frac{1}{sqrt(2pi()\sigma_1^2})e^{-(x-\mu_1)^2/2\sigma_1^2} + f2(\frac{1}{sqrt(2pi()\sigma_2^2})e^{-(x-\mu_2)^2/2\sigma_2^2}$

For polymer abundance, for the K&K2019 paper I used the fixed values from Burns & Boxall review, and went straight for density. Given these fixed relative abundances, you could calculate polymer type randomly for N particles via:

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
n = 1000
sample(pols, size = n, prob = rel.ab, replace = TRUE)

Particle volume is determined according to shape as follows:

$V = \frac{\pi}{6}L^3CSF^2$

Where CSF = corey shape factor and L = length.

Particle mass is estimated as follows:

$m = pV*\frac{1}{1e12}*1000$

Where *m* is the mass (mg), *p* is density (g/cm^3), *V* is volume (um^3) - which is calculated by the cube of the length (um) of each particle, and additional conversion factors for mg to g (x1000) and cm^3 to um^3 (1e-12).
```{r, include=FALSE}
#rm(list=ls())
n = 10000
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3))
```

```{r}
#prediction paramaters
Data <- Data %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality")
```

```{r}
##++++++++++++++++++++++++++++++++++++++++
## Calculate Volume/Volume and estimate mass/volume
##++++++++++++++++++++++++++++++++++++++++

### 10,000 particles/L
df_1L <- Data %>% 
  mutate(vol.L = 1) %>% 
  mutate(vol.mL = vol.L * 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L)

skim(df_1L)

### 1,000 particles/L
df_10L <- Data %>% 
  mutate(vol.L = 10) %>% 
  mutate(vol.mL = vol.L * 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L)

### 100 particles/L
df_100L <- Data %>% 
  mutate(vol.L = 100) %>% 
  mutate(vol.mL = vol.L * 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L)

### 10 particles/L
df_1000L <- Data %>% 
  mutate(vol.L = 1000) %>% 
  mutate(vol.mL = vol.L * 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L)
```

```{r}
sum_1L <- df_1L %>% group_by(size.category) %>% summarize(particles.L =
                                                       sum(dose.particles.mL.master) * 1000) %>% mutate(volume.L = factor("1L"))
sum_10L <- df_10L %>% group_by(size.category) %>% summarize(particles.L =
                                                       sum(dose.particles.mL.master)* 1000) %>% mutate(volume.L = factor("10L"))
sum_100L <- df_100L %>% group_by(size.category) %>% summarize(particles.L =
                                                       sum(dose.particles.mL.master)* 1000) %>% mutate(volume.L = factor("100L"))
sum_1000L <- df_1000L %>% group_by(size.category) %>% summarize(particles.L =
                                                       sum(dose.particles.mL.master)* 1000) %>% mutate(volume.L = factor("1000L"))

summary_df <- rbind(sum_1L, sum_10L, sum_100L, sum_1000L)
ggplot() +
  geom_col(data = summary_df, aes(x = volume.L, y = particles.L, fill = volume.L)) +
  labs(title = "Particles/L in Simulated Datasets",
       subtitle = "n = 10,000 particles diluted in various liters") +
  theme_minimal(base_size = 20)
```

Parameters:
Shape, density, size, 
why do 3 polymers? Leave polymer out for now. 
For mechanistic point of view- polymer does not matter.

```{r}
##++++++++++++++++++++++++++++++++++++++++
## Quick check results
##++++++++++++++++++++++++++++++++++++++++
p1 <- ggplot(df_1000L, aes(x = Size, fill = size.category)) + geom_histogram(bins = 50, alpha = 0.7) + scale_x_log10()
p2 <- ggplot(df_1000L, aes(x = Shape, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)
p3 <- ggplot(df_1000L, aes(x = Density, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)
p4 <- ggplot(df_1000L, aes(x = polymer, fill = polymer)) + geom_histogram(stat = "count")
#p5 <- ggplot(df_1000L, aes(x = mass.mg, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7 )+ scale_x_log10()
#p6 <- ggplot(df_1000L, aes(x = um3, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
p7 <- ggplot(df_1000L, aes(x = dose.um3.mL.master, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()

p8 <- df_1000L %>% group_by(size.category) %>% summarize(particles.mL =
                                                       sum(dose.particles.mL.master)) %>% 
  ggplot(aes(x = size.category, y = particles.mL, fill = size.category)) + geom_col()

p9 <- ggplot(df_1000L, aes(x = dose.mg.L.master, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
grid.arrange(p1,
             p2,
             p3,
             p4,
             #p5, 
             #p6,
             p7,
             p8,
             #p9,
             top = "Simulated Microplastics Dataset (n = 1e5)",
             ncol = 2)
```

#### Determine which variables are most available
```{r}
#determine which lvl2_f are most abundant for crustacea data
crust <- aoc_z %>% 
  dplyr::select(c(life.stage, bio.org, effect.metric, environment, exposure.route, lvl2_f, exposure.duration.d, treatments, lvl1_f, organism.group, dose.particles.mL.master, dose.mg.L.master, effect_f, effect_10, size.length.um.used.for.conversions, density.mg.um.3, csf)) %>% 
 #filter(treatments == 4) %>% 
  filter(organism.group == "Crustacea") %>% 
  filter(exposure.duration.d == 14) %>% 
  filter(lvl1_f == "Fitness") %>% 
  #filter(life.stage == "Early") %>% 
  filter(bio.org == "organism") %>% 
 # filter(effect.metric == "NOEC") %>% 
  #filter(environment == "Marine") %>% 
  filter(exposure.route == "water") %>% 
 filter(lvl2_f == "Mortality")

skim(crust)
```
```{r}
crust2 <- crust %>% drop_na(effect_f) %>%  mutate_if(~is.numeric(.) && (.) > 0, log10)

#skim(crust2)
summary(glm(data = crust2, effect_10 ~ dose.particles.mL.master + dose.mg.L.master + size.length.um.used.for.conversions, na.action = "na.exclude", family = "binomial"))
```

## Predict Toxicities
```{r}
#pull in model from DALEX
final_model <- readRDS("final_rf_model.rds")

#### prep data frames ####
df_1L_f <- df_1L %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_10L_f <- df_10L %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_100L_f <- df_100L %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_1000L_f <- df_1000L %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

#### Predict Toxicity ####
df_1L_p <- predict(final_model, newdata = df_1L_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_1L_p_f <- cbind(df_1L, df_1L_p)

df_10L_p <- predict(final_model, newdata = df_10L_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_10L_p_f <- cbind(df_10L, df_10L_p)

df_100L_p <- predict(final_model, newdata = df_100L_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_100L_p_f <- cbind(df_100L, df_100L_p)

df_1000L_p <- predict(final_model, newdata = df_1000L_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_1000L_p_f <- cbind(df_1000L, df_1000L_p)

df_pred <- rbind(df_1L_p_f, df_10L_p_f, df_100L_p_f, df_1000L_p_f)

prob_summ <- df_pred %>% mutate(volume.L = as.factor(vol.L)) %>%  group_by(volume.L) %>% summarize(tox = sum(ToxicityProbability))

prob_summ %>% ggplot(aes(x = volume.L, y = tox, fill = volume.L)) +geom_col()

df_pred %>% ggplot(aes(x = factor(vol.L), y = ToxicityProbability)) + 
  stat_ecdf(geom = "point", size = 2, color = "blue")+
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.6, size = 1.5) +
  #scale_x_log10() +
  scale_y_continuous(labels = scales::percent)+
  labs(title = "Cumulative Toxicity Probability",
       subtitle = "n = 1000 particles/L; Crustacea acute Mortality",
       caption = "Predicted Toxicity from Random Forest")
```

# Volume Constant, Particles different
## Generate Particles
### 5 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 5
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_5p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("5p"))
```

### 10 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 10
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_10p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("10p"))
```
### 20 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 20
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_20p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("20p"))
```

### 40 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 40
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_40p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>%  
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("40p"))
```

### 80 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 80
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_80p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("80p"))
```

### 160 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 160
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_160p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("160p"))
```

### 320 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 320
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_320p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
 mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("320p"))
```

### 480 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 480
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_480p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
 mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("480p"))
```

### 640 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 640
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_640p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
 mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>%  
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("640p"))
```

### 1000 particles
```{r, include=FALSE}
#rm(list=ls())
n = 1000
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_1000p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
 mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("1000p"))
```

### 1280 particles
```{r, include=FALSE}
#rm(list=ls())
n = 1280
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_1280p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
 mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("1280p"))
```

Parameters:
Shape, density, size, 
why do 3 polymers? Leave polymer out for now. 
For mechanistic point of view- polymer does not matter.

## Predict Toxicities
```{r}
#pull in model from DALEX
final_model <- readRDS("final_rf_model.rds")

#### prep data frames ####
df_5p_f <- Data_5p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_10p_f <- Data_10p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_20p_f <- Data_20p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_40p_f <- Data_40p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_80p_f <- Data_80p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_160p_f <- Data_160p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_320p_f <- Data_320p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_480p_f <- Data_480p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_640p_f <- Data_640p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_1000p_f <- Data_1000p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_1280p_f <- Data_1280p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

#### Predict Toxicity ####
df_5p_p <- predict(final_model, newdata = df_5p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_5p_p_f <- cbind(df_5p_f, df_5p_p)

df_10p_p <- predict(final_model, newdata = df_10p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_10p_p_f <- cbind(df_10p_f, df_10p_p)

df_20p_p <- predict(final_model, newdata = df_20p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_20p_p_f <- cbind(df_20p_f, df_20p_p)

df_40p_p <- predict(final_model, newdata = df_40p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_40p_p_f <- cbind(df_40p_f, df_40p_p)

df_10p_p <- predict(final_model, newdata = df_10p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_10p_p_f <- cbind(df_10p_f, df_10p_p)

df_80p_p <- predict(final_model, newdata = df_80p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_80p_p_f <- cbind(df_80p_f, df_80p_p)

df_160p_p <- predict(final_model, newdata = df_160p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_160p_p_f <- cbind(df_160p_f, df_160p_p)

df_320p_p <- predict(final_model, newdata = df_320p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_320p_p_f <- cbind(df_320p_f, df_320p_p)

df_480p_p <- predict(final_model, newdata = df_480p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_480p_p_f <- cbind(df_480p_f, df_480p_p)

df_640p_p <- predict(final_model, newdata = df_640p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_640p_p_f <- cbind(df_640p_f, df_640p_p)

df_1000p_p <- predict(final_model, newdata = df_1000p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_1000p_p_f <- cbind(df_1000p_f, df_1000p_p)

df_1280p_p <- predict(final_model, newdata = df_1280p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_1280p_p_f <- cbind(df_1280p_f, df_1280p_p)

df_pred <- rbind(df_5p_p_f, df_10p_p_f, df_20p_p_f, df_40p_p_f, df_80p_p_f, df_160p_p_f, df_320p_p_f, df_480p_p_f, df_640p_p_f, df_1000p_p_f, df_1280p_p_f)

prob_summ <- df_pred %>% group_by(particles.total) %>% summarize(tox = sum(ToxicityProbability))

####convert >1 prob to 1 ###
prob_summ <- prob_summ %>%  mutate(probability = case_when(tox < 1 ~ tox, tox >=1 ~ 1)) %>% 
  mutate(particles = as.numeric(str_sub(particles.total, end = -2))) %>% 
  mutate(log.particles = log(particles))

#plot dose response relationship
prob_summ %>% 
  ggplot(aes(x = particles, y = probability)) +
  geom_point() + 
  stat_smooth(method = drm, method.args = list(fct = LL.4()), se = FALSE) + #logit-4 paramater curve
  scale_x_log10(name = "Particles/mL (1-5,000 um)") +
  scale_y_continuous(labels = scales::percent_format(),
                     name = "Probability of NOEC") +
  labs(title = "Model-Predicted Probability of Toxicity",
       subtitle = "Crustacea; 14-d exp; Mortality; NOEC; Early; Marine",
       caption = "Random Forest Model; Simulated particle distribution (alpha = 1.6)") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# df_pred %>% 
#   group_by(particles.total) %>% 
#   summarize(tox_n = count(ToxicityProbability)) %>% 
#   mutate(cum_tox = cumsum(tox_n)) %>% 
#   ggplot(aes(x = Size, y = cum_tox)) + 
#   geom_step() +
#   facet_wrap(~particles.total)

  #scale_x_log10() +
  # scale_y_continuous(labels = scales::percent)+
  # labs(title = "Cumulative Toxicity Probability",
  #      subtitle = "n = 1000 particles/mL; Crustacea acute Mortality",
  #      caption = "Predicted Toxicity from Random Forest")
```

```{r}
dose_response <- drm(data = prob_summ, probability ~ log.particles, fct = LL.4())
summary(dose_response)
```

```{r}
summary(glm(ToxicityProbability ~ Size * Shape * polymer * um3, data = df_640p_p_f))
p1 <- df_640p_p_f %>% ggplot(aes(x = 10^Size, y = ToxicityProbability, color = polymer)) + geom_point()

p2 <- df_640p_p_f %>% ggplot(aes(x = ToxicityProbability, fill = size.category)) + geom_histogram()
grid.arrange(p1,p2)
```

```{r}
#test toxicity prediction with a single row

dataFrame <- data.frame(Doubles=double(),
                 Ints=integer(),
                 Factors=factor(),
                 Logicals=logical(),
                 Characters=character(),
                 stringsAsFactors=FALSE)
newrow =data.frame(dose.particles.mL.master = 1)
# Compare with monodisperse prediction
monoDisperse_100 <- newrow %>%  
  mutate(dose.particles.mL.master = 120000000) %>%  #120 particles/mL is about EC50 for above
  mutate(mass.mg = 0.2263657 * dose.particles.mL.master) %>% #average mass x  particles
  mutate(um3 = 50) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("120p"))

predict(final_model, newdata = monoDisperse_100, type = "prob")
```
### Uniform Distribution

It may be possible to visualize discrete relationships of toxicity by creating a uniform distribution of particles within the parameter space of the training dataset, then plotting the predicted toxicities for different predictors.

```{r}
###### determine bounds for training dataset #####
##NOTE: WILL NOT WORK IN MARKDOWN UNLESS WE PULL IN CODE FOR AOC FROM OTHER SCRIPT ###

#first extract 
particleCharacteristics <- aoc_z %>% 
  dplyr::select(c(mass.per.particle.mg, polymer, shape, density.mg.um.3, size.length.um.used.for.conversions,
                  particle.volume.um3, csf))

skim(particleCharacteristics)
```

#### Generate particles
```{r}

n = 10000

#Create uniform characteristics for base particle characteristics (size, shape, density)
csf <- runif(n, min = 0.01, 0.99)
#mass.per.particle.mg <- runif(n, min = 0, max = 1.65e+1)
density.mg.um.3 <- runif(n, min = 8.75e-10, max = 1.53e-9)
size.length.um.used.for.conversions <- runif(n, min = 1, max = 1000)

#create uniform distribution of polymer types
pols <- c("PE", "PP", "PS", "PVC", "PET")
rel.ab <- c(0.2, 0.2, 0.2, 0.2, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)

#create data frame with uniform distirbutions
uniform <- data.frame(cbind(csf, density.mg.um.3, size.length.um.used.for.conversions)) %>%
  mutate_all(funs(as.numeric(as.character(.))))
#bind polymer and convert to factor
uniform <- cbind(uniform, polymer) %>% mutate_if(is.character, as.factor)

#calculate higher order parameters and assign organism traits
unif_df <- uniform %>%  
  #particle characteristics
  mutate(size.category = factor(case_when(
    size.length.um.used.for.conversions < 10 & size.length.um.used.for.conversions >= 1 ~ "1µm < 10µm",
    size.length.um.used.for.conversions < 100 & size.length.um.used.for.conversions >= 10 ~ "10µm < 100µm",
    size.length.um.used.for.conversions < 1000 & size.length.um.used.for.conversions >= 100 ~ "100µm < 1mm",
    size.length.um.used.for.conversions < 5000 & size.length.um.used.for.conversions >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.per.particle.mg = density.mg.um.3 * size.length.um.used.for.conversions ^ 3 * 1E-9) %>% 
  mutate(particle.volume.um3 = pi/6 * (size.length.um.used.for.conversions ^ 3) * csf ^2 ) %>% #equation 4
  #volume
  mutate(vol.mL = 0.0001) %>% 
  mutate(vol.L = vol.mL / 1000) %>%  
  mutate(dose.um3.mL.master = particle.volume.um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.per.particle.mg / vol.L) %>% 
    #organism/test system characteristics
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality") %>% 
  mutate_if(is.character, as.factor)

skim(unif_df)
```

##### Inspect distributions of calculated parameters
```{r}
#create a 2 x 3 plotting matrix
par(mfrow = c(3,3))
#plot
hist(unif_df$csf)
hist(unif_df$size.length.um.used.for.conversions)
hist(unif_df$density.mg.um.3)
hist(unif_df$dose.mg.L.master)
hist(unif_df$dose.particles.mL.master)
hist(unif_df$dose.um3.mL.master)
hist(unif_df$mass.per.particle.mg)
hist(unif_df$particle.volume.um3)
#ggplot(unif_df, aes(x = size.category, y =..count..)) + geom_histogram(stat = "count")
```

##### Predict toxicities on uniform distribution of particles
```{r}
#log transform numeric variables first
unif_df <- unif_df %>%  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

#predict tox
unif_pred <- predict(final_model, newdata = unif_df, type = "prob") %>%  rename("ToxicityProbability" = `1`)
uniform_pred <- cbind(unif_pred, unif_df)

#summary(glm(ToxicityProbability ~ Size * Shape * polymer * um3, data = df_640p_p_f))
p1 <- uniform_pred %>% ggplot(aes(x = dose.particles.mL.master, y = ToxicityProbability, color = polymer)) + geom_jitter(alpha = 0.3)

p2 <- uniform_pred %>% ggplot(aes(x = dose.mg.L.master, y = ToxicityProbability, color = polymer)) + geom_point()

p3 <- uniform_pred %>% ggplot(aes(x = dose.um3.mL.master, y = ToxicityProbability, color = polymer)) + geom_point()

p4 <- uniform_pred %>% ggplot(aes(x = ToxicityProbability, fill = polymer)) + geom_histogram()

p5 <- uniform_pred %>%  ggplot(aes(x = size.length.um.used.for.conversions, y = ToxicityProbability, color = csf)) + geom_point()

p6 <- uniform_pred %>%  ggplot(aes(x = csf, y = ToxicityProbability, color = size.length.um.used.for.conversions)) + geom_point()

p7 <- uniform_pred %>%  ggplot(aes(x = particle.volume.um3, y = ToxicityProbability, color =
                                     size.length.um.used.for.conversions)) + geom_point()

p8 <- uniform_pred %>%  ggplot(aes(x = mass.per.particle.mg, y = ToxicityProbability, color =
                                     size.length.um.used.for.conversions)) + geom_point()

p9 <- uniform_pred %>%  ggplot(aes(x = density.mg.um.3, y = ToxicityProbability, color =
                                     size.length.um.used.for.conversions)) + geom_point()

grid.arrange(p1, p2, p3, p4, ncol = 2)
```
```{r}
grid.arrange(p5, p6, p7, p8, p9, ncol = 2)
```
```{r}
summary(lm(ToxicityProbability ~ size.length.um.used.for.conversions * csf * particle.volume.um3, data = uniform_pred))
```

HIGHLIGHT AREA UNDER THE CURVE
If we go down to lower number, we would not adequately mimic distribution. Stick to creating 1E5 particles, then change volume. 
Could capture variability for small bumbers by simulating small number of particles

How to calculate uncertainties? 
  Replace effect parameters by median +- SD, then predict tox for each distribution, plot, then if it is minor, FORGET about it! REsiduals in Merel's models were not randomly distributed -  so imperfect models.

What sizes should be used?
  1-5,000 um.

Wait for Merel to update interactions between shape/size/density/count.

What about max size ingest?
  Tox studies typically do not exposure animals to particles bigger than their mouth size opening, so model may not understand that relationship. 
  Could impose bioavailable fraction AFTER model. JUST do predictions for species that have bioavailable fraction data. Could estimate mouth size from prey to predator fraction. JAMS et al has mouth-size opening parameters.

Non-linear regression to get dose-response.

Mattson, Hasselof, Frontiers in Science - measures size distribution.

Final figure: SINGLE toxicity prediction for each dose. Use environmental data. Make SSD. Risk Characterization. The model's SSD is fundamentally different from the SSD made from monodisperse particles, so it is MORE reliable. 
Big assumption: all particle toxicities are ADDITIVE! Depends on effect mechanism - food dilution.

# Dose-response curves

# SSD

# Sensitivity Analysis
it would be great to use the full approach to find out which model parameters and MP subclasses the HC5 is most sensitive to.And also if that HC5 is sensitive to which of Merel’s compartment-specific MP parameterisations is used.

# Shiny App Tabs
1) ML predictor
2) Chemical data
  Bioaccumulation model, partitioning with chemicals/particles. Would add toxicity from chemicals to toxicity from particles! Use same model from Worm paper. Would show how it would compare with toxicities from particles. We have criticial body burden concept (lipids)

### 2021-03-17 Leah_Dev_Files/Concentration_Range_SSD.R

@@ -1,0 +1,746 @@
+#Title: Concentration Range by Size Fraction for SSD
+#Author: Leah Thornton Hampton
+#Created on 3/16/21
+
+##### Load Packages ####
+
+library(tidyverse)
+library(ggpubr)
+
+##### Import Data ####
+
+aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
+
+#### Tidying ####
+
+# Master dataset for scatterplots - for Heili's tab.
+aoc_v1 <- aoc %>% # start with original dataset
+  # full dataset filters.
+  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
+                                     effect == "N" ~ "No"),
+                           levels = c("No", "Yes"))) %>%
+  # removing NAs to make data set nicer
+  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported"))
+
+aoc_setup <- aoc_v1 %>% # start with original dataset
+  mutate(size_f = factor(case_when(
+    size.category == 1 ~ "1nm < 100nm",
+    size.category == 2 ~ "100nm < 1µm",
+    size.category == 3.1 ~ "1µm < 10µm",
+    size.category == 3.2 ~ "10µm < 100µm",
+    size.category == 4 ~ "100µm < 5mm",
+    size.category == 5 ~ "100µm < 5mm",
+    size.category == 0 ~ "Not Reported"),
+    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
+  # shape category data tidying.
+  mutate(shape_f = factor(case_when(
+    shape == "fiber" ~ "Fiber",
+    shape == "fragment" ~ "Fragment",
+    shape == "sphere" ~ "Sphere",
+    shape == "Not Reported" ~ "Not Reported"),
+    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% # order our different shapes.
+  # polymer category data tidying.
+  mutate(poly_f = factor(case_when(
+    polymer == "BIO" ~ "Biopolymer",
+    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
+    polymer == "LTX" ~ "Latex",
[…]

### 2021-03-17 Leah_Dev_Files/Presentation_Figures.R

@@ -144,17 +144,18 @@ aoc_work <- aoc %>%
 Fiber <- aoc_work %>%
   filter(shape_f == "Fiber") %>%
   filter(org_f != "Annelida") %>%
-  filter(life_f == "Early") %>% 
-  ggplot(aes(x = dose.particles.mL.master, y = org_f)) +
+  filter(effect_f != "NA") %>% 
+  # filter(life_f == "Early") %>% 
+  ggplot(aes(x = dose.particles.mL.master, y = org_f, fill = effect_f, color = effect_f)) +
   # geom_rect(alpha = 0.7, ymin = 0, ymax = Inf, xmin = log10(0), xmax = log10(0.03929), color = "black", fill = "gray87") + #Total particles SFEI grab sample range
   # geom_rect(alpha = 0.1, ymin = 0, ymax = Inf, xmin = log10(0), xmax = log10(0.03368), color = "black", fill = "mediumpurple1") + #Total fibers SFEI grab sample range
-  geom_boxplot(alpha = 0.8, show.legend = TRUE, aes(color = effect_f, fill = effect_f,)) +
+  geom_jitter(alpha = 0.8, size = 5, height = .2, width = .1) +
   scale_fill_manual(values = c("royalblue", "darkcyan"))+
-  scale_color_manual(values = c("black", "black"))+
+  scale_color_manual(values = c("royalblue", "darkcyan"))+
   theme_classic() +
   scale_x_log10() +
   theme(text = element_text(size = 18),
-        legend.position="none", 
+        legend.position="right", 
         legend.title = element_blank(),
         axis.text = element_text(color = "black")) +
   labs(x = "Concentration (particles/mL)", y = "Organism Group")

### 2021-03-17 Leah_Dev_Files/Team_X.R

@@ -150,15 +150,44 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
                                              risk.tier.zero == "Pass" ~ "Red Criteria Passed")))
 
 
+
+#### Correlation of Particle Mass to Volume ####
+
+VolumevMass <- aoc_setup %>%
+  mutate(mass_ug = (mass.per.particle.mg*1000)) %>%
+  group_by(particle.volume.um3, mass_ug) %>% 
+  summarise() %>% 
+  drop_na() %>% 
+  ggplot(aes(x = mass_ug, y = particle.volume.um3)) +
+  geom_point(color = "darkcyan", alpha = 0.5, size = 5, position = "jitter") + 
+  theme_bw()+
+  theme(text = element_text(size=14))+
+  scale_x_log10() +
+  scale_y_log10() +
+  labs(title = "Particle Mass (µg) by Particle Volume (µm^3)", x = "Particle Mass (µg)", y = "Particle Volume (µm^3)") +
+  stat_smooth(method = "lm", col = "black")
+
+plot(VolumevMass)
+
+x <- aoc_setup %>%
+  mutate(mass_ug = (mass.per.particle.mg*1000)) %>%
+  group_by(particle.volume.um3, mass_ug) %>% 
+  summarise() %>% 
+  drop_na() %>% 
+  lm(formula = log10(mass_ug) ~ log10(particle.volume.um3))
+
+summary(x)
+
 #### Particle Volume Histogram ####
 
 Volume <- aoc_setup %>% 
   drop_na(particle.volume.um3) %>% 
   ggplot(aes(x = particle.volume.um3)) + 
-  geom_histogram(bins = 100, color = "darkmagenta", fill = "darkmagenta")+
+  geom_histogram(bins = 50, color = "darkmagenta", fill = "darkmagenta")+
   scale_x_log10()+
   theme_minimal()+
   theme(text = element_text(size = 14))+
+  ylim(0,500)+
   labs(title = "Distribution of Particle Volume",
        subtitle = "Aquatic Organisms Database",
        x = "Particle Volume (µm^3)",
@@ -173,10 +202,11 @@ Mass <- aoc_setup %>%
   drop_na(mass.per.particle.mg) %>% 
   mutate(mass_ug = (mass.per.particle.mg*1000)) %>% 
   ggplot(aes(x = mass_ug)) + 
[…]

### 2021-04-07 app.R
Removed old warning labels

@@ -1173,10 +1173,6 @@ column(width = 12,
                                          selected = levels(aoc_setup$acute.chronic_f),
                                          options = list(`actions-box` = TRUE, style = "btn-warning"), 
                                          multiple = TRUE))),
-                      column(width = 12,
-                             
-                      column(width = 3,offset = 9,  
-                             p(strong("Warning: Using values other than defaults for yellow widgets will limit data to the following organism groups: Fish, Molluscs, Crustacea and Algae.")))),
 
                     # second row of widget headers
                     column(width=12,
@@ -1322,7 +1318,7 @@ column(width = 12,
                     br(),
                     p("Use the options below to filter the toxicity thresholds dataset. Once complete, hit the 'submit' button"),
                     br(),
-                    p(strong("Warning: Using values other than defaults for yellow widgets will limit data to the following organism groups: Fish, Molluscs, Crustacea and Algae.")),
+                    
  
                     # Row 1 of Headers
                     column(width=12,

### 2021-04-08 app.R
Fixed some minor bugs and added AF descriptions to resources tab

@@ -84,12 +84,13 @@ sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
 sizef<-as.data.frame(sizedf)%>%
   filter(effect %in% c("Y","N"))%>%
   mutate(size.category = case_when(
-    size.category == 1 ~ "<1µm",
-    size.category == 2 ~ "1µm < 10µm",
-    size.category == 3 ~ "10µm < 100µm",
+    size.category == 1 ~ "1nm < 100nm",
+    size.category == 2 ~ "100nm < 1µm",
+    size.category == 3.1 ~ "1µm < 10µm",
+    size.category == 3.2 ~ "10µm < 100µm",
     size.category == 4 ~ "100µm < 1mm",
     size.category == 5 ~ "1mm < 5mm",
-    size.category == 0 ~ "unavailable"))%>%
+    size.category == 0 ~ "Not Reported"))%>% 
   rename(Type = "size.category")%>%
   mutate_if(is.numeric, round,0)%>%
   mutate(plot="Size")
@@ -911,7 +912,7 @@ ui <- fluidPage(theme = shinytheme("flatly"), #light,
   # Title panel subtext
   tags$div("Logo created by J.C. Leapman.", tags$a(href="https://twitter.com/jcleapman", tags$img(src="twitter.png", width="2%", height="2%"))),
   tags$div("This website is only intended for use by invited participants of the Microplastics Health Effects Workshop."),
-  actionButton("database_link", label="Go to Mammalian Database", class = "btn-primary", onclick ="window.open('https://sccwrp.shinyapps.io/human_mp_tox_shiny-/', '_blank')", style = "float:right"),
+  actionButton("database_link", label="Go to Human Health Database", class = "btn-primary", onclick ="window.open('https://sccwrp.shinyapps.io/human_mp_tox_shiny-/', '_blank')", style = "float:right"),
   
   br(), # line break
   
@@ -1054,7 +1055,7 @@ column(width = 12,
                     p("Each figure displays a different metric along the y-axis - organism group, broad endpoint category, specific endpoint category, size, shape, 
                       and polymer, respectively.The values in the parentheses represent the number of measurements and studies, respectively, of each metric along the y-axis."),
                     br(),
-                    p("The data displayed in these figures are not filtered for quality and only display data from in vitro studies or in vivo studies where doses were reported 
+                    p("The data displayed in these figures only display data from in vitro studies or in vivo studies where doses were reported 
                     as mass or counts per volume - other dosing units (e.g., particle mass/food mass) 
[…]

### 2021-04-08 Shiny_Manuscript_Figures/Manuscript_Figures.Rmd
scoring plots and started script for generating manuscript figures

---
title: 'Shiny App Manuscript Figures: Aquatic Organisms'
author: "Leah Thornton Hampton"
date: "4/2/2021"
output:
  html_document: 
    toc: true
    toc_depth: 4
    number_sections: true
    toc_float: true
---

# Data Import and Tidying

```{r Data Import and Packages, echo = FALSE, results = 'hide', error = FALSE, warning = FALSE, message = FALSE}

#Import packages
library(tidyverse)

#Import Data
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)

#Data Tidying
aoc_v1 <- aoc %>% # start with original dataset
   # full dataset filters.
  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
    effect == "N" ~ "No"),
    levels = c("No", "Yes"))) %>%
  # removing NAs to make data set nicer
  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported", chem.exp.typ.nominal = "Particle Only"))

aoc_setup <- aoc_v1 %>% # start with original dataset
  mutate(size_f = factor(case_when(
    size.category == 1 ~ "1nm < 100nm",
    size.category == 2 ~ "100nm < 1µm",
    size.category == 3.1 ~ "1µm < 10µm",
    size.category == 3.2 ~ "10µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
  # shape category data tidying.
  mutate(shape_f = factor(case_when(
    shape == "fiber" ~ "Fiber",
    shape == "fragment" ~ "Fragment",
    shape == "sphere" ~ "Sphere",
    shape == "Not Reported" ~ "Not Reported"),
    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% # order our different shapes.
  # polymer category data tidying.
  mutate(poly_f = factor(case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurethane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid",
    polymer == "Not Reported" ~ "Not Reported"))) %>%
  # taxonomic category data tidying.
  mutate(org_f = factor(organism.group, levels = c("Algae", "Annelida", "Bacterium", "Cnidaria", "Crustacea",
                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% # order our different organisms.
  mutate(lvl1_f = factor(case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
    lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    lvl1 == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    lvl1 == "community" ~ "Community",
    lvl1 == "fitness" ~ "Fitness",
    lvl1 == "immune" ~ "Immune",
    lvl1 == "metabolism" ~ "Metabolism",
    lvl1 == "microbiome" ~ "Microbiome",
    lvl1 == "stress" ~ "Stress"))) %>% # creates new column with nicer names.
  # Level 2 Data tidying
  mutate(lvl2_f = factor(case_when(lvl2 == "abundance"~"Abundance",
    lvl2 == "actinobacteria" ~ "Actinobacteria",
    lvl2 == "aggressivity"~"Agressivity",
    lvl2 == "ammonia.excretion" ~ "Ammonia Excretion",
    lvl2 == "bacteroidetes"~ "Bacteriodetes",
    lvl2 == "blood"~"Blood",
    lvl2 == "body.condition"~"Body Condition",
    lvl2 == "boldness"~"Boldness",
    lvl2 == "brain.histo"~"Brain Histological Abnormalities",
    lvl2 == "burrowing"~"Burrowing",
    lvl2 == "carb.metabolism"~"Carb Metabolism",
    lvl2 == "chemokines.cytokines"~"Chemokines",
    lvl2 == "circulatory"~"Circulatory",
    lvl2 == "detoxification"~"Detoxification",
    lvl2 == "development"~"Development",
    lvl2 == "digestion"~"Digestion",
    lvl2 == "digestive.enzymes"~"Digestive Enzymes",
    lvl2 == "digestive.tract.histo"~"Digestive Tract Histological Abnormalities",
    lvl2 == "diversity"~ "Diversity",
    lvl2 == "feeding"~ "Feeding",
    lvl2 == "firmicutes"~ "Firmicutes",
    lvl2 == "gall.bladder.histo" ~ "Gall Bladder Histological Abnormalities",
    lvl2 == "gen.metabolism"~ "General Metabolism",
    lvl2 == "gill.histo"~ "Gill Histological Abnormalities",
    lvl2 == "gonad.histo"~"Gonad Histological Abnormalities",
    lvl2 == "growth"~ "Growth",
    lvl2 == "immune.cells"~"Immune Cells",
    lvl2 == "immune.other"~"Immune Other ",
    lvl2 == "intestinal.permeability"~"Intestinal Permeability",
    lvl2 == "kidney.histo"~"Kidney Histological abnormalities",
    lvl2 == "lipid.metabolism"~"Lipid Metabolism",
    lvl2 == "liver.histo"~"Liver Histological Abnormalities",
    lvl2 == "liver.kidney.products" ~ "Liver and Kidney Products",
    lvl2 == "locomotion"~"Locomotion",
    lvl2 == "mortality"~"Mortality",
    lvl2 == "nervous.system"~"Nervous System",
    lvl2 == "oxidative.stress"~"Oxidative Stress",
    lvl2 == "photosynthesis"~ "Photosynthesis",
    lvl2 == "proteobacteria"~"Protebacteria",
    lvl2 == "reproduction"~"Reproduction",
    lvl2 == "respiration"~"Respiration",
    lvl2 == "sexhormones"~"Sex Hormones",
    lvl2 == "shoaling"~"Shoaling",
    lvl2 == "stress"~"Stress",
    lvl2 == "vision.system"~"Vision System"))) %>% #Renames for widget
  #Level 3 endpoint tidying
  mutate(lvl3_f = factor(case_when(lvl3 == "11bhsd.mrnaexpression"~"11β HSD mRNA expression",
                                   lvl3 == "17bhsd.mrnaexpression"~"17β HSD mRNA expression",
                                   lvl3 == "1433e.mrnaexpression"~"1433e mRNA expression",
                                   lvl3 == "3bhsd.mrnaexpression"~"3β HSD mRNA expression",
                                   lvl3 == "5ht1.mrnaexpression"~"5ht1 mRNA expression",
                                   lvl3 == "aadd45a.mrnaexpression"~"aadd45a mRNA expression",
                                   lvl3 == "abcb1.mrnaexpression"~"abcb1 mRNA expression",
                                   lvl3 == "abcc.mrnaexpression"~"abcc mRNA expression",
                                   lvl3 == "abcc5.mrnaexpression"~"abcc5 mRNA expression",
                                   lvl3 == "abcd.mrnaexpression"~"abcd mRNA expression",
                                   lvl3 == "abnormality.occurrence"~"Abnormality Occurrence",
                                   lvl3 == "abs.rc"~"Absorption Flux/Reaction Center",
                                   lvl3 == "absorption.efficiency"~"Absorption Efficiency",
                                   lvl3 == "abundance"~"Abundance",
                                   lvl3 == "acc1.mrnaexpression"~"acc1 mRNA expression",
                                   lvl3 == "ace.diversity.index"~"ACE Diversity Index",
                                   lvl3 == "ache.activity"~"AcHE Activity",
                                   lvl3 == "ache.mrnaexpression"~"AcHE mRNA expression",
                                   lvl3 == "acid.phosphatase.activity"~"Acid Phosphatase Activity",
                                   lvl3 == "aco.mrnaexpression"~"aco mRNAexpression",
                                   lvl3 == "acrosomal.integrity"~"Acrosomal Integrity",
                                   lvl3 == "actinobacteria"~"Actinobacteria",
                                   lvl3 == "activity"~"Activity",
                                   lvl3 == "activity.during.feeding"~"Activity During Feeding",
                                   lvl3 == "aeromonas"~"Aeromonas",
                                   lvl3 == "agal.consumption"~"Agal Consumption",
                                   lvl3 == "agression.index"~"Aggression Index",
                                   lvl3 == "ahra.mrnaexpression"~"ahrα mRNA expression",
                                   lvl3 == "ahrb.mrnaexpression"~"ahrβ mRNA expression",
                                   lvl3 == "akt.mrnaexpression"~"akt mRNA expression",
                                   lvl3 == "albumin.con"~"Albumin Concentration",
                                   lvl3 == "albumin.globulin.ratio"~"Albumin to Globulin Ratio",
                                   lvl3 == "algal.density"~"Algal Density",
                                   lvl3 == "alkaline.phosphatase.activity"~"Alkaline Phosphatase Activity",
                                   lvl3 == "aproteobacteria"~"α proteobacteria",
                                   lvl3 == "alp.activity"~"Alkaline Phosphatase Activity",
                                   lvl3 == "alp.con"~"Alkaline Phosphatase Concentration",
                                   lvl3 == "alt.activity"~"Alanine Aminotransferase Activity",
                                   lvl3 == "alt.con"~"Alanine Aminotransferase Concentration",
                                   lvl3 == "amino.acid.trans"~"Amino Acid Transport",
                                   lvl3 == "ammonia.con"~"Ammonia Concentration",
                                   lvl3 == "ampka.mrnaexpression"~"ampka mRNA expression",
                                   lvl3 == "ampkb.mrnaexpression"~"ampkb mRNA expression",
                                   lvl3 == "ampkc.mrnaexpression"~"ampkc mRNA expression",
                                   lvl3 == "amylase.mrnaexpression"~"Amylanse mRNA expression",
                                   lvl3 == "antioxidant.act"~"Antioxidant Activity",
                                   lvl3 == "aox.activity"~"Alternative Oxidase Activity",
                                   lvl3 == "apo.mrnaexpression"~"Apolipoprotein mRNA expression",
                                   lvl3 == "ar.mrnaexpression"~"Androgen Receptor mRNA expression",
                                   lvl3 == "ara.mrnaexpression"~"Androgen Receptor α mRNA expression",
                                   lvl3 == "arb.mrnaexpression"~"Androgen Receptor β mRNA expression",
                                   lvl3 == "area"~"Area",
                                   lvl3 == "area.traveled"~"Area Traveled",
                                   lvl3 == "aspartate.aminotransferase.con"~"Aspartate Aminotransferase Concentration",
                                   lvl3 == "assemblage"~"Assemblage",
                                   lvl3 == "assimilation.efficiency"~"Assimilation Efficiency",
                                   lvl3 == "ast.activity"~"Aspartate Aminotransferase Activity",
                                   lvl3 == "ast.con"~"Aspartate Aminotransferase Concentration",
                                   lvl3 == "attachment"~"Attachment",
                                   lvl3 == "atubulin.mrnaexpression"~"α tubulin mRNA expression",
                                   lvl3 == "atubulin.proteinexpression"~"α tubulin Protein expression",
                                   lvl3 == "bacteroidetes"~"Bacteroidetes",
                                   lvl3 == "bcl2.mrnaexpression"~"bcl2 mRNA expression",
                                   lvl3 == "bfcod.activity"~"BFCOD Activity",
                                   lvl3 == "bile.acid.con"~"Bile Acid Concentration",
                                   lvl3 == "bilirubin.con"~"Bilirubin Concentration",
                                   lvl3 == "biomass"~"Biomass",
                                   lvl3 == "bioturbation"~"Bioturbation",
                                   lvl3 == "body.length"~"Body Length",
                                   lvl3 == "body.mass"~"Body Mass",
                                   lvl3 == "body.width"~"Body Width",
                                   lvl3 == "bproteobacteria"~"βproteobacteria",
                                   lvl3 == "brain.histo"~"Brain Histology",
                                   lvl3 == "brain.mass.index"~"Brain Mass Index",
                                   lvl3 == "brain.water.content"~"Brain Water Content",
                                   lvl3 == "burrow.length"~"Burrow Length",
                                   lvl3 == "burrow.litter"~"Burrow Litter",
                                   lvl3 == "burrow.no"~"Burrow Number",
                                   lvl3 == "burrow.vol"~"Burrow Volume",
                                   lvl3 == "burrow.walldensity"~"Burrow Wall Density",
                                   lvl3 == "burrow.wallweight"~"Burrow Wall Weight",
                                   lvl3 == "burrowing"~"Burrowing",
                                   lvl3 == "ca.mrnaexpression"~"ca mRNA expression",
                                   lvl3 == "calcification"~"Calcification",
                                   lvl3 == "calcium.con"~"Calcium Concentration",
                                   lvl3 == "capture.rate"~"Capture Rate",
                                   lvl3 == "carbohydrate.reserves"~"Carbohydrate Reserves",
                                   lvl3 == "cardiolipin.con"~"Cardiolipin Concentration",
                                   lvl3 == "cas8.mrnaexpression"~"cas8 mRNA expression",
                                   lvl3 == "casp3.mrnaexpression"~"casp3 mRNA expression",
                                   lvl3 == "casp373.mrnaexpression"~"casp373 mRNA expression",
                                   lvl3 == "casp3a.mrnaexpression"~"casp3a mRNA expression",
                                   lvl3 == "casp8.mrnaexpression"~"casp8 mRNA expression",
                                   lvl3 == "casp9.mrnaexpression"~"casp9 mRNA expression",
                                   lvl3 == "caspase3.7.mrnaexpression"~"caspase 3/7 mRNA expression",
                                   lvl3 == "cat.activity"~"Catalase Activity",
                                   lvl3 == "cat.mrnaexpression"~"Catalase mRNA expression",
                                   lvl3 == "cell.autoflourescence"~"Cell Autoflourescence",
                                   lvl3 == "cell.density"~"Cell Density",
                                   lvl3 == "cell.granularity"~"Cell Granularity",
                                   lvl3 == "cell.morphology"~"Cell Morphology",
                                   lvl3 == "cell.necrosis"~"Cell Necrosis",
                                   lvl3 == "cell.size"~"Cell Size",
                                   lvl3 == "cell.viability"~"Cell Viability",
                                   lvl3 == "cellular.energy.allocation"~"Cellular Energy Allocation",
                                   lvl3 == "chg.mrnaexpression"~"Choriogenin mRNA expression",
                                   lvl3 == "chgh.mrnaexpression"~"Choriogenin H mRNA expression",
                                   lvl3 == "chgl.mrnaexpression"~"Choriogenin L mRNA expression",
                                   lvl3 == "chlorophyll"~"Chlorophyll",
                                   lvl3 == "cholesterol.con"~"Cholesterol Concentration",
                                   lvl3 == "chrna1.mrnaexpression"~"chrna1 mRNA expression",
                                   lvl3 == "chymotrypsin.activity"~"Chymotrypsin Activity",
                                   lvl3 == "clearance.rate"~"Clearance Rate",
                                   lvl3 == "clutch.num"~"Number of Clutches",
                                   lvl3 == "clutch.size"~"Clutch Size",
                                   lvl3 == "clutch.size.first"~"First Clutch Size",
                                   lvl3 == "clutch.size.firstthree"~"First Three Clutch Sizes",
                                   lvl3 == "co2.consumption"~"CO2 Consumption",
                                   lvl3 == "condition.index"~"Condition Index",
                                   lvl3 == "congestion"~"Congestion",
                                   lvl3 == "copper.con.hepatopancreas"~"Hepatopancreas Copper Concentration",
                                   lvl3 == "cox.activity"~"Cyclooxygenase Activity",
                                   lvl3 == "coxIV.mrnaexpression"~"Cyclooxygenase IV mRNA expression",
                                   lvl3 == "cpt1.mrnaexpression"~"CPT1 mRNA expression",
                                   lvl3 == "creatine.kinase.con"~"Creatine Kinase Concentration",
                                   lvl3 == "crude.ash"~"Crude Ash",
                                   lvl3 == "crude.protein"~"Curde Protein",
                                   lvl3 == "cs.mrnaexpression"~"Chitin Synthase mRNA expression",
                                   lvl3 == "ctsl.mrnaexpression"~"Cathepsin-L mRNA expression",
                                   lvl3 == "cuznsod.mrnaexpression"~"CuZn Superoxide Dismutase mRNA expression",
                                   lvl3 == "cxcr5.mrnaexpression"~"cxcr5 mRNA expression",
                                   lvl3 == "cyp11.mrnaexpression"~"cyp11 mRNA expression",
                                   lvl3 == "cyp11a2.mrnaexpression"~"cyp11a2 mRNA expression",
                                   lvl3 == "cyp17a1.mrnaexpression"~"cyp17a1 mRNA expression",
                                   lvl3 == "cyp19a.mrnaexpression"~"cyp19a mRNA expression",
                                   lvl3 == "cyp19a2.mrnaexpression"~"cyp19a2 mRNA expression",
                                   lvl3 == "cyp19b.mrnaexpression"~"cyp19b mRNA expression",
                                   lvl3 == "cyp1a.mrnaexpression"~"cyp1a  mRNA expression",
                                   lvl3 == "cyp1a1.mrnaexpression"~"cyp1a1 mRNA expression",
                                   lvl3 == "cyp32.mrnaexpression"~"cyp32 mRNA expression",
                                   lvl3 == "cyp450.proteinexpression"~"cyp450 mRNA expression",
                                   lvl3 == "cyp4an1.mrnaexpression"~"cyp4an1 mRNA expression",
                                   lvl3 == "cyp4c33.mrnaexpression"~"cyp4c33 mRNA expression",
                                   lvl3 == "cyp4c34.mrnaexpression"~"cyp4c34 mRNA expression",
                                   lvl3 == "cytoplasmic.membrane.potential"~"Cytoplasmic Membrane Potential",
                                   lvl3 == "cytotoxicity"~"Cytotoxicity",
                                   lvl3 == "d.lactate"~"D Lactate Concentration",
                                   lvl3 == "dao.activity"~"Diamine Oxidase Activity",
                                   lvl3 == "degranulation"~"Degranulation",
                                   lvl3 == "development"~"Development",
                                   lvl3 == "developmental.abnormalities"~"Developmental Abnormalities",
                                   lvl3 == "dgat.mrnaexpression"~"Diglyceride acyltransferase mRNA expression",
                                   lvl3 == "digestive.tract.histo"~"Digestive Tract Histology",
                                   lvl3 == "dio.rc"~"Energy Dissipated/Reaction Center",
                                   lvl3 == "distance.between.fish"~"Distance Between Fish",
                                   lvl3 == "distance.during.feeding"~"Distance During Feeding",
                                   lvl3 == "diversity"~"Diversity",
                                   lvl3 == "dlac.activity"~"DLac Activity",
                                   lvl3 == "dna.damage"~"DNA Damage",
                                   lvl3 == "dopamine.con"~"Dopamine Concentration",
                                   lvl3 == "egg.size"~"Egg Size",
                                   lvl3 == "embryotoxicity"~"Embryotoxicity",
                                   lvl3 == "emergence"~"Emergence",
                                   lvl3 == "endopeptidase.activity"~"Endopeptidase Activity",
                                   lvl3 == "energy.balance"~"Energy Balance",
                                   lvl3 == "energy.reserves"~"Energy Reserves",
                                   lvl3 == "ep.mrnaexpression"~"ep mRNA expression",
                                   lvl3 == "era.mrnaexpression"~"Estrogen Receptor α mRNA expression",
                                   lvl3 == "erb.mrnaexpression"~"Estrogen Receptor β mRNA expression",
                                   lvl3 == "erk.mrnaexpression"~"erk mRNA expression",
                                   lvl3 == "erod.activity"~"EROD Activity",
                                   lvl3 == "esr2.mrnaexpression"~"esr2 mRNA expression",
                                   lvl3 == "esterase.activity"~"Esterase Activity",
                                   lvl3 == "estradiol.level"~"Estradiol Levels",
                                   lvl3 == "estradiol.testosterone.ratio"~"Estradiol to Testosterone Ratio",
                                   lvl3 == "eto.abs"~"ETO/ABS",
                                   lvl3 == "eto.rc"~"ETO/RC",
                                   lvl3 == "exopeptidase.activity"~"Exopeptidase Activity",
                                   lvl3 == "exploration"~"Exploration",
                                   lvl3 == "f0.fm"~"f0/fm",
                                   lvl3 == "f0.fv"~"f0/fv",
                                   lvl3 == "fabp6.mrnaexpression"~"fabp6 mRNA expression",
                                   lvl3 == "fadd.mrnaexpression"~"fadd mRNA expression",
                                   lvl3 == "fas.mrnaexpression"~"fas mRNA expression",
                                   lvl3 == "fatty.vac"~"Fatty Vacuoles",
                                   lvl3 == "fecundity"~"Fecundity",
                                   lvl3 == "feeding.activity"~"Feeding Activity",
                                   lvl3 == "feeding.rate"~"Feeding Rate",
                                   lvl3 == "feeding.time"~"Feeding Time",
                                   lvl3 == "fertilization"~"Fertilization",
                                   lvl3 == "fertilization.rate"~"Fertilization Rate",
                                   lvl3 == "fertilization.yield"~"Fertilization Yield",
                                   lvl3 == "fibrosis"~"Fiborsis",
                                   lvl3 == "filter.feeding"~"Filter Feeding",
                                   lvl3 == "firmicutes"~"Firmicutes",
                                   lvl3 == "food.consumption"~"Food Consumption",
                                   lvl3 == "foraging.activity"~"Foraging Activity",
                                   lvl3 == "foraging.time"~"Foraging Time",
                                   lvl3 == "foxl2.mrnaexpression"~"fox12 mRNA expression",
                                   lvl3 == "free.fatty.acid.con"~"Free Fatty Acid Concentration",
                                   lvl3 == "fshb.mrnaexpression"~"fshβ mRNA expression",
                                   lvl3 == "fshr.mrnaexpression"~"fshr mRNA expression",
                                   lvl3 == "ftzf1.mrnaexpression"~"ftzf1 mRNA expression",
                                   lvl3 == "fv'.fm'"~"fv'/fm",
                                   lvl3 == "fv.f0"~"fv/f0",
                                   lvl3 == "fv.fm"~"fv/fm",
                                   lvl3 == "fv.fm.sm.tfm"~"fv/fm/sm/tfm",
                                   lvl3 == "gadph.mrnaexpression"~"gadhph mRNA expression",
                                   lvl3 == "gall.bladder.histo"~"Gall Bladder Histology",
                                   lvl3 == "gallery.vol"~"Gallery Volume",
                                   lvl3 == "gcl.mrnaexpression"~"gcl mRNA expression",
                                   lvl3 == "gcl1.mrnaexpression"~"gcl1 mRNA expression",
                                   lvl3 == "gclcs.mrnaexpression"~"gclcs mRNA expression",
                                   lvl3 == "gclcu.mrnaexpression"~"gclcu mRNA expression",
                                   lvl3 == "genera"~"Genera",
                                   lvl3 == "gfap.mrnaexpression"~"gfap mRNA expression",
                                   lvl3 == "ggt.activity"~"Gamma-Glutamyl Transferase Activity",
                                   lvl3 == "ggt.con"~"Gamma-Glutamyl Transferase Concentration",
                                   lvl3 == "gill.histo"~"Gill Histology",
                                   lvl3 == "gk.mrnaexpression"~"gk mRNA expression",
                                   lvl3 == "globulin.con"~"Globulin Concentration",
                                   lvl3 == "glucose.con"~"Glucose Concentration",
                                   lvl3 == "glutamate.con"~"Glutamate Concentration",
                                   lvl3 == "glutathione.con"~"Glutathione Concentration",
                                   lvl3 == "glycogen.con"~"Glycogen Concentration",
                                   lvl3 == "glycogen.deplet"~"Glycogen Depletion",
                                   lvl3 == "gnrh.mrnaexpression"~"gnrh mRNA expression",
                                   lvl3 == "gnrhr.mrnaexpression"~"gnrhr mRNA expression",
                                   lvl3 == "gonad.development.index"~"Gonad Development Index",
                                   lvl3 == "gonad.histo"~"Gonad Histology",
                                   lvl3 == "gonad.regression"~"Gonad Regression",
                                   lvl3 == "gonad.somatic.index"~"Gonad Somatic Index",
                                   lvl3 == "gordonia"~"Gordonia",
                                   lvl3 == "got.activity"~"Glutamic Oxaloacetic Transaminase Activity",
                                   lvl3 == "gpt.activity"~"Glutamic Pyruvic Transaminase Activity",
                                   lvl3 == "gpx.activity"~"Glutathione Peroxidase Activity",
                                   lvl3 == "gpx.mrnaexpression"~"Glutathione Peroxidase mRNA expression",
                                   lvl3 == "gpx4b.mrnaexpression"~"gpx4b mRNA expression",
                                   lvl3 == "gr.activity"~"Glutathione Reductase Activity",
                                   lvl3 == "gr.mrnaexpression"~"Glutathione Reductase mRNA expression",
                                   lvl3 == "granulocyte.con"~"Granulocyte Concentration",
                                   lvl3 == "granulocyte.hyalinocyte.ratio"~"Granulocyte to Hyalinocyte Ratio",
                                   lvl3 == "granulocyte.oxidative.activity"~"Granulocyte Oxidative Activity",
                                   lvl3 == "granulocyte.size"~"Granulocyte Size",
                                   lvl3 == "gross.energy"~"Gross Energy",
                                   lvl3 == "growth"~"Growth",
                                   lvl3 == "growth.rate"~"Growth Rate",
                                   lvl3 == "gs.mrnaexpression"~"Glutathione Synthetase mRNA expression",
                                   lvl3 == "gsh-px.activity"~"Glutathione Peroxidase Activity",
                                   lvl3 == "gsh.activity"~"Glutathione Peroxidase Activity",
                                   lvl3 == "gsh.con"~"Glutathione Peroxidase Concentration",
                                   lvl3 == "gsp.activity"~"Glutathione Peroxidase Activity",
                                   lvl3 == "gst.activity"~"Glutathione S-Transferase Activity",
                                   lvl3 == "gst.mrnaexpression"~"Glutathione S-Transferase mRNA expression",
                                   lvl3 == "gst4.mrnaexpression"~"gst4 mRNA expression",
                                   lvl3 == "gstd.mrnaexpression"~"gstd mRNA expression",
                                   lvl3 == "gstp1.mrnaexpression"~"Glutathione S-Transferase mRNA expression",
                                   lvl3 == "gt.con"~"Glutathione Concentration",
                                   lvl3 == "gtha.mrnaexpression"~"gtha mRNA expression",
                                   lvl3 == "gusb.mrnaexpression"~"gusb mRNA expression",
                                   lvl3 == "gut.mucus.vol"~"Gut Mucus Volume",
                                   lvl3 == "gyrus.size"~"Gyrus Size",
                                   lvl3 == "h2o2.con"~"Hydorgen Peroxide Concentration",
                                   lvl3 == "h3.mrnaexpression"~"h3 mRNA expression",
                                   lvl3 == "haemocyanin.con"~"Haemocyanin Concentration",
                                   lvl3 == "hemocyte.con"~"Hemocyte Concentration",
                                   lvl3 == "haemolymph.con"~"Haemolymph Concentration",
                                   lvl3 == "hatching.success"~"Hatching Success",
                                   lvl3 == "hdl.con"~"HDL Concentration",
                                   lvl3 == "heart.rate"~"Heart Rate",
                                   lvl3 == "hepatic.somatic.index"~"Hepatic Somatic Index",
                                   lvl3 == "hepatopancreas.water.con"~"Hepatopancreas Water Concentration",
                                   lvl3 == "hex.mrnaexpression"~"hex mRNA expression",
                                   lvl3 == "hk.mrnaexpression"~"hk mRNA expression",
                                   lvl3 == "hk1.mrnaexpression"~"hk1 mRNA expression",
                                   lvl3 == "ho1.mrnaexpression"~"ho1 mRNA expression",
                                   lvl3 == "hsp70.mrnaexpression"~"hsp70 mRNA expression",
                                   lvl3 == "hsp90.mrnaexpression"~"hsp90 mRNA expression",
                                   lvl3 == "hyalinocyte.size"~"Hyalinocyte Size",
                                   lvl3 == "hylaniocyte.oxidative.activity"~"Hylaniocyte Oxidative Activity",
                                   lvl3 == "id.mrnaexpression"~"Isocitrate Dehydrogenase mRNA expression",
                                   lvl3 == "idh.activity"~"Isocitrate Dehydrogenase Activity",
                                   lvl3 == "idp.mrnaexpression"~"idp mRNA expression",
                                   lvl3 == "ifn.mrnaexpression"~"Interferon mRNA expression",
                                   lvl3 == "ifny.mrnaexpression"~"Interferon y mRNA expression",
                                   lvl3 == "igm.con"~"Immunoglobulin M Concentration",
                                   lvl3 == "igm.mrnaexpression"~"Immunoglobulin M mRNA expression",
                                   lvl3 == "il10.mrnaexpression"~"Interleukin 10 mRNA expression",
                                   lvl3 == "il17.mrnaexpression"~"Interleukin 17 mRNA expression",
                                   lvl3 == "il1a.mrnaexpression"~"Interleukin 1α mRNA expression",
                                   lvl3 == "il1b.mrnaexpression"~"Interleukin 1β mRNA expression",
                                   lvl3 == "il413a.mrnaexpression"~"Interleukin 413α mRNA expression",
                                   lvl3 == "il6.mrnaexpression"~"Interleukin 6 mRNA expression",
                                   lvl3 == "il8.mrnaexpression"~"Interluekin 8 mRNA expression",
                                   lvl3 == "ila.con"~"Interleukin 1α Concentration",
                                   lvl3 == "immobilization"~"Immobilization",
                                   lvl3 == "inflam.infiltrate"~"Inflammatory Filtrate",
                                   lvl3 == "ingestion.rate"~"Ingestion Rate",
                                   lvl3 == "inhibitory.rate"~"Inhibitory Rate",
                                   lvl3 == "instar.develop.time"~"Instar Development Time",
                                   lvl3 == "intermolt.time"~"Intermolt Time",
                                   lvl3 == "intestinal.ca.con"~"Intestinal Calcium Concentration",
                                   lvl3 == "ion.exch.ca"~"Ion Exchange (Calcium)",
                                   lvl3 == "ion.exch.k"~"Ion Exchange (Potassium)",
                                   lvl3 == "ion.exch.na"~"Ion Exchange (Sodium)",
                                   lvl3 == "ion.flow"~"Ion Flow",
                                   lvl3 == "ion.permeability"~"Ion Permeability",
                                   lvl3 == "ion.trans"~"Ion Transport",
                                   lvl3 == "jaw.tissue.damage"~"Jaw Tissue Damage",
                                   lvl3 == "jnk.mrnaexpression"~"jnk mRNA expression",
                                   lvl3 == "jump.freq"~"Jump Frequency",
                                   lvl3 == "jump.height"~"Jump Height",
                                   lvl3 == "kidney.histo"~"Kidney Histology",
                                   lvl3 == "labilization.period"~"Labilization Period",
                                   lvl3 == "lactate.con"~"Lactate Concentration",
                                   lvl3 == "ldh.activity"~"Lactate Dehydrogenase Activity",
                                   lvl3 == "ldh.con"~"Lactate Dehydrogenase Concentration",
                                   lvl3 == "ldl.con"~"LDL Concentration",
                                   lvl3 == "leaf.growth.rate"~"Leaf Growth Rate",
                                   lvl3 == "length"~"Length",
                                   lvl3 == "leukocyte.con"~"Leukocyte Concentration",
                                   lvl3 == "lhb.mrnaexpression"~"Leutinizing Hormone β mRNA expression",
                                   lvl3 == "lhr.mrnaexpression"~"Leutinizing Hormone Receptor mRNA expression",
                                   lvl3 == "lifespan"~"Lifespan",
                                   lvl3 == "light.capture"~"Light Capture",
                                   lvl3 == "light.dark.activity"~"Light/Dark Activity",
                                   lvl3 == "line.crossing"~"Line Crossing",
                                   lvl3 == "lipase.activity"~"Lipase Activity",
                                   lvl3 == "lipase.con"~"Lipase Concentration",
                                   lvl3 == "lipid.content"~"Lipid Content",
                                   lvl3 == "lipid.membrane.con"~"Lipid Membrane Concentration",
                                   lvl3 == "lipid.perox"~"Lipid Peroxidation",
                                   lvl3 == "lipid.storage.con"~"Lipid Storage Concentration",
                                   lvl3 == "lipofuscin.accum"~"Lipofuscin Accumulation",
                                   lvl3 == "liver.histo"~"Liver Histology",
                                   lvl3 == "liver.protein"~"Liver Protein",
                                   lvl3 == "lymphocyte.con"~"Lymphocyte Concentration",
                                   lvl3 == "lys.mrnaexpression"~"Lysozyme mRNA expression",
                                   lvl3 == "lysosomal.membrane.stab"~"Lysosomal Membrane Stability",
                                   lvl3 == "lysozyme.activity"~"Lysozyme Activity",
                                   lvl3 == "manf.mrnaexpression"~"manf mRNA expression",
                                   lvl3 == "mao.activity"~"Monoamine Oxidase Activity",
                                   lvl3 == "mapk.phosphorylation"~"MAPK Phosphorylation",
                                   lvl3 == "mbp.mrnaexpression"~"Myelin Basic Protein mRNA expression",
                                   lvl3 == "mbp.proteinexpression"~"Myelin Basic Protein Expression",
                                   lvl3 == "mda.con"~"Malondialdehyde Concentration",
                                   lvl3 == "meer1.mrnaexpression"~"meer 1 mRNA expression",
                                   lvl3 == "meer2.mrnaexpression"~"meer 2 mRNA expression",
                                   lvl3 == "mek.mrnaexpression"~"mek mRNA expression",
                                   lvl3 == "metamorphosis.yield"~"Metamorphosis Yield",
                                   lvl3 == "mgnrh.mrnaexpression"~"mgnrh mRNA expression",
                                   lvl3 == "min.sat.irradiance"~"Minimum Saturation Irradiance",
                                   lvl3 == "mirconuclei.freq"~"Micronuclei Frequency",
                                   lvl3 == "mitochondrial.mem.integrity"~"Mitochondrial Membrane Integrity",
                                   lvl3 == "mnsod.mrnaexpression"~"Mn Superoxide Dismutase mRNA expression",
                                   lvl3 == "moisture"~"Moisture",
                                   lvl3 == "molecular.permeability"~"Molecular Permeability",
                                   lvl3 == "molt.period.length"~"Molt Period Length",
                                   lvl3 == "mortality"~"Mortality",
                                   lvl3 == "mt10.mrnaexpression"~"mt10 mRNA expression",
                                   lvl3 == "mt20.mrnaexpression"~"mt20 mRNA expression",
                                   lvl3 == "mta.mrnaexpression"~"mta mRNA expression",
                                   lvl3 == "mtb.mrnaexpression"~"mtb mRNA expression",
                                   lvl3 == "mxr.efflux"~"MXR Efflux",
                                   lvl3 == "mytc.mrnaexpression"~"mytc mRNA expression",
                                   lvl3 == "myticina.mrnaexpression"~"myticin α mRNA expression",
                                   lvl3 == "mytilinb.mrnaexpression"~"mytilin β mRNA expression",
                                   lvl3 == "nauplius.to.adult"~"Nauplius to Adult",
                                   lvl3 == "nauplius.to.copepodid"~"Nauplius to Copepodid",
                                   lvl3 == "neutrophil.extracell.trap"~"Neutrophil Extracellular Trap",
                                   lvl3 == "ngn1.mrnaexpression"~"ngn1 mRNA expression",
                                   lvl3 == "nitrogen.conversion"~"Nitrogen Conversion",
                                   lvl3 == "no.prod"~"Nitric Oxide Production",
                                   lvl3 == "nrf2.mrnaexpression"~"nrf2 mRNA expression",
                                   lvl3 == "num.indiv.nosnails"~"Number of Individuals (No Snails)",
                                   lvl3 == "num.indiv.snails"~"Number of Individuals (With Snails)",
                                   lvl3 == "num.offspring"~"Number of Offspring",
                                   lvl3 == "number.of.molts"~"Number of Molts",
                                   lvl3 == "number.of.roots"~"Number of Roots",
                                   lvl3 == "occludin.mrnaexpression"~"occludin mRNA expression",
                                   lvl3 == "offspring.size"~"Offspring Size",
                                   lvl3 == "oocyte.diameter"~"Oocyte Diameter",
                                   lvl3 == "oocyte.num"~"Oocyte Number",
                                   lvl3 == "oxygen.consumption"~"Oxygen Consumption",
                                   lvl3 == "p38.mrnaexpression"~"p38 mRNA expression",
                                   lvl3 == "p38mapk.mrnaexpression"~"p38mapk mRNA expression",
                                   lvl3 == "p53.mrnaexpression"~"p53 mRNA expression",
                                   lvl3 == "pc.pe.ratio"~"PC/PE Ratio",
                                   lvl3 == "pcc.level"~"Protein Carbonylation Content",
                                   lvl3 == "pche.activity"~"Pseudocholinesterase Activity",
                                   lvl3 == "pepckc.mrnaexpression"~"pepckc mRNA expression",
                                   lvl3 == "percent.dstage.larvae"~"Pertengage of D-stage Larvae",
                                   lvl3 == "percent.dveliger"~"Percentage of Veliger Larvae",
                                   lvl3 == "percent.motile.sperm"~"Percentage of Motile Sperm",
                                   lvl3 == "percent.tank.used"~"Percentage of Tank Used",
                                   lvl3 == "percent.trocophore.larvae"~"Percentage of Trocophore Larvae",
                                   lvl3 == "pericardial.area"~"Pericardial Area",
                                   lvl3 == "perox.activity"~"Peroxidase Activity",
                                   lvl3 == "perox.mrnaexpression"~"Peroxidase mRNA expression",
                                   lvl3 == "petb.mrnaexpression"~"petb mRNA expression",
                                   lvl3 == "pglycoprotein.activity"~"Pglycoprotein Activity",
                                   lvl3 == "pgp.mrnaexpression"~"pgp mRNA expression",
                                   lvl3 == "pgrp.mrnaexpression"~"pgrp mRNA expression",
                                   lvl3 == "phagocytosis"~"Phagocytosis",
                                   lvl3 == "phenoloxydase.activity"~"Phenoloxydase Activity",
                                   lvl3 == "phosphorus.con"~"Phosphorus Concentration",
                                   lvl3 == "phosphatase.activity"~"Phosphatase Activity",
                                   lvl3 == "phosphatidylcholine.con"~"Phosphatidylcholine Concentration",
                                   lvl3 == "phosphatidylethanolamine.con"~"Phosphatidylethanolamine Concentration",
                                   lvl3 == "phosphatidylinositol.ceramide.amino.ethylphosphonate.con"~"PCAE Concentration",
                                   lvl3 == "phosphatidylserine.con"~"Phosphatidylserine Concentration",
                                   lvl3 == "photochemical.efficiency"~"Photochemical Efficiency",
                                   lvl3 == "pk.mrnaexpression"~"pk mRNA expression",
                                   lvl3 == "polyp.activity"~"Polyp Activity",
                                   lvl3 == "population.size"~"Population Size",
                                   lvl3 == "potassium.con"~"Potassium Concentration",
                                   lvl3 == "ppara.mrnaexpression"~"pparα mRNA expression",
                                   lvl3 == "ppary.mrnaexpression"~"ppary mRNA expression",
                                   lvl3 == "prdx1.mrnaexpression"~"prdx1 mRNA expression",
                                   lvl3 == "prdx2.mrnaexpression"~"prdx2 mRNA expression",
                                   lvl3 == "prdx3.mrnaexpression"~"prdx3 mRNA expression",
                                   lvl3 == "prdx5.mrnaexpression"~"prdx5 mRNA expression",
                                   lvl3 == "predation.risk"~"Predation Risk",
                                   lvl3 == "predatory.performance"~"Predatory Performance",
                                   lvl3 == "pregnancy.rate"~"Pregnancy Rate",
                                   lvl3 == "protein.carbonyl.con"~"Protein Carbonyl Concentration",
                                   lvl3 == "protein.con"~"Protein Concentration",
                                   lvl3 == "protein.reserves"~"Protein Reserves",
                                   lvl3 == "proteobacteria"~"proteobacteria",
                                   lvl3 == "ps2"~"Photosystem II",
                                   lvl3 == "psaa.mrnaexpression"~"psaa mRNA expression",
                                   lvl3 == "psab.mrnaexpression"~"psab mRNA expression",
                                   lvl3 == "psba.mrnaexpression"~"psba mRNA expression",
                                   lvl3 == "psbl.mrnaexpression"~"psbl mRNA expression",
                                   lvl3 == "pseudomonas"~"pseudomonas",
                                   lvl3 == "rapid.light.resp.curve"~"Rapid Light Response Curve",
                                   lvl3 == "rbc.con"~"Red Blood Cell Concentration",
                                   lvl3 == "rbcl.mrnaexpression"~"rbcl mRNA expression",
                                   lvl3 == "rc.abs"~"RC/ABS",
                                   lvl3 == "redox.state"~"Redox State",
                                   lvl3 == "relative.e.transport.rate"~"Relative e Transport State",
                                   lvl3 == "relocate"~"Relocation",
                                   lvl3 == "reproduction.factor"~"Reproduction Factor",
                                   lvl3 == "reproduction.inhibition"~"Reproduction Inhibition",
                                   lvl3 == "reproductive.freq"~"Reproductive Frequency",
                                   lvl3 == "resistance.time"~"Resistance Time",
                                   lvl3 == "respiration.rate"~"Respiration Rate",
                                   lvl3 == "respiratory.burst"~"Respiratory Burst",
                                   lvl3 == "retrmax"~"Maximal Election Transport Rate",
                                   lvl3 == "root.length"~"Root Length",
                                   lvl3 == "ros.prod"~"Reactive Oxygen Species Production",
                                   lvl3 == "s100a1.mrnaexpression"~"s100a1 mRNA expression",
                                   lvl3 == "saa.mrnaexpression"~"saa mRNA expression",
                                   lvl3 == "sd.mrnaexpression"~"sd mRNA expression",
                                   lvl3 == "segment.regeneration"~"Segment Regeneration",
                                   lvl3 == "serotonin.con"~"Serotonin Concentration",
                                   lvl3 == "settling.rate"~"Settling Rate",
                                   lvl3 == "sex.ratio"~"Sex Ratio",
                                   lvl3 == "sexual.maturity"~"Sexual Maturity",
                                   lvl3 == "sgst.mrnaexpression"~"sgst mRNA expression",
                                   lvl3 == "shannon.diversity.index"~"Shannon Diversity Index",
                                   lvl3 == "shell.growth"~"Shell Growth",
                                   lvl3 == "shell.length"~"Shell Length",
                                   lvl3 == "shell.weight"~"Shell Weight",
                                   lvl3 == "size"~"Size",
                                   lvl3 == "sod.activity"~"Superoxide Dismutase Activity",
                                   lvl3 == "sod.mrnaexpression"~"sod mRNA expression",
                                   lvl3 == "sod1.mrnaexpression"~"sod1 mRNA expression",
                                   lvl3 == "sod3.mrnaexpression"~"sod3 mRNA expression",
                                   lvl3 == "sodium.con"~"Sodium Concentration",
                                   lvl3 == "species.diversity"~"Species Diversity",
                                   lvl3 == "species.richness"~"Species Richness",
                                   lvl3 == "specific.growth.rate"~"Specific Growth Rate",
                                   lvl3 == "sperm.velocity"~"Sperm Velocity",
                                   lvl3 == "srebp1a.mrnaexpression"~"srebp1a mRNA expression",
                                   lvl3 == "star.mrnaexpression"~"star mRNA expression",
                                   lvl3 == "sterol.con"~"Sterol Concentration",
                                   lvl3 == "surface.area"~"Surface Area",
                                   lvl3 == "swim.distance"~"Swim Distance",
                                   lvl3 == "swim.speed"~"Swim Speed",
                                   lvl3 == "swimbladder.inflation"~"Swimbladder Inflation",
                                   lvl3 == "synapsin.mrnaexpression"~"synapsin mRNA expression",
                                   lvl3 == "taxa.num"~"Number of Taxa",
                                   lvl3 == "testosterone.level"~"Testosterone Levels",
                                   lvl3 == "tgfb.mrnaexpression"~"tgfb mRNA expression",
                                   lvl3 == "thiol.con"~"Thiol Concentration",
                                   lvl3 == "thrombocyte.con"~"Thrombocyte Concentration",
                                   lvl3 == "time.to.hatch"~"Time to Hatch",
                                   lvl3 == "time.to.maturation"~"Time to Maturation",
                                   lvl3 == "time.to.nauplius"~"Time to Nauplius",
                                   lvl3 == "time.to.offspring"~"Time to Offspring",
                                   lvl3 == "tlr.mrnaexpression"~"tlr mRNA expression",
                                   lvl3 == "tnfa.mrnaexpression"~"tnfα mRNA expression",
                                   lvl3 == "tnfsf13b.mrnaexpression"~"tnfsf13b mRNA expression",
                                   lvl3 == "total.distance.traveled"~"Total Distance Travelled",
                                   lvl3 == "tp53.con"~"tp53 Concentration",
                                   lvl3 == "tp53.mrnaexpression"~"tp53 mRNA expression",
                                   lvl3 == "tph2.mrnaexpression"~"tph2 mRNA expression",
                                   lvl3 == "tricellulin.mrnaexpression"~"tricellulin mRNA expression",
                                   lvl3 == "triglyceride.cholesterol.ratio"~"Tryglyceride Cholesterol Ratio",
                                   lvl3 == "triglycerides.con"~"Triglyceride Concentration",
                                   lvl3 == "tro.rc"~"TRO/RC",
                                   lvl3 == "trypsin.activity"~"Trypsin Activity",
                                   lvl3 == "tumor.promotion"~"Tumor Promotion",
                                   lvl3 == "turn.angle.light.dark.activity"~"Turn Angle, Light/Dark Activity",
                                   lvl3 == "ucp1.mrnaexpression"~"ucp1 mRNA expression",
                                   lvl3 == "ugst.mrnaexpression"~"ugst mRNA expression",
                                   lvl3 == "uricacid.con"~"Uric Acid Concentration",
                                   lvl3 == "viserosomatic.index"~"Viserosomatic Index",
                                   lvl3 == "volume"~"Volume",
                                   lvl3 == "vtg.mrnaexpression"~"vtg mRNA expression",
                                   lvl3 == "vtg.proteinexpression"~"vtg protein expression",
                                   lvl3 == "vtg1.mrnaexpression"~"vtg1 mRNA expression",
                                   lvl3 == "vtg2.mrnaexpression"~"vtg2 mRNA expression",
                                   lvl3 == "weight.change"~"Weight Change",
                                   lvl3 == "wgst.mrnaexpression"~"wgst mRNA expression",
                                   lvl3 == "wnt16.mrnaexpression"~"wnt16 mRNA expression",
                                   lvl3 == "wnt4.mrnaexpression"~"wnt4 mRNA expression",
                                   lvl3 == "yproteobacteria"~"y proteobacteria",
                                   lvl3 == "zfblue.mrnaexpression"~"zfblue mRNA expression",
                                   lvl3 == "zfrho.mrnaexpression"~"zfrho mRNA expression",
                                   lvl3 == "zo1.mrnaexpression"~"zo1 mRNA expression",
                                   lvl3 == "zoozanthelle.density"~"Zoozanthelle Density"))) %>%
  mutate(bio_f = factor(case_when( #Bio Org Data Tidying
    bio.org == "subcell"~"Subcell",
    bio.org == "cell"~"Cell",
    bio.org == "tissue" ~ "Tissue",
    bio.org == "organism"~"Organism",
    bio.org == "population"~ "Population")))%>%
  mutate(vivo_f = factor(case_when(invitro.invivo == "invivo"~"In Vivo",
    invitro.invivo == "invitro"~"In Vitro")))%>% ##Renames for widget (Not using a widget right now, but saving for human health database)
  mutate(life_f = factor(case_when(life.stage == "Early"~"Early",
    life.stage == "Juvenile"~"Juvenile",
    life.stage == "Adult"~"Adult",
    life.stage == "Not Reported"~"Not Reported")))%>% #Renames for widget
  mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
    environment == "Marine" ~ "Marine",
    environment == "Terrestrial" ~ "Terrestrial"))) %>%
  mutate(species_f = as.factor(paste(genus,species))) %>% 
  mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
  mutate(effect.metric = factor(effect.metric)) %>% #factorize
  mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) %>%  #calculate volume/mL
  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
  mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
                                            af.time_noNA == 1 ~ "Chronic",
                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>% #factorize assesment factor time into chronic/acute
  mutate(tier_zero_tech_f = factor(case_when(tech.tier.zero == "Fail" ~ "Red Criteria Failed",
                                             tech.tier.zero == "Pass" ~ "Red Criteria Passed"))) %>% 
  mutate(tier_zero_risk_f = factor(case_when(risk.tier.zero == "Fail" ~ "Red Criteria Failed",
                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed")))
```

```{r Organism Groups, echo = FALSE, results = 'hide', error = FALSE, warning = FALSE, message = FALSE}

a <- aoc_setup %>% 
  filter(environment != "Terrestrial") %>% 
  ggplot(aes(x = org_f, fill = environment)) +
  geom_bar()

plot(a)
```

### 2021-05-20 Shiny_Manuscript_Figures/Manuscript_Figures.Rmd
Minor data set changes (just character to numerical values to make plotting easier) + implementation of screening plot + random scripts for possible figure generation for paper

@@
 ---
-title: 'Shiny App Manuscript Figures: Aquatic Organisms'
+title: 'Team X Manuscript Figures: Aquatic Organisms'
 author: "Leah Thornton Hampton"
-date: "4/2/2021"
+date: "4/26/2021"
 output:
@@
 library(tidyverse)
+library(collapsibleTree) #plot type for endpoint category tree
+library(calecopal) #colors 
+library(wesanderson) #colors
 
@@
 #Data Tidying
+# Master dataset for scatterplots - for Heili's tab.
 aoc_v1 <- aoc %>% # start with original dataset
@@
     size.category == 5 ~ "1mm < 5mm",
-    size.category == 0 ~ "Not Reported"),
-    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
+    size.category == 0 ~ "Size Not Reported"),
+    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Size Not Reported"))) %>% # creates new column with nicer names and order by size levels.
   # shape category data tidying.
@@
   mutate(tier_zero_risk_f = factor(case_when(risk.tier.zero == "Fail" ~ "Red Criteria Failed",
-                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed")))
+                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed"))) %>% 
+  #Remove leachate and additive/chemical transfer experiments
+  dplyr::filter(leachate.only != "Y") %>%
+  mutate(chem.exp.typ.nominal_f = factor(case_when(chem.exp.typ.nominal == "Particle Only" ~ "Particle Only",
+                                                   chem.exp.typ.nominal == "co.exp" ~ "Chemical Co-Exposure",
+                                                   chem.exp.typ.nominal == "sorbed" ~ "Chemical Transfer"))) %>% 
+  dplyr::filter(chem.exp.typ.nominal_f == "Particle Only") %>% 
+  #calculate maximum ingestible size (if not already in database)
+  mutate(max.size.ingest.mm = ifelse(is.na(max.size.ingest.mm), 
+    10^(0.9341 * log10(body.length.cm - 1.1200)) * 10,  #(Jamm et al 2020 Nature paper)correction for cm to mm
+    max.size.ingest.mm)) # if already present, just use that
 ```
 
-```{r Organism Groups, echo = FALSE, results = 'hide', error = FALSE, warning = FALSE, message = FALSE}
+```{r collapsible tree example}
 
+collapsibleTree(aoc_setup, root = "Aquatic Organisms Database", hierarchy = c("lvl1_f", "lvl2_f", "lvl3_f", "bio_f"),
+                    fontSize = 16, zoomable = FALSE,    
+    fill = c(
+      # The root
+      "seashell",
+      # lvl1
+      rep("turquoise", length(unique(aoc_setup$lvl1_f))),
+      # lvl2
+      rep("palegreen", length(unique(paste(aoc_setup$lvl1_f, aoc_setup$lvl2_f)))),
+      # lvl3
+      rep("hotpink", length(unique(paste(aoc_setup$lvl1_f, aoc_setup$lvl2_f, aoc_setup$lvl3_f)))),
+      # bio org
+      rep("orchid", length(unique(paste(aoc_setup$lvl1_f, aoc_setup$lvl2_f, aoc_setup$lvl3_f, aoc_setup$bio_f))))))
+
+```
+
+```{r endpoints and organism groups, fig.width = 15}
+library(viridis)
 a <- aoc_setup %>% 
-  filter(environment != "Terrestrial") %>% 
-  ggplot(aes(x = org_f, fill = environment)) +
-  geom_bar()
+  filter(effect_f != "NA") %>%
+  count(lvl1_f, org_f) %>% #how many obs of each organism group within each lvl1
+  group_by(lvl1_f) %>%
+  mutate(n_lvl1 = sum(n)) %>% #how many obs within each lvl1
+  ungroup() %>%   
+  ggplot(aes(x = fct_reorder(lvl1_f, n_lvl1), y = n, group = n, fill = org_f)) +
+  geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), color = "black", width = 1) +
+  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 10)) +
+  scale_y_continuous(expand = c(0,0)) +
+  scale_fill_manual(values= wes_palette("Zissou1", type = "continuous", 13)) +
+  theme_classic() +
+  theme(text = element_text(size = 16))+
+  labs(y = "Number of Endpoints in Database", x = "Endpoint Category", fill = "Organism\nGroup")
 
 plot(a)
+
 ```
+
+```{r particle characteristics}
+
+particle <- aoc_setup %>%
+  group_by(size_f) %>%
+  #count the number of data points in each size category
+  mutate(size_count = n()) %>% 
+  #percentage that report polymer
+  mutate(poly_pct = (sum(poly_f != "Not Reported")/size_count)*100) %>% 
+  #percentage that report shape
+  mutate(shape_pct = (sum(shape_f != "Not Reported")/size_count)*100) %>%
+  #percentage that report density
+  replace_na(list(density.reported.estimated = "Not Reported")) %>%   
+  mutate(density_f = as.factor(density.reported.estimated)) %>% 
+  mutate(density_pct = (sum(density_f == "reported")/size_count)*100) %>%  
+  #percentage that report charge
+  mutate(charge_pct = (sum(!is.na(charge))/size_count)*100) %>%
+  #percentage that report functional group
+  mutate(fxngroup_pct = (sum(!is.na(functional.group))/size_count)*100) %>%
+  #validate size
+  mutate(sizevalid_f = as.factor(size.valid)) %>% 
+  mutate(sizevalid_pct = (sum(sizevalid_f == "Y")/size_count)*100) %>% 
+  #validate shape
+  mutate(shapevalid_f = as.factor(shape.valid)) %>% 
+  mutate(shapevalid_pct = (sum(shapevalid_f == "Y")/size_count)*100) %>% 
+  #validate polymer
+  mutate(polyvalid_f = as.factor(polymer.valid)) %>% 
+  mutate(polyvalid_pct = (sum(polyvalid_f == "Y")/size_count)*100) %>% 
+  ungroup() %>% 
+  group_by(size_f, poly_pct, shape_pct, density_pct, charge_pct, fxngroup_pct, sizevalid_pct, shapevalid_pct, polyvalid_pct) %>% 
+  summarise() %>% 
+  rename("Polymer Type" = poly_pct, "Morphology" = shape_pct, "Density" = density_pct, "Charge" = charge_pct, "Functional Group" = fxngroup_pct, "Size Validated" = sizevalid_pct, "Morphology Validated" = shapevalid_pct, "Polymer Validated" = polyvalid_pct)
+
+require(pheatmap)
+require(RColorBrewer)
+
+# Make column 1 equal to the row name and remove data from column 1 and arrange size levels
+
+All_data <- data.frame(particle, row.names = 1)%>% 
+  arrange("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Size Not Reported")
+
+#Turn data into a matrix format
+
+All_matrix <- data.matrix(All_data)
+
+color <- brewer.pal(n=8, name = "Blues")
+
+Pretty_heatmap <- pheatmap(All_matrix,
+        main = "Frequency (%) of Particle Characteristics Reported",
+        fontsize = 12,
+        treeheight_row = 0, treeheight_col = 0,
+        cluster_rows = FALSE, cluster_cols = FALSE,
+        display_numbers = TRUE,
+        color = color,
+        number_format = "%.0f", fontsize_number = 10, number_color = "black")
+
+```

### 2021-05-20 TeamX_Manuscript_Figures/Manuscript_Figures_TeamX.Rmd
Minor data set changes (just character to numerical values to make plotting easier) + implementation of screening plot + random scripts for possible figure generation for paper

--- 2021-04-08 Shiny_Manuscript_Figures/Manuscript_Figures.Rmd
@@
 author: "Leah Thornton Hampton"
-date: "4/2/2021"
+date: "4/26/2021"
 output:
@@
     toc_float: true
+    code_folding: hide
 ---
@@
 
-```{r Data Import and Packages, echo = FALSE, results = 'hide', error = FALSE, warning = FALSE, message = FALSE}
+```{r Data Import and Packages, warning = FALSE, message = FALSE}
 
@@
   mutate(tier_zero_risk_f = factor(case_when(risk.tier.zero == "Fail" ~ "Red Criteria Failed",
-                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed")))
+                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed"))) %>% 
+  #Remove leachate and additive/chemical transfer experiments
+  dplyr::filter(leachate.only != "Y") %>%
+  mutate(chem.exp.typ.nominal_f = factor(case_when(chem.exp.typ.nominal == "Particle Only" ~ "Particle Only",
+                                                   chem.exp.typ.nominal == "co.exp" ~ "Chemical Co-Exposure",
+                                                   chem.exp.typ.nominal == "sorbed" ~ "Chemical Transfer"))) %>% 
+  dplyr::filter(chem.exp.typ.nominal_f == "Particle Only") %>% 
+  #calculate maximum ingestible size (if not already in database)
+  mutate(max.size.ingest.mm = ifelse(is.na(max.size.ingest.mm), 
+    10^(0.9341 * log10(body.length.cm - 1.1200)) * 10,  #(Jamm et al 2020 Nature paper)correction for cm to mm
+    max.size.ingest.mm)) # if already present, just use that
 ```
 
-```{r Organism Groups, echo = FALSE, results = 'hide', error = FALSE, warning = FALSE, message = FALSE}
+# Ludo
+
+```{r Ludo, warning = FALSE, message = FALSE, fig.height = 7, fig.width = 12}
 
 a <- aoc_setup %>% 
-  filter(environment != "Terrestrial") %>% 
-  ggplot(aes(x = org_f, fill = environment)) +
-  geom_bar()
+  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+  filter(shape_f != "Not Reported") %>% #drop studies where shapes aren't reported
+  filter(effect_f != "NA") %>% #drop ECX values
+  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
+  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
+  ggplot(aes(y = shape_f, x = dose.mg.L.master, fill = effect_f)) +
+  geom_boxplot(alpha = 0.8) +
+  scale_x_log10() +
+  labs(y = "Morphology", x = "mg/L", fill = "Effect") +
+  theme_classic() +
+  theme(text = element_text(size = 16))
 
 plot(a)
+
+b <- aoc_setup %>% 
+  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+  filter(shape_f != "Not Reported") %>% #drop studies where shapes aren't reported
+  filter(effect_f != "NA") %>% #drop ECX values
+  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
+  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
+  ggplot(aes(y = shape_f, x = dose.particles.mL.master, fill = effect_f)) +
+  geom_boxplot(alpha = 0.8) +
+  scale_x_log10() +
+  labs(y = "Morphology", x = "particles/mL", fill = "Effect") +
+  theme_classic() +
+  theme(text = element_text(size = 16))
+
+plot(b)
+
+c <- aoc_setup %>% 
+  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+  filter(!poly_f %in% c("Not Reported", "Polyethylene Vinyl Acetate", "Polycarbonate", "Latex")) %>% 
+  filter(effect_f != "NA") %>% #drop ECX values
+  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
+  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
+  ggplot(aes(y = poly_f, x = dose.mg.L.master, fill = effect_f)) +
+  geom_boxplot(alpha = 0.8) +
+  scale_x_log10() +
+  labs(y = "Polymer", x = "mg/L", fill = "Effect") +
+  theme_classic() +
+  theme(text = element_text(size = 16)) +
+  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 10))
+
+plot(c)
+
+d <- aoc_setup %>% 
+  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+  filter(!poly_f %in% c("Not Reported", "Polyethylene Vinyl Acetate", "Polycarbonate", "Latex")) %>%
+  filter(effect_f != "NA") %>% #drop ECX values
+  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
+  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
+  ggplot(aes(y = poly_f, x = dose.particles.mL.master, fill = effect_f)) +
+  geom_boxplot(alpha = 0.8) +
+  scale_x_log10() +
+  labs(y = "Polymer", x = "particles/mL", fill = "Effect") +
+  theme_classic() +
+  theme(text = element_text(size = 16)) +
+  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 10))
+
+plot(d)
 ```

### 2021-05-20 app.R

@@ -31,6 +31,7 @@ library(ggsci) #color palettes
 # library(bslib) #better themes. required for dark mode
 # library(thematic) #complete control over themes (including plots) required for dark mode
 library(collapsibleTree) #plot type for endpoint category tree
+library(hrbrthemes) #theme for screening plot
 
 # Load finalized dataset.
 aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
@@ -40,20 +41,6 @@ aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
 # All text inputs below.
 
 #### Overview AO Setup ####
-
-#Final_effect_dataset <- read_csv("Final_effect_dataset.csv")%>%
-  #mutate(plot_f = case_when(
-    #plot_f == "Polymer" ~ "Polymer",
-    #plot_f == "Size" ~ "Size",
-    #plot_f == "Shape" ~ "Shape",
-    #plot_f == "Organism" ~ "Organism",
-    #plot_f == "Lvl1" ~ "Endpoint Category",
-    #plot_f == "Life.stage" ~ "Life Stage",
-    #plot_f == "Invivo.invivo" ~ "In Vivo or In Vitro",
-    #plot_f == "Exposure.route" ~ "Exposure Route"))%>%
-  #mutate(plot_f = factor(plot_f))%>%
-  #mutate(logEndpoints = log(Endpoints))%>%
-  #rename(Percent = Freq)
 
 polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
 polyf<-as.data.frame(polydf)%>% #Makes data frame 
@@ -861,7 +848,11 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
   mutate(tier_zero_tech_f = factor(case_when(tech.tier.zero == "Fail" ~ "Red Criteria Failed",
                                              tech.tier.zero == "Pass" ~ "Red Criteria Passed"))) %>% 
   mutate(tier_zero_risk_f = factor(case_when(risk.tier.zero == "Fail" ~ "Red Criteria Failed",
-                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed"))) %>% 
+                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed"))) %>%
+  #Paste the range of treatments used within each study - there is often a range because this is different depending on the endpoint
+  group_by(doi) %>% 
[…]

### 2021-05-20 app.R
Added quality filter widgets to screening tab

@@ -1443,7 +1443,7 @@ column(width = 12,
                     # "Download Data" is the title that appears on the button
                       
                         column(width = 3,
-                        actionButton("reset_input", "Reset Filters"))), # adds update button
+                        actionButton("reset_input", "Reset Filters", class = "btn-primary"))), # adds update button
                       
                       # "Reset_input" is the internal name
                       # "Reset Filter" is the title that appears on the button  
@@ -1910,8 +1910,10 @@ tabPanel("6: Study Screening",
          h3("Study Screening Results", align = "center"),
          br(),
          p("This plot displays scores from the quality screening exercise developed by", a(href ="https://pubs.acs.org/doi/abs/10.1021/acs.est.0c03057", 'de Ruijter et al. (2020)', .noOWs = "outside"), "with some modification. For more information, including the scoring rubric used, see the document 'Study Screening Rubric' under the Resources tab."),
-         br(),
+         b(),
          p("Interact with the Data: Use your cursor to zoom and hover over the plot to view additional information about each study."),
+         br(),
+         p("The range of treatment groups used with each study is displayed in the hover over text box."),
          br(),
          
          # widget headers
@@ -2022,19 +2024,41 @@ tabPanel("6: Study Screening",
                                    options = list(`actions-box` = TRUE), 
                                    multiple = TRUE))),
          
-         #Go Button and Reset Button
-         
+         #Second row of widget headers
+         column(width=12,
+                
+                column(width = 3,
+                       h4("Quality Criteria"))),
+
+         #New row of widgets
          column(width = 12,
-                br(),
+                
+                column(width = 3,
+                       pickerInput(inputId = "tech_tier_zero_quality", # chronic/acute checklist
+                                   label = "Technical Quality:", 
+                                   choices = levels(aoc_setup$tier_zero_tech_f),
+                                   selected = levels(aoc_setup$tier_zero_tech_f),
+                                   options = list(`actions-box` = TRUE), 
+                                   multiple = TRUE)),
+                
                 column(width = 3,
                        actionButton("go_quality", "Update Filters", class = "btn-success")),
                 
                 column(width = 3,
[…]

### 2021-05-20 app.R
Minor changes and deployment

@@ -1910,7 +1910,7 @@ tabPanel("6: Study Screening",
          h3("Study Screening Results", align = "center"),
          br(),
          p("This plot displays scores from the quality screening exercise developed by", a(href ="https://pubs.acs.org/doi/abs/10.1021/acs.est.0c03057", 'de Ruijter et al. (2020)', .noOWs = "outside"), "with some modification. For more information, including the scoring rubric used, see the document 'Study Screening Rubric' under the Resources tab."),
-         b(),
+         br(),
          p("Interact with the Data: Use your cursor to zoom and hover over the plot to view additional information about each study."),
          br(),
          p("The range of treatment groups used with each study is displayed in the hover over text box."),
@@ -4472,7 +4472,7 @@ output$downloadSsdPlot <- downloadHandler(
             panel.grid.major=element_blank(),
             axis.text.x = element_text(size = 7, angle = 90, vjust = 0.5, hjust = .5),
             plot.title = element_text(hjust = 0.5)) %>% 
-      req(nrow(quality_filtered()) > 0) #Supresses warning message text before submit button is clicked
+      req(nrow(quality_filtered()) > 0) #Suppresses warning message text before submit button is clicked
       
   })
   

### 2021-06-21 simulated-distributions_FishDev.rmd
bits and bobs for migration

--- 2021-03-17 Scott_Modeling/Modelling/simulated distributions.Rmd
@@
 ---
-title: "Simulated Probabilistic Distributions"
+title: "Simulated Probabilistic Distributions_ Algae DEV"
 author: "Scott Coffin"
@@
   mutate(treatments = 4) %>% 
-  mutate(organism.group = "Crustacea") %>% 
+  mutate(organism.group = "Algae") %>% 
   mutate(exposure.duration.d = 14) %>% 
@@
   mutate(exposure.route = "water") %>% 
-  mutate(lvl2_f = "Mortality")
+  mutate(lvl2_f = "Development")
 ```
@@
 ```{r}
-#determine which lvl2_f are most abundant for crustacea data
+#determine which lvl2_f are most abundant for Algae data
 crust <- aoc_z %>% 
@@
  #filter(treatments == 4) %>% 
-  filter(organism.group == "Crustacea") %>% 
+  filter(organism.group == "Algae") %>% 
   filter(exposure.duration.d == 14) %>% 
@@
   filter(exposure.route == "water") %>% 
- filter(lvl2_f == "Mortality")
+ filter(lvl2_f == "Development")
 
@@
 ```
-```{r}
+
+```{r eval=FALSE, include=FALSE}
 crust2 <- crust %>% drop_na(effect_f) %>%  mutate_if(~is.numeric(.) && (.) > 0, log10)
@@
   labs(title = "Cumulative Toxicity Probability",
-       subtitle = "n = 1000 particles/L; Crustacea acute Mortality",
+       subtitle = "n = 1000 particles/L; Algae acute Development",
        caption = "Predicted Toxicity from Random Forest")
@@
   mutate(exposure.route = "water") %>% 
-  mutate(lvl2_f = "Mortality") %>% 
+  mutate(lvl2_f = "Development") %>% 
   mutate(vol.mL = 1) %>% 
@@
   mutate(exposure.route = "water") %>% 
-  mutate(lvl2_f = "Mortality") %>% 
+  mutate(lvl2_f = "Development") %>% 
  mutate(vol.mL = 1) %>% 
@@
   labs(title = "Model-Predicted Probability of Toxicity",
-       subtitle = "Crustacea; 14-d exp; Mortality; NOEC; Early; Marine",
+       subtitle = "Algae; 14-d exp; Development; NOEC; Early; Marine",
        caption = "Random Forest Model; Simulated particle distribution (alpha = 1.6)") +
@@
   # labs(title = "Cumulative Toxicity Probability",
-  #      subtitle = "n = 1000 particles/mL; Crustacea acute Mortality",
+  #      subtitle = "n = 1000 particles/mL; Algae acute Development",
   #      caption = "Predicted Toxicity from Random Forest")
@@
 monoDisperse_100 <- newrow %>%  
-  mutate(dose.particles.mL.master = 120000000) %>%  #120 particles/mL is about EC50 for above
-  mutate(mass.mg = 0.2263657 * dose.particles.mL.master) %>% #average mass x  particles
-  mutate(um3 = 50) %>% 
+ # mutate(dose.particles.mL.master = 120) %>%  #120 particles/mL is about EC50 for above
+  mutate(mass.mg = 0.2263657 * 120) %>% #average mass x 120 particles
+  mutate(um3 = 242182027) %>% 
   mutate(treatments = 4) %>% 
-  mutate(organism.group = "Crustacea") %>% 
+  mutate(organism.group = "Algae") %>% 
   mutate(exposure.duration.d = 14) %>% 
@@
   mutate(exposure.route = "water") %>% 
-  mutate(lvl2_f = "Mortality") %>% 
+  mutate(lvl2_f = "Development") %>% 
   mutate_if(is.character, as.factor)
@@
 #summary(glm(ToxicityProbability ~ Size * Shape * polymer * um3, data = df_640p_p_f))
-p1 <- uniform_pred %>% ggplot(aes(x = dose.particles.mL.master, y = ToxicityProbability, color = polymer)) + geom_jitter(alpha = 0.3)
+p1 <- uniform_pred %>% ggplot(aes(x = dose.particles.mL.master, y = ToxicityProbability, color = polymer)) + geom_point()
 
@@
 # SSD
+
+Create simulated SSD
+Test time.
+
+BOOTSTRAP to see if it fits the same each time.
+
+TRY DIFFERENT MODEL TO PROVE THAT OVERFITTING IS NOT OCCURING.
 

### 2021-06-21 Modelling/Generate_random_size_CSF_density.R

@@ -1,0 +1,103 @@
+rm(list=ls())
+
+n = 10000
+## Generate values for the three distributions
+
+
+#install.packages("msm")
+library(msm) ## rtnorm - get upper and lower limit of shape distribution
+
+#install.packages("GeneralizedHyperbolic")
+library(GeneralizedHyperbolic) ## normal-inverse Gaussian
+
+##++++++++++++++++++++++++++++++++++++++++
+## SIZE DISTRIBUTION
+##++++++++++++++++++++++++++++++++++++++++
+
+xmin = 1 #UM
+alpha = 1.6
+
+X.func <- function (X){
+  success <- FALSE
+  while (!success){
+    U = runif(1, 0, 1)
+    X = xmin*(1-U)^(1/(1-alpha))
+    success <- X < 5000} ##should be smaller than 5000 um 
+  return(X)
+}
+
+Data <- data.frame(Size = numeric(0));
+
+for(i in 1:n){
+  X <- X.func()
+  Data <- rbind(Data, X)
+}
+
+colnames(Data) <- c("Size")
+
+min(Data$Size) ##20 um
+max(Data$Size) ##5000 um
+
+##++++++++++++++++++++++++++++++++++++++++
+## SHAPE DISTRIBUTION
+##++++++++++++++++++++++++++++++++++++++++
+
+mu1 <- 0.08
+mu2 <- 0.44
+sd1 <- 0.03
+sd2 <- 0.19
+lambda1 <- 0.06
+lambda2 <- 0.94
+
+
+#Sample N random uniforms U
+U =runif(n)
+
+#Sampling from the mixture
+for(i in 1:n){
+  if(U[i]<lambda1){
+    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
+  }else{
+    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
+  }
+}
+
+min(Data$Shape) ##0
+max(Data$Shape) ##1
+
+##++++++++++++++++++++++++++++++++++++++++
+## DENSITY DISTRIBUTION
+##++++++++++++++++++++++++++++++++++++++++
+
+d.alpha = 73.8 #tail heaviness
+d.beta = 69.9  #asymmetry
+d.mu = 0.840   #location
+d.delta = 0.0972 #scale
+
+D.func <- function (D){
+  success <- FALSE
+  while (!success){
[…]

### 2021-06-21 test/app.R

@@ -1,0 +1,30 @@
+library(shiny)
+library(shinyjs)
+library(bslib)
+library(thematic)
+#automatically theme plots
+thematic_shiny()
+
+light <- bs_theme()
+dark <- bs_theme(bg = "black", fg = "white", primary = "purple")
+ui <- fluidPage(
+  theme = light, 
+  div(
+    class = "custom-control custom-switch", 
+    tags$input(
+      id = "dark_mode", type = "checkbox", class = "custom-control-input",
+      onclick = HTML("Shiny.setInputValue('dark_mode', document.getElementById('dark_mode').value);")
+    ),
+    tags$label(
+      "Dark mode", `for` = "dark_mode", class = "custom-control-label"
+    )
+  )
+)
+server <- function(input, output, session) {
+  observe({
+    session$setCurrentTheme(
+      if (isTRUE(input$dark_mode)) dark else light
+    )
+  })
+}
+shinyApp(ui, server)

### 2021-06-21 testApp.R

@@ -1,0 +1,25 @@
+library(shiny)
+library(shinyjs)
+light <- bs_theme()
+dark <- bs_theme(bg = "black", fg = "white", primary = "purple")
+ui <- fluidPage(
+  theme = light, 
+  div(
+    class = "custom-control custom-switch", 
+    tags$input(
+      id = "dark_mode", type = "checkbox", class = "custom-control-input",
+      onclick = HTML("Shiny.setInputValue('dark_mode', document.getElementById('dark_mode').value);")
+    ),
+    tags$label(
+      "Dark mode", `for` = "dark_mode", class = "custom-control-label"
+    )
+  )
+)
+server <- function(input, output, session) {
+  observe({
+    session$setCurrentTheme(
+      if (isTRUE(input$dark_mode)) dark else light
+    )
+  })
+}
+shinyApp(ui, server)

### 2021-06-21 umap_demo.R

@@ -1,0 +1,65 @@
+library(uwot)
+
+# See function man page for help
+?umap
+
+# Non-numeric columns are ignored, so in a lot of cases you can pass a data
+# frame directly to umap
+iris_umap <- umap(iris, n_neighbors = 50, learning_rate = 0.5, init = "random")
+
+# Load mnist from somewhere, e.g.
+#devtools::install_github("jlmelville/snedata")
+mnist <- snedata::download_mnist()
+mnist_umap <- umap(mnist, n_neighbors = 15, min_dist = 0.001, verbose = TRUE)
+
+# For high dimensional datasets (> 100-1000 columns) using PCA to reduce
+# dimensionality is highly recommended to avoid the nearest neighbor search
+# taking a long time. Keeping only 50 dimensions can speed up calculations
+# without affecting the visualization much
+mnist_umap <- umap(mnist, pca = 50)
+
+# Use a specific number of threads
+mnist_umap <- umap(mnist, n_neighbors = 15, min_dist = 0.001, verbose = TRUE, n_threads = 8)
+
+# Use a different metric
+mnist_umap_cosine <- umap(mnist, n_neighbors = 15, metric = "cosine", min_dist = 0.001, verbose = TRUE, n_threads = 8)
+
+# If you are only interested in visualization, `fast_sgd = TRUE` gives a much faster optimization
+mnist_umap_fast_sgd <- umap(mnist, n_neighbors = 15, metric = "cosine", min_dist = 0.001, verbose = TRUE, fast_sgd = TRUE)
+
+# Supervised dimension reduction
+mnist_umap_s <- umap(mnist, n_neighbors = 15, min_dist = 0.001, verbose = TRUE, n_threads = 8,
+                     y = mnist$Label, target_weight = 0.5)
+
+# Add new points to an existing embedding
+mnist_train <- head(mnist, 60000)
+mnist_test <- tail(mnist, 10000)
+
+# You must set ret_model = TRUE to return extra data we need
+# coordinates are in mnist_train_umap$embedding
+mnist_train_umap <- umap(mnist_train, verbose = TRUE, ret_model = TRUE)
+mnist_test_umap <- umap_transform(mnist_test, mnist_train_umap, verbose = TRUE)
+
+# Save the nearest neighbor data
+mnist_nn <- umap(mnist, ret_nn = TRUE)
+# coordinates are now in mnist_nn$embedding
+
+# Re-use the nearest neighor data and save a lot of time
[…]

### 2021-07-26 app.R
Conversion to shiny dashboard and UI revision for intro, overview and exploration tabs

@@ -4,8 +4,6 @@
 
 #### Setup ####
 
-# Anything that should only happen ONCE should be placed in this setup section, prior to the actual shiny structure.
-
 # Load packages
 library(tidyverse) #General everything
 library(RColorBrewer) #color palette
@@ -32,6 +30,7 @@ library(ggsci) #color palettes
 # library(thematic) #complete control over themes (including plots) required for dark mode
 library(collapsibleTree) #plot type for endpoint category tree
 library(hrbrthemes) #theme for screening plot
+library(shinydashboard)
 
 # Load finalized dataset.
 aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
@@ -955,7 +954,7 @@ aoc_ERM_default <- aoc_setup  %>%
   mutate(EC_env_p.particles.mL = EC_poly_p.particles.mL * CF_bio) %>%  #aligned particle effect concentraiton (1-5000 um)
   
   #### Surface area ERM ####
-mutate(mu.sa.mono = particle.surface.area.um2) %>% #define mu_x_mono for alignment to ERM
+mutate(mu.sa.mono = as.numeric(particle.surface.area.um2)) %>% #define mu_x_mono for alignment to ERM
   #calculate lower ingestible surface area
   mutate(x_LL_sa = SAfnx(a = 0.5 * x1D_set, 
                          b = 0.5 * R.ave * x1D_set, 
@@ -972,7 +971,7 @@ mutate(mu.sa.mono = particle.surface.area.um2) %>% #define mu_x_mono for alignme
   mutate(EC_env_sa.particles.mL = EC_poly_sa.particles.mL * CF_bio) %>% 
   #### volume ERM ####
 #define mu_x_mono for alignment to ERM
-mutate(mu.v.mono = particle.volume.um3) %>% 
+mutate(mu.v.mono = as.numeric(particle.volume.um3)) %>% 
   #calculate lower ingestible volume 
   mutate(x_LL_v = volumefnx(R = R.ave,
                             L = x1D_set)) %>%
@@ -1044,9 +1043,10 @@ aoc_endpoint <- aoc_setup %>%
 # light <- bs_theme() # DARK MODE SWITCH
 # dark <- bs_theme(bg = "black", fg = "white", primary = "purple") # DARK MODE SWITCH
 
-# build UI
-ui <- fluidPage(theme = shinytheme("flatly"), #light, 
+ui <- dashboardPage(
+  # theme = shinytheme("flatly"), #light, 
 
[…]

### 2021-07-27 app.R
ssd tab makeover

@@ -1044,9 +1044,8 @@ aoc_endpoint <- aoc_setup %>%
 # dark <- bs_theme(bg = "black", fg = "white", primary = "purple") # DARK MODE SWITCH
 
 ui <- dashboardPage(
-  # theme = shinytheme("flatly"), #light, 
 
-      dashboardHeader(title = "Toxicity of Microplastics Explorer", titleWidth = 400),
+  dashboardHeader(title = "Toxicity of Microplastics Explorer", titleWidth = 400),
                 # #### DARK MODE SWITCH ####
                 #   div(class = "custom-control custom-switch",
                 #     tags$input(id = "dark_mode", type = "checkbox", class = "custom-control-input",
@@ -1054,38 +1053,22 @@ ui <- dashboardPage(
                 #     tags$label("Dark mode", `for` = "dark_mode", class = "custom-control-label")
                 #     ),
                  ####
-  
-  # App title
-  # titlePanel(title=div(img(src = "main_logo.png", width = "10%", height = "10%"), "Toxicity of Microplastics Explorer: Aquatic Organisms")),
-  # 
-  # # Title panel subtext
-  # tags$div("Logo created by J.C. Leapman.", tags$a(href="https://twitter.com/jcleapman", tags$img(src="twitter.png", width="2%", height="2%"))),
-  # tags$div("This website is only intended for use by invited participants of the Microplastics Health Effects Workshop."),
-  # actionButton("database_link", label="Go to Human Health Database", class = "btn-primary", onclick ="window.open('https://sccwrp.shinyapps.io/human_mp_tox_shiny-/', '_blank')", style = "float:right"),
-  # 
-  # br(), # line break
-  
-  # Main panel for displaying outputs
-  # mainPanel(width = 12,
-  #   
-  #     # Output: set of 6 tabs
-  #     tabsetPanel(type = "tabs",
-  
-  dashboardSidebar(width = 150,
+
+  dashboardSidebar(width = 175,
                    
                    sidebarMenu(
                      
                      #Logo image
                      br(),
                      tags$img(src="main_logo_drop.png", width = "100%", height = "100%", style = 'display: block; margin-left: auto; margin-right: auto;'),
-                     br(),
-                     
+                     tags$div("Logo created by J.C. Leapman.", align = 'center', style = 'font-size: 10px; display: block; margin-left: auto; margin-right: auto;'), 
+                     br(),         
                      #List of tabs
                      menuItem("Welcome", tabName = "Welcome", icon = icon("home")),
[…]

### 2021-07-27 app.R
Add widgets for experiment type, search table

@@ -833,6 +833,8 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
   mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
     environment == "Marine" ~ "Marine",
     environment == "Terrestrial" ~ "Terrestrial"))) %>%
+  mutate(weather.biofoul_f = factor(case_when(weather.biofoul == "Y" ~ "Yes",
+                                              weather.biofoul == "N" ~ "No"))) %>% 
   mutate(species_f = as.factor(paste(genus,species))) %>% 
   mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
   mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
@@ -850,18 +852,16 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
   group_by(doi) %>% 
   mutate(treatment_range = ifelse(min(treatments) == max(treatments), paste0(treatments), paste0(min(treatments),"-",max(treatments)))) %>% 
   ungroup() %>% 
-  #Remove leachate and additive/chemical transfer experiments
-  dplyr::filter(leachate.only != "Y") %>%
-  mutate(chem.exp.typ.nominal_f = factor(case_when(chem.exp.typ.nominal == "Particle Only" ~ "Particle Only",
-                                                   chem.exp.typ.nominal == "co.exp" ~ "Chemical Co-Exposure",
-                                                   chem.exp.typ.nominal == "sorbed" ~ "Chemical Transfer"))) %>% 
-  dplyr::filter(chem.exp.typ.nominal_f == "Particle Only") %>% 
+  #Create factors for leachate and additive/chemical transfer experiments
+  mutate(exp_type_f = factor(case_when(leachate.only == "Y" ~ "Leachate",
+                                       chem.exp.typ.nominal == "Particle Only" ~ "Particle Only",
+                                       chem.exp.typ.nominal == "co.exp" ~ "Chemical Co-Exposure",
+                                       chem.exp.typ.nominal == "sorbed" ~ "Chemical Transfer"), levels = c("Particle Only", "Chemical Transfer", "Chemical Co-Exposure", "Leachate"))) %>% 
   #calculate maximum ingestible size (if not already in database)
   mutate(max.size.ingest.mm = ifelse(is.na(max.size.ingest.mm), 
[…]

### 2021-07-27 app.R
set up study screening tab, suppressed erorr messages on ssd plots and added filter to alignmnet to only include particle only data

@@ -931,7 +931,9 @@ R.ave = 0.77 #average width to length ratio for microplastics in marine enviornm
 p.ave = 1.10 #average density in marine surface water
 
 # calculate ERM for each species
-aoc_ERM_default <- aoc_setup  %>% 
+aoc_ERM_default <- aoc_setup  %>%
+  #filter the data to only include particle only data
+  dplyr::filter(exp_type_f == "Particle Only") %>% 
   # define upper size WIDTH for ingestion (based on average width:length ratio)
   mutate(x2M = case_when(is.na(max.size.ingest.um) ~ (1/R.ave) * x2D_set, #all calculations below occur for length. Width is R.ave * length, so correcting here makes width the max size ingest below
                          (max.size.ingest.um * (1/R.ave)) < x2D_set ~ ((1/R.ave) * max.size.ingest.um),
@@ -1249,172 +1251,166 @@ tabItem(tabName = "Screening",
         
         box(title = "Data Selection", status = "primary", width = 12, collapsible = TRUE,
             
+            p("This plot displays scores from the quality screening exercise developed by", a(href ="https://pubs.acs.org/doi/abs/10.1021/acs.est.0c03057", 'de Ruijter et al. (2020)', .noOWs = "outside"), "with some modification. 
+            For more information, including the scoring rubric used, see Resources."),
             
-        )#close box
-        #          h3("Study Screening Results", align = "center"),
-        #          br(),
-        #          p("This plot displays scores from the quality screening exercise developed by", a(href ="https://pubs.acs.org/doi/abs/10.1021/acs.est.0c03057", 'de Ruijter et al. (2020)', .noOWs = "outside"), "with some modification. For more information, including the scoring rubric used, see the document 'Study Screening Rubric' under the Resources tab."),
-        #          br(),
-        #          p("Interact with the Data: Use your cursor to zoom and hover over the plot to view additional information about each study."),
-        #          br(),
-        #          p("The range of treatment groups used with each study is displayed in the hover over text box."),
-        #          br(),
-        #          
-        #          # widget headers
-        #          column(width=12,
-        #                 
-        #                 column(width = 3,
-        #                        h4("Effects")),
-        #                 
-        #                 column(width = 3,
-        #                        h4("Particle Characteristics")),
-        #                 
-        #                 column(width = 3,
-        #                        h4("Biological Factors"))),
-        #          
-        #          # widgets
-        #          
-        #          column(width = 12,
-        #                 
[…]

### 2021-07-27 app.R
Now the error messages are suppressed on diagnostic ssd plots

@@ -4263,10 +4263,10 @@ output$downloadSsdPlot <- downloadHandler(
                 caption = "Predicted species sensitivity distribution concentrations with uncertanties."
                 )
   })
-  
+
   # Cullen and Frey Graph
   output$ssd_CF_plot <- renderPlot({
-    req(aoc_filter_ssd() > 0)
+    req(nrow(aoc_filter_ssd())>0)
     #reactive to static
     aoc_SSD <- aoc_filter_ssd()
     
@@ -4313,7 +4313,7 @@ output$downloadSsdPlot <- downloadHandler(
   #QQ plot
   output$ssd_qq_plot <- renderPlot({
     #req
-    req(aoc_filter_ssd()>0)
+    req(nrow(aoc_filter_ssd())>0)
     
     #reactive to static
     aocFitLNorm <- aocSSDFitLNorm()
@@ -4330,7 +4330,7 @@ output$downloadSsdPlot <- downloadHandler(
   #pp plot
   output$ssd_pp_plot <- renderPlot({
     #req
-    req(aoc_filter_ssd()>0)
+    req(nrow(aoc_filter_ssd())>0)
     
     #reactive to static
     aocFitLNorm <- aocSSDFitLNorm()
@@ -4348,7 +4348,7 @@ output$downloadSsdPlot <- downloadHandler(
   #Histogram
   output$ssd_dens_plot <- renderPlot({
     #req
-    req(aoc_filter_ssd()>0)
+    req(nrow(aoc_filter_ssd())>0)
     
     #reactive report x-axis
     particle_mass_check_ssd <- input$particle_mass_check_ssd

### 2021-07-27 app.R
minor edit

@@ -891,7 +891,7 @@ mux.polyfnx.2 = function(x_UL,x_LL){
 ## function to calcualte min and max ingestible surface area ##
 SAfnx = function(a = a, # a = 0.5 * length
                  b = b, # b = 0.5 * width
-                 c = c # c = 0.5 * height (note that hieght is 0.67 * width)
+                 c = c # c = 0.5 * height (note that height is 0.67 * width)
 ){
   SA = 4*pi*(((a*b)^1.6 + (a*c)^1.6 + (b*c)^1.6) / 3)^(1/1.6)
   return(SA)}

### 2021-07-27 app.R
Added authors and year to data table

@@ -2352,7 +2352,7 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
 #### Search S ####
    
    output$databaseDataTable <- renderDataTable(
-     aoc_setup[,c("doi", "exp_type_f", "env_f", "org_f", "species_f", "life_f", "acute.chronic_f", "size_f", 
+     aoc_setup[,c("doi", "authors", "year","exp_type_f", "env_f", "org_f", "species_f", "life_f", "acute.chronic_f", "size_f", 
                   "size.length.um.used.for.conversions", "shape_f", "poly_f", "weather.biofoul_f", "lvl1_f", "lvl2_f", "lvl3_f", "effect_f")],
      filter = "top",
      extensions = c('Buttons'),
@@ -2361,7 +2361,7 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
        buttons = list(I('colvis'), c('copy', 'csv', 'excel')),
        scrollY = 400,
        scroller = TRUE),
-     colnames = c('DOI', 'Experiment Type', 'Environment', 'Organism Group', 'Species', 'Life Stage', 'Exposure Duration',
+     colnames = c('DOI', 'Authors', 'Year', 'Experiment Type', 'Environment', 'Organism Group', 'Species', 'Life Stage', 'Exposure Duration',
                   'Size Category', 'Mean/Median Particle Size', 'Shape', 'Polymer', 'Weathering/Biofoul', 'Broad Endpoint Category',
                   'Specific Endpoint Category', 'Endpoint', 'Effect'))
    

### 2021-07-27 app.R
Removed dependent checklists for widgets

@@ -1452,7 +1452,13 @@ tabItem(tabName = "Exploration",
                         multiple = TRUE)), # allows for multiple inputs
                       
                       #Specific endpoint selection
-                      column(width = 4, htmlOutput("secondSelection")), # dependent endpoint checklist
+                      column(width = 4,
+                        pickerInput(inputId = "lvl2_check", 
+                        label = "Specific Endpoint Category:", 
+                        choices = levels(aoc_setup$lvl2_f),
+                        selected = levels(aoc_setup$lvl2_f),
+                        options = list(`actions-box` = TRUE),
+                        multiple = TRUE)),
                       
                       #Effect y/n selection
                       column(width = 4,
@@ -1486,7 +1492,12 @@ tabItem(tabName = "Exploration",
                       
                       #species selection
                       column(width = 4,
-                        htmlOutput("SpeciesSelection_exp"), 
+                      pickerInput(inputId = "species_check", 
+                                  label = "Species:", 
+                                  choices = levels(aoc_setup$species_f),
+                                  selected = levels(aoc_setup$species_f),
+                                  options = list(`actions-box` = TRUE),
+                                  multiple = TRUE),
                         
                       #biological organization selection
                         pickerInput(inputId = "bio_check", 
@@ -2367,47 +2378,46 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
    
 #### Exploration S ####
   
-  #Create dependent dropdown checklists: select lvl2 by lvl1.
-  output$secondSelection <- renderUI({
-    
-    lvl1_c <- input$lvl1_check # assign level values to "lvl1_c"
-    
-    aoc_new <- aoc_setup %>% # take original dataset
-      filter(lvl1_f %in% lvl1_c) %>% # filter by level inputs
-      mutate(lvl2_f_new = factor(as.character(lvl2_f))) # new subset of factors
-      
-    pickerInput(inputId = "lvl2_check", 
-      label = "Specific Endpoint Category:", 
-      choices = levels(aoc_new$lvl2_f_new),
-      selected = levels(aoc_new$lvl2_f_new),
-      options = list(`actions-box` = TRUE),
-      multiple = TRUE)})
+  # #Create dependent dropdown checklists: select lvl2 by lvl1.
+  # output$secondSelection <- renderUI({
+  #   
[…]

### 2021-07-28 app.R
Make widgets not reactive to each other and fixing welcome page

@@ -933,7 +933,7 @@ p.ave = 1.10 #average density in marine surface water
 # calculate ERM for each species
 aoc_ERM_default <- aoc_setup  %>%
   #filter the data to only include particle only data
-  dplyr::filter(exp_type_f == "Particle Only") %>% 
+  # dplyr::filter(exp_type_f == "Particle Only") %>% 
   # define upper size WIDTH for ingestion (based on average width:length ratio)
   mutate(x2M = case_when(is.na(max.size.ingest.um) ~ (1/R.ave) * x2D_set, #all calculations below occur for length. Width is R.ave * length, so correcting here makes width the max size ingest below
                          (max.size.ingest.um * (1/R.ave)) < x2D_set ~ ((1/R.ave) * max.size.ingest.um),
@@ -1019,7 +1019,7 @@ mutate(mu.ssa.mono = mu.sa.mono/mu.m.mono) %>% #define mu_x_mono for alignment t
   mutate(EC_env_ssa.particles.mL = EC_poly_ssa.particles.mL * CF_bio)
 
 
-#### SSD AO Setup ####
+#### SSD Setup ####
 
 # Master dataset for SSDs
 aoc_z <- aoc_setup %>% # start with Heili's altered dataset (no filtration for terrestrial data)
@@ -1072,32 +1072,56 @@ ui <- dashboardPage(
                      menuItem("Study Screening", tabName = "Screening", icon = icon("check-circle")),
                      menuItem("Exploration", tabName = "Exploration", icon = icon("chart-bar")),
                      menuItem("SSD", tabName = "SSD", icon = icon("fish")),
-                     menuItem("Resources", tabName = "Resources"),
-                     menuItem("Contact", tabName = "Contact"))
+                     menuItem("Resources", tabName = "Resources", icon = icon("question-circle")),
+                     menuItem("Contact", tabName = "Contact", icon = icon("envelope")),
+                     br(),
+                     br(),
+                     #Twitter icon
+                     menuItem("Follow Us on Twitter!", href = "https://twitter.com/ToMExApp", icon = icon("twitter")))
   
                    ), #End dashboard sidebar
 
   dashboardBody(
     
+    #extends background color automatically
+    tags$head(tags$style(HTML('.content-wrapper { overflow: auto; }'))),
+    
     tabItems(
       
 #### Welcome UI ####        
-                  tabItem(tabName = "Welcome",
-                    br(), 
-                    h3("What is the Microplastics Toxicity Database?", align = "center"), #Section 1
[…]

### 2021-07-28 app.R
Odds and ends formatting

@@ -1097,10 +1097,6 @@ ui <- dashboardPage(
               
               
               box(status = "primary", width = 12,
-                     
-                    #top left box 
-                    column(width = 5, 
-                    p(tags$img(src="welcome.png", width = "100%", height = "100%"))),      
                     
                     #top right box
                     column(width = 7, 
@@ -1115,9 +1111,12 @@ ui <- dashboardPage(
                     toxicity data pertaining to microplastics."),
                   
                     p("Use the side panel on the left of the page to navigate to each section. Each section provides different information or data visualization options. 
-                      More specific instructions may be found within each section."))),
+                      More specific instructions may be found within each section.")),
+                  
+                    #top left box 
+                    column(width = 5, 
+                           p(tags$img(src="welcome.png", width = "100%", height = "100%")))),
               
-              column(width = 12,
                     
                     #bottom left box  
                     box(status = "primary", width = 12, 
@@ -1158,7 +1157,7 @@ ui <- dashboardPage(
                     p(align = "center", a(href = "https://rochmanlab.com/", 'Dr. Chelsea Rochman'),", University of Toronto",
                       tags$a(href="https://twitter.com/ChelseaRochman", icon("twitter"))),
                     p(align = "center", a(href = "https://www.sccwrp.org/about/staff/alvina-mehinto/", 'Dr. Alvina Mehinto'),", Southern California Coastal Water Research Project"), 
-                    p(align = "center", a(href = "https://www.sccwrp.org/about/staff/steve-weisberg/", 'Dr. Steve Weisberg'),", Southern California Coastal Water Research Project"))), 
+                    p(align = "center", a(href = "https://www.sccwrp.org/about/staff/steve-weisberg/", 'Dr. Steve Weisberg'),", Southern California Coastal Water Research Project")), 
                   
[…]

### 2021-07-28 app.R
Minor modification of alignment description

@@ -1641,7 +1641,8 @@ tabItem(tabName = "Exploration",
                       conditionalPanel(condition = "input.dose_check == 'Particles/mL'",
                               br(),
                               h4("Ecologically Relevant Metric Alignment"),
-                              p("A monodisperse effect concentration (e.g. 5 micron spheres) may be re-scaled to a default size range (e.g. 1 - 5,000 microns) using methods described in Kooi et al (2021). Re-scaling to a default size range allows direct comparison to exposure concentrations for a default size range (which may also be re-scaled). The following radio buttons apply corrections for bioavailability (i.e. limiting available particles to max ingestable size), and a further correction for the ecologically relevant metric (ERM). For a given ERM, the threshold may be related to both mono- or polydisperse particles interchangeably so long as the total magnitude of ERM remains the same (Koelmans et al, 2020). If, for example, 'volume' is chosen below as an ERM, the monodisperse effect concentration is first corrected for bioavailability and aligned to whichever default size range the user chooses below. This aligned threshold (in particles/mL) is then multiplied by a correction for polydisperse volume based on the average volumes for the given range of microplastics in the environment "),
+                              p("A monodisperse effect concentration (e.g. 5 micron spheres) may be re-scaled to a default size range (e.g. 1 - 5,000 microns) using methods described in Kooi et al (2021). Re-scaling to a default size range allows direct comparison to exposure concentrations for a default size range (which may also be re-scaled). The following radio buttons apply corrections for bioavailability (i.e. limiting available particles to max ingestable size), and a further correction for the ecologically relevant metric (ERM). For a given ERM, the threshold may be related to both mono- or polydisperse particles interchangeably so long as the total magnitude of ERM remains the same (Koelmans et al, 2020). If, for example, 'volume' is chosen below as an ERM, the monodisperse effect concentration is first corrected for bioavailability and aligned to whichever default size range the user chooses below. This aligned threshold (in particles/mL) is then multiplied by a correction for polydisperse volume based on the average volumes for the given range of microplastics in the environment.",
+                                strong("This analysis will only include 'Particle Only' data.")),
                               
                               #ERM Checkbox
                               column(width = 3,
@@ -1922,13 +1923,8 @@ tabItem(tabName = "SSD",
                           conditionalPanel(condition = "input.particle_mass_check_ssd == 'Particles/mL'",
                                            br(),
                                            h4("Ecologically Relevant Metric Alignment"),
[…]

### 2021-07-28 app.R
Fixed issues with tables and put in filters for particle only data for alignment

@@ -932,8 +932,7 @@ p.ave = 1.10 #average density in marine surface water
 
 # calculate ERM for each species
 aoc_ERM_default <- aoc_setup  %>%
-  #filter the data to only include particle only data
-  # dplyr::filter(exp_type_f == "Particle Only") %>% 
+   
   # define upper size WIDTH for ingestion (based on average width:length ratio)
   mutate(x2M = case_when(is.na(max.size.ingest.um) ~ (1/R.ave) * x2D_set, #all calculations below occur for length. Width is R.ave * length, so correcting here makes width the max size ingest below
                          (max.size.ingest.um * (1/R.ave)) < x2D_set ~ ((1/R.ave) * max.size.ingest.um),
@@ -2442,7 +2441,8 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
        dom = 'Brtip',
        buttons = list(I('colvis'), c('copy', 'csv', 'excel')),
        scrollY = 400,
-       scroller = TRUE),
+       scrollH = TRUE,
+       sScrollX = TRUE),
      colnames = c('DOI', 'Authors', 'Year', 'Experiment Type', 'Environment', 'Organism Group', 'Species', 'Life Stage', 'Exposure Duration',
                   'Size Category', 'Mean/Median Particle Size', 'Shape', 'Polymer', 'Weathering/Biofoul', 'Broad Endpoint Category',
                   'Specific Endpoint Category', 'Endpoint', 'Effect'))
@@ -2749,6 +2749,8 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
     
     # calculate ERM for each species
     aoc_setup <- aoc_setup %>% 
+      #filter the data to only include particle only data
+      dplyr::filter(exp_type_f == "Particle Only") %>%
       # define upper size length for ingestion
       mutate(x2M = max.size.ingest.mm * 1000) %>% #max size ingest in um
       # calculate effect threshold for particles
@@ -3719,7 +3721,9 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
     p.ave = input$p.ave_ssd#1.10 #average density in marine surface water
     
     # calculate ERM for each species
-    aoc_z <- aoc_z %>% 
+    aoc_z <- aoc_z %>%
+      #filter the data to only include particle only data
+      dplyr::filter(exp_type_f == "Particle Only") %>%
       # define upper size WIDTH for ingestion (based on average width:length ratio)
[…]

### 2021-07-28 App_Nonessential/Maybe_Delete/Concentration data/Risk_Characterization_GlobalData_MultipleThresholds.Rmd
Moved all scripts that are non essential to the app to a seperate folder since they are interferring with deployment

---
title: "Global Microplastics Risk Characterization"
author: "Scott Coffin"
date: "3/10/2021"
output:   
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
    includes:
     # after_body: footer.html
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE,time_it = TRUE) #report
```

```{r}
library(tidyverse)
library(calecopal)
library(ssdtools)
library(DT)
library(plotly)
library(gridExtra)
library(grid)
library(wesanderson)
library(ggdark)
library(broom)
library(knitr)
```

# Hazard Concentration and Alignment Parameters
```{r}
# Choose hazard concentration and confidence intervals
#particles/L
Threshold1 = 0.18 #90000 
Threshold2 = 0.38 #383000
Threshold3 = 0.37 #391000
Threshold4 = 2.08 #1e6

kable(data.frame(
  Category = c("Hazard Concentrations (particles/L)"),
  "Threshold1" = Threshold1,
  "Threshold 2" = Threshold2,
  "Threshold 3" = Threshold3,
  "Threshold 4" = Threshold4))
```

```{r}
# Choose concentration alignment parameters
alpha = 2.5 #(use 1.6 or 2.5)
x2D_set = 5000#100 #upper size range (microns)
x1D_set = 100#1 #lower size range (microns)

kable(data.frame(
  Category = c("Concentration Alignment Parameters"),
  alpha = alpha,
  "lower_size_range_microns" = x1D_set,
  "upper_size_range_microns" = x2D_set))
```

```{r}
#data import
# Adam et al (2019) data needs correcting
adam <- read.csv("Concentration data/adam2019.csv", na.strings = "N.A.") %>% 
  mutate(x1M = Min.particle.size.um,
         x2M = Max.particle.size.um) %>% 
  mutate(Sample.Type = "sample")
```

```{r}
# Align Data

#function to derive correction factor (CF) from Koelmans et al (equation 2)
CFfnx = function(a = alpha, #default alpha from Koelmans et al (2020)
                 x2D = x2D_set, #set detault values to convert ranges to (1-5,000 um) #5mm is upper defuault 
                 x1D = x1D_set, #1 um is lower default size
                 x2M, x1M){

  CF = (x2D^(1-a)-x1D^(1-a))/(x2M^(1-a)-x1M^(1-a))

  return(CF)
}
#verify it works (expected answer is 40.37)
#CFfnx(x1M = 333, x2M = 5000)

adam <- adam %>% 
  mutate(CF = CFfnx(x1M = x1M, x2M = x2M)) %>%  #create new column with correction factor 
  mutate(particles.m3.corrected = CF * Single.Measurement.conc....m3.) %>% #convert single concenetrations
  mutate(particles.m3.corrected_mean = CF * Mean.conc....m3.) %>%  #convert mean concentrations from distributions
  mutate(particles.m3.corrected_median = CF * Median.conc....m3.) %>%   #convert mean concentrations from distributions
  mutate(particles.single.median.m3 = ifelse(is.na(particles.m3.corrected), particles.m3.corrected_median, particles.m3.corrected)) %>% 
  mutate(particles.m3.master = ifelse(is.na(particles.single.median.m3), particles.m3.corrected_mean, particles.single.median.m3)) %>% 
  mutate(particle.L.master = particles.m3.master/1000) %>% 
  filter(particle.L.master > 0) %>% 
  mutate(System = factor(System))
```
#Plot

```{r}
# read in concentration data
samplesADAM <- adam %>% 
  mutate(Conc = particle.L.master)

#make new dataframe to plot both histograms together
sampleSimpleADAM <- samplesADAM %>%
  dplyr::select(Conc, Sample.Type) %>% 
  droplevels()

#make new dataframe to plot both histograms together
dfADAM <- rbind(sampleSimpleADAM)#,food.dilution.simple)
```

## Exceedances
```{r}
#calculate exceedance
dfADAM_exceedance <- dfADAM %>% 
  mutate(aboveThreshold1 = factor(case_when(
    Conc >= Threshold1 ~ "above Threshold 1",
    Conc < Threshold1 ~ "below Threshold 1",
  )))

#give summary stat for exceedance
exceedance <- dfADAM_exceedance  %>%
  filter(Sample.Type == "sample") %>% 
  dplyr::select(c(Conc, aboveThreshold1)) %>%
  group_by(aboveThreshold1) %>%
  dplyr::summarize(n = n()) %>% 
  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%"))

kable(exceedance)
```
## Histogram
```{r}
#generate plot
dfADAM_exceedance %>% 
  filter(Sample.Type == "sample") %>% 
  ggplot(aes(x = Conc, fill = aboveThreshold1))+
  geom_histogram(aes(y = ..count../sum(..count..)),bins = 50, alpha = 0.9, position = "identity") +
  #geom_smooth(stat = 'density') +
  geom_vline(xintercept = Threshold1, linetype = "dashed", color = "red") +
  geom_text(aes(x = Threshold1- 0.5*Threshold1, y = 0.055), label = paste(Threshold1,"particles/L"),  color = "red") +
  geom_text(aes(x = Threshold1- 0.5*Threshold1, y = 0.06), label = ("Threshold1"),  color = "red") +
  geom_text(aes(x = Threshold1+ Threshold1, y = 0.050), label = paste(exceedance$rel.freq[1], "Above Threshold1"), color = "red") +
  scale_x_log10() +
  #coord_cartesian(xlim = c(0,100000000)) +
  # scale_x_continuous(labels = scales::scientific) +
  xlab("Concentration (particles/L)")+
  scale_y_continuous(name = "Relative Density", labels = scales::percent)+
  #scale_fill_discrete(labels = c("Environmental Concentration", "LOEC")) +
  #scale_color_discrete(labels = c("Environmental Concentration", "LOEC")) +
  labs(title = "Histograms of Concentrations in Adam et al 2019 Dataset",
       caption = "Adam et al. 2019 data; all data corrected to 1-5,000 um; nominal particle/L",
       fill = "Env. Conc. or Tox. Conc.",
       color = "Env. Conc. or Tox. Conc.") +
  theme_minimal() +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank())
```
## ECDF

```{r}
#ECDF by System
adam %>% 
filter(System != "") %>% 
  ggplot(aes(x = particle.L.master)) + #, color = System))+
  stat_ecdf(geom = "point", size = 1, alpha = 0.7) +
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.3, size = 1.5, color = 'azure3') +
  scale_color_manual(values = wes_palette("Darjeeling1"))+
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'green', size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 'yellow', size = 1.3) +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'orange', size = 1.3) +
  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	'red', size = 1.3) +
  geom_text(label = "Threshold 1", color = 'green', x = Threshold1, y = 0.15, size = 6)+
  geom_text(label = "Threshold 2", color = 'yellow', x = Threshold2, y = 0.22, size = 6)+
  geom_text(label = "Threshold 3", color = 'orange', x = Threshold3, y = 0.35, size = 6)+
  geom_text(label = "Threshold 4", color = 'red', x = Threshold4, y = 0.45, size = 6)+
  ylab("Cumulative Density") +
  xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, "um)"))+
  scale_y_continuous(labels = scales::percent)+
  coord_trans(x = "log10") +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  #theme_minimal() +
  dark_theme_minimal(base_size = 15) +
  labs(title = "Global Surface Water Marine Microplastics Concentrations",
       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
       caption = paste("Concentration data from Adams et al (2019): 23 studies, 57 sampling locations, n = 377; corrected for size ranges. Alpha = ",alpha))+
  theme(
    plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

# Modelling

```{r}
sample_dists_ADAM <- ssd_fit_dists(samplesADAM, #data frame
                              left = "Conc", #string of the column in data with the concentrations
                              # right = left, #string of the column with the right concentration values. If different from left, then the data are considerd to be censored
                              dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), #char vector of distribution anmes
                              computable = FALSE, #flag specifying whether to only return fits with numerically computable standard errors
                              silent = FALSE) #flag indicating whether fits should fail silently

autoplotADAM<- autoplot(sample_dists_ADAM) #plots the distribution in ggplot2
autoplotADAM
ssd_gof(sample_dists_ADAM) #check the goodness of fit
#there are multiple fitting distributions, so check which fits best
sample_gof_ADAM <- ssd_gof(sample_dists_ADAM)
sample_gof_ADAM[order(sample_gof_ADAM$delta), ] #orders by delta. Use the aicc (Akaike's Information Criterion corrected for sample size) for model selection 
write.csv(sample_gof_ADAM,"Concentration data/sample_gof_ADAM.csv")
#choose the distribution that you want to plot
sample_dists_ADAM_choice <- ssd_fit_dists(samplesADAM, #data frame
                                   left = "Conc", #string of the column in data with the concentrations
                                   # right = left, #string of the column with the right concentration values. If different from left, then the data are considerd to be censored
                                   dists = c("lgumbel"), #char vector of distribution anmes
                                   computable = FALSE, #flag specifying whether to only return fits with numerically computable standard errors
                                   silent = FALSE) #flag indicating whether fits should fail silently
set.seed(99)
sample_pred_ADAM <- predict(sample_dists_ADAM_choice,
                       average = FALSE,
                       ic = "aicc",
                       nboot = 10,
                       ci= TRUE) #estimates model-averaged estimates based on aicc

sample_pred_ADAM # The resultant object is a data frame of the estimated concentration (est) with standard error (se) and lower (lcl) and upper (ucl) 95% confidence limits by percent of species affected (percent). The confidence limits are estimated using parametric bootstrapping.

```

```{r}
sample_pred_ADAM %>% mutate_if(is.numeric, ~ signif(., 3)) %>% 
  datatable(rownames = FALSE,
            extensions = c('Buttons', 'Scroller'),
            options = list(
              dom = 'Brftp',
              scrollY = 400,
              scroller = TRUE,
              buttons = c('copy', 'csv', 'excel')), 
            class = "compact",
            colnames = c("Percent", "Estimated Mean Concentration", "Standard Error", "Lower 95% Confidence Limit", "Upper 95% Confidence Limit", "Distribution"),
            caption = "Predicted Concentration distribution with uncertanties."
  )

#order data
samplesADAM <- samplesADAM %>% 
  filter(System != "") #take out blanks

sampleSSDADAM <- samplesADAM[order(samplesADAM$Conc), ]
sampleSSDADAM$frac <- ppoints(samplesADAM$Conc, 0.5)
```

```{r}
ECDF_model_occurrence_ADAM <- ggplot(sample_pred_ADAM,aes_string(x = "est")) +
  geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "#81a88d", fill = "#81a88d") +
  geom_line(aes_string(y = "percent/100"), linetype = 'dashed', alpha = 0.8) +
  geom_point(data = sampleSSDADAM,aes(x = Conc, y =frac, color = System), size =1) + 
  #geom_text(data = sampleSSD, aes(x = Conc, y = frac, label = Location), hjust = 1.1, size = 4) + #season labels
  scale_y_continuous("Cumulative Distribution (%)", labels = scales::percent) +
  #expand_limits(y = c(0, 1)) +
  xlab("Concentration (particles/L)")+
  labs(title = "Adam et al 2019 Microplastics Concentration Cumulative Distribution Function",
       subtitle = "Smoothing/95% CI ribbon based on average of log-logical and log-normal Distributions Fit",
       caption = "Adam et al 2019 data; sampling corrected to 1-5,000 um") +
  coord_trans(x = "log10") +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
  scale_color_manual(values = wes_palette("Darjeeling2"))

#white mode
ECDF_model_occurrence_ADAM_white <- ECDF_model_occurrence_ADAM +
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 	'red') +
  geom_text(label = "5% HC: 95% LCL", color = 'red', x = Threshold3, y = 0)+
  geom_text(label = "5% hazard concentration", color = 'red', x = 110, y = 0.03)+
  geom_text(label = "5% HC: 95% UCL", color = 'red', x = Threshold2, y = 0)+
  geom_text(x = 110, y = 0, label = paste(Threshold1,"particles/L"), color = 'red') +  #label for hazard conc
  #geom_hline(yintercept = 0.925, linetype = 'twodash', color = "#A2A475") +
  #geom_text(label = "92.5% samples below 5% HC Mean", x = 4.5, y = 0.94, color = "#A2A475") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12))
ECDF_model_occurrence_ADAM_white

```

```{r}
#dark mode
ECDF_model_occurrence_ADAM_dark <- ECDF_model_occurrence_ADAM +
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 	'red') +
  geom_text(label = "95% LCL", color = 'red', x = Threshold3, y = 0, size = 5)+
  geom_text(label = "5% HC", color = 'red', x = Threshold1, y = 0, size = 5)+
  geom_text(label = "95% UCL", color = 'red', x = Threshold2, y = 0, size = 5)+
  geom_text(x = Threshold1, y = 0.1, label = paste(Threshold1,"particles/L"), color = 'red', size =5) +  #label for hazard conc
  #geom_hline(yintercept = 0.925, linetype = 'twodash', color = "yellow") +
  #geom_text(label = "92.5%", x = 3.0, y = 0.96, color = "yellow", size = 6) +
  dark_theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12))
ECDF_model_occurrence_ADAM_dark
```

### 2021-07-28 App_Nonessential/Maybe_Delete/Risk Characterization/Risk_Characterization_GlobalData_MultipleThresholds.Rmd
Moved all scripts that are non essential to the app to a seperate folder since they are interferring with deployment

--- 2021-07-28 App_Nonessential/Maybe_Delete/Concentration data/Risk_Characterization_GlobalData_MultipleThresholds.Rmd
@@
 x2D_set = 5000#100 #upper size range (microns)
-x1D_set = 100#1 #lower size range (microns)
+x1D_set = 1#1 #lower size range (microns)
 
@@
 # Adam et al (2019) data needs correcting
-adam <- read.csv("Concentration data/adam2019.csv", na.strings = "N.A.") %>% 
+adam <- read.csv("Risk Characterization/datasets/adam2019.csv", na.strings = "N.A.", fileEncoding="UTF-8-BOM") %>% 
   mutate(x1M = Min.particle.size.um,
+         x2M = Max.particle.size.um) %>% 
+  mutate(Sample.Type = "sample")
+
+#SF bay data from Zhu et al (2021)
+sfBay <- read.csv("Risk Characterization/datasets/SFBayData.csv", na.strings = "NA", fileEncoding="UTF-8-BOM", stringsAsFactors = TRUE) %>% 
+  rename(x1M = Min.particle.size.um,
          x2M = Max.particle.size.um) %>% 
@@
   filter(particle.L.master > 0) %>% 
-  mutate(System = factor(System))
+  mutate(System = factor(System)) %>% 
+  mutate(source = "Adam")
+
+# correct aquatic concentrations for SF Bay
+sfBay_sub <- sfBay %>% 
+  mutate(CF = CFfnx(x1M = x1M, x2M = x2M)) %>%  #create new column with correction factor 
+  mutate(particle.L.master = particles.L.blank.corrected * CF) %>% 
+  mutate(source = "SFEI") %>% 
+  mutate(Reference = "Zhu et. al 2021") %>% 
+  rename(Sampling.location = general.location) %>% 
+  select(-c(particles.L.blank.corrected, particles.kg.blank.corrected, particles.fish.blank.corrected))
+
+## join datasets ##
+#first ensure that both datasets share all columns
+adam_sub <- adam %>% 
+  select(c(Sampling.apparatus, Sampling.location, x1M, x2M, Sample.Type, System, CF,particle.L.master, source, Reference)) %>% 
+  mutate(sampleID = NA,
+         specific.location = NA,
+         latitude = NA,
+         longitude = NA,
+         sample.matrix = NA,
+         sample.matrix.specific = NA
+         )
+
+#join data
+globalData <- rbind(adam_sub, sfBay_sub) %>% 
+  mutate(Conc = particle.L.master)
 ```
@@
 ```{r}
-# read in concentration data
-samplesADAM <- adam %>% 
-  mutate(Conc = particle.L.master)
-
 #make new dataframe to plot both histograms together
-sampleSimpleADAM <- samplesADAM %>%
+sampleSimpleADAM <- globalData %>%
   dplyr::select(Conc, Sample.Type) %>% 

### 2021-07-28 App_Nonessential/TeamX_Manuscript_Figures/Manuscript_Figures_TeamX.Rmd
Moved all scripts that are non essential to the app to a seperate folder since they are interferring with deployment

--- 2021-05-20 TeamX_Manuscript_Figures/Manuscript_Figures_TeamX.Rmd
@@
 library(tidyverse) 
+library(wesanderson)
 
@@
 ```
+
+# Leah
+
+```{r Leah, warning = FALSE, message = FALSE, fig.height = 7, fig.width = 12}
+
+a <- aoc_setup %>%
+  #Remove studies that don't report doses in mass OR counts per volume
+  filter_at(vars(dose.mg.L.master, dose.particles.mL.master), any_vars(!is.na(.))) %>% 
+  #Designate if doses were reported in mass, counts or both
+  mutate(mass_count = ifelse(dose.mg.L.master.converted.reported == "reported" & dose.particles.mL.master.converted.reported == "converted", 1, 0)) %>%  
+  mutate(particles_count = ifelse(dose.particles.mL.master.converted.reported == "reported" & dose.mg.L.master.converted.reported == "converted", 1, 0)) %>% 
+  mutate(both_count = ifelse(dose.particles.mL.master.converted.reported == "reported" & dose.mg.L.master.converted.reported == "reported", 1, 0)) %>% 
+  #Unique particles within each study
+  distinct(doi, year, poly_f, shape_f, size.length.um.used.for.conversions, dose.mg.L.master.converted.reported, dose.particles.mL.master.converted.reported, mass_count, particles_count, both_count) %>%
+  #Remove letters from years if they exist
+  mutate(year = str_remove(year,"[abc]")) %>% 
+  #Count of the frequency of doses reported in mass, count or both within each year
+  group_by(year) %>% 
+  mutate(`Mass Only` = sum(mass_count, na.rm = TRUE)) %>% 
+  mutate(`Particles Only` = sum(particles_count, na.rm = TRUE)) %>% 
+  mutate(Both = sum(both_count, na.rm = TRUE)) %>% 
+  ungroup() %>% 
+  distinct(year, Both, `Mass Only`, `Particles Only`) %>% 
+  pivot_longer(!year, names_to = "dose_metric", values_to = "freq")
+
+b <- a %>%
+  ggplot(aes(x = year, y = freq, fill = fct_rev(dose_metric))) +
+  geom_bar(position = "dodge", stat = "identity") +
+  theme_classic() +
+  scale_fill_manual(values = c("#3B9AB2", "#EBCC2A", "#F21A00")) +
+  scale_y_continuous(expand = c(0,0)) +
+  labs(y = "Frequency", x = "Year", fill = "Dose Metric") +
+  theme(text = element_text(size = 16))
+
+plot(b)
+
+#Find studies that use multiple sizes
+c <- aoc_setup %>% 
+  group_by(doi, size.length.um.used.for.conversions) %>% 
+  summarise() %>% 
+  ungroup() %>% 
+  count(doi)
+
+#Find studies that use multiple shapes
+d <- aoc_setup %>% 
+  group_by(doi, shape_f) %>% 
+  summarise() %>% 
+  ungroup() %>% 
+  count(doi)
+
+#Find studies that use multiple polymers
+e <- aoc_setup %>% 
+  group_by(doi, poly_f) %>% 
+  summarise() %>% 
+  ungroup() %>% 
+  count(doi)
+```

### 2021-07-28 App_Nonessential/simulated-distributions_crustDev.Rmd
Moved all scripts that are non essential to the app to a seperate folder since they are interferring with deployment

--- 2021-03-17 Scott_Modeling/Modelling/simulated distributions.Rmd
@@
   labs(title = "Cumulative Toxicity Probability",
-       subtitle = "n = 1000 particles/L; Crustacea acute Mortality",
+       subtitle = "n = 1000 particles/L; Crustacea acute Development",
        caption = "Predicted Toxicity from Random Forest")
@@
   labs(title = "Model-Predicted Probability of Toxicity",
-       subtitle = "Crustacea; 14-d exp; Mortality; NOEC; Early; Marine",
+       subtitle = "Crustacea; 14-d exp; Development; NOEC; Early; Marine",
        caption = "Random Forest Model; Simulated particle distribution (alpha = 1.6)") +
@@
   # labs(title = "Cumulative Toxicity Probability",
-  #      subtitle = "n = 1000 particles/mL; Crustacea acute Mortality",
+  #      subtitle = "n = 1000 particles/mL; Crustacea acute Development",
   #      caption = "Predicted Toxicity from Random Forest")
@@
 monoDisperse_100 <- newrow %>%  
-  mutate(dose.particles.mL.master = 120000000) %>%  #120 particles/mL is about EC50 for above
-  mutate(mass.mg = 0.2263657 * dose.particles.mL.master) %>% #average mass x  particles
-  mutate(um3 = 50) %>% 
+ # mutate(dose.particles.mL.master = 120) %>%  #120 particles/mL is about EC50 for above
+  mutate(mass.mg = 0.2263657 * 120) %>% #average mass x 120 particles
+  mutate(um3 = 242182027) %>% 
   mutate(treatments = 4) %>% 

### 2021-07-30 App_Nonessential/Maybe_Delete/Concentration data/Risk_Characterization_GlobalData.Rmd

--- 2021-07-28 App_Nonessential/Maybe_Delete/Concentration data/Risk_Characterization_GlobalData_MultipleThresholds.Rmd
@@
 # Choose hazard concentration and confidence intervals
-#particles/L
-Threshold1 = 0.18 #90000 
-Threshold2 = 0.38 #383000
-Threshold3 = 0.37 #391000
-Threshold4 = 2.08 #1e6
+HC5LCL = 11 #95% lower CLHC5
+HC5 = 75.6 #HC5
+HC5UCL = 521 #95% upper CL for HC5
 
@@
   Category = c("Hazard Concentrations (particles/L)"),
-  "Threshold1" = Threshold1,
-  "Threshold 2" = Threshold2,
-  "Threshold 3" = Threshold3,
-  "Threshold 4" = Threshold4))
+  "HC5_95_Lower_CL" = HC5LCL,
+  HazardConcentration5 = HC5,
+  "HC5_95_Upper_CL" = HC5UCL))
 ```
@@
 alpha = 2.5 #(use 1.6 or 2.5)
-x2D_set = 5000#100 #upper size range (microns)
-x1D_set = 100#1 #lower size range (microns)
+x2D = 5000 #upper size range (microns)
+x1D = 1 #lower size range (microns)
 
@@
   alpha = alpha,
-  "lower_size_range_microns" = x1D_set,
-  "upper_size_range_microns" = x2D_set))
+  "lower_size_range_microns" = x1D,
+  "upper_size_range_microns" = x2D))
 ```
@@
 CFfnx = function(a = alpha, #default alpha from Koelmans et al (2020)
-                 x2D = x2D_set, #set detault values to convert ranges to (1-5,000 um) #5mm is upper defuault 
-                 x1D = x1D_set, #1 um is lower default size
+                 x2D = 5000, #set detault values to convert ranges to (1-5,000 um) #5mm is upper defuault 
+                 x1D = 1, #1 um is lower default size
                  x2M, x1M){
@@
 
-## Exceedances
 ```{r}
@@
 dfADAM_exceedance <- dfADAM %>% 
-  mutate(aboveThreshold1 = factor(case_when(
-    Conc >= Threshold1 ~ "above Threshold 1",
-    Conc < Threshold1 ~ "below Threshold 1",
+  mutate(aboveHC5 = factor(case_when(
+    Conc >= HC5 ~ "above HC5",
+    Conc < HC5 ~ "below HC5",
   )))
@@
   filter(Sample.Type == "sample") %>% 
-  dplyr::select(c(Conc, aboveThreshold1)) %>%
-  group_by(aboveThreshold1) %>%
+  dplyr::select(c(Conc, aboveHC5)) %>%
+  group_by(aboveHC5) %>%
   dplyr::summarize(n = n()) %>% 
@@
 ```
-## Histogram
+
 ```{r}
@@
   filter(Sample.Type == "sample") %>% 
-  ggplot(aes(x = Conc, fill = aboveThreshold1))+
+  ggplot(aes(x = Conc, fill = aboveHC5))+
   geom_histogram(aes(y = ..count../sum(..count..)),bins = 50, alpha = 0.9, position = "identity") +
   #geom_smooth(stat = 'density') +
-  geom_vline(xintercept = Threshold1, linetype = "dashed", color = "red") +
-  geom_text(aes(x = Threshold1- 0.5*Threshold1, y = 0.055), label = paste(Threshold1,"particles/L"),  color = "red") +
-  geom_text(aes(x = Threshold1- 0.5*Threshold1, y = 0.06), label = ("Threshold1"),  color = "red") +
-  geom_text(aes(x = Threshold1+ Threshold1, y = 0.050), label = paste(exceedance$rel.freq[1], "Above Threshold1"), color = "red") +
+  geom_vline(xintercept = HC5, linetype = "dashed", color = "red") +
+  geom_text(aes(x = HC5- 0.5*HC5, y = 0.055), label = paste(HC5,"particles/L"),  color = "red") +
+  geom_text(aes(x = HC5- 0.5*HC5, y = 0.06), label = ("HC5"),  color = "red") +
+  geom_text(aes(x = HC5+ HC5, y = 0.050), label = paste(exceedance$rel.freq[1], "Above HC5"), color = "red") +
   scale_x_log10() +
@@
 ```
-## ECDF
 
@@
 filter(System != "") %>% 
-  ggplot(aes(x = particle.L.master)) + #, color = System))+
-  stat_ecdf(geom = "point", size = 1, alpha = 0.7) +
-  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.3, size = 1.5, color = 'azure3') +
+  ggplot(aes(x = particle.L.master, color = System))+
+  stat_ecdf(geom = "point", size = 2) +
+  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.6, size = 1.5) +
   scale_color_manual(values = wes_palette("Darjeeling1"))+
-  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'green', size = 1.3) +
-  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 'yellow', size = 1.3) +
-  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'orange', size = 1.3) +
-  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	'red', size = 1.3) +
-  geom_text(label = "Threshold 1", color = 'green', x = Threshold1, y = 0.15, size = 6)+
-  geom_text(label = "Threshold 2", color = 'yellow', x = Threshold2, y = 0.22, size = 6)+
-  geom_text(label = "Threshold 3", color = 'orange', x = Threshold3, y = 0.35, size = 6)+
-  geom_text(label = "Threshold 4", color = 'red', x = Threshold4, y = 0.45, size = 6)+
+  geom_vline(xintercept = HC5, linetype = 'dashed', color = 'red') +
+  geom_vline(xintercept = HC5LCL, linetype = 'dashed', color = 'red') +
+  geom_vline(xintercept = HC5UCL, linetype = 'dashed', color = 	'red') +
+  geom_text(label = "95% LCL", color = 'red', x = log10(13), y = 0.15)+
+  geom_text(label = paste("HC5:",HC5, "particles/L"), color = 'red', x = log10(105), y = 0.07)+
+  geom_text(label = "95% UCL", color = 'red', x = log10(440), y = 0.15)+
   ylab("Cumulative Density") +
-  xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, "um)"))+
+  xlab("Particles/L (1-5,000 um)")+
   scale_y_continuous(labels = scales::percent)+
-  coord_trans(x = "log10") +
-  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
+  scale_x_continuous(trans = "log10") +
   annotation_logticks(sides = "b")+ #log scale rick marks on bottom
-  #theme_minimal() +
-  dark_theme_minimal(base_size = 15) +
-  labs(title = "Global Surface Water Marine Microplastics Concentrations",
-       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
-       caption = paste("Concentration data from Adams et al (2019): 23 studies, 57 sampling locations, n = 377; corrected for size ranges. Alpha = ",alpha))+
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 18),
+  theme_minimal() +
+  labs(title = "Global Concentrations ECDF by System",
+       subtitle = "Particles/L corrected to 1-5,000 um",
+       caption = "Concentration data from Adams et al (2019); corrected for size via Koelmans. Hazard Concentration from Koelmans et al (2020)")+
+  theme(plot.title = element_text(hjust = 0.5, size = 18),
         plot.subtitle = element_text(hjust = 0.5, size = 14))
@@
 ECDF_model_occurrence_ADAM_white <- ECDF_model_occurrence_ADAM +
-  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'red') +
-  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 'red') +
-  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 	'red') +
-  geom_text(label = "5% HC: 95% LCL", color = 'red', x = Threshold3, y = 0)+
+  geom_vline(xintercept = HC5, linetype = 'dashed', color = 'red') +
+  geom_vline(xintercept = HC5LCL, linetype = 'dashed', color = 'red') +
+  geom_vline(xintercept = HC5UCL, linetype = 'dashed', color = 	'red') +
+  geom_text(label = "5% HC: 95% LCL", color = 'red', x = HC5LCL, y = 0)+
   geom_text(label = "5% hazard concentration", color = 'red', x = 110, y = 0.03)+
-  geom_text(label = "5% HC: 95% UCL", color = 'red', x = Threshold2, y = 0)+
-  geom_text(x = 110, y = 0, label = paste(Threshold1,"particles/L"), color = 'red') +  #label for hazard conc
+  geom_text(label = "5% HC: 95% UCL", color = 'red', x = HC5UCL, y = 0)+
+  geom_text(x = 110, y = 0, label = paste(HC5,"particles/L"), color = 'red') +  #label for hazard conc
   #geom_hline(yintercept = 0.925, linetype = 'twodash', color = "#A2A475") +
@@
 ECDF_model_occurrence_ADAM_dark <- ECDF_model_occurrence_ADAM +
-  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'red') +
-  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 'red') +
-  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 	'red') +
-  geom_text(label = "95% LCL", color = 'red', x = Threshold3, y = 0, size = 5)+
-  geom_text(label = "5% HC", color = 'red', x = Threshold1, y = 0, size = 5)+
-  geom_text(label = "95% UCL", color = 'red', x = Threshold2, y = 0, size = 5)+
-  geom_text(x = Threshold1, y = 0.1, label = paste(Threshold1,"particles/L"), color = 'red', size =5) +  #label for hazard conc
+  geom_vline(xintercept = HC5, linetype = 'dashed', color = 'red') +
+  geom_vline(xintercept = HC5LCL, linetype = 'dashed', color = 'red') +
+  geom_vline(xintercept = HC5UCL, linetype = 'dashed', color = 	'red') +
+  geom_text(label = "95% LCL", color = 'red', x = HC5LCL, y = 0, size = 5)+
+  geom_text(label = "5% HC", color = 'red', x = HC5, y = 0, size = 5)+
+  geom_text(label = "95% UCL", color = 'red', x = HC5UCL, y = 0, size = 5)+
+  geom_text(x = HC5, y = 0.1, label = paste(HC5,"particles/L"), color = 'red', size =5) +  #label for hazard conc
   #geom_hline(yintercept = 0.925, linetype = 'twodash', color = "yellow") +

### 2021-07-30 App_Nonessential/Shiny_Manuscript_Figures/Manuscript_Figures.Rmd

--- 2021-05-20 Shiny_Manuscript_Figures/Manuscript_Figures.Rmd
@@
 
-```{r collapsible tree example}
+```{r collapsible tree example, include=FALSE}
 
@@
 
-```{r endpoints and organism groups, fig.width = 15}
+```{r endpoints and organism groups, fig.width = 15, include=FALSE}
 library(viridis)
@@
 
-```{r particle characteristics}
+```{r particle characteristics, include=FALSE}
 

### 2021-07-30 App_Nonessential/Threshold_Manuscript_Figures/Confidence_Voting_Figures.Rmd

---
title: "Confidence Voting Figures"
author: "Scott Coffin"
date: "6/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(reshape2)
```

```{r}
votes <- read_excel("Threshold_Manuscript_Figures/voting_ambient.xlsx", sheet  ="Tidy") %>% 
  mutate(Value = as.numeric(Value)) %>% 
  mutate_if(is.character,as.factor)
```

```{r}
library(ggplot2) 
library(grid)
library(RColorBrewer)

make_gradient <- function(deg = 45, n = 100, cols = blues9) {
  cols <- colorRampPalette(cols)(n + 1)
  rad <- deg / (180 / pi)
  mat <- matrix(
    data = rep(seq(0, 1, length.out = n) * cos(rad), n),
    byrow = TRUE,
    ncol = n
  ) +
  matrix(
    data = rep(seq(0, 1, length.out = n) * sin(rad), n),
    byrow = FALSE,
    ncol = n
  )
  mat <- mat - min(mat)
  mat <- mat / max(mat)
  mat <- 1 + mat * n
  mat <- matrix(data = cols[round(mat)], ncol = n)
  grid::rasterGrob(
    image = mat,
    width = unit(1, "npc"),
    height = unit(1, "npc"), 
    interpolate = TRUE
  )
}
```

#prep
```{r}
#get data in wide format
cast <- votes %>% 
  #rowid_to_column(var = "rowid") %>% 
  #dcast(Threshold + ERM + Round + Person ~  Confidence.Type, value.var = "Value")
  dcast(... ~ Confidence.Type, value.var = "Value")

final <- cast %>% group_by(Person, Threshold, Round, ERM) %>%
    summarise(Confidence = sum(Confidence, na.rm = T),
    `Level of Agreement` = sum(`Level of Agreement`, na.rm = T), 
    `Weight of evidence` = sum(`Weight of evidence`, na.rm = T)) %>% 
  na_if(0)

### Repeat for confidence in threshold approach
#get data in wide format
approach <- votes %>% 
  filter(ERM == "overall") %>% 
  #rowid_to_column(var = "rowid") %>% 
  #dcast(Threshold + ERM + Round + Person ~  Confidence.Type, value.var = "Value")
  dcast(... ~ Threshold, value.var = "Value")
```

#Plot
```{r}
g <- make_gradient(
  deg = 135, n = 500, cols = rev(brewer.pal(5, "Spectral"))
)

#make plot
final %>% 
  group_by(Threshold, Round, ERM) %>% 
  summarize(mean.agreement = mean(`Level of Agreement`, na.rm = TRUE),
            sd.agreement = sd(`Level of Agreement`, na.rm = TRUE),
            mean.weight = mean(`Weight of evidence`, na.rm = TRUE),
            sd.weight = sd(`Weight of evidence`, na.rm = TRUE)) %>% 
  filter(ERM != "overall") %>% 
 # filter(Round == "Round 2") %>% 
         #Threshold == "Threshold 1") %>% 
  ggplot(aes(y = mean.agreement, x = mean.weight, color = Round)) +
  #make backgroud first
  annotation_custom(
    grob = g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
  ) +
  #layer points on top
  geom_point() +
  scale_color_manual(name = "Voting Round",
                       values = c("gray1", "gray80"),
                     labels = c("Round 2", "Round 1")) +
  # add error bars
  geom_pointrange(aes(ymin = mean.agreement - sd.agreement, ymax = mean.agreement + sd.agreement)) +
  geom_errorbarh(aes(xmin = mean.weight - sd.weight, xmax = mean.weight + sd.weight)) +
  #give range
  scale_x_continuous(limits = c(1,5),
                     breaks = 1:5,
                     labels = c("Very low", "Low", "Medium", "High", "Very high")) +
  scale_y_continuous(limits = c(1,5),
                     breaks = 1:5,
                     labels = c("Very low", "Low", "Medium", "High", "Very high")) +
 # scale_x_discrete(breaks = 1:3, labels = c("Very low", "Medium", "Very high")) +
  #labels
  labs(title = "Ambient Microplastics Framework Confidence Voting",
       subtitle = "Mean +- s.d. (n = 11)",
         x = "Weight of Evidence",
       y = "Level of Agreement in the outcome among studies") +
theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
            plot.subtitle = element_text(hjust = 0.5)) +
  facet_grid(Threshold ~ ERM)#,
             #nrow = 4)
```
```{r}
approach %>% 
  group_by(Round) %>% 
  summarize(mean.analytical = mean(`Anlaytical Process`, na.rm = TRUE),
            sd.analytical = sd(`Anlaytical Process`, na.rm = TRUE),
            mean.Framework = mean(`Framework`, na.rm = TRUE),
            sd.Framework = sd(`Framework`, na.rm = TRUE)) %>% 
 # filter(Round == "Round 2") %>% 
         #Threshold == "Threshold 1") %>% 
  ggplot(aes(y = mean.analytical, x = mean.Framework, color = Round)) +
  #make backgroud first
  annotation_custom(
    grob = g, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
  ) +
  #layer points on top
  geom_point() +
  scale_color_manual(name = "Voting Round",
                       values = c("gray10", "gray80"),
                     labels = c("Round 2", "Round 1")) +
  # add error bars
  geom_pointrange(aes(ymin = mean.analytical - sd.analytical, ymax = mean.analytical + sd.analytical)) +
  geom_errorbarh(aes(xmin = mean.Framework - sd.Framework, xmax = mean.Framework + sd.Framework)) +
  #give range
  coord_cartesian(ylim = c(1,5),
                  xlim = c(1,5)) +
  scale_x_continuous(
                     breaks = 1:5,
                     labels = c("Very low", "Low", "Medium", "High", "Very high")) +
  scale_y_continuous(
                     breaks = 1:5,
                     labels = c("Very low", "Low", "Medium", "High", "Very high")) +
 # scale_x_discrete(breaks = 1:3, labels = c("Very low", "Medium", "Very high")) +
  #labels
  labs(title = "Ambient Microplastics Framework Confidence Voting",
       subtitle = "Mean +- s.d. (n = 11)",
         x = "Confidence in Analytical Process",
       y = "Confidence in Framework") +
theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
            plot.subtitle = element_text(hjust = 0.5))
```

### 2021-07-30 app.R
moved screen section down, added download plot button, commented out particle only filters

@@ -2,6 +2,8 @@
 #### File created: September 23, 2020
 #### Code contributors: Heili Lowman, Leah Thornton Hampton, Scott Coffin, Emily Darin
 
+
+
 #### Setup ####
 
 # Load packages
@@ -1068,9 +1070,9 @@ ui <- dashboardPage(
                      menuItem("Welcome", tabName = "Welcome", icon = icon("home")),
                      menuItem("Overview", tabName = "Overview", icon = icon("globe")),
                      menuItem("Search", tabName = "Search", icon = icon("search")),
-                     menuItem("Study Screening", tabName = "Screening", icon = icon("check-circle")),
                      menuItem("Exploration", tabName = "Exploration", icon = icon("chart-bar")),
                      menuItem("SSD", tabName = "SSD", icon = icon("fish")),
+                     menuItem("Study Screening", tabName = "Screening", icon = icon("check-circle")),
                      menuItem("Resources", tabName = "Resources", icon = icon("question-circle")),
                      menuItem("Contact", tabName = "Contact", icon = icon("envelope")),
                      br(),
@@ -1709,7 +1711,10 @@ tabItem(tabName = "Exploration",
                      tabPanel("Organism Group",
                         column(width = 12,      
                         plotOutput(outputId = "organism_plot_react", height = "600px")),
-                              
+                        
+                        column(width = 3,
+                               downloadButton("downloadexploration_org", "Download Plot", icon("download"), style="color: #fff; background-color: #337ab7; border-color: #2e6da4")),
+                     
                      ),#closes tab panel
                      
                      tabPanel("Broad Endpoint Category",
@@ -2449,48 +2454,6 @@ server <- function (input, output){  #dark mode: #(input, output, session) {
    
    #### Screening S ####
    
-   # #Specific endpoint (lvl2) widget that is reactive to Broad endpoint (lvl1) widget selections
-   # output$secondSelection_quality <- renderUI({
-   #   
-   #   #Assign lvl1 selection values to variable
-   #   lvl1_c <- input$lvl1_quality 
-   #   
-   #   aoc_new <- aoc_setup %>% 
-   #     #Filter by lvl1 selection
-   #     filter(lvl1_f %in% lvl1_c) %>%
-   #     #Create new subset based on lvl1 selection
[…]

### 2021-07-30 app.R
added options to use ANY dose metric with ANY ERM and split the tab

@@ -863,7 +863,11 @@ aoc_setup <- aoc_v1 %>% # start with original dataset
   mutate(max.size.ingest.mm = ifelse(is.na(max.size.ingest.mm), 
     10^(0.9341 * log10(body.length.cm) - 1.1200) * 10,  #(Jamm et al 2020 Nature paper)correction for cm to mm
     max.size.ingest.mm)) %>%  # if already present, just use that
-  mutate(max.size.ingest.um = 1000 * max.size.ingest.mm) #makes it less confusing below
+  mutate(max.size.ingest.um = 1000 * max.size.ingest.mm) %>%  #makes it less confusing below
+  #calcualte monodisperse unaligned effect concentration for surface area
+  mutate(dose.um2.mL.master = as.numeric(particle.surface.area.um2) * dose.particles.mL.master) %>% 
+  #calculate monodisperse unaligned  specific surface area dose
+  mutate(dose.um2.ug.mL.master = dose.um2.mL.master / (mass.per.particle.mg / 1000)) #correct mg to ug
 
 #### Ecologically Relevant Metric calculations ####
 
@@ -1907,10 +1911,10 @@ tabItem(tabName = "SSD",
                      tabPanel("Dose Metric",
                               
                           column(width = 4,
-                                 radioButtons(inputId = "particle_mass_check_ssd", 
+                                 radioButtons(inputId = "dose_check_ssd", 
                                  label = "Dose Metric:",
-                                 choices = c("Particles/mL", "mg/L", "um3/mL"),
-                                 selected = "mg/L")),
+                                 choices = c("Particles/mL", "µg/mL", "µm3/mL", "µm2/mL", "µm2/µg/mL"),
+                                 selected = "µg/mL")),
                           
                           column(width = 8,
                                  radioButtons(
@@ -1918,17 +1922,11 @@ tabItem(tabName = "SSD",
                                  label = "Do you want to use just the reported, just the converted, or all exposure concentrations?",
                                  choices = list("reported", "converted", "all"),
                                  selected = "all")),
+                     ), #close tabPanel
+                    
+                     tabPanel("Ecologically Relevant Metric Alignment",
                           
[…]

### 2021-08-03 Database_CleanUp_AquaticOrg.R
Trying out dramatically reduced data clean up, will remove redundant files when confident that this hasn't introduced bugs

@@ -1,0 +1,666 @@
+#Script for one-time data clean up to eliminate unnecessary parts of setup. 
+
+
+library(tidyverse)
+
+aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
+
+aoc_v1 <- aoc %>% 
+  mutate(effect = case_when(effect == "Y" ~ "Yes",
+                                   effect == "N" ~ "No")) %>%  
+  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported", chem.exp.typ.nominal = "Particle Only"))
+
+aoc_setup <- aoc_v1 %>% 
+  mutate(size.category = case_when(
+    size.category == 1 ~ "1nm < 100nm",
+    size.category == 2 ~ "100nm < 1µm",
+    size.category == 3 ~ "1µm < 100µm",
+    size.category == 4 ~ "100µm < 1mm",
+    size.category == 5 ~ "1mm < 5mm",
+    size.category == 0 ~ "Not Reported")) %>%    
+  mutate(shape = case_when(
+    shape == "fiber" ~ "Fiber",
+    shape == "fragment" ~ "Fragment",
+    shape == "sphere" ~ "Sphere",
+    shape == "Not Reported" ~ "Not Reported")) %>% 
+  mutate(polymer = case_when(
+    polymer == "BIO" ~ "Biopolymer",
+    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
+    polymer == "LTX" ~ "Latex",
+    polymer == "PA" ~ "Polyamide",
+    polymer == "PE" ~ "Polyethylene",
+    polymer == "PC" ~ "Polycarbonate",
+    polymer == "PET" ~ "Polyethylene Terephthalate",
+    polymer == "PI" ~ "Polyisoprene",
+    polymer == "PMMA" ~ "Polymethylmethacrylate",
+    polymer == "PP" ~ "Polypropylene",
+    polymer == "PS" ~ "Polystyrene",
+    polymer == "PUR" ~ "Polyurethane",
+    polymer == "PVC" ~ "Polyvinylchloride",
+    polymer == "PLA" ~ "Polylactic Acid",
+    polymer == "Not Reported" ~ "Not Reported")) %>%
+  mutate(lvl1 = case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
+                                   lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
[…]

### 2021-08-03 app.R

@@ -8,6 +8,7 @@
 
 # Load packages
 library(tidyverse) #General everything
+library(data.table) #Faster functions for loading data set
 library(RColorBrewer) #color palette
 library(ggplot2) #plotting
 library(ggrepel) #For adding text labels that repel away from data points
@@ -35,7 +36,7 @@ library(hrbrthemes) #theme for screening plot
 library(shinydashboard)
 
 # Load finalized dataset.
-aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
+aoc <- fread("AquaticOrganisms_Clean_test.csv") 
 
 #### Introduction Setup ####
 
@@ -45,7 +46,7 @@ aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
 
 polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
 polyf<-as.data.frame(polydf)%>% #Makes data frame 
-  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
+  filter(effect %in% c("Yes","No"))%>% #Sorts into Yes and No
   mutate(polymer = case_when(
     polymer == "BIO" ~ "Biopolymer",
     polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
@@ -70,7 +71,7 @@ polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
 
 sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
 sizef<-as.data.frame(sizedf)%>%
-  filter(effect %in% c("Y","N"))%>%
+  filter(effect %in% c("Yes","No"))%>%
   mutate(size.category = case_when(
     size.category == 1 ~ "1nm < 100nm",
     size.category == 2 ~ "100nm < 1µm",
@@ -91,7 +92,7 @@ sizefinal<- data.frame(cbind(sizef, study_s))%>%
       
 shapedf<-rowPerc(xtabs(~shape + effect, aoc))
 shapef<-as.data.frame(shapedf)%>%
-  filter(effect %in% c("Y","N"))%>%
+  filter(effect %in% c("Yes","No"))%>%
   rename(Type="shape")%>%
   mutate_if(is.numeric, round,0)%>%
   mutate(plot="Shape")%>%
[…]

### 2021-08-05 App_Nonessential/Database_CleanUp_AquaticOrg.R
Temporary files for staging new database clean up

@@ -633,6 +633,31 @@ aoc_setup <- aoc_v1 %>%
                                   environment == "Terrestrial" ~ "Terrestrial")) %>%
   mutate(weather.biofoul = case_when(weather.biofoul == "Y" ~ "Yes",
                                      weather.biofoul == "N" ~ "No")) %>% 
+  mutate(mix = case_when(mix == "Y" ~ "Yes",
+                         mix == "N" ~ "No")) %>% 
+  mutate(negative.control = case_when(negative.control == "Y" ~ "Yes",
+                                     negative.control == "N" ~ "No")) %>% 
+  mutate(reference.material = case_when(reference.material == "Y" ~ "Yes",
+                                     reference.material == "N" ~ "No")) %>% 
+  mutate(size.valid = case_when(size.valid == "Y" ~ "Yes",
+                                        size.valid == "N" ~ "No")) %>%
+  mutate(shape.valid = case_when(shape.valid == "Y" ~ "Yes",
+                                shape.valid == "N" ~ "No")) %>%
+  mutate(polymer.valid = case_when(polymer.valid == "Y" ~ "Yes",
+                                        polymer.valid == "N" ~ "No")) %>%
+  mutate(sodium.azide = case_when(sodium.azide == "Y" ~ "Yes",
+                                  sodium.azide == "N" ~ "No",
+                                  sodium.azide == "unknown" ~ "unknown")) %>%
+  mutate(contaminant.screen = case_when(contaminant.screen == "Y" ~ "Yes",
+                                        contaminant.screen == "N" ~ "No")) %>%
+  mutate(concentration.valid = case_when(concentration.valid == "Y" ~ "Yes",
+                                        concentration.valid == "N" ~ "No")) %>%
+  mutate(uptake.valid = case_when(uptake.valid == "Y" ~ "Yes",
+                                         uptake.valid == "N" ~ "No")) %>%
+  mutate(fed = case_when(fed == "Y" ~ "Yes",
+                                         fed == "N" ~ "No")) %>%
+  mutate(background.plastics = case_when(background.plastics == "Y" ~ "Yes",
+                                         background.plastics == "N" ~ "No")) %>%
   # mutate(species_f = as.factor(paste(genus,species))) %>% 
[…]

### 2021-08-05 app.R
Major changes across the app - more concise setup and general debugging

@@ -2,47 +2,49 @@
 #### File created: September 23, 2020
 #### Code contributors: Heili Lowman, Leah Thornton Hampton, Scott Coffin, Emily Darin
 
-
-
 #### Setup ####
 
 # Load packages
-library(tidyverse) #General everything
-library(data.table) #Faster functions for loading data set
-library(RColorBrewer) #color palette
-library(ggplot2) #plotting
-library(ggrepel) #For adding text labels that repel away from data points
-library(calecopal) #Color palette
-library(shiny) #Runs shiny
-library(shinythemes) #Shiny theme for the page
-library(shinyWidgets) #Widgets
-library(scales) #SSD - Use the percent format
-library(reshape2) #Overview tab - melts bars together
-library(ssdtools) #SSD package
-library(DT) #Build HTML data tables
-library(plotly) #Make plots interactive
-library(viridis) #Colors
-library(scales) #To use "percent" function
-library(shinyjs) #Exploration tab - reset button
-library(tigerstats) #turns things into percents
-library(ggbeeswarm) #plot all points
-library(fitdistrplus) #alt SSD 
-library(ggdark) #dark mode ggplot
-library(ggsci) #color palettes
-# library(bslib) #better themes. required for dark mode
-# library(thematic) #complete control over themes (including plots) required for dark mode
-library(collapsibleTree) #plot type for endpoint category tree
-library(hrbrthemes) #theme for screening plot
+
+#General Purpose
+library(tidyverse)
+library(shiny)
 library(shinydashboard)
+library(shinyWidgets)
+library(data.table) #Data import
+library(shinyjs) #Reset button
+
+#Plotting
+library(ggplot2) 
+library(reshape2) #Overview - melt bars together
+library(collapsibleTree) #Overview - endpoint categorization
+library(ggbeeswarm) #Exploration - Beeswarm plots
+library(plotly) #Screening - interactive plots
+library(hrbrthemes) #Screening - theme
+
+#Data Tables
+library(DT) #Build HTML data tables
+
+#Color Packages
+library(RColorBrewer)
+library(calecopal)
+library(viridis) 
+library(ggdark) 
+library(ggsci) 
+
+#Data Analysis
[…]

### 2021-08-05 app.R
fill function on collapsible tree just died on deployment (though it works locally) - removing it

@@ -6,7 +6,7 @@
 
 # Load packages
 
-#General Purpose
+# #General Purpose
 library(tidyverse)
 library(shiny)
 library(shinydashboard)
@@ -15,7 +15,7 @@ library(data.table) #Data import
 library(shinyjs) #Reset button
 
 #Plotting
-library(ggplot2) 
+library(ggplot2)
 library(reshape2) #Overview - melt bars together
 library(collapsibleTree) #Overview - endpoint categorization
 library(ggbeeswarm) #Exploration - Beeswarm plots
@@ -28,9 +28,9 @@ library(DT) #Build HTML data tables
 #Color Packages
 library(RColorBrewer)
 library(calecopal)
-library(viridis) 
-library(ggdark) 
-library(ggsci) 
+library(viridis)
+library(ggdark)
+library(ggsci)
 
 #Data Analysis
 library(ssdtools) #SSD package
@@ -406,7 +406,7 @@ aoc_ERM_default <- aoc_setup  %>%
 
 aoc_search <- aoc_setup %>%
          #general
-  select(doi, authors, year, species_f, org_f, env_f, life_f, vivo_f, sex, body.length.cm, max.size.ingest.mm,
+  dplyr::select(doi, authors, year, species_f, org_f, env_f, life_f, vivo_f, sex, body.length.cm, max.size.ingest.mm,
          #experimental parameters
          exp_type_f, exposure.route, mix, negative.control, reference.material, exposure.media, solvent, detergent,
          media.ph, media.sal.ppt, media.temp, media.temp.min, media.temp.max, exposure.duration.d, acute.chronic_f,
@@ -720,8 +720,8 @@ tabItem(tabName = "Overview",
         box(title = "Biological Endpoint Catgorization", status = "primary", width = 12, collapsible = TRUE,
         
         br(),
-        p("This plot displays the categorization of measured endpoints in the database. Nodes correspond to the Broad Endpoint Category (blue), 
-        the Specific Endpoint Category (green), Endpoints (pink) and the level of biological organization (purple). Alternatively, the widget 
+        p("This plot displays the categorization of measured endpoints in the database. Nodes correspond to the Broad Endpoint Category, 
+        the Specific Endpoint Category, Endpoints and the level of biological organization from left to right. The widget 
         below may be used to select endpoints at various Biological Levels of Organization. Click nodes to expand and collapse the plot."),
         br(),
             
[…]

### 2021-08-06 app.R
App is unbroken!

@@ -6,39 +6,32 @@
 
 # Load packages
 
-# #General Purpose
-library(tidyverse)
-library(shiny)
+# Load packages
+library(tidyverse) #General everything
 library(shinydashboard)
-library(shinyWidgets)
-library(data.table) #Data import
-library(shinyjs) #Reset button
-
-#Plotting
-library(ggplot2)
-library(reshape2) #Overview - melt bars together
-library(collapsibleTree) #Overview - endpoint categorization
-library(ggbeeswarm) #Exploration - Beeswarm plots
-library(plotly) #Screening - interactive plots
-library(hrbrthemes) #Screening - theme
-
-#Data Tables
-library(DT) #Build HTML data tables
-
-#Color Packages
-library(RColorBrewer)
-library(calecopal)
-library(viridis)
-library(ggdark)
-library(ggsci)
-
-#Data Analysis
-library(ssdtools) #SSD package
-library(tigerstats) #SSD - Use the percent format
+library(RColorBrewer) #color palette
+library(ggplot2) #plotting
+library(calecopal) #Color palette
+library(shiny) #Runs shiny
+library(shinythemes) #Shiny theme for the page
+library(shinyWidgets) #Widgets
 library(scales) #SSD - Use the percent format
+library(reshape2) #Overview tab - melts bars together
+library(ssdtools) #SSD package
+library(DT) #Build HTML data tables
+library(plotly) #Make plots interactive
+library(viridis) #Colors
+library(shinyjs) #Exploration tab - reset button
+library(tigerstats) #turns things into percents
+library(ggbeeswarm) #plot all points
+# library(fitdistrplus) #alt SSD 
+library(ggdark) #dark mode ggplot
+library(ggsci) #color palettes
+library(collapsibleTree) #plot type for endpoint category tree
+library(hrbrthemes) #theme for screening plot
 
 # Load finalized dataset.
-aoc <- fread("AquaticOrganisms_Clean_Final.csv") 
+aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) 
 
 #### Welcome Setup ####
 
@@ -47,9 +40,11 @@ aoc <- fread("AquaticOrganisms_Clean_Final.csv")
 #### Overview Setup ####
 
[…]

### 2021-09-29 App_Nonessential/Database_CleanUp_AquaticOrg.R
moved all nonessential files onto local machine

@@ -1,707 +1,0 @@
-#Script for one-time data clean up to eliminate unnecessary parts of setup. 
-
-
-library(tidyverse)
-
-aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
-
-aoc_v1 <- aoc %>% 
-  mutate(effect = case_when(effect == "Y" ~ "Yes",
-                                   effect == "N" ~ "No")) %>%  
-  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported", chem.exp.typ.nominal = "Particle Only"))
-
-aoc_setup <- aoc_v1 %>% 
-  mutate(size.category = case_when(
-    size.category == 1 ~ "1nm < 100nm",
-    size.category == 2 ~ "100nm < 1µm",
-    size.category == 3 ~ "1µm < 100µm",
-    size.category == 4 ~ "100µm < 1mm",
-    size.category == 5 ~ "1mm < 5mm",
-    size.category == 0 ~ "Not Reported")) %>%    
-  mutate(shape = case_when(
-    shape == "fiber" ~ "Fiber",
-    shape == "fragment" ~ "Fragment",
-    shape == "sphere" ~ "Sphere",
-    shape == "Not Reported" ~ "Not Reported")) %>% 
-  mutate(polymer = case_when(
-    polymer == "BIO" ~ "Biopolymer",
-    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
-    polymer == "LTX" ~ "Latex",
-    polymer == "PA" ~ "Polyamide",
-    polymer == "PE" ~ "Polyethylene",
-    polymer == "PC" ~ "Polycarbonate",
-    polymer == "PET" ~ "Polyethylene Terephthalate",
-    polymer == "PI" ~ "Polyisoprene",
-    polymer == "PMMA" ~ "Polymethylmethacrylate",
-    polymer == "PP" ~ "Polypropylene",
-    polymer == "PS" ~ "Polystyrene",
-    polymer == "PUR" ~ "Polyurethane",
-    polymer == "PVC" ~ "Polyvinylchloride",
-    polymer == "PLA" ~ "Polylactic Acid",
-    polymer == "Not Reported" ~ "Not Reported")) %>%
-  mutate(lvl1 = case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
-                                   lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
[…]

### 2021-09-29 App_Nonessential/Maybe_Delete/Endpoints .R

@@ -1,202 +1,0 @@
-
-
-
-
-
-
-
-
-
-
-
-polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
-polyf<-as.data.frame(polydf)%>% #Makes data frame 
-  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
-  mutate(polymer = case_when(
-    polymer == "BIO" ~ "Biopolymer",
-    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
-    polymer == "LTX" ~ "Latex",
-    polymer == "PA" ~ "Polyamide",
-    polymer == "PE" ~ "Polyethylene",
-    polymer == "PC" ~ "Polycarbonate",
-    polymer == "PET" ~ "Polyethylene Terephthalate",
-    polymer == "PI" ~ "Polyisoprene",
-    polymer == "PMMA" ~ "Polymethylmethacrylate",
-    polymer == "PP" ~ "Polypropylene",
-    polymer == "PS" ~ "Polystyrene",
-    polymer == "PUR" ~ "Polyurathane",
-    polymer == "PVC" ~ "Polyvinylchloride",
-    polymer == "PLA" ~ "Polylactic Acid"))%>%
-  mutate_if(is.numeric, round,0) #rounds percents 
-Endpoints<-xtabs(~polymer +effect ,aoc) #Pulls all study obs. for polymer from dataset
-polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
-  rename(Endpoints='Freq.1')%>% #renames column
-  mutate(logEndpoints = log(Endpoints))%>%
-  rename(Percent = Freq)#renames column
-
-sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
-sizef<-as.data.frame(sizedf)%>%
-  filter(effect %in% c("Y","N"))%>%
-  mutate(size.category = case_when(
-    size.category == 1 ~ "<1µm",
-    size.category == 2 ~ "1µm < 10µm",
-    size.category == 3 ~ "10µm < 100µm",
-    size.category == 4 ~ "100µm < 1mm",
-    size.category == 5 ~ "1mm < 5mm",
-    size.category == 0 ~ "unavailable"))%>%
-  rename(Type = "size.category")%>%
-  mutate_if(is.numeric, round,0)%>%
-  mutate(plot="Size")
[…]

### 2021-09-29 App_Nonessential/Maybe_Delete/Final_effect_dataset2.0.R

@@ -1,193 +1,0 @@
-
-
-aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
-polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
-polyf<-as.data.frame(polydf)%>% #Makes data frame 
-  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
-  rename(Type = "polymer")%>%#rename so future columns have same name 
-  mutate(Type = case_when(
-    Type == "BIO" ~ "Biopolymer",
-    Type == "EVA" ~ "Polyethylene Vinyl Acetate",
-    Type == "LTX" ~ "Latex",
-    Type == "PA" ~ "Polyamide",
-    Type == "PE" ~ "Polyethylene",
-    Type == "PC" ~ "Polycarbonate",
-    Type == "PET" ~ "Polyethylene Terephthalate",
-    Type == "PI" ~ "Polyisoprene",
-    Type == "PMMA" ~ "Polymethylmethacrylate",
-    Type == "PP" ~ "Polypropylene",
-    Type == "PS" ~ "Polystyrene",
-    Type == "PUR" ~ "Polyurathane",
-    Type == "PVC" ~ "Polyvinylchloride",
-    Type == "PLA" ~ "Polyactic Acid"))%>%
-  mutate_if(is.numeric, round,0)%>% #rounds percents 
-  mutate(plot="Polymer") # change column name for check list
-Endpoints<-xtabs(~polymer +effect ,aoc) #Pulls all study obs. for polymer from dataset
-polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
-  rename(Endpoints='Freq.1')%>% #renames column
-  rename(category='polymer')%>%#renames column
-  mutate(logEndpoints = log(Endpoints))%>%
-  rename(Percent = Freq)#renames column
-
-polyfinal
-
-count()
-
-sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
-sizef<-as.data.frame(sizedf)%>%
-  filter(effect %in% c("Y","N"))%>%
-  mutate(size.category = case_when(
-    size.category == 1 ~ "<1µm",
-    size.category == 2 ~ "1µm < 10µm",
-    size.category == 3 ~ "10µm < 100µm",
-    size.category == 4 ~ "100µm < 1mm",
[…]

### 2021-09-29 App_Nonessential/New_Data_Tidying.R

@@ -1,65 +1,0 @@
-
-aoc_explore <- aoc %>% 
-  
-  #Factors
-  mutate(effect_f = as.factor(effect)) %>% 
-  
-  mutate(size_f = as.factor(size.category)) %>% 
-  
-  mutate(shape_f = as.factor(shape)) %>% 
-  
-  mutate(poly_f = as.factor(polymer)) %>% 
-  
-  mutate(org_f = as.factor(organism.group)) %>% 
-  
-  mutate(lvl1_f = as.factor(lvl1)) %>% 
-  
-  mutate(lvl2_f = as.factor(lvl2)) %>% 
-  
-  mutate(lvl3_f = as.factor(lvl3)) %>% 
-  
-  mutate(bio_f = as.factor(bio.org)) %>% 
-  
-  mutate(vivo_f = as.factor(invitro.invivo)) %>% 
-           
-  mutate(life_f = as.factor(life.stage)) %>% 
-    
-  mutate(env_f = as.factor(environment)) %>% 
-    
-  mutate(weather.biofoul_f = as.factor(weather.biofoul)) %>% 
-    
-  mutate(species_f = as.factor(paste(genus,species))) %>% 
-    
-  mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
-  
-  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
-  
-  mutate(effect.metric = factor(effect.metric)) %>% 
-    
-  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>%
-    
-  mutate(acute.chronic_f = factor(acute.chronic)) %>%
-    
-  mutate(tier_zero_tech_f = factor(tech.tier.zero)) %>%
-
-  mutate(tier_zero_risk_f = factor(risk.tier.zero)) %>%
-  
-  mutate(exp_type_f = factor(experiment.type)) %>% 
-  
-  mutate(max.size.ingest.mm = ifelse(is.na(max.size.ingest.mm), 
-                                     10^(0.9341 * log10(body.length.cm) - 1.1200) * 10,  #(Jamm et al 2020 Nature paper)correction for cm to mm
-                                     max.size.ingest.mm)) %>%  # if already present, just use that
-  mutate(max.size.ingest.um = 1000 * max.size.ingest.mm) %>%  #makes it less confusing below
[…]

### 2021-09-29 App_Nonessential/Workshop_Analyses/Quality_Score_Plots.R

@@ -1,789 +1,0 @@
-#Quality Score Plots for meeting on 4/8/21
-#Created by: Leah Thornton Hampton
-#Date: 4/7/21
-
-#Load Packages
-library(tidyverse) #General everything
-library(calecopal) #Color palette
-
-#Import Data
-
-aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
-
-#Data Tidying
-
-aoc_v1 <- aoc %>% # start with original dataset
-  # full dataset filters.
-  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
-                                     effect == "N" ~ "No"),
-                           levels = c("No", "Yes"))) %>%
-  # removing NAs to make data set nicer
-  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported", chem.exp.typ.nominal = "Particle Only"))
-
-aoc_setup <- aoc_v1 %>% # start with original dataset
-  mutate(size_f = factor(case_when(
-    size.category == 1 ~ "1nm < 100nm",
-    size.category == 2 ~ "100nm < 1µm",
-    size.category == 3.1 ~ "1µm < 10µm",
-    size.category == 3.2 ~ "10µm < 100µm",
-    size.category == 4 ~ "100µm < 1mm",
-    size.category == 5 ~ "1mm < 5mm",
-    size.category == 0 ~ "Not Reported"),
-    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
-  # shape category data tidying.
-  mutate(shape_f = factor(case_when(
-    shape == "fiber" ~ "Fiber",
-    shape == "fragment" ~ "Fragment",
-    shape == "sphere" ~ "Sphere",
-    shape == "Not Reported" ~ "Not Reported"),
-    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% # order our different shapes.
-  # polymer category data tidying.
-  mutate(poly_f = factor(case_when(
-    polymer == "BIO" ~ "Biopolymer",
-    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
-    polymer == "LTX" ~ "Latex",
[…]

### 2021-09-29 App_Nonessential/custom plots/custom plots.R

@@ -1,2 +1,0 @@
-library(tidyverse)
-env <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)

### 2021-09-29 App_Nonessential/test/app.R

@@ -1,30 +1,0 @@
-library(shiny)
-library(shinyjs)
-library(bslib)
-library(thematic)
-#automatically theme plots
-thematic_shiny()
-
-light <- bs_theme()
-dark <- bs_theme(bg = "black", fg = "white", primary = "purple")
-ui <- fluidPage(
-  theme = light, 
-  div(
-    class = "custom-control custom-switch", 
-    tags$input(
-      id = "dark_mode", type = "checkbox", class = "custom-control-input",
-      onclick = HTML("Shiny.setInputValue('dark_mode', document.getElementById('dark_mode').value);")
-    ),
-    tags$label(
-      "Dark mode", `for` = "dark_mode", class = "custom-control-label"
-    )
-  )
-)
-server <- function(input, output, session) {
-  observe({
-    session$setCurrentTheme(
-      if (isTRUE(input$dark_mode)) dark else light
-    )
-  })
-}
-shinyApp(ui, server)

### 2021-09-29 App_Nonessential/testApp.R

@@ -1,25 +1,0 @@
-library(shiny)
-library(shinyjs)
-light <- bs_theme()
-dark <- bs_theme(bg = "black", fg = "white", primary = "purple")
-ui <- fluidPage(
-  theme = light, 
-  div(
-    class = "custom-control custom-switch", 
-    tags$input(
-      id = "dark_mode", type = "checkbox", class = "custom-control-input",
-      onclick = HTML("Shiny.setInputValue('dark_mode', document.getElementById('dark_mode').value);")
-    ),
-    tags$label(
-      "Dark mode", `for` = "dark_mode", class = "custom-control-label"
-    )
-  )
-)
-server <- function(input, output, session) {
-  observe({
-    session$setCurrentTheme(
-      if (isTRUE(input$dark_mode)) dark else light
-    )
-  })
-}
-shinyApp(ui, server)

### 2021-09-29 App_Nonessential/PCA.Rmd
readded files because error

--- 2021-02-23 SVM,kNN.Rmd
@@
 ---
-title: "SVM, K-NN"
+title: "PCA and UMAP"
 author: "Scott Coffin"
@@
 library(quantregForest)
-library(caret) #support vector machines
+library(caret)
 library(tidyverse)
@@
 library(randomForest)
-
-# Attach Packages
-library(kernlab)      # SVM methodology
-library(e1071)        # SVM methodology
 ```
@@
 
-# Support Vector Machine
-## Kitchen Sink
-### Visualization
+# PCA
 
-#### Support Vector Machine - Non-linear
-
-Support Vector Classifiers are a subset of the group of classification structures known as Support Vector Machines.
-Support Vector Machines can construct classification boundaries that are nonlinear in shape.
-The options for classification structures using the svm() command from the e1071 package are linear, polynomial, radial, and sigmoid.
-Toobserve relationships of classifiers, we plot the data in two dimensions.
+# UMAP
 
 ```{r}
-# Plot data
-multiVar %>% 
-ggplot(aes(x = log(size.length.um.used.for.conversions), y = log(dose.mg.L.master), color = effect_f, shape = effect_f)) + 
-  geom_point(size = 2) +
-  scale_color_manual(values=c("#000000", "#FF0000")) +
-  theme(legend.position = "none")
+require(uwot)
+require(Rtsne)
+require(vizier) #devtools::install_github("jlmelville/vizier")
+
+# For some functions we need to strip out non-numeric columns and convert data to matrix
+x2m <- function(X) {
+  if (!methods::is(X, "matrix")) {
+    m <- as.matrix(X[, which(vapply(X, is.numeric, logical(1)))])
+  }
+  else {m <- X} 
+  m}
+
+#choose values with most completeness
+multiVar2 <- multiVar %>% 
+  filter(!environment == "Terrestrial") %>% 
+  dplyr::select(size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
+  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
+  drop_na() #drop missing
+
+#convert discrete variables to numeric
+multiVar2[] <- data.matrix(multiVar2)
+
+# build umap for small dataset (<10,000 points)
+multiVar_map  <- umap(multiVar2, pca = 10)
+
+# Remove duplicates for t-SNE
+#multiVar2_noNa_dup <- multiVar2_noNa[-which(duplicated(x2m(multiVar2_noNa))), ]
+
+#build t-SNE
+#multiVar_tsne <- Rtsne::Rtsne(multiVar2_noNa_dup, perplexity = 15, initial_dims = 100, partical_pca = TRUE, exaggeration_factor = 4)
+
+# Non-numeric columns are ignored, so in a lot of cases you can pass a data
+# frame directly to umap
+#iris_umap <- umap(iris, n_neighbors = 50, learning_rate = 0.5, init = "random")
+
+#visualize umap
+embed_img <- function(X, Y, k = 15, ...) {
+  args <- list(...)
+  args$coords <- Y
+  args$x <- X
+
+  do.call(vizier::embed_plot, args)
+}
+
+#plot
+embed_img(multiVar2, multiVar_map, pc_axes = TRUE, equal_axes = TRUE, alpha_scale = 0.5, title = "Tox UMAP", cex = 1)
 ```
-Notice that the data is not linearly separable, and furthermore, isn’t all clustered together in a single group.
-There are two sections of class 1 observations with a cluster of class 2 observations in between.
-To demonstrate the power of SVMs, we’ll take 100 random observations from the set and use them to construct our boundary.
-We set kernel = "radial" based on the shape of our data and plot the results.
 
 ```{r}
-# set pseudorandom number generator
-set.seed(123)
-# sample training data
-set.seed(4)
-kitchenSink_split <- kitchenSink %>%
-  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
-kitchenSink_train <- training(kitchenSink_split)
-kitchenSink_test <- testing(kitchenSink_split)
-# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
-count_kitchenSink <- paste0('n = ',nrow(kitchenSink))
-skim(kitchenSink)
+#PCA
+pca <- stats::prcomp(multiVar2[,-5], retx = TRUE, rank. = 2)
+#build color pallete
+my_colors = colorRampPalette(c("red", "yellow", "green"))(nrow(multiVar2))
+#plot
+embed_plot(pca$x, multiVar2$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), #turbo, #rainbow, #my_colors, 
+           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE)
+```
 
-subTrain <- kitchenSink_test %>% 
-  dplyr::select(c(effect_f, dose.mg.L.master, size.length.um.used.for.conversions)) %>% 
+```{r}
+require(plotly)
+#PCA for discrete variables
+multiVar_discrete <- multiVar %>% 
+  select(size.length.um.used.for.conversions, polymer, dose.mg.L.master, exposure.duration.d) %>% 
+  drop_na %>% 
   mutate_if(~is.numeric(.) && (.) > 0, log10)
 
-#fit model
-svmfit_kitchenSink <- svm(effect_f ~ ., 
-                          data = subTrain, kernel = "radial", gamma = 1, cost = 1)
-summary(svmfit_kitchenSink)
-# plot classifier
-plot(svmfit_kitchenSink, data = subTrain)
+#buildPCA
+pca_discrete <- stats::prcomp(multiVar_discrete[,-2], retx = TRUE, rank. = 2)
+
+embed_plotly(pca_discrete$x, multiVar_discrete$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), 
+           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE,
+           tooltip = paste("Polymer:", multiVar_discrete$polymer))
 ```
-The same procedure can be run using the kernlab package, which has far more kernel options than the corresponding function in e1071.
-In addition to the four choices in e1071, this package allows use of a hyperbolic tangent, Laplacian, Bessel, Spline, String, or ANOVA RBF kernel.
-To fit this data, we set the cost to be the same as it was before, 1.
-```{r}
-# Fit radial-based SVM in kernlab
-kernfit <- ksvm(effect_f ~ ., data = subTrain, type = "C-svc", kernel = 'rbfdot', C = 1, scaled = c())
-# Plot training data
-plot(kernfit, data = subTrain)
-```
-
-### Data Split
-```{r}
-#choose relevant predictors and log-transform
-kitchenSink <- aoc_z %>% 
-  mutate_if(is.character, as.factor) %>% 
-  mutate(doi = as.character(doi)) %>% 
-  dplyr::select(size_f,
-                #doi, #need to split studies
-                size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
-  # mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
- # mutate(effect_10 = case_when( #convert ordinal to numeric
- #     effect_f == "Y" ~ 1,
- #     effect_f == "N" ~ 0
- #   ))
-drop_na() #drop missing
-
-skim(kitchenSink)
-```
-#### Drop NA
-```{r}
-# Create calibration and validation splits with tidymodels initial_split() function.
-set.seed(4)
-kitchenSink_split <- kitchenSink %>%
-  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
-# default is 3/4ths split (but 75% training, 25% testing).
-# Stratification (strata) = grouping training/testing sets by region, state, etc.
-# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
-
-# Create a training data set with the training() function
-# Pulls from training and testing sets created by initial_split()
-kitchenSink_train <- training(kitchenSink_split)
-kitchenSink_test <- testing(kitchenSink_split)
-# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
-
-count_kitchenSink <- paste0('n = ',nrow(kitchenSink))
-
-skim(kitchenSink)
-```
-
-#### Rough Fix NA
-```{r}
-#rough fix
-kitchenSink_roughfix <- na.roughfix(kitchenSink)
-
-# Create calibration and validation splits with tidymodels initial_split() function.
-set.seed(4)
-kitchenSink_split_roughfix <- kitchenSink_roughfix %>%
-  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
-# default is 3/4ths split (but 75% training, 25% testing).
-# Stratification (strata) = grouping training/testing sets by region, state, etc.
-# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
-
-# Create a training data set with the training() function
-# Pulls from training and testing sets created by initial_split()
-kitchenSink_train_roughfix <- training(kitchenSink_split_roughfix)
-kitchenSink_test_roughfix <- testing(kitchenSink_split_roughfix)
-# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
-
-count_roughfix <- paste0('n = ',nrow(kitchenSink_roughfix))
-
-skim(kitchenSink_roughfix)
-```
-
-#### Multiple Imputation NA
-
-The algorithm starts by imputing NAs using na.roughfix.
-Then randomForest is called with the completed data.
-The proximity matrix from the randomForest is used to update the imputation of the NAs.
-For continuous predictors, the imputed value is the weighted average of the non-missing obervations, where the weights are the proximities.
-For categorical predictors, the imputed value is the category with the largest average proximity.
-This process is iterated iter times.
-
-Note: Imputation has not (yet) been implemented for the unsupervised case.
-Also, Breiman (2003) notes that the OOB estimate of error from randomForest tend to be optimistic when run on the data matrix with imputed values.
-```{r}
-# impute values
-#drop NA's in response
-kitchenSink_noNa <- kitchenSink %>% drop_na(effect_f)
-
-#impute
-set.seed(111)
-kitchenSink_rfImpute <- rfImpute(data = kitchenSink_noNa, 
-                                    effect_f ~., #response value, cannot contain NA's
-                                    iter = 4,
-                                    ntree =75)
-
-# Create calibration and validation splits with tidymodels initial_split() function.
-set.seed(4)
-kitchenSink_split_imputed <- kitchenSink_rfImpute %>%
-  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
-# default is 3/4ths split (but 75% training, 25% testing).
-# Stratification (strata) = grouping training/testing sets by region, state, etc.
-# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)
-
-# Create a training data set with the training() function
-# Pulls from training and testing sets created by initial_split()
-kitchenSink_train_imputed <- training(kitchenSink_split_imputed)
-kitchenSink_test_imputed <- testing(kitchenSink_split_imputed)
-# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
-
-count_imputed <- paste0('n = ',nrow(kitchenSink_rfImpute))
-
-skim(kitchenSink_rfImpute)
-```
-
-### Train
-#### Drop Na
-```{r}
-#examine train and test sets
-dim(kitchenSink_train); dim(kitchenSink_test)
-```
-
-Before we train our model, we’ll first implement the trainControl() method.
-This will control all the computational overheads so that we can use the train() function provided by the caret package.
-The training method will train our data on different algorithms.
-
-First, let’s focus on the traincontrol() method:
-
-```{r}
-trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
-# The trainControl() method here, is taking 3 parameters.
-# The “method” parameter defines the resampling method, in this demo we’ll be using the repeatedcv or the repeated cross-validation method.
-# The next parameter is the “number”, this basically holds the number of resampling iterations.
-# The “repeats ” parameter contains the sets to compute for our repeated cross-validation. We are using setting number =10 and repeats =3
-# This trainControl() method returns a list. We are going to pass this on our train() method.
-
-svm_Linear <- train(effect_f ~., 
-                    data = kitchenSink_train, 
-                    method = "bayesglm", #Bayesian Model
-                    trControl = trctrl, 
-                    preProcess = c("center", "scale"), 
-                    tuneLength = 10)
-#The “preProcess” parameter is for preprocessing our training data.
-# We are passing 2 values in our “pre-process” parameter “center” & “scale”. These two help for centering and scaling the data.
-# After pre-processing, these convert our training data with mean value as approximately “0” and standard deviation as “1”.
-The “tuneLength” parameter holds an integer value.
-This is for tuning our algorithm.
-
-svm_Linear
-```
-It’s a linear model therefore, it just tested at value “C” = 1.
-
-Now, our model is trained with C value as 1. We are ready to predict classes for our test set. We can use predict() method.
-
-The caret package provides predict() method for predicting results.
-We are passing 2 arguments.
-Its first parameter is our trained model and second parameter “newdata” holds our testing data frame.
-The predict() method returns a list, we are saving it in a test_pred variable.
-
-```{r}
-test_pred <- predict(svm_Linear, newdata = kitchenSink_test)
-confusionMatrix(table(test_pred, kitchenSink_test$effect_f))
-```
-
-The output shows that our model accuracy for test set is %
-
-By following the above procedure, we can build our svmLinear classifier.
-
-We can also do some customization for selecting C value(Cost) in Linear classifier. This can be done by inputting values in grid search.
-
-The next code snippet will show you, building & tuning of an SVM classifier with different values of C.
-
-We are going to put some values of C using expand.grid() into “grid” dataframe.
-Next step is to use this dataframe for testing our classifier at specific C values.
-It needs to be put in train() method with tuneGrid parameter.
-
-```{r}
-grid <- expand.grid(C = c(0,0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2,5))
-svm_Linear_Grid <- train(effect_f ~., data = kitchenSink_train, method = "svmLinear",
-                         trControl = trctrl,
-                         preProcess = c("center", "scale"),
-                         tuneGrid = grid,
-                         tuneLength = 10)
-svm_Linear_Grid
-plot(svm_Linear_Grid)
-```
-The above plot is showing that our classifier is giving best accuracy on C = 1.75. Let’s try to make predictions using this model for our test set.
-
-```{r}
-test_pred_grid <- predict(svm_Linear_Grid, newdata = kitchenSink_train)
-confusionMatrix(table(test_pred_grid, kitchenSink_train$effect_f))
-```
-
-# ROC Curves
-
-Prep for plotting.
-```{r}
-
-#### kitchenSink
-kitchenSinkpredictions <- as.data.frame(predict(kitchenSinkrf, kitchenSink_test%>% dplyr::select(-effect_f), type = "prob"))
-# predict class and then attach test class
-kitchenSinkpredictions$predict <- names(kitchenSinkpredictions)[1:2][apply(kitchenSinkpredictions[,1:2], 1, which.max)]
-kitchenSinkpredictions$observed <- kitchenSink_test$effect_f
-head(kitchenSinkpredictions)
-
-#### kitchenSink_roughfix_roughfix
-kitchenSink_roughfixpredictions <- as.data.frame(predict(kitchenSink_roughfixrf, kitchenSink_roughfix_test%>% dplyr::select(-effect_f), type = "prob"))
-# predict class and then attach test class
-kitchenSink_roughfixpredictions$predict <- names(kitchenSink_roughfixpredictions)[1:2][apply(kitchenSink_roughfixpredictions[,1:2], 1, which.max)]
-kitchenSinkpredictions$observed <- kitchenSink_test$effect_f
-head(kitchenSinkpredictions)
-
-###CHRONIC
-chronicpredictions <- as.data.frame(predict(chronicrf, chronic_test %>% dplyr::select(-effect_f), type = "prob"))
-# predict class and then attach test class
-chronicpredictions$predict <- names(chronicpredictions)[1:2][apply(chronicpredictions[,1:2], 1, which.max)]
-chronicpredictions$observed <- chronic_test$effect_f
-
-####ACUTE
-predictions <- as.data.frame(predict(acuterf, acute_test%>% dplyr::select(-effect_f), type = "prob"))
-# predict class and then attach test class
-predictions$predict <- names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
-predictions$observed <- acute_test$effect_f
-head(predictions)
-
-###nofilter
-nofilterpredictions <- as.data.frame(predict(nofilterrf, nofilter_test%>% dplyr::select(-effect_f), type = "prob"))
-# predict class and then attach test class
-nofilterpredictions$predict <- names(nofilterpredictions)[1:2][apply(nofilterpredictions[,1:2], 1, which.max)]
-nofilterpredictions$observed <- nofilter_test$effect_f
-
-###nofilterOptimized
-nofilter.optimizedpredictions <- as.data.frame(predict(myrf_optimized, multiVar_smkitchenSink_test %>%  dplyr::select(-effect_f), type = "prob"))
-# predict class and then attach test class
-nofilter.optimizedpredictions$predict <- names(nofilter.optimizedpredictions)[1:2][apply(nofilter.optimizedpredictions[,1:2], 1, which.max)]
-nofilter.optimizedpredictions$observed <- multiVar_smkitchenSink_test$effect_f
-
-###nofilterOptimizedImputed
-nofilter.optimized.imputedpredictions <- as.data.frame(predict(myrf_optimized_imputed, multiVar_smkitchenSink_test_imputed %>%  dplyr::select(-effect_f), type = "prob"))
-# predict class and then attach test class
-nofilter.optimized.imputedpredictions$predict <- names(nofilter.optimized.imputedpredictions)[1:2][apply(nofilter.optimized.imputedpredictions[,1:2], 1, which.max)]
-nofilter.optimized.imputedpredictions$observed <- multiVar_smkitchenSink_test_imputed$effect_f
-```
-Plot.
-```{r}
-require(ggdark)
-# 1 ROC curve, yes vs no for acute
-roc.acute <- roc(ifelse(predictions$observed=="Y", "Y", "N"), as.numeric(predictions$Y))
-
-#chronic
-roc.chronic <- roc(ifelse(chronicpredictions$observed=="Y", "Y", "N"), as.numeric(chronicpredictions$Y))
-#kitchenSink
-roc.kitchenSink <- roc(ifelse(kitchenSinkpredictions$observed=="Y", "Y", "N"), as.numeric(kitchenSinkpredictions$Y))
-
-#nofilter
-roc.nofilter <- roc(ifelse(nofilterpredictions$observed=="Y", "Y", "N"), as.numeric(nofilterpredictions$Y))
-
-#no filter (optimized)
-roc.nofilter.optimized <- roc(ifelse(nofilter.optimizedpredictions$observed=="Y", "Y", "N"), as.numeric(nofilter.optimizedpredictions$Y))
-
-#no filter (optimized; imputed)
-roc.nofilter.optimized.imputed <- roc(ifelse(nofilter.optimized.imputedpredictions$observed=="Y", "Y", "N"), as.numeric(nofilter.optimized.imputedpredictions$Y))
-
-##make ROC curves
-
-#kitchenSink
-kitchenSinkROC <- ggroc(roc.kitchenSink, col = "yellow") + 
-  labs(title = "Chronic and Acute",
-       subtitle = paste0(accuracy_kitchenSink,', ',count_kitchenSink)) + 
-  dark_theme_bw()
-
-#acute
-acuteROC <- ggroc(roc.acute, col = "green") + 
-  labs(title = "Acute",
-       subtitle = paste0(accuracy_acute,', ',count_acute)) + 
-  dark_theme_bw()
-
-#chronic
-chronicROC <- ggroc(roc.chronic, col = "blue") + 
-  labs(title = "Chronic",
-       subtitle = paste0(accuracy_chronic,', ',count_chronic)) + #auto label
-  dark_theme_bw()
-
-#no filter
-nofilterROC <- ggroc(roc.nofilter, col = "red3") + 
-  labs(title = "Entire Dataset",
-       subtitle = paste0(accuracy_nofilter,', ',count_nofilter)) + 
-  dark_theme_bw()
-
-#optimized (rough fix)
-nofilteroptimizedROC <- ggroc(roc.nofilter.optimized, col = "orangered") + 
-  labs(title = "Entire Dataset (optimized)",
-       subtitle = paste0(accuracy_optimized,', ',count_optimized)) + 
-  dark_theme_bw()
-
-#optimized (multiple imputation)
-nofilteroptimizedimputedROC <- ggroc(roc.nofilter.optimized.imputed, col = "orange") + 
-  labs(title = "Entire Dataset (optimized; imputed)",
-       subtitle = paste0(accuracy_optimized_imputed,', ',count_optimized_imputed)) + 
-  dark_theme_bw()
-
-#arrange together and print
-require(gridExtra)
-grid.arrange(nofilterROC, nofilteroptimizedROC, nofilteroptimizedimputedROC,kitchenSinkROC, acuteROC, chronicROC,
-             ncol = 3)
-```
-ROC curves may also be visualized together
-```{r}
-require(pROC)
-require(tidyverse)
-require(ggdark)
-require(ggsci)
-ggroc(list(kitchenSink = roc.nofilter, optimized = roc.nofilter.optimized, optimized.imputed = roc.nofilter.optimized.imputed, organisms = roc.kitchenSink, acute = roc.acute, chronic = roc.chronic), aes = c("linetype", "color")) +
-  labs(title = "ROC Curves for Aquatic Toxicity Random Forest",
-       subtitle = "n = 4615",
-       color = "Dataset",
-       linetype = "Dataset") +
-   scale_color_tron() +
-  # theme_bw(base_size = 20)
- dark_theme_bw(base_size = 20)# +
- theme(plot.title.position = element_text(hjust = 0.5),
-     plot.subtitle.position = element_text(hjust = 0.5))
-```

### 2021-09-29 App_Nonessential/Maybe_Delete/Endpoints .R

@@ -1,0 +1,202 @@
+
+
+
+
+
+
+
+
+
+
+
+polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
+polyf<-as.data.frame(polydf)%>% #Makes data frame 
+  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
+  mutate(polymer = case_when(
+    polymer == "BIO" ~ "Biopolymer",
+    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
+    polymer == "LTX" ~ "Latex",
+    polymer == "PA" ~ "Polyamide",
+    polymer == "PE" ~ "Polyethylene",
+    polymer == "PC" ~ "Polycarbonate",
+    polymer == "PET" ~ "Polyethylene Terephthalate",
+    polymer == "PI" ~ "Polyisoprene",
+    polymer == "PMMA" ~ "Polymethylmethacrylate",
+    polymer == "PP" ~ "Polypropylene",
+    polymer == "PS" ~ "Polystyrene",
+    polymer == "PUR" ~ "Polyurathane",
+    polymer == "PVC" ~ "Polyvinylchloride",
+    polymer == "PLA" ~ "Polylactic Acid"))%>%
+  mutate_if(is.numeric, round,0) #rounds percents 
+Endpoints<-xtabs(~polymer +effect ,aoc) #Pulls all study obs. for polymer from dataset
+polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
+  rename(Endpoints='Freq.1')%>% #renames column
+  mutate(logEndpoints = log(Endpoints))%>%
+  rename(Percent = Freq)#renames column
+
+sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
+sizef<-as.data.frame(sizedf)%>%
+  filter(effect %in% c("Y","N"))%>%
+  mutate(size.category = case_when(
+    size.category == 1 ~ "<1µm",
+    size.category == 2 ~ "1µm < 10µm",
+    size.category == 3 ~ "10µm < 100µm",
+    size.category == 4 ~ "100µm < 1mm",
+    size.category == 5 ~ "1mm < 5mm",
+    size.category == 0 ~ "unavailable"))%>%
+  rename(Type = "size.category")%>%
+  mutate_if(is.numeric, round,0)%>%
+  mutate(plot="Size")
[…]

### 2021-09-29 App_Nonessential/Maybe_Delete/Final_effect_dataset2.0.R

@@ -1,0 +1,193 @@
+
+
+aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
+polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
+polyf<-as.data.frame(polydf)%>% #Makes data frame 
+  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
+  rename(Type = "polymer")%>%#rename so future columns have same name 
+  mutate(Type = case_when(
+    Type == "BIO" ~ "Biopolymer",
+    Type == "EVA" ~ "Polyethylene Vinyl Acetate",
+    Type == "LTX" ~ "Latex",
+    Type == "PA" ~ "Polyamide",
+    Type == "PE" ~ "Polyethylene",
+    Type == "PC" ~ "Polycarbonate",
+    Type == "PET" ~ "Polyethylene Terephthalate",
+    Type == "PI" ~ "Polyisoprene",
+    Type == "PMMA" ~ "Polymethylmethacrylate",
+    Type == "PP" ~ "Polypropylene",
+    Type == "PS" ~ "Polystyrene",
+    Type == "PUR" ~ "Polyurathane",
+    Type == "PVC" ~ "Polyvinylchloride",
+    Type == "PLA" ~ "Polyactic Acid"))%>%
+  mutate_if(is.numeric, round,0)%>% #rounds percents 
+  mutate(plot="Polymer") # change column name for check list
+Endpoints<-xtabs(~polymer +effect ,aoc) #Pulls all study obs. for polymer from dataset
+polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
+  rename(Endpoints='Freq.1')%>% #renames column
+  rename(category='polymer')%>%#renames column
+  mutate(logEndpoints = log(Endpoints))%>%
+  rename(Percent = Freq)#renames column
+
+polyfinal
+
+count()
+
+sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
+sizef<-as.data.frame(sizedf)%>%
+  filter(effect %in% c("Y","N"))%>%
+  mutate(size.category = case_when(
+    size.category == 1 ~ "<1µm",
+    size.category == 2 ~ "1µm < 10µm",
+    size.category == 3 ~ "10µm < 100µm",
+    size.category == 4 ~ "100µm < 1mm",
[…]

### 2021-09-29 App_Nonessential/Maybe_Delete/Risk Characterization/Concentration data.R

@@ -1,0 +1,970 @@
+## Concentration Data
+library(tidyverse)
+library(calecopal)
+library(ssdtools)
+library(DT)
+library(plotly)
+library(gridExtra)
+library(grid)
+library(wesanderson)
+library(ggdark)
+library(broom)
+
+##### Concentration data prep ####
+#SFEI data is already prepped
+SFEI <- read.csv("Concentration data/SFEI.csv", stringsAsFactors = TRUE) %>% 
+  mutate(Location = case_when(
+    Location == "Bay\n" ~ "Bay",
+    Location == "Sanctuary" ~ "Sanctuary"
+  ))
+
+# Adam et al (2019) data needs correcting
+adam <- read.csv("Concentration data/adam2019.csv", na.strings = "N.A.") %>% 
+  mutate(x1M = Min.particle.size.um,
+         x2M = Max.particle.size.um) %>% 
+  mutate(Sample.Type = "sample")
+
+#function to derive correction factor (CF) from Koelmans et al (equation 2)
+CFfnx = function(a = 2.5, #default alpha from Koelmans et al (2020)
+                 x2D = 5000, #set detault values to convert ranges to (1-5,000 um) #5mm is upper defuault 
+                 x1D = 1, #1 um is lower default size
+                 x2M, x1M){
+  
+  CF = (x2D^(1-a)-x1D^(1-a))/(x2M^(1-a)-x1M^(1-a))
+  
+  return(CF)
+}
+#verify it works (expected answer is 40.37)
+CFfnx(x1M = 333, x2M = 5000)
+
+adam <- adam %>% 
+  mutate(CF = CFfnx(x1M = x1M, x2M = x2M)) %>%  #create new column with correction factor 
+  mutate(particles.m3.corrected = CF * Single.Measurement.conc....m3.) %>% #convert single concenetrations
+  mutate(particles.m3.corrected_mean = CF * Mean.conc....m3.) %>%  #convert mean concentrations from distributions
+  mutate(particles.m3.corrected_median = CF * Median.conc....m3.) %>%   #convert mean concentrations from distributions
+  mutate(particles.single.median.m3 = ifelse(is.na(particles.m3.corrected), particles.m3.corrected_median, particles.m3.corrected)) %>% 
+  mutate(particles.m3.master = ifelse(is.na(particles.single.median.m3), particles.m3.corrected_mean, particles.single.median.m3)) %>% 
+  mutate(particle.L.master = particles.m3.master/1000) %>% 
[…]

### 2021-09-29 App_Nonessential/Maybe_Delete/Simplifying via continuous probability distributions.R

@@ -1,0 +1,103 @@
+## Simplifying Microplastics Via Continuous Probability Distributions ##
+# Scott Coffin
+# 12-1-2020
+# Resources: https://pubs.acs.org/doi/suppl/10.1021/acs.est.0c02982/suppl_file/es0c02982_si_001.pdf
+# https://pubs.acs.org/doi/suppl/10.1021/acs.estlett.9b00379/suppl_file/ez9b00379_si_001.pdf
+# https://pubs.acs.org/doi/10.1021/acs.estlett.9b00379
+
+# Bimodal universal shape distribution for environmental microplastic from Kooi and Koelmans (2019)
+library(truncnorm)
+library(tidyverse)
+
+#average parameters from distribution
+f1 = 0.06
+f2 = 0.94
+delta1 = 0.03
+delta2 = 0.19
+mu1 = 0.08
+mu2 = 0.44
+x = runif(n=100)
+
+
+#define list from above
+param_list = list("f1" = f1, "f2" = f2, "delta1" = delta1, "delta2" = delta2, "mu1" = mu1, "mu2" = mu2, "x" = x)
+
+#permanently define euler's number
+e <- exp(1)
+lockBinding("e", globalenv())
+
+#starting conditions for x
+x = runif(n = 100) #particle size in micrometers
+
+#equation 4
+y = f1 * (1/sqrt(2 * pi * delta1 ^ 2)) * (e^-(x - mu1)^2 /2 *delta1 ^ 2)+ f2 * (1/sqrt(2 * pi * delta2 ^ 2)) * (e^-(x - mu2)^2 /2 *delta2 ^ 2)
+
+
+# Monte Carlo Simulation 
+library(MonteCarlo)
+#define equation 4
+binom <- function(x, f1, f2, delta1, delta2, mu1, mu2){
+  #generate sample
+  #x = runif(n=1000)
+  #calculate CSF
+  y <- f1 * (1/sqrt(2 * pi * delta1 ^ 2)) * (e^-(x - mu1)^2 /2 *delta1 ^ 2)+ f2 * (1/sqrt(2 * pi * delta2 ^ 2)) * (e^-(x - mu2)^2 /2 *delta2 ^ 2)
+  #return result
+  return(list("y" = y))
+}
+
+MC_result <- MonteCarlo(func=binom, nrep=1, param_list=param_list)
+
+df<-MakeFrame(MC_result)
+head(df)
+library(dplyr)
[…]

### 2021-09-29 App_Nonessential/Workshop_Analyses/Team_X.R

@@ -1,0 +1,378 @@
+# Microplastics Toxicity Literature Review
+# Southern California Coastal Water Research Project
+# Team X Data Analysis
+# Created on: January 17, 2021 by Leah Thornton Hampton
+
+#### Packages ####
+
+# Attach packages necessary for the actions below.
+library(tidyverse) # For data tidying/viz purposes.
+library(patchwork) # For knitting separate figures together.
+library(devtools) # For installing packages from github.
+library(calecopal) # For palettes used in figures below.
+# Install by running devtools::install_github("an-bui/calecopal") in console.
+# See available palettes by running names(cal_palettes) in console.
+# Display a palette by running cal_palette("sbchannel") in console.
+
+#### Data Import ####
+
+aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
+
+#### Set Up #### 
+
+#This is the exact same setup used in the app as of 3/2/21
+
+aoc_v1 <- aoc %>% # start with original dataset
+  # full dataset filters.
+  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
+                                     effect == "N" ~ "No"),
+                           levels = c("No", "Yes"))) %>%
+  # removing NAs to make data set nicer
+  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported"))
+
+aoc_setup <- aoc_v1 %>% # start with original dataset
+  mutate(size_f = factor(case_when(
+    size.category == 1 ~ "1nm < 100nm",
+    size.category == 2 ~ "100nm < 1µm",
+    size.category == 3 ~ "1µm < 100µm",
+    size.category == 4 ~ "100µm < 1mm",
+    size.category == 5 ~ "1mm < 5mm",
+    size.category == 0 ~ "Not Reported"),
+    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
+  # shape category data tidying.
+  mutate(shape_f = factor(case_when(
+    shape == "fiber" ~ "Fiber",
+    shape == "fragment" ~ "Fragment",
+    shape == "sphere" ~ "Sphere",
+    shape == "Not Reported" ~ "Not Reported"),
[…]

### 2021-09-29 App_Nonessential/custom plots/custom plots.R

@@ -1,0 +1,2 @@
+library(tidyverse)
+env <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)

### 2021-09-30 app.R
Updated contributors list

@@ -1313,7 +1313,7 @@ ui <- dashboardPage(
                         p(align = "center", a(href = "https://agency.calepa.ca.gov/staffdirectory/detail.asp?UID=69294&BDO=7&VW=DET&SL=S", 'Dr. Scott Coffin'),", California State Water Resources Control Board", 
                           tags$a(href="https://twitter.com/DrSCoffin", icon("twitter")), tags$a(href="https://github.com/ScottCoffin", icon("github"))),
                         
-                        p(align = "center", "Emily Darin, Southern California Coastal Water Research Project",
+                        p(align = "center", a(href = "https://www.sccwrp.org/about/staff/emily-darin/", 'Emily Darin'),", Southern California Coastal Water Research Project",
                           tags$a(href="https://github.com/EmilyDarin", icon("github"))),
                         
                         p(align = "center", a(href = "https://www.sfei.org/users/liz-miller", 'Dr. Ezra Miller'),", Aquatic Science Center"),
@@ -1335,6 +1335,9 @@ ui <- dashboardPage(
                         p(align = "center", "Laura Monclús Anglada, Norwegian University of Science and Technology"),
                         
                         p(align = "center", a(href = "https://www.sccwrp.org/about/staff/syd-kotar/", "Sydney Kotar"),", Southern California Coastal Water Research Project"),
+                        
+                        p(align = "center", a(href = "https://branderlab.net/", 'Dr. Susanne Brander'),", Oregon State University",
+                          tags$a(href="https://twitter.com/smbrander", icon("twitter"))),
                         
                         p(align = "center", a(href = "https://www.ntnu.edu/employees/martin.wagner", 'Dr. Martin Wagner'),", Norwegian University of Science and Technology",
                           tags$a(href="https://twitter.com/martiwag", icon("twitter"))),

### 2021-10-02 app.R
Updated Heili's affiliation

@@ -1307,8 +1307,8 @@ ui <- dashboardPage(
                         p(align = "center", a(href = "https://www.sccwrp.org/about/staff/leah-thornton-hampton/", 'Dr. Leah Thornton Hampton'),", Southern California Coastal Water Research Project ", 
                           tags$a(href="https://twitter.com/DrLeahTH", icon("twitter")), tags$a(href="https://github.com/leahth", icon("github"))),
                         
-                        p(align = "center", "Heili Lowman, Southern California Coastal Water Research Project ",
-                          tags$a(href="https://twitter.com/heili_lowman", icon("twitter")), tags$a(href="https://github.com/hlowman", icon("github"))), 
+                        p(align = "center", a(href = "https://www.heililowman.com/", 'Dr. Heili Lowman'),", University of Nevada Reno ", 
+                          tags$a(href="https://twitter.com/heili_lowman", icon("twitter")), tags$a(href="https://github.com/hlowman", icon("github"))),
                         
                         p(align = "center", a(href = "https://agency.calepa.ca.gov/staffdirectory/detail.asp?UID=69294&BDO=7&VW=DET&SL=S", 'Dr. Scott Coffin'),", California State Water Resources Control Board", 
                           tags$a(href="https://twitter.com/DrSCoffin", icon("twitter")), tags$a(href="https://github.com/ScottCoffin", icon("github"))),
@@ -1326,7 +1326,7 @@ ui <- dashboardPage(
                         
                         p(align = "center", "Vera de Ruitjer, Wageningen University"),
                         
-                        p(align = "center", "Samreen Siddiqui, Oregon State University"),
+                        p(align = "center", "Dr. Samreen Siddiqui, Oregon State University"),
                         
                         p(align = "center", "Andrea Faltynkova, Norwegian University of Science and Technology"),
                         

### 2021-10-02 app.R
updated affiliation

@@ -1316,7 +1316,7 @@ ui <- dashboardPage(
                         p(align = "center", a(href = "https://www.sccwrp.org/about/staff/emily-darin/", 'Emily Darin'),", Southern California Coastal Water Research Project",
                           tags$a(href="https://github.com/EmilyDarin", icon("github"))),
                         
-                        p(align = "center", a(href = "https://www.sfei.org/users/liz-miller", 'Dr. Ezra Miller'),", Aquatic Science Center"),
+                        p(align = "center", a(href = "https://www.sfei.org/users/liz-miller", 'Dr. Ezra Miller'),", San Francisco Estuary Institute"),
                         
                         p(align = "center", a(href = "https://rochmanlab.com/people/", 'Dr. Ludovic Hermabessiere'),", University of Toronto", 
                           tags$a(href="https://twitter.com/HermabessiereL", icon("twitter"))),

### 2022-01-08 App_Nonessential/TeamX_Manuscript_Figures/Manuscript_Figures_TeamX.Rmd
team x manuscript figures

@@
 library(wesanderson)
+library(pheatmap)
+library(RColorBrewer)
+library(ggpubr)
+library(ssdtools)
+library(scales)
+
+library(ggdark)
+library(ggsci)
 
 #Import Data
-aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)
-
-#Data Tidying
-# Master dataset for scatterplots - for Heili's tab.
-aoc_v1 <- aoc %>% # start with original dataset
-   # full dataset filters.
-  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
-    effect == "N" ~ "No"),
-    levels = c("No", "Yes"))) %>%
-  # removing NAs to make data set nicer
-  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported", chem.exp.typ.nominal = "Particle Only"))
-
-aoc_setup <- aoc_v1 %>% # start with original dataset
-  mutate(size_f = factor(case_when(
-    size.category == 1 ~ "1nm < 100nm",
-    size.category == 2 ~ "100nm < 1µm",
-    size.category == 3.1 ~ "1µm < 10µm",
-    size.category == 3.2 ~ "10µm < 100µm",
-    size.category == 4 ~ "100µm < 1mm",
-    size.category == 5 ~ "1mm < 5mm",
-    size.category == 0 ~ "Not Reported"),
-    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 10µm", "10µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
-  # shape category data tidying.
-  mutate(shape_f = factor(case_when(
-    shape == "fiber" ~ "Fiber",
-    shape == "fragment" ~ "Fragment",
-    shape == "sphere" ~ "Sphere",
-    shape == "Not Reported" ~ "Not Reported"),
-    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% # order our different shapes.
-  # polymer category data tidying.
-  mutate(poly_f = factor(case_when(
-    polymer == "BIO" ~ "Biopolymer",
-    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
-    polymer == "LTX" ~ "Latex",
-    polymer == "PA" ~ "Polyamide",
-    polymer == "PE" ~ "Polyethylene",
-    polymer == "PC" ~ "Polycarbonate",
-    polymer == "PET" ~ "Polyethylene Terephthalate",
-    polymer == "PI" ~ "Polyisoprene",
-    polymer == "PMMA" ~ "Polymethylmethacrylate",
-    polymer == "PP" ~ "Polypropylene",
-    polymer == "PS" ~ "Polystyrene",
-    polymer == "PUR" ~ "Polyurethane",
-    polymer == "PVC" ~ "Polyvinylchloride",
-    polymer == "PLA" ~ "Polylactic Acid",
-    polymer == "Not Reported" ~ "Not Reported"))) %>%
-  # taxonomic category data tidying.
-  mutate(org_f = factor(organism.group, levels = c("Algae", "Annelida", "Bacterium", "Cnidaria", "Crustacea",
-                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% # order our different organisms.
-  mutate(lvl1_f = factor(case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
-    lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
-    lvl1 == "circulatory.respiratory" ~ "Circulatory, Respiratory",
-    lvl1 == "community" ~ "Community",
-    lvl1 == "fitness" ~ "Fitness",
-    lvl1 == "immune" ~ "Immune",
-    lvl1 == "metabolism" ~ "Metabolism",
-    lvl1 == "microbiome" ~ "Microbiome",
-    lvl1 == "stress" ~ "Stress"))) %>% # creates new column with nicer names.
-  # Level 2 Data tidying
-  mutate(lvl2_f = factor(case_when(lvl2 == "abundance"~"Abundance",
-    lvl2 == "actinobacteria" ~ "Actinobacteria",
-    lvl2 == "aggressivity"~"Agressivity",
-    lvl2 == "ammonia.excretion" ~ "Ammonia Excretion",
-    lvl2 == "bacteroidetes"~ "Bacteriodetes",
-    lvl2 == "blood"~"Blood",
-    lvl2 == "body.condition"~"Body Condition",
-    lvl2 == "boldness"~"Boldness",
-    lvl2 == "brain.histo"~"Brain Histological Abnormalities",
-    lvl2 == "burrowing"~"Burrowing",
-    lvl2 == "carb.metabolism"~"Carb Metabolism",
-    lvl2 == "chemokines.cytokines"~"Chemokines",
-    lvl2 == "circulatory"~"Circulatory",
-    lvl2 == "detoxification"~"Detoxification",
-    lvl2 == "development"~"Development",
-    lvl2 == "digestion"~"Digestion",
-    lvl2 == "digestive.enzymes"~"Digestive Enzymes",
-    lvl2 == "digestive.tract.histo"~"Digestive Tract Histological Abnormalities",
-    lvl2 == "diversity"~ "Diversity",
-    lvl2 == "feeding"~ "Feeding",
-    lvl2 == "firmicutes"~ "Firmicutes",
-    lvl2 == "gall.bladder.histo" ~ "Gall Bladder Histological Abnormalities",
-    lvl2 == "gen.metabolism"~ "General Metabolism",
-    lvl2 == "gill.histo"~ "Gill Histological Abnormalities",
-    lvl2 == "gonad.histo"~"Gonad Histological Abnormalities",
-    lvl2 == "growth"~ "Growth",
-    lvl2 == "immune.cells"~"Immune Cells",
-    lvl2 == "immune.other"~"Immune Other ",
-    lvl2 == "intestinal.permeability"~"Intestinal Permeability",
-    lvl2 == "kidney.histo"~"Kidney Histological abnormalities",
-    lvl2 == "lipid.metabolism"~"Lipid Metabolism",
-    lvl2 == "liver.histo"~"Liver Histological Abnormalities",
-    lvl2 == "liver.kidney.products" ~ "Liver and Kidney Products",
-    lvl2 == "locomotion"~"Locomotion",
-    lvl2 == "mortality"~"Mortality",
-    lvl2 == "nervous.system"~"Nervous System",
-    lvl2 == "oxidative.stress"~"Oxidative Stress",
-    lvl2 == "photosynthesis"~ "Photosynthesis",
-    lvl2 == "proteobacteria"~"Protebacteria",
-    lvl2 == "reproduction"~"Reproduction",
-    lvl2 == "respiration"~"Respiration",
-    lvl2 == "sexhormones"~"Sex Hormones",
-    lvl2 == "shoaling"~"Shoaling",
-    lvl2 == "stress"~"Stress",
-    lvl2 == "vision.system"~"Vision System"))) %>% #Renames for widget
-  #Level 3 endpoint tidying
-  mutate(lvl3_f = factor(case_when(lvl3 == "11bhsd.mrnaexpression"~"11β HSD mRNA expression",
-                                   lvl3 == "17bhsd.mrnaexpression"~"17β HSD mRNA expression",
-                                   lvl3 == "1433e.mrnaexpression"~"1433e mRNA expression",
-                                   lvl3 == "3bhsd.mrnaexpression"~"3β HSD mRNA expression",
-                                   lvl3 == "5ht1.mrnaexpression"~"5ht1 mRNA expression",
-                                   lvl3 == "aadd45a.mrnaexpression"~"aadd45a mRNA expression",
-                                   lvl3 == "abcb1.mrnaexpression"~"abcb1 mRNA expression",
-                                   lvl3 == "abcc.mrnaexpression"~"abcc mRNA expression",
-                                   lvl3 == "abcc5.mrnaexpression"~"abcc5 mRNA expression",
-                                   lvl3 == "abcd.mrnaexpression"~"abcd mRNA expression",
-                                   lvl3 == "abnormality.occurrence"~"Abnormality Occurrence",
-                                   lvl3 == "abs.rc"~"Absorption Flux/Reaction Center",
-                                   lvl3 == "absorption.efficiency"~"Absorption Efficiency",
-                                   lvl3 == "abundance"~"Abundance",
-                                   lvl3 == "acc1.mrnaexpression"~"acc1 mRNA expression",
-                                   lvl3 == "ace.diversity.index"~"ACE Diversity Index",
-                                   lvl3 == "ache.activity"~"AcHE Activity",
-                                   lvl3 == "ache.mrnaexpression"~"AcHE mRNA expression",
-                                   lvl3 == "acid.phosphatase.activity"~"Acid Phosphatase Activity",
-                                   lvl3 == "aco.mrnaexpression"~"aco mRNAexpression",
-                                   lvl3 == "acrosomal.integrity"~"Acrosomal Integrity",
-                                   lvl3 == "actinobacteria"~"Actinobacteria",
-                                   lvl3 == "activity"~"Activity",
-                                   lvl3 == "activity.during.feeding"~"Activity During Feeding",
-                                   lvl3 == "aeromonas"~"Aeromonas",
-                                   lvl3 == "agal.consumption"~"Agal Consumption",
-                                   lvl3 == "agression.index"~"Aggression Index",
-                                   lvl3 == "ahra.mrnaexpression"~"ahrα mRNA expression",
-                                   lvl3 == "ahrb.mrnaexpression"~"ahrβ mRNA expression",
-                                   lvl3 == "akt.mrnaexpression"~"akt mRNA expression",
-                                   lvl3 == "albumin.con"~"Albumin Concentration",
-                                   lvl3 == "albumin.globulin.ratio"~"Albumin to Globulin Ratio",
-                                   lvl3 == "algal.density"~"Algal Density",
-                                   lvl3 == "alkaline.phosphatase.activity"~"Alkaline Phosphatase Activity",
-                                   lvl3 == "aproteobacteria"~"α proteobacteria",
-                                   lvl3 == "alp.activity"~"Alkaline Phosphatase Activity",
-                                   lvl3 == "alp.con"~"Alkaline Phosphatase Concentration",
-                                   lvl3 == "alt.activity"~"Alanine Aminotransferase Activity",
-                                   lvl3 == "alt.con"~"Alanine Aminotransferase Concentration",
-                                   lvl3 == "amino.acid.trans"~"Amino Acid Transport",
-                                   lvl3 == "ammonia.con"~"Ammonia Concentration",
-                                   lvl3 == "ampka.mrnaexpression"~"ampka mRNA expression",
-                                   lvl3 == "ampkb.mrnaexpression"~"ampkb mRNA expression",
-                                   lvl3 == "ampkc.mrnaexpression"~"ampkc mRNA expression",
-                                   lvl3 == "amylase.mrnaexpression"~"Amylanse mRNA expression",
-                                   lvl3 == "antioxidant.act"~"Antioxidant Activity",
-                                   lvl3 == "aox.activity"~"Alternative Oxidase Activity",
-                                   lvl3 == "apo.mrnaexpression"~"Apolipoprotein mRNA expression",
-                                   lvl3 == "ar.mrnaexpression"~"Androgen Receptor mRNA expression",
-                                   lvl3 == "ara.mrnaexpression"~"Androgen Receptor α mRNA expression",
-                                   lvl3 == "arb.mrnaexpression"~"Androgen Receptor β mRNA expression",
-                                   lvl3 == "area"~"Area",
-                                   lvl3 == "area.traveled"~"Area Traveled",
-                                   lvl3 == "aspartate.aminotransferase.con"~"Aspartate Aminotransferase Concentration",
-                                   lvl3 == "assemblage"~"Assemblage",
-                                   lvl3 == "assimilation.efficiency"~"Assimilation Efficiency",
-                                   lvl3 == "ast.activity"~"Aspartate Aminotransferase Activity",
-                                   lvl3 == "ast.con"~"Aspartate Aminotransferase Concentration",
-                                   lvl3 == "attachment"~"Attachment",
-                                   lvl3 == "atubulin.mrnaexpression"~"α tubulin mRNA expression",
-                                   lvl3 == "atubulin.proteinexpression"~"α tubulin Protein expression",
-                                   lvl3 == "bacteroidetes"~"Bacteroidetes",
-                                   lvl3 == "bcl2.mrnaexpression"~"bcl2 mRNA expression",
-                                   lvl3 == "bfcod.activity"~"BFCOD Activity",
-                                   lvl3 == "bile.acid.con"~"Bile Acid Concentration",
-                                   lvl3 == "bilirubin.con"~"Bilirubin Concentration",
-                                   lvl3 == "biomass"~"Biomass",
-                                   lvl3 == "bioturbation"~"Bioturbation",
-                                   lvl3 == "body.length"~"Body Length",
-                                   lvl3 == "body.mass"~"Body Mass",
-                                   lvl3 == "body.width"~"Body Width",
-                                   lvl3 == "bproteobacteria"~"βproteobacteria",
-                                   lvl3 == "brain.histo"~"Brain Histology",
-                                   lvl3 == "brain.mass.index"~"Brain Mass Index",
-                                   lvl3 == "brain.water.content"~"Brain Water Content",
-                                   lvl3 == "burrow.length"~"Burrow Length",
-                                   lvl3 == "burrow.litter"~"Burrow Litter",
-                                   lvl3 == "burrow.no"~"Burrow Number",
-                                   lvl3 == "burrow.vol"~"Burrow Volume",
-                                   lvl3 == "burrow.walldensity"~"Burrow Wall Density",
-                                   lvl3 == "burrow.wallweight"~"Burrow Wall Weight",
-                                   lvl3 == "burrowing"~"Burrowing",
-                                   lvl3 == "ca.mrnaexpression"~"ca mRNA expression",
-                                   lvl3 == "calcification"~"Calcification",
-                                   lvl3 == "calcium.con"~"Calcium Concentration",
-                                   lvl3 == "capture.rate"~"Capture Rate",
-                                   lvl3 == "carbohydrate.reserves"~"Carbohydrate Reserves",
-                                   lvl3 == "cardiolipin.con"~"Cardiolipin Concentration",
-                                   lvl3 == "cas8.mrnaexpression"~"cas8 mRNA expression",
-                                   lvl3 == "casp3.mrnaexpression"~"casp3 mRNA expression",
-                                   lvl3 == "casp373.mrnaexpression"~"casp373 mRNA expression",
-                                   lvl3 == "casp3a.mrnaexpression"~"casp3a mRNA expression",
-                                   lvl3 == "casp8.mrnaexpression"~"casp8 mRNA expression",
-                                   lvl3 == "casp9.mrnaexpression"~"casp9 mRNA expression",
-                                   lvl3 == "caspase3.7.mrnaexpression"~"caspase 3/7 mRNA expression",
-                                   lvl3 == "cat.activity"~"Catalase Activity",
-                                   lvl3 == "cat.mrnaexpression"~"Catalase mRNA expression",
-                                   lvl3 == "cell.autoflourescence"~"Cell Autoflourescence",
-                                   lvl3 == "cell.density"~"Cell Density",
-                                   lvl3 == "cell.granularity"~"Cell Granularity",
-                                   lvl3 == "cell.morphology"~"Cell Morphology",
-                                   lvl3 == "cell.necrosis"~"Cell Necrosis",
-                                   lvl3 == "cell.size"~"Cell Size",
-                                   lvl3 == "cell.viability"~"Cell Viability",
-                                   lvl3 == "cellular.energy.allocation"~"Cellular Energy Allocation",
-                                   lvl3 == "chg.mrnaexpression"~"Choriogenin mRNA expression",
-                                   lvl3 == "chgh.mrnaexpression"~"Choriogenin H mRNA expression",
-                                   lvl3 == "chgl.mrnaexpression"~"Choriogenin L mRNA expression",
-                                   lvl3 == "chlorophyll"~"Chlorophyll",
-                                   lvl3 == "cholesterol.con"~"Cholesterol Concentration",
-                                   lvl3 == "chrna1.mrnaexpression"~"chrna1 mRNA expression",
-                                   lvl3 == "chymotrypsin.activity"~"Chymotrypsin Activity",
-                                   lvl3 == "clearance.rate"~"Clearance Rate",
-                                   lvl3 == "clutch.num"~"Number of Clutches",
-                                   lvl3 == "clutch.size"~"Clutch Size",
-                                   lvl3 == "clutch.size.first"~"First Clutch Size",
-                                   lvl3 == "clutch.size.firstthree"~"First Three Clutch Sizes",
-                                   lvl3 == "co2.consumption"~"CO2 Consumption",
-                                   lvl3 == "condition.index"~"Condition Index",
-                                   lvl3 == "congestion"~"Congestion",
-                                   lvl3 == "copper.con.hepatopancreas"~"Hepatopancreas Copper Concentration",
-                                   lvl3 == "cox.activity"~"Cyclooxygenase Activity",
-                                   lvl3 == "coxIV.mrnaexpression"~"Cyclooxygenase IV mRNA expression",
-                                   lvl3 == "cpt1.mrnaexpression"~"CPT1 mRNA expression",
-                                   lvl3 == "creatine.kinase.con"~"Creatine Kinase Concentration",
-                                   lvl3 == "crude.ash"~"Crude Ash",
-                                   lvl3 == "crude.protein"~"Curde Protein",
-                                   lvl3 == "cs.mrnaexpression"~"Chitin Synthase mRNA expression",
-                                   lvl3 == "ctsl.mrnaexpression"~"Cathepsin-L mRNA expression",
-                                   lvl3 == "cuznsod.mrnaexpression"~"CuZn Superoxide Dismutase mRNA expression",
-                                   lvl3 == "cxcr5.mrnaexpression"~"cxcr5 mRNA expression",
-                                   lvl3 == "cyp11.mrnaexpression"~"cyp11 mRNA expression",
-                                   lvl3 == "cyp11a2.mrnaexpression"~"cyp11a2 mRNA expression",
-                                   lvl3 == "cyp17a1.mrnaexpression"~"cyp17a1 mRNA expression",
-                                   lvl3 == "cyp19a.mrnaexpression"~"cyp19a mRNA expression",
-                                   lvl3 == "cyp19a2.mrnaexpression"~"cyp19a2 mRNA expression",
-                                   lvl3 == "cyp19b.mrnaexpression"~"cyp19b mRNA expression",
-                                   lvl3 == "cyp1a.mrnaexpression"~"cyp1a  mRNA expression",
-                                   lvl3 == "cyp1a1.mrnaexpression"~"cyp1a1 mRNA expression",
-                                   lvl3 == "cyp32.mrnaexpression"~"cyp32 mRNA expression",
-                                   lvl3 == "cyp450.proteinexpression"~"cyp450 mRNA expression",
-                                   lvl3 == "cyp4an1.mrnaexpression"~"cyp4an1 mRNA expression",
-                                   lvl3 == "cyp4c33.mrnaexpression"~"cyp4c33 mRNA expression",
-                                   lvl3 == "cyp4c34.mrnaexpression"~"cyp4c34 mRNA expression",
-                                   lvl3 == "cytoplasmic.membrane.potential"~"Cytoplasmic Membrane Potential",
-                                   lvl3 == "cytotoxicity"~"Cytotoxicity",
-                                   lvl3 == "d.lactate"~"D Lactate Concentration",
-                                   lvl3 == "dao.activity"~"Diamine Oxidase Activity",
-                                   lvl3 == "degranulation"~"Degranulation",
-                                   lvl3 == "development"~"Development",
-                                   lvl3 == "developmental.abnormalities"~"Developmental Abnormalities",
-                                   lvl3 == "dgat.mrnaexpression"~"Diglyceride acyltransferase mRNA expression",
-                                   lvl3 == "digestive.tract.histo"~"Digestive Tract Histology",
-                                   lvl3 == "dio.rc"~"Energy Dissipated/Reaction Center",
-                                   lvl3 == "distance.between.fish"~"Distance Between Fish",
-                                   lvl3 == "distance.during.feeding"~"Distance During Feeding",
-                                   lvl3 == "diversity"~"Diversity",
-                                   lvl3 == "dlac.activity"~"DLac Activity",
-                                   lvl3 == "dna.damage"~"DNA Damage",
-                                   lvl3 == "dopamine.con"~"Dopamine Concentration",
-                                   lvl3 == "egg.size"~"Egg Size",
-                                   lvl3 == "embryotoxicity"~"Embryotoxicity",
-                                   lvl3 == "emergence"~"Emergence",
-                                   lvl3 == "endopeptidase.activity"~"Endopeptidase Activity",
-                                   lvl3 == "energy.balance"~"Energy Balance",
-                                   lvl3 == "energy.reserves"~"Energy Reserves",
-                                   lvl3 == "ep.mrnaexpression"~"ep mRNA expression",
-                                   lvl3 == "era.mrnaexpression"~"Estrogen Receptor α mRNA expression",
-                                   lvl3 == "erb.mrnaexpression"~"Estrogen Receptor β mRNA expression",
-                                   lvl3 == "erk.mrnaexpression"~"erk mRNA expression",
-                                   lvl3 == "erod.activity"~"EROD Activity",
-                                   lvl3 == "esr2.mrnaexpression"~"esr2 mRNA expression",
-                                   lvl3 == "esterase.activity"~"Esterase Activity",
-                                   lvl3 == "estradiol.level"~"Estradiol Levels",
-                                   lvl3 == "estradiol.testosterone.ratio"~"Estradiol to Testosterone Ratio",
-                                   lvl3 == "eto.abs"~"ETO/ABS",
-                                   lvl3 == "eto.rc"~"ETO/RC",
-                                   lvl3 == "exopeptidase.activity"~"Exopeptidase Activity",
-                                   lvl3 == "exploration"~"Exploration",
-                                   lvl3 == "f0.fm"~"f0/fm",
-                                   lvl3 == "f0.fv"~"f0/fv",
-                                   lvl3 == "fabp6.mrnaexpression"~"fabp6 mRNA expression",
-                                   lvl3 == "fadd.mrnaexpression"~"fadd mRNA expression",
-                                   lvl3 == "fas.mrnaexpression"~"fas mRNA expression",
-                                   lvl3 == "fatty.vac"~"Fatty Vacuoles",
-                                   lvl3 == "fecundity"~"Fecundity",
-                                   lvl3 == "feeding.activity"~"Feeding Activity",
-                                   lvl3 == "feeding.rate"~"Feeding Rate",
-                                   lvl3 == "feeding.time"~"Feeding Time",
-                                   lvl3 == "fertilization"~"Fertilization",
-                                   lvl3 == "fertilization.rate"~"Fertilization Rate",
-                                   lvl3 == "fertilization.yield"~"Fertilization Yield",
-                                   lvl3 == "fibrosis"~"Fiborsis",
-                                   lvl3 == "filter.feeding"~"Filter Feeding",
-                                   lvl3 == "firmicutes"~"Firmicutes",
-                                   lvl3 == "food.consumption"~"Food Consumption",
-                                   lvl3 == "foraging.activity"~"Foraging Activity",
-                                   lvl3 == "foraging.time"~"Foraging Time",
-                                   lvl3 == "foxl2.mrnaexpression"~"fox12 mRNA expression",
-                                   lvl3 == "free.fatty.acid.con"~"Free Fatty Acid Concentration",
-                                   lvl3 == "fshb.mrnaexpression"~"fshβ mRNA expression",
-                                   lvl3 == "fshr.mrnaexpression"~"fshr mRNA expression",
-                                   lvl3 == "ftzf1.mrnaexpression"~"ftzf1 mRNA expression",
-                                   lvl3 == "fv'.fm'"~"fv'/fm",
-                                   lvl3 == "fv.f0"~"fv/f0",
-                                   lvl3 == "fv.fm"~"fv/fm",
-                                   lvl3 == "fv.fm.sm.tfm"~"fv/fm/sm/tfm",
-                                   lvl3 == "gadph.mrnaexpression"~"gadhph mRNA expression",
-                                   lvl3 == "gall.bladder.histo"~"Gall Bladder Histology",
-                                   lvl3 == "gallery.vol"~"Gallery Volume",
-                                   lvl3 == "gcl.mrnaexpression"~"gcl mRNA expression",
-                                   lvl3 == "gcl1.mrnaexpression"~"gcl1 mRNA expression",
-                                   lvl3 == "gclcs.mrnaexpression"~"gclcs mRNA expression",
-                                   lvl3 == "gclcu.mrnaexpression"~"gclcu mRNA expression",
-                                   lvl3 == "genera"~"Genera",
-                                   lvl3 == "gfap.mrnaexpression"~"gfap mRNA expression",
-                                   lvl3 == "ggt.activity"~"Gamma-Glutamyl Transferase Activity",
-                                   lvl3 == "ggt.con"~"Gamma-Glutamyl Transferase Concentration",
-                                   lvl3 == "gill.histo"~"Gill Histology",
-                                   lvl3 == "gk.mrnaexpression"~"gk mRNA expression",
-                                   lvl3 == "globulin.con"~"Globulin Concentration",
-                                   lvl3 == "glucose.con"~"Glucose Concentration",
-                                   lvl3 == "glutamate.con"~"Glutamate Concentration",
-                                   lvl3 == "glutathione.con"~"Glutathione Concentration",
-                                   lvl3 == "glycogen.con"~"Glycogen Concentration",
-                                   lvl3 == "glycogen.deplet"~"Glycogen Depletion",
-                                   lvl3 == "gnrh.mrnaexpression"~"gnrh mRNA expression",
-                                   lvl3 == "gnrhr.mrnaexpression"~"gnrhr mRNA expression",
-                                   lvl3 == "gonad.development.index"~"Gonad Development Index",
-                                   lvl3 == "gonad.histo"~"Gonad Histology",
-                                   lvl3 == "gonad.regression"~"Gonad Regression",
-                                   lvl3 == "gonad.somatic.index"~"Gonad Somatic Index",
-                                   lvl3 == "gordonia"~"Gordonia",
-                                   lvl3 == "got.activity"~"Glutamic Oxaloacetic Transaminase Activity",
-                                   lvl3 == "gpt.activity"~"Glutamic Pyruvic Transaminase Activity",
-                                   lvl3 == "gpx.activity"~"Glutathione Peroxidase Activity",
-                                   lvl3 == "gpx.mrnaexpression"~"Glutathione Peroxidase mRNA expression",
-                                   lvl3 == "gpx4b.mrnaexpression"~"gpx4b mRNA expression",
-                                   lvl3 == "gr.activity"~"Glutathione Reductase Activity",
-                                   lvl3 == "gr.mrnaexpression"~"Glutathione Reductase mRNA expression",
-                                   lvl3 == "granulocyte.con"~"Granulocyte Concentration",
-                                   lvl3 == "granulocyte.hyalinocyte.ratio"~"Granulocyte to Hyalinocyte Ratio",
-                                   lvl3 == "granulocyte.oxidative.activity"~"Granulocyte Oxidative Activity",
-                                   lvl3 == "granulocyte.size"~"Granulocyte Size",
-                                   lvl3 == "gross.energy"~"Gross Energy",
-                                   lvl3 == "growth"~"Growth",
-                                   lvl3 == "growth.rate"~"Growth Rate",
-                                   lvl3 == "gs.mrnaexpression"~"Glutathione Synthetase mRNA expression",
-                                   lvl3 == "gsh-px.activity"~"Glutathione Peroxidase Activity",
-                                   lvl3 == "gsh.activity"~"Glutathione Peroxidase Activity",
-                                   lvl3 == "gsh.con"~"Glutathione Peroxidase Concentration",
-                                   lvl3 == "gsp.activity"~"Glutathione Peroxidase Activity",
-                                   lvl3 == "gst.activity"~"Glutathione S-Transferase Activity",
-                                   lvl3 == "gst.mrnaexpression"~"Glutathione S-Transferase mRNA expression",
-                                   lvl3 == "gst4.mrnaexpression"~"gst4 mRNA expression",
-                                   lvl3 == "gstd.mrnaexpression"~"gstd mRNA expression",
-                                   lvl3 == "gstp1.mrnaexpression"~"Glutathione S-Transferase mRNA expression",
-                                   lvl3 == "gt.con"~"Glutathione Concentration",
-                                   lvl3 == "gtha.mrnaexpression"~"gtha mRNA expression",
-                                   lvl3 == "gusb.mrnaexpression"~"gusb mRNA expression",
-                                   lvl3 == "gut.mucus.vol"~"Gut Mucus Volume",
-                                   lvl3 == "gyrus.size"~"Gyrus Size",
-                                   lvl3 == "h2o2.con"~"Hydorgen Peroxide Concentration",
-                                   lvl3 == "h3.mrnaexpression"~"h3 mRNA expression",
-                                   lvl3 == "haemocyanin.con"~"Haemocyanin Concentration",
-                                   lvl3 == "hemocyte.con"~"Hemocyte Concentration",
-                                   lvl3 == "haemolymph.con"~"Haemolymph Concentration",
-                                   lvl3 == "hatching.success"~"Hatching Success",
-                                   lvl3 == "hdl.con"~"HDL Concentration",
-                                   lvl3 == "heart.rate"~"Heart Rate",
-                                   lvl3 == "hepatic.somatic.index"~"Hepatic Somatic Index",
-                                   lvl3 == "hepatopancreas.water.con"~"Hepatopancreas Water Concentration",
-                                   lvl3 == "hex.mrnaexpression"~"hex mRNA expression",
-                                   lvl3 == "hk.mrnaexpression"~"hk mRNA expression",
-                                   lvl3 == "hk1.mrnaexpression"~"hk1 mRNA expression",
-                                   lvl3 == "ho1.mrnaexpression"~"ho1 mRNA expression",
-                                   lvl3 == "hsp70.mrnaexpression"~"hsp70 mRNA expression",
-                                   lvl3 == "hsp90.mrnaexpression"~"hsp90 mRNA expression",
-                                   lvl3 == "hyalinocyte.size"~"Hyalinocyte Size",
-                                   lvl3 == "hylaniocyte.oxidative.activity"~"Hylaniocyte Oxidative Activity",
-                                   lvl3 == "id.mrnaexpression"~"Isocitrate Dehydrogenase mRNA expression",
-                                   lvl3 == "idh.activity"~"Isocitrate Dehydrogenase Activity",
-                                   lvl3 == "idp.mrnaexpression"~"idp mRNA expression",
-                                   lvl3 == "ifn.mrnaexpression"~"Interferon mRNA expression",
-                                   lvl3 == "ifny.mrnaexpression"~"Interferon y mRNA expression",
-                                   lvl3 == "igm.con"~"Immunoglobulin M Concentration",
-                                   lvl3 == "igm.mrnaexpression"~"Immunoglobulin M mRNA expression",
-                                   lvl3 == "il10.mrnaexpression"~"Interleukin 10 mRNA expression",
-                                   lvl3 == "il17.mrnaexpression"~"Interleukin 17 mRNA expression",
-                                   lvl3 == "il1a.mrnaexpression"~"Interleukin 1α mRNA expression",
-                                   lvl3 == "il1b.mrnaexpression"~"Interleukin 1β mRNA expression",
-                                   lvl3 == "il413a.mrnaexpression"~"Interleukin 413α mRNA expression",
-                                   lvl3 == "il6.mrnaexpression"~"Interleukin 6 mRNA expression",
-                                   lvl3 == "il8.mrnaexpression"~"Interluekin 8 mRNA expression",
-                                   lvl3 == "ila.con"~"Interleukin 1α Concentration",
-                                   lvl3 == "immobilization"~"Immobilization",
-                                   lvl3 == "inflam.infiltrate"~"Inflammatory Filtrate",
-                                   lvl3 == "ingestion.rate"~"Ingestion Rate",
-                                   lvl3 == "inhibitory.rate"~"Inhibitory Rate",
-                                   lvl3 == "instar.develop.time"~"Instar Development Time",
-                                   lvl3 == "intermolt.time"~"Intermolt Time",
-                                   lvl3 == "intestinal.ca.con"~"Intestinal Calcium Concentration",
-                                   lvl3 == "ion.exch.ca"~"Ion Exchange (Calcium)",
-                                   lvl3 == "ion.exch.k"~"Ion Exchange (Potassium)",
-                                   lvl3 == "ion.exch.na"~"Ion Exchange (Sodium)",
-                                   lvl3 == "ion.flow"~"Ion Flow",
-                                   lvl3 == "ion.permeability"~"Ion Permeability",
-                                   lvl3 == "ion.trans"~"Ion Transport",
-                                   lvl3 == "jaw.tissue.damage"~"Jaw Tissue Damage",
-                                   lvl3 == "jnk.mrnaexpression"~"jnk mRNA expression",
-                                   lvl3 == "jump.freq"~"Jump Frequency",
-                                   lvl3 == "jump.height"~"Jump Height",
-                                   lvl3 == "kidney.histo"~"Kidney Histology",
-                                   lvl3 == "labilization.period"~"Labilization Period",
-                                   lvl3 == "lactate.con"~"Lactate Concentration",
-                                   lvl3 == "ldh.activity"~"Lactate Dehydrogenase Activity",
-                                   lvl3 == "ldh.con"~"Lactate Dehydrogenase Concentration",
-                                   lvl3 == "ldl.con"~"LDL Concentration",
-                                   lvl3 == "leaf.growth.rate"~"Leaf Growth Rate",
-                                   lvl3 == "length"~"Length",
-                                   lvl3 == "leukocyte.con"~"Leukocyte Concentration",
-                                   lvl3 == "lhb.mrnaexpression"~"Leutinizing Hormone β mRNA expression",
-                                   lvl3 == "lhr.mrnaexpression"~"Leutinizing Hormone Receptor mRNA expression",
-                                   lvl3 == "lifespan"~"Lifespan",
-                                   lvl3 == "light.capture"~"Light Capture",
-                                   lvl3 == "light.dark.activity"~"Light/Dark Activity",
-                                   lvl3 == "line.crossing"~"Line Crossing",
-                                   lvl3 == "lipase.activity"~"Lipase Activity",
-                                   lvl3 == "lipase.con"~"Lipase Concentration",
-                                   lvl3 == "lipid.content"~"Lipid Content",
-                                   lvl3 == "lipid.membrane.con"~"Lipid Membrane Concentration",
-                                   lvl3 == "lipid.perox"~"Lipid Peroxidation",
-                                   lvl3 == "lipid.storage.con"~"Lipid Storage Concentration",
-                                   lvl3 == "lipofuscin.accum"~"Lipofuscin Accumulation",
-                                   lvl3 == "liver.histo"~"Liver Histology",
-                                   lvl3 == "liver.protein"~"Liver Protein",
-                                   lvl3 == "lymphocyte.con"~"Lymphocyte Concentration",
-                                   lvl3 == "lys.mrnaexpression"~"Lysozyme mRNA expression",
-                                   lvl3 == "lysosomal.membrane.stab"~"Lysosomal Membrane Stability",
-                                   lvl3 == "lysozyme.activity"~"Lysozyme Activity",
-                                   lvl3 == "manf.mrnaexpression"~"manf mRNA expression",
-                                   lvl3 == "mao.activity"~"Monoamine Oxidase Activity",
-                                   lvl3 == "mapk.phosphorylation"~"MAPK Phosphorylation",
-                                   lvl3 == "mbp.mrnaexpression"~"Myelin Basic Protein mRNA expression",
-                                   lvl3 == "mbp.proteinexpression"~"Myelin Basic Protein Expression",
-                                   lvl3 == "mda.con"~"Malondialdehyde Concentration",
-                                   lvl3 == "meer1.mrnaexpression"~"meer 1 mRNA expression",
-                                   lvl3 == "meer2.mrnaexpression"~"meer 2 mRNA expression",
-                                   lvl3 == "mek.mrnaexpression"~"mek mRNA expression",
-                                   lvl3 == "metamorphosis.yield"~"Metamorphosis Yield",
-                                   lvl3 == "mgnrh.mrnaexpression"~"mgnrh mRNA expression",
-                                   lvl3 == "min.sat.irradiance"~"Minimum Saturation Irradiance",
-                                   lvl3 == "mirconuclei.freq"~"Micronuclei Frequency",
-                                   lvl3 == "mitochondrial.mem.integrity"~"Mitochondrial Membrane Integrity",
-                                   lvl3 == "mnsod.mrnaexpression"~"Mn Superoxide Dismutase mRNA expression",
-                                   lvl3 == "moisture"~"Moisture",
-                                   lvl3 == "molecular.permeability"~"Molecular Permeability",
-                                   lvl3 == "molt.period.length"~"Molt Period Length",
-                                   lvl3 == "mortality"~"Mortality",
-                                   lvl3 == "mt10.mrnaexpression"~"mt10 mRNA expression",
-                                   lvl3 == "mt20.mrnaexpression"~"mt20 mRNA expression",
-                                   lvl3 == "mta.mrnaexpression"~"mta mRNA expression",
-                                   lvl3 == "mtb.mrnaexpression"~"mtb mRNA expression",
-                                   lvl3 == "mxr.efflux"~"MXR Efflux",
-                                   lvl3 == "mytc.mrnaexpression"~"mytc mRNA expression",
-                                   lvl3 == "myticina.mrnaexpression"~"myticin α mRNA expression",
-                                   lvl3 == "mytilinb.mrnaexpression"~"mytilin β mRNA expression",
-                                   lvl3 == "nauplius.to.adult"~"Nauplius to Adult",
-                                   lvl3 == "nauplius.to.copepodid"~"Nauplius to Copepodid",
-                                   lvl3 == "neutrophil.extracell.trap"~"Neutrophil Extracellular Trap",
-                                   lvl3 == "ngn1.mrnaexpression"~"ngn1 mRNA expression",
-                                   lvl3 == "nitrogen.conversion"~"Nitrogen Conversion",
-                                   lvl3 == "no.prod"~"Nitric Oxide Production",
-                                   lvl3 == "nrf2.mrnaexpression"~"nrf2 mRNA expression",
-                                   lvl3 == "num.indiv.nosnails"~"Number of Individuals (No Snails)",
-                                   lvl3 == "num.indiv.snails"~"Number of Individuals (With Snails)",
-                                   lvl3 == "num.offspring"~"Number of Offspring",
-                                   lvl3 == "number.of.molts"~"Number of Molts",
-                                   lvl3 == "number.of.roots"~"Number of Roots",
-                                   lvl3 == "occludin.mrnaexpression"~"occludin mRNA expression",
-                                   lvl3 == "offspring.size"~"Offspring Size",
-                                   lvl3 == "oocyte.diameter"~"Oocyte Diameter",
-                                   lvl3 == "oocyte.num"~"Oocyte Number",
-                                   lvl3 == "oxygen.consumption"~"Oxygen Consumption",
-                                   lvl3 == "p38.mrnaexpression"~"p38 mRNA expression",
-                                   lvl3 == "p38mapk.mrnaexpression"~"p38mapk mRNA expression",
-                                   lvl3 == "p53.mrnaexpression"~"p53 mRNA expression",
-                                   lvl3 == "pc.pe.ratio"~"PC/PE Ratio",
-                                   lvl3 == "pcc.level"~"Protein Carbonylation Content",
-                                   lvl3 == "pche.activity"~"Pseudocholinesterase Activity",
-                                   lvl3 == "pepckc.mrnaexpression"~"pepckc mRNA expression",
-                                   lvl3 == "percent.dstage.larvae"~"Pertengage of D-stage Larvae",
-                                   lvl3 == "percent.dveliger"~"Percentage of Veliger Larvae",
-                                   lvl3 == "percent.motile.sperm"~"Percentage of Motile Sperm",
-                                   lvl3 == "percent.tank.used"~"Percentage of Tank Used",
-                                   lvl3 == "percent.trocophore.larvae"~"Percentage of Trocophore Larvae",
-                                   lvl3 == "pericardial.area"~"Pericardial Area",
-                                   lvl3 == "perox.activity"~"Peroxidase Activity",
-                                   lvl3 == "perox.mrnaexpression"~"Peroxidase mRNA expression",
-                                   lvl3 == "petb.mrnaexpression"~"petb mRNA expression",
-                                   lvl3 == "pglycoprotein.activity"~"Pglycoprotein Activity",
-                                   lvl3 == "pgp.mrnaexpression"~"pgp mRNA expression",
-                                   lvl3 == "pgrp.mrnaexpression"~"pgrp mRNA expression",
-                                   lvl3 == "phagocytosis"~"Phagocytosis",
-                                   lvl3 == "phenoloxydase.activity"~"Phenoloxydase Activity",
-                                   lvl3 == "phosphorus.con"~"Phosphorus Concentration",
-                                   lvl3 == "phosphatase.activity"~"Phosphatase Activity",
-                                   lvl3 == "phosphatidylcholine.con"~"Phosphatidylcholine Concentration",
-                                   lvl3 == "phosphatidylethanolamine.con"~"Phosphatidylethanolamine Concentration",
-                                   lvl3 == "phosphatidylinositol.ceramide.amino.ethylphosphonate.con"~"PCAE Concentration",
-                                   lvl3 == "phosphatidylserine.con"~"Phosphatidylserine Concentration",
-                                   lvl3 == "photochemical.efficiency"~"Photochemical Efficiency",
-                                   lvl3 == "pk.mrnaexpression"~"pk mRNA expression",
-                                   lvl3 == "polyp.activity"~"Polyp Activity",
-                                   lvl3 == "population.size"~"Population Size",
-                                   lvl3 == "potassium.con"~"Potassium Concentration",
-                                   lvl3 == "ppara.mrnaexpression"~"pparα mRNA expression",
-                                   lvl3 == "ppary.mrnaexpression"~"ppary mRNA expression",
-                                   lvl3 == "prdx1.mrnaexpression"~"prdx1 mRNA expression",
-                                   lvl3 == "prdx2.mrnaexpression"~"prdx2 mRNA expression",
-                                   lvl3 == "prdx3.mrnaexpression"~"prdx3 mRNA expression",
-                                   lvl3 == "prdx5.mrnaexpression"~"prdx5 mRNA expression",
-                                   lvl3 == "predation.risk"~"Predation Risk",
-                                   lvl3 == "predatory.performance"~"Predatory Performance",
-                                   lvl3 == "pregnancy.rate"~"Pregnancy Rate",
-                                   lvl3 == "protein.carbonyl.con"~"Protein Carbonyl Concentration",
-                                   lvl3 == "protein.con"~"Protein Concentration",
-                                   lvl3 == "protein.reserves"~"Protein Reserves",
-                                   lvl3 == "proteobacteria"~"proteobacteria",
-                                   lvl3 == "ps2"~"Photosystem II",
-                                   lvl3 == "psaa.mrnaexpression"~"psaa mRNA expression",
-                                   lvl3 == "psab.mrnaexpression"~"psab mRNA expression",
-                                   lvl3 == "psba.mrnaexpression"~"psba mRNA expression",
-                                   lvl3 == "psbl.mrnaexpression"~"psbl mRNA expression",
-                                   lvl3 == "pseudomonas"~"pseudomonas",
-                                   lvl3 == "rapid.light.resp.curve"~"Rapid Light Response Curve",
-                                   lvl3 == "rbc.con"~"Red Blood Cell Concentration",
-                                   lvl3 == "rbcl.mrnaexpression"~"rbcl mRNA expression",
-                                   lvl3 == "rc.abs"~"RC/ABS",
-                                   lvl3 == "redox.state"~"Redox State",
-                                   lvl3 == "relative.e.transport.rate"~"Relative e Transport State",
-                                   lvl3 == "relocate"~"Relocation",
-                                   lvl3 == "reproduction.factor"~"Reproduction Factor",
-                                   lvl3 == "reproduction.inhibition"~"Reproduction Inhibition",
-                                   lvl3 == "reproductive.freq"~"Reproductive Frequency",
-                                   lvl3 == "resistance.time"~"Resistance Time",
-                                   lvl3 == "respiration.rate"~"Respiration Rate",
-                                   lvl3 == "respiratory.burst"~"Respiratory Burst",
-                                   lvl3 == "retrmax"~"Maximal Election Transport Rate",
-                                   lvl3 == "root.length"~"Root Length",
-                                   lvl3 == "ros.prod"~"Reactive Oxygen Species Production",
-                                   lvl3 == "s100a1.mrnaexpression"~"s100a1 mRNA expression",
-                                   lvl3 == "saa.mrnaexpression"~"saa mRNA expression",
-                                   lvl3 == "sd.mrnaexpression"~"sd mRNA expression",
-                                   lvl3 == "segment.regeneration"~"Segment Regeneration",
-                                   lvl3 == "serotonin.con"~"Serotonin Concentration",
-                                   lvl3 == "settling.rate"~"Settling Rate",
-                                   lvl3 == "sex.ratio"~"Sex Ratio",
-                                   lvl3 == "sexual.maturity"~"Sexual Maturity",
-                                   lvl3 == "sgst.mrnaexpression"~"sgst mRNA expression",
-                                   lvl3 == "shannon.diversity.index"~"Shannon Diversity Index",
-                                   lvl3 == "shell.growth"~"Shell Growth",
-                                   lvl3 == "shell.length"~"Shell Length",
-                                   lvl3 == "shell.weight"~"Shell Weight",
-                                   lvl3 == "size"~"Size",
-                                   lvl3 == "sod.activity"~"Superoxide Dismutase Activity",
-                                   lvl3 == "sod.mrnaexpression"~"sod mRNA expression",
-                                   lvl3 == "sod1.mrnaexpression"~"sod1 mRNA expression",
-                                   lvl3 == "sod3.mrnaexpression"~"sod3 mRNA expression",
-                                   lvl3 == "sodium.con"~"Sodium Concentration",
-                                   lvl3 == "species.diversity"~"Species Diversity",
-                                   lvl3 == "species.richness"~"Species Richness",
-                                   lvl3 == "specific.growth.rate"~"Specific Growth Rate",
-                                   lvl3 == "sperm.velocity"~"Sperm Velocity",
-                                   lvl3 == "srebp1a.mrnaexpression"~"srebp1a mRNA expression",
-                                   lvl3 == "star.mrnaexpression"~"star mRNA expression",
-                                   lvl3 == "sterol.con"~"Sterol Concentration",
-                                   lvl3 == "surface.area"~"Surface Area",
-                                   lvl3 == "swim.distance"~"Swim Distance",
-                                   lvl3 == "swim.speed"~"Swim Speed",
-                                   lvl3 == "swimbladder.inflation"~"Swimbladder Inflation",
-                                   lvl3 == "synapsin.mrnaexpression"~"synapsin mRNA expression",
-                                   lvl3 == "taxa.num"~"Number of Taxa",
-                                   lvl3 == "testosterone.level"~"Testosterone Levels",
-                                   lvl3 == "tgfb.mrnaexpression"~"tgfb mRNA expression",
-                                   lvl3 == "thiol.con"~"Thiol Concentration",
-                                   lvl3 == "thrombocyte.con"~"Thrombocyte Concentration",
-                                   lvl3 == "time.to.hatch"~"Time to Hatch",
-                                   lvl3 == "time.to.maturation"~"Time to Maturation",
-                                   lvl3 == "time.to.nauplius"~"Time to Nauplius",
-                                   lvl3 == "time.to.offspring"~"Time to Offspring",
-                                   lvl3 == "tlr.mrnaexpression"~"tlr mRNA expression",
-                                   lvl3 == "tnfa.mrnaexpression"~"tnfα mRNA expression",
-                                   lvl3 == "tnfsf13b.mrnaexpression"~"tnfsf13b mRNA expression",
-                                   lvl3 == "total.distance.traveled"~"Total Distance Travelled",
-                                   lvl3 == "tp53.con"~"tp53 Concentration",
-                                   lvl3 == "tp53.mrnaexpression"~"tp53 mRNA expression",
-                                   lvl3 == "tph2.mrnaexpression"~"tph2 mRNA expression",
-                                   lvl3 == "tricellulin.mrnaexpression"~"tricellulin mRNA expression",
-                                   lvl3 == "triglyceride.cholesterol.ratio"~"Tryglyceride Cholesterol Ratio",
-                                   lvl3 == "triglycerides.con"~"Triglyceride Concentration",
-                                   lvl3 == "tro.rc"~"TRO/RC",
-                                   lvl3 == "trypsin.activity"~"Trypsin Activity",
-                                   lvl3 == "tumor.promotion"~"Tumor Promotion",
-                                   lvl3 == "turn.angle.light.dark.activity"~"Turn Angle, Light/Dark Activity",
-                                   lvl3 == "ucp1.mrnaexpression"~"ucp1 mRNA expression",
-                                   lvl3 == "ugst.mrnaexpression"~"ugst mRNA expression",
-                                   lvl3 == "uricacid.con"~"Uric Acid Concentration",
-                                   lvl3 == "viserosomatic.index"~"Viserosomatic Index",
-                                   lvl3 == "volume"~"Volume",
-                                   lvl3 == "vtg.mrnaexpression"~"vtg mRNA expression",
-                                   lvl3 == "vtg.proteinexpression"~"vtg protein expression",
-                                   lvl3 == "vtg1.mrnaexpression"~"vtg1 mRNA expression",
-                                   lvl3 == "vtg2.mrnaexpression"~"vtg2 mRNA expression",
-                                   lvl3 == "weight.change"~"Weight Change",
-                                   lvl3 == "wgst.mrnaexpression"~"wgst mRNA expression",
-                                   lvl3 == "wnt16.mrnaexpression"~"wnt16 mRNA expression",
-                                   lvl3 == "wnt4.mrnaexpression"~"wnt4 mRNA expression",
-                                   lvl3 == "yproteobacteria"~"y proteobacteria",
-                                   lvl3 == "zfblue.mrnaexpression"~"zfblue mRNA expression",
-                                   lvl3 == "zfrho.mrnaexpression"~"zfrho mRNA expression",
-                                   lvl3 == "zo1.mrnaexpression"~"zo1 mRNA expression",
-                                   lvl3 == "zoozanthelle.density"~"Zoozanthelle Density"))) %>%
-  mutate(bio_f = factor(case_when( #Bio Org Data Tidying
-    bio.org == "subcell"~"Subcell",
-    bio.org == "cell"~"Cell",
-    bio.org == "tissue" ~ "Tissue",
-    bio.org == "organism"~"Organism",
-    bio.org == "population"~ "Population")))%>%
-  mutate(vivo_f = factor(case_when(invitro.invivo == "invivo"~"In Vivo",
-    invitro.invivo == "invitro"~"In Vitro")))%>% ##Renames for widget (Not using a widget right now, but saving for human health database)
-  mutate(life_f = factor(case_when(life.stage == "Early"~"Early",
-    life.stage == "Juvenile"~"Juvenile",
-    life.stage == "Adult"~"Adult",
-    life.stage == "Not Reported"~"Not Reported")))%>% #Renames for widget
-  mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
-    environment == "Marine" ~ "Marine",
-    environment == "Terrestrial" ~ "Terrestrial"))) %>%
-  mutate(species_f = as.factor(paste(genus,species))) %>% 
-  mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
-  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
-  mutate(effect.metric = factor(effect.metric)) %>% #factorize
-  mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) %>%  #calculate volume/mL
-  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
-  mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
-                                            af.time_noNA == 1 ~ "Chronic",
-                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>% #factorize assesment factor time into chronic/acute
-  mutate(tier_zero_tech_f = factor(case_when(tech.tier.zero == "Fail" ~ "Red Criteria Failed",
-                                             tech.tier.zero == "Pass" ~ "Red Criteria Passed"))) %>% 
-  mutate(tier_zero_risk_f = factor(case_when(risk.tier.zero == "Fail" ~ "Red Criteria Failed",
-                                             risk.tier.zero == "Pass" ~ "Red Criteria Passed"))) %>% 
-  #Remove leachate and additive/chemical transfer experiments
-  dplyr::filter(leachate.only != "Y") %>%
-  mutate(chem.exp.typ.nominal_f = factor(case_when(chem.exp.typ.nominal == "Particle Only" ~ "Particle Only",
-                                                   chem.exp.typ.nominal == "co.exp" ~ "Chemical Co-Exposure",
-                                                   chem.exp.typ.nominal == "sorbed" ~ "Chemical Transfer"))) %>% 
-  dplyr::filter(chem.exp.typ.nominal_f == "Particle Only") %>% 
-  #calculate maximum ingestible size (if not already in database)
-  mutate(max.size.ingest.mm = ifelse(is.na(max.size.ingest.mm), 
-    10^(0.9341 * log10(body.length.cm - 1.1200)) * 10,  #(Jamm et al 2020 Nature paper)correction for cm to mm
-    max.size.ingest.mm)) # if already present, just use that
+aoc_setup <- readRDS("aoc_setup.RDS")
+aoc_z <- readRDS("aoc_z.RDS")
 ```
 
-# Ludo
+# Figure 1
 
-```{r Ludo, warning = FALSE, message = FALSE, fig.height = 7, fig.width = 12}
+```{r}
 
-a <- aoc_setup %>% 
-  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
-  filter(shape_f != "Not Reported") %>% #drop studies where shapes aren't reported
-  filter(effect_f != "NA") %>% #drop ECX values
-  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
-  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
-  ggplot(aes(y = shape_f, x = dose.mg.L.master, fill = effect_f)) +
-  geom_boxplot(alpha = 0.8) +
-  scale_x_log10() +
-  labs(y = "Morphology", x = "mg/L", fill = "Effect") +
-  theme_classic() +
-  theme(text = element_text(size = 16))
+#Select particle characteristics
+selected_variables <- aoc_setup %>% 
+  dplyr::select(
+    size_f,
+    shape_f,
+    poly_f,
+    density.reported.estimated,
+    charge,
+    functional.group,
+    particle.behavior,
+    weather.biofoul) %>% 
+  filter(size_f != "Not Reported")
 
-plot(a)
+selected_variables$shape_f[selected_variables$shape_f == "Not Reported" ] <- NA
+selected_variables$poly_f[selected_variables$poly_f == "Not Reported" ] <- NA
+selected_variables$density.reported.estimated[selected_variables$density.reported.estimated == "estimated" ] <- NA
+selected_variables$particle.behavior[selected_variables$particle.behavior == "Not Evaluated" ] <- NA
+selected_variables$weather.biofoul[selected_variables$weather.biofoul == "N" ] <- NA
 
-b <- aoc_setup %>% 
-  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
-  filter(shape_f != "Not Reported") %>% #drop studies where shapes aren't reported
-  filter(effect_f != "NA") %>% #drop ECX values
-  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
-  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
-  ggplot(aes(y = shape_f, x = dose.particles.mL.master, fill = effect_f)) +
-  geom_boxplot(alpha = 0.8) +
-  scale_x_log10() +
-  labs(y = "Morphology", x = "particles/mL", fill = "Effect") +
-  theme_classic() +
-  theme(text = element_text(size = 16))
+CompletenessSummary_size <- selected_variables %>%
+  group_by(size_f) %>% 
+  summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
+  mutate(across(is.numeric,~round(., 2))) %>% 
+  mutate(across(is.numeric, ~100 *(1 -.))) %>% 
+  rename("Size Category" = size_f, "Shape" = shape_f, "Polymer" = poly_f, "Density" = density.reported.estimated,
+         "Charge" = charge, "Functional Group" = functional.group, "Particle Behavior" = particle.behavior,
+         "Weathering" = weather.biofoul)
 
-plot(b)
+#Heatmap
 
-c <- aoc_setup %>% 
-  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
-  filter(!poly_f %in% c("Not Reported", "Polyethylene Vinyl Acetate", "Polycarbonate", "Latex")) %>% 
-  filter(effect_f != "NA") %>% #drop ECX values
-  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
-  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
-  ggplot(aes(y = poly_f, x = dose.mg.L.master, fill = effect_f)) +
-  geom_boxplot(alpha = 0.8) +
-  scale_x_log10() +
-  labs(y = "Polymer", x = "mg/L", fill = "Effect") +
-  theme_classic() +
-  theme(text = element_text(size = 16)) +
-  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 10))
+#Transpose data
+transposed_size <- as.data.frame(t(as.matrix(CompletenessSummary_size[2:8]))) %>% #1 is category, 2-6 are variables
+  arrange('1nm < 100nm')
+#Reassign column names
+colnames(transposed_size) <- c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm")
+#Format as matrix
+MissingMatrix_size <- data.matrix(transposed_size)
+#build heatmap
+map <- pheatmap(MissingMatrix_size,
+         main = "Particle Characteristics Reported", #title
+         fontsize = 12,
+         angle_col = 0,
+         cluster_rows = FALSE, cluster_cols = FALSE,
+         display_numbers = TRUE,
+         fontsize_number = 12,
+         number_color = "black",
+         number_format = "%.0f%%",
+         treeheight_row = 0, treeheight_col = 0, 
+         color = c("#EEFFFF", "#BAFFFF", "#88E7FF", "#55B4CC", "#3B9AB2", "#228199"))
 
-plot(c)
+ggsave(filename = "Figure1.jpg",
+       path = "App_Nonessential/TeamX_Manuscript_Figures/",
+       plot = map, width = 8, height = 4, units = "in")
 
-d <- aoc_setup %>% 
-  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
-  filter(!poly_f %in% c("Not Reported", "Polyethylene Vinyl Acetate", "Polycarbonate", "Latex")) %>%
-  filter(effect_f != "NA") %>% #drop ECX values
-  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
-  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
-  ggplot(aes(y = poly_f, x = dose.particles.mL.master, fill = effect_f)) +
-  geom_boxplot(alpha = 0.8) +
-  scale_x_log10() +
-  labs(y = "Polymer", x = "particles/mL", fill = "Effect") +
-  theme_classic() +
-  theme(text = element_text(size = 16)) +
-  scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 10))
-
-plot(d)
 ```
 
-# Leah
+# Figure 2
 
@@
   ggplot(aes(x = year, y = freq, fill = fct_rev(dose_metric))) +
-  geom_bar(position = "dodge", stat = "identity") +
+  geom_bar(position = "dodge", stat = "identity", color = "black") +
+  theme_classic()+
+  scale_fill_manual(values = c("#3B9AB2", "#EBCC2A", "#F21A00")) +
+  scale_y_continuous(expand = c(0,0), limits = c(0,40)) +
+  labs(y = "Number of Studies", x = "Year", fill = "Concentration Metric") +
+  theme(text = element_text(size = 12), legend.position = "top")
+
+ggsave(filename = "Figure2.jpg",
+       path = "App_Nonessential/TeamX_Manuscript_Figures/",
+       plot = b, width = 12, height = 8, units = "in")
+
+```
+
+# Polymer Type Percentages
+
+```{r}
+
+a <- aoc_setup %>% 
+  group_by(doi, poly_f) %>% 
+  summarise() %>%  
+  ungroup() %>%
+  group_by(poly_f) %>% 
+  mutate(poly_count = n()) %>% 
+  ungroup() %>%
+  mutate(study_count = n_distinct(doi)) %>% 
+  mutate(pct_poly = poly_count/study_count*100)
+
+```
+
+# Figure 3
+
+```{r}
+
+a <- aoc_setup %>% 
+  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+  filter(shape_f != "Not Reported") %>% #drop studies where shapes aren't reported
+  filter(effect_f != "NA") %>% #drop ECX values
+  filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
+  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
+  ggplot(aes(y = shape_f, x = dose.particles.mL.master, fill = effect_f)) +
+  geom_boxplot(alpha = 0.8) +
+  scale_x_log10() +
+  labs(y = "Morphology", x = "Particles/mL", fill = "Effect") +
   theme_classic() +
-  scale_fill_manual(values = c("#3B9AB2", "#EBCC2A", "#F21A00")) +
-  scale_y_continuous(expand = c(0,0)) +
-  labs(y = "Frequency", x = "Year", fill = "Dose Metric") +
-  theme(text = element_text(size = 16))
+  theme(text = element_text(size = 12))+
+  scale_fill_manual(values = c("#3B9AB2", "#EBCC2A"))
+
+plot(a)
+
+ggsave(filename = "Figure3.jpg",
+       path = "App_Nonessential/TeamX_Manuscript_Figures/",
+       plot = a, width = 8, height = 4, units = "in")
+
+```
+
+# Figure 4
+
+```{r}
+
+#Panel A (count)
+a <- aoc_setup %>% 
+  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+  filter(size_f != "Not Reported") %>% #drop studies where shapes aren't reported
+  filter(effect_f != "NA") %>% #drop ECX values
+  drop_na(dose.mg.L.master|dose.particles.mL.master) %>%
+  ggplot(aes(y = size_f, x = dose.particles.mL.master, fill = effect_f)) +
+  geom_boxplot(alpha = 0.8) +
+  scale_x_log10() +
+  labs(y = "Size Category", x = "Particles/mL", fill = "Effect") +
+  theme_classic() +
+  theme(text = element_text(size = 12))+
+  scale_fill_manual(values = c("#3B9AB2", "#EBCC2A"))
+
+plot(a)
+
+#Panel B (mass)
+b <- aoc_setup %>% 
+  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+  filter(size_f != "Not Reported") %>% #drop studies where shapes aren't reported
+  filter(effect_f != "NA") %>% #drop ECX values
+  drop_na(dose.mg.L.master|dose.particles.mL.master) %>%
+  ggplot(aes(y = size_f, x = dose.mg.L.master, fill = effect_f)) +
+  geom_boxplot(alpha = 0.8) +
+  scale_x_log10() +
+  labs(y = "Size Category", x = "µg/mL", fill = "Effect") +
+  theme_classic() +
+  theme(text = element_text(size = 12))+
+  scale_fill_manual(values = c("#3B9AB2", "#EBCC2A"))
 
@@
 
-#Find studies that use multiple sizes
-c <- aoc_setup %>% 
-  group_by(doi, size.length.um.used.for.conversions) %>% 
-  summarise() %>% 
-  ungroup() %>% 
-  count(doi)
+ab <- ggarrange(a,b, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO")
 
-#Find studies that use multiple shapes
-d <- aoc_setup %>% 
-  group_by(doi, shape_f) %>% 
-  summarise() %>% 
-  ungroup() %>% 
-  count(doi)
+plot(ab)
 
-#Find studies that use multiple polymers
-e <- aoc_setup %>% 
-  group_by(doi, poly_f) %>% 
-  summarise() %>% 
-  ungroup() %>% 
-  count(doi)
+ggsave(filename = "Figure4.jpg",
+       path = "App_Nonessential/TeamX_Manuscript_Figures/",
+       plot = ab, width = 8, height = 8, units = "in")
+
+#Alternate Option(facet wrap)
+# c <- aoc_setup %>% 
+#   filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+#   filter(size_f != "Not Reported") %>% #drop studies where shapes aren't reported
+#   filter(effect_f != "NA") %>% #drop ECX values
+#   select(size_f, effect_f, dose.particles.mL.master, dose.mg.L.master) %>% 
+#   drop_na(dose.mg.L.master|dose.particles.mL.master) %>% 
+#   pivot_longer(cols = c(dose.particles.mL.master, dose.mg.L.master), names_to = "dose_metric", values_to = "dose") %>%   
+#   mutate(dose_metric = factor(case_when(
+#     dose_metric == "dose.particles.mL.master" ~ "particles/mL",
+#     dose_metric == "dose.mg.L.master" ~ "µg/mL"))) %>% 
+#   ggplot(aes(y = size_f, x = dose, fill = effect_f)) +
+#   geom_boxplot(alpha = 0.8) +
+#   scale_x_log10() +
+#   facet_wrap(~dose_metric, scales = "free_x", nrow = 2) +
+#   labs(y = "Size Category", x = "Dose", fill = "Effect") +
+#   theme_classic() +
+#   theme(text = element_text(size = 12))+
+#   scale_fill_manual(values = c("#3B9AB2", "#EBCC2A"))
+# 
+# plot(c)
+# 
+# ggsave(filename = "Figure4_Alt.jpg",
+#        path = "App_Nonessential/TeamX_Manuscript_Figures/",
+#        plot = c, width = 8, height = 8, units = "in")
+
 ```
+
+#Figure 5
+
+```{r ERM}
+###define sizes for filtering and alignment##
+# smaller size bin
+small_tier_lower_size <- 1 #um
+small_tier_upper_size <- 5000 #um
+upper.tissue.trans.size.um <- 100 #10 #um #set size for filtering data and x2M
+# larger size bin
+large_tier_lower_size <- 1 #um
+large_tier_upper_size <- 5000 #um
+```
+
+```{r}
+## First filter data with global filters
+aoc_intermediate <- aoc_z %>% 
+  filter(!environment %in% c("Terrestrial", "Not Reported"),
+         org_f != "Bacterium",
+         org_f != "Plant",
+         effect.metric != "HONEC",
+         tier_zero_tech_f == "Red Criteria Passed",
+         risk.13 != 0 #Drop studies that received a score of 0 for endpoints criteria (this also drops studies that have not yet been scored) - KEEP THIS AFTER THE RED CRITERIA FILTERS  
+         ) %>%  
+  mutate(max.size.ingest.um = 1000 * max.size.ingest.mm)
+```
+
+## Count
+
+```{r size category selection}
+#1 - 100 nanometers
+tier1_2_nano <- aoc_intermediate %>% 
+         mutate(dose_new = dose.particles.mL.master / (af.time * af.noec)) %>%  
+         drop_na(dose_new) %>% 
+         mutate(dose_new = dose_new * 1000) %>% 
+  filter(between(size.length.um.used.for.conversions, 0.001, 0.1))
+
+### 100 - 1000 nm
+tier1_2_sub <- aoc_intermediate %>% 
+         mutate(dose_new = dose.particles.mL.master / (af.time * af.noec)) %>%  
+         drop_na(dose_new) %>% 
+         mutate(dose_new = dose_new * 1000) %>% 
+  filter(between(size.length.um.used.for.conversions, 0.1, 1))
+
+### 1 - 100 um
+tier1_2_small <- aoc_intermediate %>% 
+         mutate(dose_new = dose.particles.mL.master / (af.time * af.noec)) %>%  
+         drop_na(dose_new) %>% 
+         mutate(dose_new = dose_new * 1000) %>% 
+  filter(between(size.length.um.used.for.conversions, 1, 100))
+
+### 100 - 1000 um
+tier1_2_large <- aoc_intermediate %>% 
+         mutate(dose_new = dose.particles.mL.master / (af.time * af.noec)) %>%  
+         drop_na(dose_new) %>% 
+         mutate(dose_new = dose_new * 1000) %>% 
+  filter(between(size.length.um.used.for.conversions, 100, 1000)) 
+```
+
+```{r data collapse}
+
+#1 - 100 nanometers
+tier1_2_nano_1q <- tier1_2_nano %>% 
+  group_by(Species, Group) %>% 
+  summarize(Conc = quantile(dose_new, 0.25))
+
+### 100 - 1000 nm
+tier1_2_sub_1q <- tier1_2_sub %>% 
+  group_by(Species, Group) %>% 
+  summarize(Conc = quantile(dose_new, 0.25))
+
+### 1 - 100 um
+tier1_2_small_1q <- tier1_2_small %>% 
+  group_by(Species, Group) %>% 
+  summarize(Conc = quantile(dose_new, 0.25))
+
+### 100 - 1000 um
+tier1_2_large_1q <- tier1_2_large %>% 
+  group_by(Species, Group) %>% 
+  summarize(Conc = quantile(dose_new, 0.25))
+```
+
+### SSD Build
+
+#### Nano
+```{r include=TRUE, warning=FALSE}
+###### --modelling ####
+dists_tier1_2_nano_1q <- ssd_fit_dists(tier1_2_nano_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE)
+
+#autoplot(dists_tier1_2_nano_1q) #plots the distribution in ggplotier2
+#### Goodness of fit ####
+gof_tier1_2_nano_1q <- as.data.frame(ssd_gof(dists_tier1_2_nano_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))
+
+#### Prediction ####
+set.seed(99)
+pred_tier1_2_nano_1q <- predict(dists_tier1_2_nano_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE)
+
+#order data
+SSD_tier1_2_nano_1q <- tier1_2_nano_1q[order(tier1_2_nano_1q$Conc), ]
+SSD_tier1_2_nano_1q$frac <- ppoints(tier1_2_nano_1q$Conc, 0.5)
+
+#add unique ID for multiplot
+pred_tier1_2_nano_1q_nano <- pred_tier1_2_nano_1q
+SSD_tier1_2_nano_1q_nano <- SSD_tier1_2_nano_1q
+
+tier1_2_nano_1q_lcl <- c(pred_tier1_2_nano_1q$lcl[5]) #CI95
+tier1_2_nano_1q_hc5 <- c(pred_tier1_2_nano_1q$est[5]) #HC5
+pred_tier1_2_nano_1q$est_format <-format(pred_tier1_2_nano_1q$est, digits = 3, scientific = TRUE)
+```
+
+#### Sub
+```{r include=TRUE, warning=FALSE}
+###### --modelling ####
+dists_tier1_2_sub_1q <- ssd_fit_dists(tier1_2_sub_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE)
+
+#autoplot(dists_tier1_2_sub_1q) #plots the distribution in ggplotier2
+#### Goodness of fit ####
+gof_tier1_2_sub_1q <- as.data.frame(ssd_gof(dists_tier1_2_sub_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))
+
+#### Prediction ####
+set.seed(99)
+pred_tier1_2_sub_1q <- predict(dists_tier1_2_sub_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE)
+
+#order data
+SSD_tier1_2_sub_1q <- tier1_2_sub_1q[order(tier1_2_sub_1q$Conc), ]
+SSD_tier1_2_sub_1q$frac <- ppoints(tier1_2_sub_1q$Conc, 0.5)
+
+#add unique ID for multiplot
+pred_tier1_2_sub_1q_sub <- pred_tier1_2_sub_1q
+SSD_tier1_2_sub_1q_sub <- SSD_tier1_2_sub_1q
+
+tier1_2_sub_1q_lcl <- c(pred_tier1_2_sub_1q$lcl[5]) #CI95
+tier1_2_sub_1q_hc5 <- c(pred_tier1_2_sub_1q$est[5]) #HC5
+pred_tier1_2_sub_1q$est_format <-format(pred_tier1_2_sub_1q$est, digits = 3, scientific = TRUE)
+```
+
+#### Small
+```{r include=TRUE, warning=FALSE}
+###### --modelling ####
+dists_tier1_2_small_1q <- ssd_fit_dists(tier1_2_small_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE)
+
+#autoplot(dists_tier1_2_small_1q) #plots the distribution in ggplotier2
+#### Goodness of fit ####
+gof_tier1_2_small_1q <- as.data.frame(ssd_gof(dists_tier1_2_small_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))
+
+#### Prediction ####
+set.seed(99)
+pred_tier1_2_small_1q <- predict(dists_tier1_2_small_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE)
+
+#order data
+SSD_tier1_2_small_1q <- tier1_2_small_1q[order(tier1_2_small_1q$Conc), ]
+SSD_tier1_2_small_1q$frac <- ppoints(tier1_2_small_1q$Conc, 0.5)
+
+#add unique ID for multiplot
+pred_tier1_2_small_1q_small <- pred_tier1_2_small_1q
+SSD_tier1_2_small_1q_small <- SSD_tier1_2_small_1q
+
+tier1_2_small_1q_lcl <- c(pred_tier1_2_small_1q$lcl[5]) #CI95
+tier1_2_small_1q_hc5 <- c(pred_tier1_2_small_1q$est[5]) #HC5
+pred_tier1_2_small_1q$est_format <-format(pred_tier1_2_small_1q$est, digits = 3, scientific = TRUE)
+```
+
+#### Large
+```{r include=TRUE, warning=FALSE}
+###### --modelling ####
+dists_tier1_2_large_1q <- ssd_fit_dists(tier1_2_large_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE)
+
+#autoplot(dists_tier1_2_large_1q) #plots the distribution in ggplotier2
+#### Goodness of fit ####
+gof_tier1_2_large_1q <- as.data.frame(ssd_gof(dists_tier1_2_large_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))
+
+#### Prediction ####
+set.seed(99)
+pred_tier1_2_large_1q <- predict(dists_tier1_2_large_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE)
+
+#order data
+SSD_tier1_2_large_1q <- tier1_2_large_1q[order(tier1_2_large_1q$Conc), ]
+SSD_tier1_2_large_1q$frac <- ppoints(tier1_2_large_1q$Conc, 0.5)
+
+#add unique ID for multiplot
+pred_tier1_2_large_1q_large <- pred_tier1_2_large_1q
+SSD_tier1_2_large_1q_large <- SSD_tier1_2_large_1q
+
+tier1_2_large_1q_lcl <- c(pred_tier1_2_large_1q$lcl[5]) #CI95
+tier1_2_large_1q_hc5 <- c(pred_tier1_2_large_1q$est[5]) #HC5
+pred_tier1_2_large_1q$est_format <-format(pred_tier1_2_large_1q$est, digits = 3, scientific = TRUE)
+```
+
+```{r}
+# Build dataframe with all SSDs so we can get a color scale
+p_t1_l_p <- pred_tier1_2_nano_1q_nano %>% mutate(size = "0.001 - 0.1 µm")
+p_t1_l_u <- pred_tier1_2_sub_1q_sub %>% mutate(size = "0.1 - 1 µm")
+p_t1_l_v <- pred_tier1_2_small_1q_small %>% mutate(size = "1 - 100 µm")
+p_t1_l_sa <- pred_tier1_2_large_1q_large %>% mutate(size = "100 - 1,000 µm")
+
+predictions_t1_large <- rbind.data.frame(p_t1_l_p, 
+                                p_t1_l_u,
+                                p_t1_l_v,
+                                p_t1_l_sa) %>% 
+  mutate(size = factor(size))
+predictions_t1_large
+
+#do the same for individual points
+s_t1_l_p <- SSD_tier1_2_nano_1q_nano %>% mutate(size = "0.001 - 0.1 µm")
+s_t1_l_u <- SSD_tier1_2_sub_1q_sub %>% mutate(size = "0.1 - 1 µm")
+s_t1_l_v <- SSD_tier1_2_small_1q_small %>% mutate(size = "1 - 100 µm")
+s_t1_l_sa <- SSD_tier1_2_large_1q_large %>% mutate(size = "100 - 1,000 µm")
+
+points_t1_large <- rbind.data.frame(s_t1_l_p, 
+                                s_t1_l_u,
+                                s_t1_l_v,
+                                s_t1_l_sa)
+
+points_t1_large$size <- as.factor(points_t1_large$size)
+```
+#### Plot
+```{r}
+
+count_plot <- predictions_t1_large %>% 
+ggplot(aes(x = est, y = percent/100, color = size)) +
+  geom_line(size = 1) +
+  geom_xribbon(aes(xmin = lcl, xmax = ucl, y = percent/100, fill = size),color = "black", alpha = 0.15) +
+  geom_point(data = points_t1_large,aes(x = Conc, y =frac, color = size)) + 
+  # scale_color_tron(name = "Size Bins") +
+  scale_fill_manual(values = c("#FF0000", "#00A08A", "#F2AD00", "#5BBCD6"),name = "Size Bins") +
+  scale_color_manual(values = c("#FF0000", "#00A08A", "#F2AD00", "#5BBCD6"),name = "Size Bins") +
+  # scale_fill_tron(name = "Size Bins") +
+  guides(fill = FALSE) + # suppress different legend for colors and lines
+  coord_trans(x = "log10") +
+      scale_x_continuous(limits = c(1e-5, 1e25),
+                         breaks = scales::trans_breaks(
+                           "log10", function(x) 10^x, n = 10),
+                         labels = trans_format(
+                           "log10", scales::math_format(10^.x))) +
+  scale_y_continuous("Species Affected (%)", 
+                     labels = scales::percent, 
+                     limits = c(0,1)) +
+  labs(x = "Particles/mL") +
+  theme_classic()+
+  theme(text = element_text(size = 12))
+
+plot(count_plot)
+```
+
+## Mass
+
+```{r size category selection}
+#1 - 100 nanometers
+tier1_2_nano <- aoc_intermediate %>% 
+         mutate(dose_new = dose.mg.L.master / (af.time * af.noec)) %>%  
+         drop_na(dose_new) %>% 
+         mutate(dose_new = dose_new * 1000) %>% 
+  filter(between(size.length.um.used.for.conversions, 0.001, 0.1))
+
+### 100 - 1000 nm
+tier1_2_sub <- aoc_intermediate %>% 
+         mutate(dose_new = dose.mg.L.master / (af.time * af.noec)) %>%  
+         drop_na(dose_new) %>% 
+         mutate(dose_new = dose_new * 1000) %>% 
+  filter(between(size.length.um.used.for.conversions, 0.1, 1))
+
+### 1 - 100 um
+tier1_2_small <- aoc_intermediate %>% 
+         mutate(dose_new = dose.mg.L.master / (af.time * af.noec)) %>%  
+         drop_na(dose_new) %>% 
+         mutate(dose_new = dose_new * 1000) %>% 
+  filter(between(size.length.um.used.for.conversions, 1, 100))
+
+### 100 - 1000 um
+tier1_2_large <- aoc_intermediate %>% 
+         mutate(dose_new = dose.mg.L.master / (af.time * af.noec)) %>%  
+         drop_na(dose_new) %>% 
+         mutate(dose_new = dose_new * 1000) %>% 
+  filter(between(size.length.um.used.for.conversions, 100, 1000)) 
+```
+
+```{r data collapse}
+
+#1 - 100 nanometers
+tier1_2_nano_1q <- tier1_2_nano %>% 
+  group_by(Species, Group) %>% 
+  summarize(Conc = quantile(dose_new, 0.25))
+
+### 100 - 1000 nm
+tier1_2_sub_1q <- tier1_2_sub %>% 
+  group_by(Species, Group) %>% 
+  summarize(Conc = quantile(dose_new, 0.25))
+
+### 1 - 100 um
+tier1_2_small_1q <- tier1_2_small %>% 
+  group_by(Species, Group) %>% 
+  summarize(Conc = quantile(dose_new, 0.25))
+
+### 100 - 1000 um
+tier1_2_large_1q <- tier1_2_large %>% 
+  group_by(Species, Group) %>% 
+  summarize(Conc = quantile(dose_new, 0.25))
+```
+
+### SSD Build
+
+#### Nano
+```{r include=TRUE, warning=FALSE}
+###### --modelling ####
+dists_tier1_2_nano_1q <- ssd_fit_dists(tier1_2_nano_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE)
+
+#autoplot(dists_tier1_2_nano_1q) #plots the distribution in ggplotier2
+#### Goodness of fit ####
+gof_tier1_2_nano_1q <- as.data.frame(ssd_gof(dists_tier1_2_nano_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))
+
+#### Prediction ####
+set.seed(99)
+pred_tier1_2_nano_1q <- predict(dists_tier1_2_nano_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE)
+
+#order data
+SSD_tier1_2_nano_1q <- tier1_2_nano_1q[order(tier1_2_nano_1q$Conc), ]
+SSD_tier1_2_nano_1q$frac <- ppoints(tier1_2_nano_1q$Conc, 0.5)
+
+#add unique ID for multiplot
+pred_tier1_2_nano_1q_nano <- pred_tier1_2_nano_1q
+SSD_tier1_2_nano_1q_nano <- SSD_tier1_2_nano_1q
+
+tier1_2_nano_1q_lcl <- c(pred_tier1_2_nano_1q$lcl[5]) #CI95
+tier1_2_nano_1q_hc5 <- c(pred_tier1_2_nano_1q$est[5]) #HC5
+pred_tier1_2_nano_1q$est_format <-format(pred_tier1_2_nano_1q$est, digits = 3, scientific = TRUE)
+```
+
+#### Sub
+```{r include=TRUE, warning=FALSE}
+###### --modelling ####
+dists_tier1_2_sub_1q <- ssd_fit_dists(tier1_2_sub_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE)
+
+#autoplot(dists_tier1_2_sub_1q) #plots the distribution in ggplotier2
+#### Goodness of fit ####
+gof_tier1_2_sub_1q <- as.data.frame(ssd_gof(dists_tier1_2_sub_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))
+
+#### Prediction ####
+set.seed(99)
+pred_tier1_2_sub_1q <- predict(dists_tier1_2_sub_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE)
+
+#order data
+SSD_tier1_2_sub_1q <- tier1_2_sub_1q[order(tier1_2_sub_1q$Conc), ]
+SSD_tier1_2_sub_1q$frac <- ppoints(tier1_2_sub_1q$Conc, 0.5)
+
+#add unique ID for multiplot
+pred_tier1_2_sub_1q_sub <- pred_tier1_2_sub_1q
+SSD_tier1_2_sub_1q_sub <- SSD_tier1_2_sub_1q
+
+tier1_2_sub_1q_lcl <- c(pred_tier1_2_sub_1q$lcl[5]) #CI95
+tier1_2_sub_1q_hc5 <- c(pred_tier1_2_sub_1q$est[5]) #HC5
+pred_tier1_2_sub_1q$est_format <-format(pred_tier1_2_sub_1q$est, digits = 3, scientific = TRUE)
+```
+
+#### Small
+```{r include=TRUE, warning=FALSE}
+###### --modelling ####
+dists_tier1_2_small_1q <- ssd_fit_dists(tier1_2_small_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE)
+
+#autoplot(dists_tier1_2_small_1q) #plots the distribution in ggplotier2
+#### Goodness of fit ####
+gof_tier1_2_small_1q <- as.data.frame(ssd_gof(dists_tier1_2_small_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))
+
+#### Prediction ####
+set.seed(99)
+pred_tier1_2_small_1q <- predict(dists_tier1_2_small_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE)
+
+#order data
+SSD_tier1_2_small_1q <- tier1_2_small_1q[order(tier1_2_small_1q$Conc), ]
+SSD_tier1_2_small_1q$frac <- ppoints(tier1_2_small_1q$Conc, 0.5)
+
+#add unique ID for multiplot
+pred_tier1_2_small_1q_small <- pred_tier1_2_small_1q
+SSD_tier1_2_small_1q_small <- SSD_tier1_2_small_1q
+
+tier1_2_small_1q_lcl <- c(pred_tier1_2_small_1q$lcl[5]) #CI95
+tier1_2_small_1q_hc5 <- c(pred_tier1_2_small_1q$est[5]) #HC5
+pred_tier1_2_small_1q$est_format <-format(pred_tier1_2_small_1q$est, digits = 3, scientific = TRUE)
+```
+
+#### Large
+```{r include=TRUE, warning=FALSE}
+###### --modelling ####
+dists_tier1_2_large_1q <- ssd_fit_dists(tier1_2_large_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE)
+
+#autoplot(dists_tier1_2_large_1q) #plots the distribution in ggplotier2
+#### Goodness of fit ####
+gof_tier1_2_large_1q <- as.data.frame(ssd_gof(dists_tier1_2_large_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))
+
+#### Prediction ####
+set.seed(99)
+pred_tier1_2_large_1q <- predict(dists_tier1_2_large_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE)
+
+#order data
+SSD_tier1_2_large_1q <- tier1_2_large_1q[order(tier1_2_large_1q$Conc), ]
+SSD_tier1_2_large_1q$frac <- ppoints(tier1_2_large_1q$Conc, 0.5)
+
+#add unique ID for multiplot
+pred_tier1_2_large_1q_large <- pred_tier1_2_large_1q
+SSD_tier1_2_large_1q_large <- SSD_tier1_2_large_1q
+
+tier1_2_large_1q_lcl <- c(pred_tier1_2_large_1q$lcl[5]) #CI95
+tier1_2_large_1q_hc5 <- c(pred_tier1_2_large_1q$est[5]) #HC5
+pred_tier1_2_large_1q$est_format <-format(pred_tier1_2_large_1q$est, digits = 3, scientific = TRUE)
+```
+
+```{r}
+# Build dataframe with all SSDs so we can get a color scale
+p_t1_l_p <- pred_tier1_2_nano_1q_nano %>% mutate(size = "0.001 - 0.1 µm")
+p_t1_l_u <- pred_tier1_2_sub_1q_sub %>% mutate(size = "0.1 - 1 µm")
+p_t1_l_v <- pred_tier1_2_small_1q_small %>% mutate(size = "1 - 100 µm")
+p_t1_l_sa <- pred_tier1_2_large_1q_large %>% mutate(size = "100 - 1,000 µm")
+
+predictions_t1_large <- rbind.data.frame(p_t1_l_p, 
+                                p_t1_l_u,
+                                p_t1_l_v,
+                                p_t1_l_sa) %>% 
+  mutate(size = factor(size))
+predictions_t1_large
+
+#do the same for individual points
+s_t1_l_p <- SSD_tier1_2_nano_1q_nano %>% mutate(size = "0.001 - 0.1 µm")
+s_t1_l_u <- SSD_tier1_2_sub_1q_sub %>% mutate(size = "0.1 - 1 µm")
+s_t1_l_v <- SSD_tier1_2_small_1q_small %>% mutate(size = "1 - 100 µm")
+s_t1_l_sa <- SSD_tier1_2_large_1q_large %>% mutate(size = "100 - 1,000 µm")
+
+points_t1_large <- rbind.data.frame(s_t1_l_p, 
+                                s_t1_l_u,
+                                s_t1_l_v,
+                                s_t1_l_sa)
+
+points_t1_large$size <- as.factor(points_t1_large$size)
+```
+#### Plot
+```{r}
+
+mass_plot <- predictions_t1_large %>% 
+ggplot(aes(x = est, y = percent/100, color = size)) +
+  geom_line(size = 1) +
+  geom_xribbon(aes(xmin = lcl, xmax = ucl, y = percent/100, fill = size),color = "black", alpha = 0.15) +
+  geom_point(data = points_t1_large,aes(x = Conc, y =frac, color = size)) +
+  scale_fill_manual(values = c("#FF0000", "#00A08A", "#F2AD00", "#5BBCD6"),name = "Size Bins") +
+  scale_color_manual(values = c("#FF0000", "#00A08A", "#F2AD00", "#5BBCD6"),name = "Size Bins") +
+  guides(fill = FALSE) + # suppress different legend for colors and lines
+  coord_trans(x = "log10") +
+      scale_x_continuous(limits = c(1e-15, 1e15),
+                         breaks = scales::trans_breaks(
+                           "log10", function(x) 10^x, n = 10),
+                         labels = trans_format(
+                           "log10", scales::math_format(10^.x))) +
+  scale_y_continuous("Species Affected (%)", 
+                     labels = scales::percent, 
+                     limits = c(0,1)) +
+  labs(x = "µg/mL") +
+  theme_classic()+
+  theme(text = element_text(size = 12))
+
+plot(mass_plot)
+```
+
+## Combine Panels
+```{r}
+
+count_mass_ssd <- ggarrange(count_plot,mass_plot, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO")
+
+plot(count_mass_ssd)
+
+ggsave(filename = "Figure5.jpg",
+       path = "App_Nonessential/TeamX_Manuscript_Figures/",
+       plot = count_mass_ssd, width = 8, height = 8, units = "in")
+
+```
+
+#Figure 7
+
+```{r}
+
+Group_size <- aoc_setup %>% 
+  ggplot(aes(x = size.length.um.used.for.conversions, fill = org_f)) +
+  geom_histogram(bins = 20, color = "black") +
+  scale_fill_manual(values = c('#a50026','#d73027','#f46d43','#fdae61','#fee090','#ffffbf','#e0f3f8','#abd9e9','#74add1','#4575b4','#313695', '#641e16')) +
+  ylab("Number of Observations")+
+  xlab("Particle Length (µm)") +
+  theme_classic() +
+  theme(text = element_text(size = 12)) +
+  scale_x_log10()+
+  labs(fill = "Organism Group")
+
+plot(Group_size)
+
+Endpoint_size <- aoc_setup %>% 
+  ggplot(aes(x = size.length.um.used.for.conversions, fill = lvl1_f)) +
+  geom_histogram(bins = 20, color = "black") +
+  scale_fill_manual(values = c('#d73027','#f46d43','#fdae61','#fee090','#ffffbf','#e0f3f8','#abd9e9','#74add1','#4575b4')) +
+  xlab("Particle Length (µm)") +
+  ylab("Number of Observations")+
+  theme_classic() +
+  theme(text = element_text(size = 12)) +
+  scale_x_log10()+
+  labs(fill = "Endpoint Category")
+
+plot(Endpoint_size)
+
+size_group_endpoint <- ggarrange(Group_size,Endpoint_size, nrow = 2, legend = "right", labels = "AUTO")
+
+plot(size_group_endpoint)
+
+ggsave(filename = "Figure7.jpg",
+       path = "App_Nonessential/TeamX_Manuscript_Figures/",
+       plot = size_group_endpoint, width = 8, height = 8, units = "in")
+
+```

### 2022-01-08 App_Nonessential/TeamX_Manuscript_Figures/Manuscript_Figures_TeamX.Rmd
changed colors to match

@@
          treeheight_row = 0, treeheight_col = 0, 
-         color = c("#EEFFFF", "#BAFFFF", "#88E7FF", "#55B4CC", "#3B9AB2", "#228199"))
+         color = c('#d73027','#fc8d59','#fee090','#e0f3f8','#91bfdb','#4575b4'))
 
@@
   theme_classic()+
-  scale_fill_manual(values = c("#3B9AB2", "#EBCC2A", "#F21A00")) +
+  scale_fill_manual(values = c('#d73027','#fee090','#4575b4')) +
   scale_y_continuous(expand = c(0,0), limits = c(0,40)) +
   labs(y = "Number of Studies", x = "Year", fill = "Concentration Metric") +
-  theme(text = element_text(size = 12), legend.position = "top")
+  theme(text = element_text(size = 20), legend.position = "top")
 
@@
   theme(text = element_text(size = 12))+
-  scale_fill_manual(values = c("#3B9AB2", "#EBCC2A"))
+  scale_fill_manual(values = c("#fee090", '#4575b4'))
 
@@
   # scale_color_tron(name = "Size Bins") +
-  scale_fill_manual(values = c("#FF0000", "#00A08A", "#F2AD00", "#5BBCD6"),name = "Size Bins") +
-  scale_color_manual(values = c("#FF0000", "#00A08A", "#F2AD00", "#5BBCD6"),name = "Size Bins") +
+  scale_fill_manual(values = c('#d7191c','#fdae61','#abd9e9','#2c7bb6'),name = "Size Bins") +
+  scale_color_manual(values = c('#d7191c','#fdae61','#abd9e9','#2c7bb6'),name = "Size Bins") +
   # scale_fill_tron(name = "Size Bins") +
@@
   geom_point(data = points_t1_large,aes(x = Conc, y =frac, color = size)) +
-  scale_fill_manual(values = c("#FF0000", "#00A08A", "#F2AD00", "#5BBCD6"),name = "Size Bins") +
-  scale_color_manual(values = c("#FF0000", "#00A08A", "#F2AD00", "#5BBCD6"),name = "Size Bins") +
+  scale_fill_manual(values = c('#d7191c','#fdae61','#abd9e9','#2c7bb6'),name = "Size Bins") +
+  scale_color_manual(values = c('#d7191c','#fdae61','#abd9e9','#2c7bb6'),name = "Size Bins") +
   guides(fill = FALSE) + # suppress different legend for colors and lines

### 2022-01-11 App_Nonessential/TeamX_Manuscript_Figures/Manuscript_Figures_TeamX.Rmd
reformat figure 7 x axis

@@
   theme_classic() +
-  theme(text = element_text(size = 12)) +
-  scale_x_log10()+
+  theme(text = element_text(size = 12),
+        axis.text.x = element_text(margin = margin(t = 10))) +
+  scale_x_log10(labels = scales::label_number(accuracy = 1))+
+  annotation_logticks(sides = "b", outside = TRUE)+
+  coord_cartesian(clip = "off")+
+  scale_y_continuous(expand = c(0,0))+
   labs(fill = "Organism Group")
@@
   theme_classic() +
-  theme(text = element_text(size = 12)) +
-  scale_x_log10()+
+  theme(text = element_text(size = 12),
+        axis.text.x = element_text(margin = margin(t = 10))) +
+  scale_x_log10(labels = scales::label_number(accuracy = 1))+
+  annotation_logticks(sides = "b", outside = TRUE)+
+  coord_cartesian(clip = "off")+
+  scale_y_continuous(expand = c(0,0))+
   labs(fill = "Endpoint Category")

### 2022-05-03 App_Nonessential/TeamX_Manuscript_Figures/Manuscript_Figures_TeamX.Rmd

--- 2022-01-08 App_Nonessential/TeamX_Manuscript_Figures/Manuscript_Figures_TeamX.Rmd
@@
 library(scales)
-
+library(DT)
 library(ggdark)
@@
 
-ggsave(filename = "Figure1.jpg",
-       path = "App_Nonessential/TeamX_Manuscript_Figures/",
-       plot = map, width = 8, height = 4, units = "in")
+# ggsave(filename = "Figure1.jpg",
+#        path = "App_Nonessential/TeamX_Manuscript_Figures/",
+#        plot = map, width = 8, height = 4, units = "in")
 
@@
   ggplot(aes(x = year, y = freq, fill = fct_rev(dose_metric))) +
-  geom_bar(position = "dodge", stat = "identity", color = "black") +
+  geom_bar(stat = "identity", color = "black") +
   theme_classic()+
   scale_fill_manual(values = c('#d73027','#fee090','#4575b4')) +
-  scale_y_continuous(expand = c(0,0), limits = c(0,40)) +
-  labs(y = "Number of Studies", x = "Year", fill = "Concentration Metric") +
+  scale_y_continuous(expand = c(0,0)) +
+  labs(y = "Number of Test Concentrations Reported", x = "Year", fill = "Concentration Metric") +
   theme(text = element_text(size = 20), legend.position = "top")
@@
 
+#Studies and measurements labels
 a <- aoc_setup %>% 
+      drop_na(dose.particles.mL.master) %>% #Drop data not reported in counts
+      drop_na(effect_f) %>% #drop ECX values
+      filter(shape_f != "Not Reported") %>% #drop studies where shapes aren't reported
+      group_by(shape_f, effect_f) %>% # need to include so there's a recognized "y"
+      summarize(dose.particles.mL.master= quantile(dose.particles.mL.master, .1), # need for recognized "x"
+                measurements = n(),
+                studies = n_distinct(doi))
+
+#Select data
+b <- aoc_setup %>%
+  drop_na(dose.particles.mL.master) %>%
   filter(environment != "Terrestrial") %>% #drop terrestrial organisms
@@
   filter(org_f %in% c("Crustacea", "Fish", "Mollusca")) %>%  #Select relevant organism groups
-  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction")) %>% 
+  filter(lvl2_f %in% c("Growth", "Body Condition", "Development", "Mortality", "Reproduction"))
+
+#Plot
+c <- b %>% 
   ggplot(aes(y = shape_f, x = dose.particles.mL.master, fill = effect_f)) +
   geom_boxplot(alpha = 0.8) +
-  scale_x_log10() +
+  geom_text(data = a, 
+                aes(label = paste("(",measurements,",",studies,")"),
+                  y = shape_f,
+                  x = Inf,
+                  hjust = 1),
+                position = position_dodge(.9),
+                size = 3)+
+  scale_x_log10(limits = c(-1,10000000000000000)) +
   labs(y = "Morphology", x = "Particles/mL", fill = "Effect") +
@@
 
-plot(a)
+plot(c)
 
@@
        path = "App_Nonessential/TeamX_Manuscript_Figures/",
-       plot = a, width = 8, height = 4, units = "in")
+       plot = c, width = 8, height = 4, units = "in")
+
+#Table
+
+d <- b %>%
+  select(doi, authors, year, species_f, life_f, shape_f, lvl1_f, lvl2_f, lvl3_f, dose.particles.mL.master, effect_f) %>% 
+  rename("DOI" = doi, "Authors" = authors, "Publication Year" = year, 
+         "Species" = species_f, "Life Stage" = life_f, "Shape" = shape_f, 
+         "Broad Endpoint Category" = lvl1_f, "Specific Endpoint Category" = lvl2_f,
+         "Endpoint" = lvl3_f, "Concentration (Particles/mL)" = dose.particles.mL.master, 
+         "Effect Detected?" = effect_f)
+
+write.csv(d, "App_Nonessential/TeamX_Manuscript_Figures/Figure3_Data_SI.csv", row.names = FALSE)
 
@@
 #Panel A (count)
+
+#Studies and measurements labels
+a_labs <- aoc_setup %>% 
+      drop_na(dose.mg.L.master|dose.particles.mL.master) %>%
+      drop_na(effect_f) %>% #drop ECX values
+      filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+      filter(size_f != "Not Reported") %>% #drop studies where shapes aren't reported
+      filter(effect_f != "NA") %>% #drop ECX values
+      group_by(size_f, effect_f) %>% # need to include so there's a recognized "y"
+      summarize(dose.particles.mL.master= quantile(dose.particles.mL.master, .1), # need for recognized "x"
+                measurements = n(),
+                studies = n_distinct(doi))
+
 a <- aoc_setup %>% 
@@
   geom_boxplot(alpha = 0.8) +
-  scale_x_log10() +
+  geom_text(data = a_labs, 
+                aes(label = paste("(",measurements,",",studies,")"),
+                  y = size_f,
+                  x = Inf,
+                  hjust = 1),
+                position = position_dodge(.9),
+                size = 3)+
+  scale_x_log10(limits = c(-1,1E25)) +
   labs(y = "Size Category", x = "Particles/mL", fill = "Effect") +
@@
 #Panel B (mass)
+
+#Studies and measurements labels
+b_labs <- aoc_setup %>% 
+      drop_na(dose.mg.L.master|dose.particles.mL.master) %>%
+      drop_na(effect_f) %>% #drop ECX values
+      filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+      filter(size_f != "Not Reported") %>% #drop studies where shapes aren't reported
+      filter(effect_f != "NA") %>% #drop ECX values
+      group_by(size_f, effect_f) %>% # need to include so there's a recognized "y"
+      summarize(dose.mg.L.master= quantile(dose.mg.L.master, .1), # need for recognized "x"
+                measurements = n(),
+                studies = n_distinct(doi))
+
 b <- aoc_setup %>% 
@@
   geom_boxplot(alpha = 0.8) +
-  scale_x_log10() +
+  geom_text(data = b_labs, 
+                aes(label = paste("(",measurements,",",studies,")"),
+                  y = size_f,
+                  x = Inf,
+                  hjust = 1),
+                position = position_dodge(.9),
+                size = 3)+
+  scale_x_log10(limits = c(-1,1E25)) +
   labs(y = "Size Category", x = "µg/mL", fill = "Effect") +
@@
        plot = ab, width = 8, height = 8, units = "in")
+
+#Table
+
+d <- aoc_setup %>%
+  filter(environment != "Terrestrial") %>% #drop terrestrial organisms
+  filter(size_f != "Not Reported") %>% #drop studies where shapes aren't reported
+  filter(effect_f != "NA") %>% #drop ECX values
+  drop_na(dose.mg.L.master|dose.particles.mL.master) %>%
+  select(doi, authors, year, species_f, life_f, size_f, lvl1_f, lvl2_f, lvl3_f, 
+         dose.particles.mL.master, dose.mg.L.master, effect_f) %>% 
+  rename("DOI" = doi, "Authors" = authors, "Publication Year" = year, 
+         "Species" = species_f, "Life Stage" = life_f, "Size Category" = size_f, 
+         "Broad Endpoint Category" = lvl1_f, "Specific Endpoint Category" = lvl2_f,
+         "Endpoint" = lvl3_f, "Concentration (Particles/mL)" = dose.particles.mL.master, 
+         "Concentration (µg/mL)" = dose.mg.L.master,
+         "Effect Detected?" = effect_f)
+
+write.csv(d, "App_Nonessential/TeamX_Manuscript_Figures/Figure4_Data_SI.csv", row.names = FALSE)
 
@@
   mutate(max.size.ingest.um = 1000 * max.size.ingest.mm)
+```
+
+```{r}
+
+#Data table for export
+
+ssd_table <- aoc_intermediate %>% 
+  mutate(dose_filter = ifelse(is.na(dose.particles.mL.master)&is.na(dose.mg.L.master), "N", "Y")) %>% 
+  filter(dose_filter == "Y") %>% 
+  select(doi, authors, year, species_f, life_f, size_f, lvl1_f, lvl2_f, lvl3_f, 
+         dose.particles.mL.master, dose.mg.L.master, effect.metric) %>% 
+  rename("DOI" = doi, "Authors" = authors, "Publication Year" = year, 
+         "Species" = species_f, "Life Stage" = life_f, "Size Category" = size_f, 
+         "Broad Endpoint Category" = lvl1_f, "Specific Endpoint Category" = lvl2_f,
+         "Endpoint" = lvl3_f, "Concentration (Particles/mL)" = dose.particles.mL.master, 
+         "Concentration (µg/mL)" = dose.mg.L.master, "Effect Metric" = effect.metric)
+
+write.csv(ssd_table, "App_Nonessential/TeamX_Manuscript_Figures/Figure5_Data_SI.csv", row.names = FALSE)
+
 ```
@@
 
+#Group size plot
 Group_size <- aoc_setup %>% 
-  ggplot(aes(x = size.length.um.used.for.conversions, fill = org_f)) +
-  geom_histogram(bins = 20, color = "black") +
-  scale_fill_manual(values = c('#a50026','#d73027','#f46d43','#fdae61','#fee090','#ffffbf','#e0f3f8','#abd9e9','#74add1','#4575b4','#313695', '#641e16')) +
+  filter(org_f != "Mixed") %>% 
+  filter(size_f != "Not Reported") %>% 
+  group_by(org_f, size_f) %>% 
+  mutate(count = n()) %>% 
+  ungroup() %>% 
+  group_by(org_f, size_f, count) %>% 
+  summarise() %>% 
+  ggplot(aes(x = reorder(org_f, count), y = count, fill = size_f)) +
+  geom_col(position = position_stack(reverse = TRUE), color = "black") +
+  scale_fill_manual(values = c('#d73027','#fc8d59','#fee090','#e0f3f8','#4575b4')) +
   ylab("Number of Observations")+
-  xlab("Particle Length (µm)") +
+  xlab("Organism Group") +
   theme_classic() +
-  theme(text = element_text(size = 12)) +
-  scale_x_log10()+
-  labs(fill = "Organism Group")
+  theme(text = element_text(size = 12),
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  guides(fill = guide_legend(reverse = TRUE))+
+  scale_y_continuous(expand = c(0,0))+
+  labs(fill = "Size Category")
 
@@
 
+#Group size table
+Group_size_table <- aoc_setup %>%
+  filter(org_f != "Mixed") %>% 
+  filter(size_f != "Not Reported") %>% 
+  group_by(org_f, size_f) %>% 
+  mutate(count = n()) %>% 
+  ungroup() %>% 
+  group_by(org_f, size_f, count) %>% 
+  summarise() %>%
+  rename("Organism Group" = org_f, "Size Cateogry" = size_f, "Number of Observations" = count)
+
+write.csv(Group_size_table, "App_Nonessential/TeamX_Manuscript_Figures/Figure7A_Data_SI.csv", row.names = FALSE)
+
+#Endpoint size plot
 Endpoint_size <- aoc_setup %>%
-  ggplot(aes(x = size.length.um.used.for.conversions, fill = lvl1_f)) +
-  geom_histogram(bins = 20, color = "black") +
-  scale_fill_manual(values = c('#d73027','#f46d43','#fdae61','#fee090','#ffffbf','#e0f3f8','#abd9e9','#74add1','#4575b4')) +
-  xlab("Particle Length (µm)") +
+  mutate(lvl1_f = ifelse(lvl1_f == "Behavioral, Sensory, Neurological", "Behavioral, Neurological", paste(lvl1_f))) %>% 
+  filter(size_f != "Not Reported") %>% 
+  group_by(lvl1_f, size_f) %>% 
+  mutate(count = n()) %>% 
+  ungroup() %>% 
+  group_by(lvl1_f, size_f, count) %>% 
+  summarise() %>% 
+  ggplot(aes(x = reorder(lvl1_f, count), y = count, fill = size_f)) +
+  geom_col(position = position_stack(reverse = TRUE), color = "black") +
+  scale_fill_manual(values = c('#d73027','#fc8d59','#fee090','#e0f3f8','#4575b4')) +
   ylab("Number of Observations")+
+  xlab("Endpoint Category") +
   theme_classic() +
-  theme(text = element_text(size = 12)) +
-  scale_x_log10()+
-  labs(fill = "Endpoint Category")
+  theme(text = element_text(size = 12),
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  guides(fill = guide_legend(reverse = TRUE))+
+  scale_y_continuous(expand = c(0,0))+
+  labs(fill = "Size Category")+
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
 
 plot(Endpoint_size)
+
+#Endpoint size table
+Endpoint_size_table <- aoc_setup %>%
+  mutate(lvl1_f = ifelse(lvl1_f == "Behavioral, Sensory, Neurological", "Behavioral, Neurological", paste(lvl1_f))) %>% 
+  filter(size_f != "Not Reported") %>% 
+  group_by(lvl1_f, size_f) %>% 
+  mutate(count = n()) %>% 
+  ungroup() %>% 
+  group_by(lvl1_f, size_f, count) %>% 
+  summarise() %>%
+  rename("Endpoint Category" = lvl1_f, "Size Cateogry" = size_f, "Number of Observations" = count)
+
+write.csv(Endpoint_size_table, "App_Nonessential/TeamX_Manuscript_Figures/Figure7B_Data_SI.csv", row.names = FALSE)
 

### 2022-05-03 app.R
Added data submission tab and updated contact email

@@ -265,6 +265,7 @@ ui <- dashboardPage(
                      menuItem("Calculators", tabName = "Calculators", icon = icon("calculator")),
                      menuItem("Predictions", tabName = "Predictions", icon = icon("brain")),
                      menuItem("Resources", tabName = "Resources", icon = icon("question-circle")),
+                     menuItem("Data Submission", tabName = "Submission", icon = icon("fas fa-file-upload")),
                      menuItem("Contact", tabName = "Contact", icon = icon("envelope")),
                      br(),
                      br(),
@@ -1918,14 +1919,32 @@ tabItem(tabName = "Resources",
          
         ), #close tab
 
+#### Data Submission UI ####
+
+tabItem(tabName = "Submission", 
+        
+        box(title = "Data Submission", width = 6, status = "primary",
+            p("To submit new data to ToMEx, download the data submission template using the link below. Complete the submission template
+              using the embedded descriptions and the ", 
+              a(href ="https://sccwrp-my.sharepoint.com/:b:/g/personal/leahth_sccwrp_org/EeyE7n7JZdJPi_EYUD_D-dsBxNv5qlBtzwihmr9SbxH_Og?e=Crfu6Z",
+                'Data Category Descriptions', .noOWs = "outside"),
+              ". Once the data submission template is completed, use the submission form to complete the data upload."),
+            br(),
+            p("For questions regarding data submission or to check to see if data from a specific study has already been uploaded to ToMEx, please email tomex@sccwrp.org"),
+            br(),
+            p(align = "center",a(href = "https://sccwrp-my.sharepoint.com/:b:/g/personal/leahth_sccwrp_org/EeyE7n7JZdJPi_EYUD_D-dsBxNv5qlBtzwihmr9SbxH_Og?e=Crfu6Z", 'Data Submission Template')),
+            br(),
+            p(align = "center",a(href = "https://sccwrp-my.sharepoint.com/:b:/g/personal/leahth_sccwrp_org/EZ0r1AdQqsJGnPuvDJFNyxMBQ60ibEbAiCqrjNqRtlX7gg?e=18fTUr", 'Submission Form'))),
+        
+        ), #close tab
+
 #### Contact UI ####
 
 tabItem(tabName = "Contact", 
          
-        box(title = "Contact", width = 12, status = "primary",
-         p("For scientific questions, please contact Dr. Leah Thornton Hampton (leahth@sccwrp.org)."),
-         br(),
[…]

### 2022-05-05 app.R
Added data template and submission form links

@@ -1942,9 +1942,9 @@ tabItem(tabName = "Submission",
             br(),
             p("For questions regarding data submission or to check to see if data from a specific study has already been uploaded to ToMEx, please email tomex@sccwrp.org"),
             br(),
-            p(align = "center",a(href = "https://sccwrp-my.sharepoint.com/:b:/g/personal/leahth_sccwrp_org/EeyE7n7JZdJPi_EYUD_D-dsBxNv5qlBtzwihmr9SbxH_Og?e=Crfu6Z", 'Data Submission Template')),
+            p(align = "center",a(href = "https://sccwrp-my.sharepoint.com/:x:/g/personal/leahth_sccwrp_org/EZ0Gvnn4BkdHsQ2VKVEQQnkBFToXfPW9r4qNX-qcZ5z_LA?e=eISa57", 'Data Submission Template')),
             br(),
-            p(align = "center",a(href = "https://sccwrp-my.sharepoint.com/:b:/g/personal/leahth_sccwrp_org/EZ0r1AdQqsJGnPuvDJFNyxMBQ60ibEbAiCqrjNqRtlX7gg?e=18fTUr", 'Submission Form'))),
+            p(align = "center",a(href = "https://forms.office.com/r/7syA1rnNgH", 'Submission Form'))),
         
         ), #close tab
 

### 2022-05-05 app.R
Increasing height of collapsible tree

@@ -484,7 +484,7 @@ tabItem(tabName = "Overview",
           
           column(width = 12,
           #collapsible tree plot
-          collapsibleTree::collapsibleTreeOutput("plot", height = "400px"),
+          collapsibleTree::collapsibleTreeOutput("plot", height = "700px"),
           
           ), #closes out column
           

### 2022-05-05 app.R
increased height on collapsible tree

@@ -485,7 +485,7 @@ tabItem(tabName = "Overview",
           
           column(width = 12,
           #collapsible tree plot
-          collapsibleTree::collapsibleTreeOutput("plot", height = "400px"),
+          collapsibleTree::collapsibleTreeOutput("plot", height = "700px"),
           
           ), #closes out column
           
